### 2016-05-25 R/ITHIM.Rmd
Set eval = FALSE for archived code

## Integrated Transport and Health Impacts Model (ITHIM)

This document will eventually serve as a compendium to the R package
_ITHIM_.  It will describe how to use the package to implement the
ITHIM model, as well as a detailed description of the model itself.

As a working document it will serve as a place for SGY to post
results, questions, interpreations, notes, etc.

```{r date}
date()
```

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE}
options(repos='http://cran.rstudio.com/')
library("ggplot2")
library("reshape2")
library("plyr")
library("ITHIM")
```

## R Implementation

```{r instantiate}
ITHIM <- list(parameters = parList <- createParameterList(), means = computeMeanMatrices(parList))
```
Then we compute quintiles for active transport time.
```{r quints}
ITHIM <- within(ITHIM, {
quintilesActiveTransportTime <- computeQuintiles(means$meanActiveTransportTime, means$sdActiveTransportTime)
})
```

Then the METs by walking and cycling.
```{r quintMETs}
ITHIM <- within(ITHIM, {
  quintWalkingTime = list(M = quintilesActiveTransportTime[["M"]] * (1-means$propTimeCycling[,"M"]), F = quintilesActiveTransportTime[["F"]] * (1-means$propTimeCycling[,"F"]))
  quintCyclingTime = list(M = quintilesActiveTransportTime[["M"]] * (means$propTimeCycling[,"M"]), F = quintilesActiveTransportTime[["F"]] * (means$propTimeCycling[,"F"]))
  quintWalkingMET = list(M = means$meanWalkMET[,"M"]*quintWalkingTime[["M"]]/60, F = means$meanWalkMET[,"F"]*quintWalkingTime[["F"]]/60)
  quintCyclingMET = list(M = means$meanCycleMET[,"M"]*quintCyclingTime[["M"]]/60, F = means$meanCycleMET[,"F"]*quintCyclingTime[["F"]]/60)
})
```

And the METs for active travel.
```{r totalTravelMETS}
ITHIM <- within(ITHIM, {
quintTotalTravelMET = list(M = quintWalkingMET[["M"]] + quintCyclingMET[["M"]], F = quintWalkingMET[["F"]] + quintCyclingMET[["F"]])
})
```
Now we print all of the parameters and results.
```{r ITHIM}
names(ITHIM)
```
### Example
To run the example contained in the _ITHIM_ package run the following.

```{r package}
library("ITHIM")
example("ITHIM-package")
```

```{r plot1}
D <- melt(ITHIM$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
p <- ggplot(D, aes(age,  activeTransportTime)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
p
```

```{r plot2}
D <- melt(ITHIM$parameters$Rwt, c("age","sex"), value.name = "rhowt")
p <- ggplot(D, aes(age,  rhowt)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
p
```

## Archive
### Active Transport Estimates (Speed and Distance by Transport Mode)
These data were ripped out of the EXCEL workbook and vary per scenario.
```{r transportData, eval = FALSE, echo = FALSE}

transportMode <- c("walk","cycle","bus","minibus","train","car.driver","car.passenger","mbike","elec.bike.scooter","LGV","HGV.rigid","HGV.articulated")

speed <- c(1.35299774571063,5.39374775150058,5.22298847623309,NA,9.94147658251227,17.36906017290580,14.51012134295660,NA,NA,NA,NA,NA)

dist <- c(2.44072161538462,1.19969769230769,4.37851867307692,NA,5.58545353846154,111.91985351923100,39.12264265384620,0.00000000000000,0.00000000000000,NA,1.05479452054795,NA) # miles per week

Data.Speed <- data.frame(speed,dist, row.names = transportMode)
```

```{r headDataSpeed, eval = FALSE}
Data.Speed
```

```{r transportTime, echo = FALSE, results = "hide", eval = FALSE}
head(transportTime <- Data.Speed$dist/Data.Speed$speed*60)
```

### Emissions Data

```{r roadAndCO2Data, echo = FALSE, eval = FALSE, results = "hide"}

CO2 <- c(957.686509527760,221.578783667454,203.624526028894,247.869842164617,333.468424533125,90.000000000000,835.014773125786)

Data.CO2 <- data.frame(CO2, row.names = c("bus","Cars.urban","Cars.mway","LGVs.urban","LGVs.mway","motorbike","HGV"))

Data.RoadType <- data.frame(local=c(0.75,0.53,0.0,0.09,0.29,NA,0.09,NA), arterial=c(0.25,0.47,0.96,0.31,0.42,NA,0.31,NA), highway=c(0.0,0,0.04,0.60,0.29,NA,0.60,NA),row.names=c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike") )

Data.Occupancy <- data.frame(occupancy = c(1.0,1.0,18.9,1.3,1.0,NA,1.0,NA), row.names = c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike"))

```

The amount of carbon emissions varies by transport mode and road type.
To accomodate this we need the distribution of road type by transport
mode.  At some point we need the expected number of passengers per
transport mode and we create it here, arbitrarily perhaps.

```{r showData, echo = TRUE, eval = FALSE, results = "show"}
Data.CO2
Data.RoadType
Data.Occupancy
```

### Non-(Active Transport) METs

The file NonTravelMETs.csv was created from the tripleWin EXCEL file.
These data are used in a table look-up for a given county.  The data
represent the non-travel-related energy expenditures as measured in
METs.  The data are stratified by age and sex.  Variable names were
not given for the 7 numerical variables listed.  Here we name them x1
through x7.

```{r nonTravelMETs, eval = FALSE, echo = FALSE}
nonTravelMETs <- data.frame(scan(file="./data/NonTravelMETs.csv", sep = ",", what = list("","","","",1,1,1,1,1,1,1), skip = 1))
names(nonTravelMETs) <- c("ID","sexcat","agecat2","GESTFIPS", paste0("x",1:7))
```

```{r ggplot, fig.width=12, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE, results = "hide", eval = FALSE}
D <- melt(nonTravelMETs)
p <- ggplot(D, aes(x = variable, y = value, colour = agecat2), outlier.colour = "red", outlier.shape = 1 )
p + geom_boxplot()
```

Who is the outlier?  Perhaps it is a cohort of laborers, e.g., a county of coal miners?

```{r foo, eval = FALSE, echo = TRUE}
nonTravelMETs[which.max(nonTravelMETs$x5),]
```

Strange that it is group of middle-aged men.  I see this as a red
flag.  I remove this datum and redraw the boxplots, but this time only
for x5, x6 and x7, the others seemed less interesting.

```{r fips, eval = FALSE, echo = FALSE, results = "hide"}
fips <- read.csv("./data/fips_codes_website.csv")
any(fips$County.FIPS.Code==38)
```

```{r Nontravelmets, echo = FALSE, message = FALSE, warning = FALSE, results = "hide", fig.width=12, fig.height=6, eval = FALSE}
D <- melt(nonTravelMETs[-which.max(nonTravelMETs$x5),-(5:8)])
p <- ggplot(D, aes(x = variable, y = value, colour = agecat2), outlier.colour = "red", outlier.shape = 1 ) + geom_boxplot()

p + facet_grid(. ~ sexcat)
```

Thoughts?

### Relative Risk of MET Exposure

In the worksheet _Phy activity RRs_ we see estimates for relative risk
for various disease outcomes.  My guess is that some transformation is
being made to convert the exposure from active transport time to METs.
I don't know why this is needed yet.  I can't make much progress here
until I understand that better.

### Coefficient of Variation

An estimate for the coefficient of variation is provided without much
explanation.  The literature cites an empirical result.  The feeling
is that as mean active transport increases in the population people
are fitter and there is less variance in active transport time.  I
guess this has been shown in the same refernce for the coefficient of
variation estimate.

### Vital Statistics Data

#### 2001 Data with Vargo Script

```{r vitalstats, eval = FALSE, echo = FALSE}
source("../vargo/process_2001_vital_stats.r")
saveRDS(mortality2001, file = "./data/mortality2001.rds")
```

#### 2011 (VS11MORT.DUSMCPUB)

Lets read in the CDC data using the file VS11MORT.DUSMCPUB and see whats up.

```{r VS11, echo = FALSE, eval = FALSE}
tapePosition <- list(state = c(29:30), skippedVars = c(31:34), county = c(35:37), skippedVars = c(38:68), sex = 69, ageUnit = 70, ageValue = c(71:73), skippedVars = c(74:145), ICD = c(146:149), skippedVars = c(150:444), race = c(445:446)) # Magic number!!!
colClasses <- c("factor","factor","factor", "factor","integer", "character", "factor")
fileWidth <- 488 # Magic number!!!
```

```{r VS11Classes, echo = FALSE, eval = FALSE}
names(colClasses) <- names(tapePosition)[-which(names(tapePosition)=="skippedVars")]
colClasses
```

To include more variables edit the above lines.  The variable listed
must be contiguous.

```{r convertAge, warning = FALSE, eval = FALSE, echo = FALSE}

varWidths <- unlist(lapply(tapePosition, "length"))
varWidths <- ifelse(names(tapePosition) == "skippedVars", -varWidths, varWidths)
widths <- c(-(min(tapePosition[[1]])-1),c(varWidths))
widths <- c(widths,-(fileWidth-sum(abs(widths))))

Data.Mortality <- read.fwf(file = "~/ITHIM/R/data/VS11MORT.DUSMCPUB.first100", widths = widths, col.names = names(colClasses), header = FALSE, colClasses = colClasses)
Data.Mortality <- data.frame(Data.Mortality, convertedAge = with(Data.Mortality, convertAge(value = ageValue, unit = ageUnit)))

p <- ggplot(Data.Mortality, aes(x = convertedAge)) + geom_histogram(binwidth = 1)
p + facet_grid(. ~ sex)

```

```{r ICD, fig.width = 12, echo = FALSE, eval = FALSE}
Data.IDC.Codes <- read.csv(file="./data/IDC_Codes_key.csv", header = TRUE, colClasses = c("character","character","character"), na.string = "")
Data.IDC.Codes <- subset(Data.IDC.Codes, Include == "1" & !is.na(ITHIMCause))
ITHIMCauseOfDeath <- Data.IDC.Codes$ITHIMCause
names(ITHIMCauseOfDeath) <- Data.IDC.Codes$cause.of.death
Data.Mortality <- data.frame(Data.Mortality, CauseofDeathITHIM = ITHIMCauseOfDeath[Data.Mortality$ICD])
Data.Mortality <- subset(Data.Mortality, !is.na(CauseofDeathITHIM) & !is.na(convertedAge))
D <- data.frame(disease = Data.Mortality$CauseofDeathITHIM)
p <- ggplot(D, aes(disease))
p + geom_bar()

```

```{r CoDTable, eval = FALSE}
with(Data.Mortality, table(CauseofDeathITHIM))
```

```{r ageClass, fig.width = 12, echo = FALSE, eval = FALSE}
ageClass <- c(rep("00-04",5),rep("05-14",10),rep("15-29",15),rep("30-44",15),rep("45-59",15),rep("60-69",10),rep("70-79",10),rep("80+",121))
names(ageClass) <- as.character(0:200)
Data.Mortality <- data.frame(Data.Mortality, ageClass = as.factor(ageClass[as.character(as.integer(Data.Mortality$convertedAge))]))
with(Data.Mortality, table(ageClass))

D <- data.frame(ageClass = Data.Mortality$ageClass, disease = Data.Mortality$CauseofDeathITHIM)
p <- ggplot(D, aes(disease))
p + geom_bar(aes(fill=ageClass))
```

```{r histbyDisease, fig.width = 12, echo = FALSE, eval = FALSE}
D <- data.frame(convertedAge = Data.Mortality$convertedAge, disease = Data.Mortality$CauseofDeathITHIM)
p <- ggplot(D, aes(convertedAge, ..density.., colour = disease)) + geom_histogram(binwidth=5, show.legend = TRUE)
p + facet_grid( . ~ disease)
```

```{r counties, eval = FALSE}
Data.AgeSex <- read.csv("./data/county_age_sex2002.csv", header=T, col.names = c("county.name","state.name", "MSA","sex","all.ages","age0-4","age5-14","age15-29","age30-44","age45-59","age60-69","age70-79","age80+", "STATE","COUNTY"), colClasses = c("factor","factor","factor","factor","integer","integer","integer","integer","integer","integer","integer","integer","integer", "factor", "factor"))

age.sex.2 <- melt(age.sex[,c(14,15,4,6:13)], id=c("STATE","COUNTY","sex"))

#age.sex.3 <- dcast(age.sex.2, STATE + COUNTY + variable ~ sex, mean)

mort.age <- merge(mort.ITHIM, age.sex.3, by.x=c("state","county", "ITHIM.age"), by.y=c("STATE","COUNTY","variable"))

mort.age$year <- 2001

mort.age$app.pop <- ifelse(mort.age$sex ==1, mort.age$male, mort.age$female)

```

```{r tabulate, eval = FALSE}
Data.Mortality.Table <- as.data.frame(ftable(with(Data.Mortality, table(sex,ageClass,CauseofDeathITHIM))))
```

```{r saveMortality, echo = FALSE, eval = FALSE}
saveRDS(object = Data.Mortality, file = "./data/Data.Mortality.rds")
saveRDS(object = Data.Mortality.Table, file = "./data/Data.Mortality.Table.rds")
```

### Appendix

```{r appendix}
date()
sessionInfo()
```

### 2016-05-25 R/ITHIM.Rmd
Added install_github to ITHIM.Rmd

@@
 
-```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE}
+```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
 options(repos='http://cran.rstudio.com/')
@@
 library("plyr")
-library("ITHIM")
+library("devtools")
+install_github("syounkin/ITHIM")
 ```

### 2016-05-25 R/ITHIM.Rmd
Forgot to load ITHIM

--- 2016-05-25 R/ITHIM.Rmd
@@
 
-```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE}
+```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
 options(repos='http://cran.rstudio.com/')
@@
 library("plyr")
+```
+```{r loadITHIM}
+library("devtools")
+install_github("syounkin/ITHIM")
 library("ITHIM")

### 2016-05-25 R/ITHIM.Rmd
Update to ITHIM.Rmd
Introduced getQuintiles into the instanciation of the ITHIM model

--- 2016-05-25 R/ITHIM.Rmd
@@
-## Integrated Transport and Health Impacts Model (ITHIM)
+# Integrated Transport and Health Impacts Model
 
+## R Package _ITHIM_
 This document will eventually serve as a compendium to the R package
@@
 
-```{r date}
-date()
-```
-
-```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE}
+```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
 options(repos='http://cran.rstudio.com/')
@@
 library("plyr")
+```
+```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE}
+library("devtools")
+install_github("syounkin/ITHIM")
 library("ITHIM")
+date()
 ```
 
-## R Implementation
+### Instanciation
 
 ```{r instantiate}
-ITHIM <- list(parameters = parList <- createParameterList(), means = computeMeanMatrices(parList))
-```
-Then we compute quintiles for active transport time.
-```{r quints}
-ITHIM <- within(ITHIM, {
-quintilesActiveTransportTime <- computeQuintiles(means$meanActiveTransportTime, means$sdActiveTransportTime)
-})
+ITHIM <- list(
+  parameters = parameters <- createParameterList(),
+  means = means <- computeMeanMatrices(parameters),
+  quintiles = quintiles <- getQuintiles(means)
+)
 ```
 
-Then the METs by walking and cycling.
-```{r quintMETs}
-ITHIM <- within(ITHIM, {
-  quintWalkingTime = list(M = quintilesActiveTransportTime[["M"]] * (1-means$propTimeCycling[,"M"]), F = quintilesActiveTransportTime[["F"]] * (1-means$propTimeCycling[,"F"]))
-  quintCyclingTime = list(M = quintilesActiveTransportTime[["M"]] * (means$propTimeCycling[,"M"]), F = quintilesActiveTransportTime[["F"]] * (means$propTimeCycling[,"F"]))
-  quintWalkingMET = list(M = means$meanWalkMET[,"M"]*quintWalkingTime[["M"]]/60, F = means$meanWalkMET[,"F"]*quintWalkingTime[["F"]]/60)
-  quintCyclingMET = list(M = means$meanCycleMET[,"M"]*quintCyclingTime[["M"]]/60, F = means$meanCycleMET[,"F"]*quintCyclingTime[["F"]]/60)
-})
-```
-
-And the METs for active travel.
-```{r totalTravelMETS}
-ITHIM <- within(ITHIM, {
-quintTotalTravelMET = list(M = quintWalkingMET[["M"]] + quintCyclingMET[["M"]], F = quintWalkingMET[["F"]] + quintCyclingMET[["F"]])
-})
-```
-Now we print all of the parameters and results.
 ```{r ITHIM}
@@
 ```
+
+For help run the following.
+```{r help, eval = FALSE}
+help(createParameterList)
+help(computeMeansMatrices)
+help(getQuintiles)
+```
+
 ### Example
@@
 
-```{r package}
-library("ITHIM")
+```{r package, eval = FALSE}
 example("ITHIM-package")
 ```
+### Figures
 
-```{r plot1}
+```{r plot1, echo = FALSE}
 D <- melt(ITHIM$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
@@
 
-```{r plot2}
+```{r plot2, echo = FALSE}
 D <- melt(ITHIM$parameters$Rwt, c("age","sex"), value.name = "rhowt")
@@
 
-## Archive
+# Code Graveyard
+
 ### Active Transport Estimates (Speed and Distance by Transport Mode)

### 2016-05-26 R/ITHIM.Rmd
Compute total METs and RR

@@
 )
+
+RR <- createActiveTransportRRs()
+RR <- matrix(RR$BreastCancer[,"F"],nrow=nAgeClass,ncol = 5, byrow = FALSE)
+totalMETs <- mapply("+",ITHIM$quintiles$TotalTravelMET,computeNonTravelMETs(),SIMPLIFY=FALSE)
+
+RR^(totalMETs$F^0.5)
+
 ```

### 2016-05-26 R/ITHIM.Rmd
Update to ITHIM.Rmd

--- 2016-05-25 R/ITHIM.Rmd
@@
 ```{r instantiate}
-ITHIM <- list(
-  parameters = parameters <- createParameterList(),
+ITHIM.baseline <- list(
+  parameters = parameters <- createParameterList(baseline = TRUE),
   means = means <- computeMeanMatrices(parameters),
@@
 )
+ITHIM.scenario <- list(
+  parameters = parameters <- createParameterList(baseline = FALSE),
+  means = means <- computeMeanMatrices(parameters),
+  quintiles = quintiles <- getQuintiles(means)
+)
+
+RR <- createActiveTransportRRs()
+RR <- matrix(RR$BreastCancer[,"F"],nrow=nAgeClass,ncol = 5, byrow = FALSE)
+
+totalMETs.baseline <- mapply("+",ITHIM.baseline$quintiles$TotalTravelMET,computeNonTravelMETs(),SIMPLIFY=FALSE)
+totalMETs.scenario <- mapply("+",ITHIM.scenario$quintiles$TotalTravelMET,computeNonTravelMETs(),SIMPLIFY=FALSE)
+
+RR^(totalMETs$F^0.5)
+
 ```

### 2016-05-27 R/ITHIM.Rmd
Introduced AF to script

@@
 library("devtools")
-install_github("syounkin/ITHIM")
+install_github("syounkin/ITHIM", ref = "devel", force = TRUE)
 library("ITHIM")
@@
 RR <- createActiveTransportRRs()
-RR <- matrix(RR$BreastCancer[,"F"],nrow=nAgeClass,ncol = 5, byrow = FALSE)
+
+nAgeClass <- 8
 
@@
 
-RR^(totalMETs$F^0.5)
+totalMETs.baseline <- lapply(totalMETs.baseline, function(x) ifelse(x<2.5,0.1,x))
+totalMETs.scenario <- lapply(totalMETs.scenario, function(x) ifelse(x<2.5,0.1,x))
 
+## RR for Females and Breast Cancer
+RR <- matrix(RR$BreastCancer[,"F"],nrow=nAgeClass,ncol = 5, byrow = FALSE)
+RR.baseline <- RR^(totalMETs.baseline$F^0.5)
+RR.scenario <- RR^(totalMETs.scenario$F^0.5)
+diseaseBurden <- RR.baseline/RR.scenario
+burden <- 1 - (rowSums(diseaseBurden)-ncol(diseaseBurden))/rowSums(diseaseBurden)
+(AF <- 1 - burden)
 ```
@@
 ```{r ITHIM}
+ITHIM <- ITHIM.baseline
 names(ITHIM)

### 2016-05-27 R/ITHIM.Rmd
Update to ITHIM.Rmd

@@
 
-### Instanciation
+### Running in R
 
@@
 
-RR <- createActiveTransportRRs()
-
-nAgeClass <- 8
-
 totalMETs.baseline <- mapply("+",ITHIM.baseline$quintiles$TotalTravelMET,computeNonTravelMETs(),SIMPLIFY=FALSE)
 totalMETs.scenario <- mapply("+",ITHIM.scenario$quintiles$TotalTravelMET,computeNonTravelMETs(),SIMPLIFY=FALSE)
-
 totalMETs.baseline <- lapply(totalMETs.baseline, function(x) ifelse(x<2.5,0.1,x))
 totalMETs.scenario <- lapply(totalMETs.scenario, function(x) ifelse(x<2.5,0.1,x))
+```
+Now we can compute the attributable fraction (AF).
 
+```{r RRs}
 ## RR for Females and Breast Cancer
+RR <- createActiveTransportRRs()
+nAgeClass <- 8
 RR <- matrix(RR$BreastCancer[,"F"],nrow=nAgeClass,ncol = 5, byrow = FALSE)
@@
 RR.scenario <- RR^(totalMETs.scenario$F^0.5)
-diseaseBurden <- RR.baseline/RR.scenario
-burden <- 1 - (rowSums(diseaseBurden)-ncol(diseaseBurden))/rowSums(diseaseBurden)
-(AF <- 1 - burden)
+diseaseBurden.scenario <- RR.baseline/RR.scenario
+diseaseBurden.baseline <- RR.baseline/RR.baseline # What!
+(AF <- (rowSums(diseaseBurden.scenario)-rowSums(diseaseBurden.baseline))/rowSums(diseaseBurden.scenario))
 ```
 
-```{r ITHIM}
-ITHIM <- ITHIM.baseline
-names(ITHIM)
+```{r explore}
+names(ITHIM.baseline$parameters)
 ```
 
-For help run the following.
+We use Rws to estimate mean walking speed and from that mean walking
+METs.  Mean cycling METs, however, is constant and is not a function
+of mean cycling speed or Rcs.  So, I do not think Rcs is currently
+used.
+
+To see the help page for a function use help().
 ```{r help, eval = FALSE}

### 2016-05-27 R/ITHIM.Rmd
Almost have AF replicating...

--- 2016-05-26 R/ITHIM.Rmd
@@
 library("devtools")
-install_github("syounkin/ITHIM")
+install("~/ITHIM/")
+# install_github("syounkin/ITHIM", ref = "devel", force = TRUE)
 library("ITHIM")
@@
 )
+```
+Now we can compute the attributable fraction (AF).
 
+```{r RRs}
 RR <- createActiveTransportRRs()
-RR <- matrix(RR$BreastCancer[,"F"],nrow=nAgeClass,ncol = 5, byrow = FALSE)
+RR.baseline <- lapply(RR, MET2RR, ITHIM.baseline$quintiles$TotalMET)
+RR.scenario <- lapply(RR, MET2RR, ITHIM.scenario$quintiles$TotalMET)
 
-totalMETs.baseline <- mapply("+",ITHIM.baseline$quintiles$TotalTravelMET,computeNonTravelMETs(),SIMPLIFY=FALSE)
-totalMETs.scenario <- mapply("+",ITHIM.scenario$quintiles$TotalTravelMET,computeNonTravelMETs(),SIMPLIFY=FALSE)
-
-RR^(totalMETs$F^0.5)
-
+diseaseBurden.scenario <- mapply(ratioForList,RR.baseline, RR.scenario, SIMPLIFY = FALSE)
+diseaseBurden.baseline <- mapply(ratioForList,RR.baseline, RR.baseline, SIMPLIFY = FALSE) # What!
+AF <- mapply(AFForList, diseaseBurden.scenario,diseaseBurden.baseline, SIMPLIFY = FALSE)
 ```
 
-```{r ITHIM}
-names(ITHIM)
+Note that some of the AF values will disagree with the EXCEL
+spreadsheet because Diabetes, HHD_Air, CVD_Air and Stroke_Air do not
+combine travel METs with non-travel METs.
+
+```{r AF}
+100*round(AF$BreastCancer[-(1:2),],4)
+
+```{r explore}
+names(ITHIM.baseline$parameters)
 ```
 
-For help run the following.
+We use Rws to estimate mean walking speed and from that mean walking
+METs.  Mean cycling METs, however, is constant and is not a function
+of mean cycling speed or Rcs.  So, I do not think Rcs is currently
+used.
+
+To see the help page for a function use help().
 ```{r help, eval = FALSE}

### 2016-05-28 R/ITHIM.Rmd
Update to ITHIM.Rmd

@@
 
-```{r AF}
-100*round(AF$BreastCancer[-(1:2),],4)
-
+```{r AF, echo = FALSE}
+lapply(AF[-(which(names(AF)%in%c("CVD","Diabetes")))], function(x) 100*round(x[-(1:2),],4))
+```
+### Relative Risk Model
 ```{r explore}
@@
 ```
-
-We use Rws to estimate mean walking speed and from that mean walking
-METs.  Mean cycling METs, however, is constant and is not a function
-of mean cycling speed or Rcs.  So, I do not think Rcs is currently
-used.
+```{r RRFigure, echo = FALSE}
+phi <- function(x,k=0.5,phi0=RR$BreastCancer$F[1,1]) phi0^{x^k}
+plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
+legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
+```
 
@@
 ```{r plot1, echo = FALSE}
-D <- melt(ITHIM$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
+D <- melt(ITHIM.baseline$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
 p <- ggplot(D, aes(age,  activeTransportTime)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
@@
 ```{r plot2, echo = FALSE}
-D <- melt(ITHIM$parameters$Rwt, c("age","sex"), value.name = "rhowt")
+D <- melt(ITHIM.baseline$parameters$Rwt, c("age","sex"), value.name = "rhowt")
 p <- ggplot(D, aes(age,  rhowt)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")

### 2016-05-28 R/ITHIM.Rmd
Update to ITHIM.Rmd

@@
 ```{r instantiate}
-ITHIM.baseline <- list(
+ITHIM <- list(
   parameters = parameters <- createParameterList(baseline = TRUE),

### 2016-05-29 R/ITHIM.Rmd
Update to ITHIM.Rmd

--- 2016-05-27 R/ITHIM.Rmd
@@
 ```{r RRs}
-RR <- createActiveTransportRRs()
-RR.baseline <- lapply(RR, MET2RR, ITHIM.baseline$quintiles$TotalMET)
-RR.scenario <- lapply(RR, MET2RR, ITHIM.scenario$quintiles$TotalMET)
+comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+names(comparitiveRisk)
+```
 
-diseaseBurden.scenario <- mapply(ratioForList,RR.baseline, RR.scenario, SIMPLIFY = FALSE)
-diseaseBurden.baseline <- mapply(ratioForList,RR.baseline, RR.baseline, SIMPLIFY = FALSE) # What!
-AF <- mapply(AFForList, diseaseBurden.scenario,diseaseBurden.baseline, SIMPLIFY = FALSE)
+```{r show}
+100*round(comparitiveRisk$AF$BreastCancer,4)
 ```
@@
 
-```{r AF}
-100*round(AF$BreastCancer[-(1:2),],4)
-
+```{r AF, echo = FALSE}
+lapply(comparitiveRisk$AF[-(which(names(comparitiveRisk$AF)%in%c("CVD","Diabetes")))], function(x) 100*round(x[-(1:2),],4))
+```
+### Relative Risk Model
 ```{r explore}
@@
 ```
-
-We use Rws to estimate mean walking speed and from that mean walking
-METs.  Mean cycling METs, however, is constant and is not a function
-of mean cycling speed or Rcs.  So, I do not think Rcs is currently
-used.
+```{r RRFigure, echo = FALSE}
+phi <- function(x,k=0.5,phi0=comparitiveRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
+plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
+legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
+```
 
@@
 help(createParameterList)
-help(computeMeansMatrices)
-help(getQuintiles)
 ```
@@
 
-```{r plot1, echo = FALSE}
-D <- melt(ITHIM$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
-p <- ggplot(D, aes(age,  activeTransportTime)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
+```{r plot2, echo = FALSE, fig.width = 12}
+plotRR(comparitiveRisk$RR.baseline$Depression,comparitiveRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
+```
+
+```{r plot3, echo = FALSE, fig.width = 12, eval = FALSE}
+plotRR(comparitiveRisk$RR.baseline$BreastCancer,comparitiveRisk$RR.scenario$BreastCancer) + coord_cartesian(ylim = c(0.75, 1.05)) + ggtitle("Breast Cancer")
+```
+
+```{r plot4, echo = FALSE, fig.height = 12}
+D<-melt(comparitiveRisk$AF)
+names(D)<- c("age","sex","AF","disease")
+p <- ggplot(D, aes(age,AF)) + geom_bar(aes(fill=sex), stat = "identity") + facet_grid( disease ~ .) + coord_cartesian(ylim = c(0, 0.1))
 p
 ```
-
-```{r plot2, echo = FALSE}
-D <- melt(ITHIM$parameters$Rwt, c("age","sex"), value.name = "rhowt")
-p <- ggplot(D, aes(age,  rhowt)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
-p
-```
-
 # Code Graveyard
@@
 ### Active Transport Estimates (Speed and Distance by Transport Mode)
+```{r plot1, echo = FALSE, eval = FALSE}
+D <- melt(ITHIM.baseline$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
+p <- ggplot(D, aes(age,  activeTransportTime)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
+p + coord_cartesian(ylim = c(0.75, 1.05))
+```
+
 These data were ripped out of the EXCEL workbook and vary per scenario.

### 2016-05-29 R/ITHIM.Rmd
Tweaked slides and ITHIM.Rmd

@@
 comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-names(comparitiveRisk)
 ```
@@
 ```{r show}
-100*round(comparitiveRisk$AF$BreastCancer,4)
+with(comparitiveRisk, 100*round(AF$BreastCancer,4))
 ```
@@
 ```{r AF, echo = FALSE}
-lapply(comparitiveRisk$AF[-(which(names(comparitiveRisk$AF)%in%c("CVD","Diabetes")))], function(x) 100*round(x[-(1:2),],4))
+with(comparitiveRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
 ```

### 2016-05-31 R/ITHIM.Rmd
Updated ITHIM.Rmd

@@
 ```
+```{r ap}
+k <- createAirPollutionRRs()
+```
+
 ### Relative Risk Model

### 2016-06-06 R/ITHIM.Rmd
Added a little road injury code to ITHIM.Rmd

@@
 p
+```
+### Road Injury Array
+```{r road}
+baselineInjuries <- array(dim=c(6,7,3,2))
+injury <- c("fatal","serious")
+road <- c("minor road","major road","motorway")
+vehicle <- c("walk","cycle","bus","car","HGV","mbike")
+dimnames(baselineInjuries) <- list(vehicle,c(vehicle,"NOV"),road,injury)
+
+Data.RoadType <- data.frame(local=c(0.75,0.53,0.0,0.09,0.29,NA,0.09,NA), arterial=c(0.25,0.47,0.96,0.31,0.42,NA,0.31,NA), highway=c(0.0,0,0.04,0.60,0.29,NA,0.60,NA),row.names=c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike") )
+
+Data.Occupancy <- data.frame(occupancy = c(1.0,1.0,18.9,1.3,1.0,NA,1.0,NA), row.names = c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike"))
+
 ```

### 2016-06-07 R/ITHIM.Rmd
Added code to loop through values of muwt

@@
 combine travel METs with non-travel METs.
+
+```{r parlist}
+AF.sum <- numeric(1e3); i <- 0;
+muwt.vec <- seq(107.1,150,length.out=length(AF.pseudo))
+#muct.vec <- seq(2.7,3.0,length.out=length(AF.pseudo))
+for(muwt in muwt.vec){
+i <- i+1
+parameters <- createParameterList(baseline = FALSE)
+
+parameters <- setParameter(parName="muwt", parValue = muwt, parList = parameters)
+ITHIM.scenario <- list(
+  parameters = parameters <- parameters,
+  means = means <- computeMeanMatrices(parameters),
+  quintiles = quintiles <- getQuintiles(means)
+)
+comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+AF.sum[i] <- with(comparitiveRisk, sum(100*AF$BreastCancer))
+}
+D <- data.frame(mu.wt = muwt.vec,AF.sum=AF.sum)
+p <- ggplot(D, aes(x = mu.wt, y = AF.sum))
+p + geom_point()
+```
 

### 2016-06-09 R/shiny/server.R
Began working with Shiny

@@ -1,0 +1,56 @@
+library(shiny)
+library(datasets)
+library(ITHIM)
+library(ggplot2)
+library(reshape2)
+
+# We tweak the "am" field to have nicer factor labels. Since this doesn't
+# rely on any user inputs we can do this once at startup and then use the
+# value throughout the lifetime of the application
+#mpgData <- mtcars
+#mpgData$am <- factor(mpgData$am, labels = c("Automatic", "Manual"))
+
+# Define server logic required to plot various variables against mpg
+shinyServer(function(input, output) {
+
+  # Compute the forumla text in a reactive expression since it is
+  # shared by the output$caption and output$mpgPlot expressions
+#  formulaText <- reactive({
+#    paste("mpg ~", input$variable)
+#  })
+
+  # Return the formula text for printing as a caption
+ # output$caption <- renderText({
+ #   formulaText()
+  #})
+
+  # Generate a plot of the requested variable against mpg and only
+                                        # include outliers if requested
+
+    parameters <- createParameterList(baseline = TRUE)
+    means <- computeMeanMatrices(parameters)
+    quintiles <- getQuintiles(means)
+
+    ITHIM.baseline <- list( parameters = parameters, means = means, quintiles = quintiles )
+
+    parameters <- createParameterList(baseline = FALSE)
+    means <- computeMeanMatrices(parameters)
+    quintiles <- getQuintiles(means)
+
+    ITHIM.scenario <- list( parameters = parameters, means = means, quintiles = quintiles )
+
+#)
+## ITHIM.scenario <- list(
+##   parameters = parameters <- createParameterList(baseline = FALSE),
+##   means = means <- computeMeanMatrices(parameters),
+##   quintiles = quintiles <- getQuintiles(means)
+## )
+    comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+
+
+    output$ITHIMPlot <- renderPlot({
+      #  hist(comparitiveRisk$RR.baseline[[as.character(input$variable)]]$M[,1])
+      plotRR(comparitiveRisk$RR.baseline[[input$variable]],comparitiveRisk$RR.scenario[[input$variable]]) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle(as.character(input$variable))
+
+  })
+})

### 2016-06-09 R/shiny/ui.R

@@ -1,0 +1,25 @@
+library(shiny)
+
+# Define UI for miles per gallon application
+shinyUI(pageWithSidebar(
+
+  # Application title
+  headerPanel("ITHIM"),
+
+  # Sidebar with controls to select the variable to plot against mpg
+  # and to specify whether outliers should be included
+  sidebarPanel(
+    selectInput("variable", "Variable:",
+                list("Depression" = "Depression",
+                     "Dementia" = "Dementia"))
+
+#    checkboxInput("outliers", "Show outliers", FALSE)
+  ),
+
+  # Show the caption and plot of the requested variable against mpg
+  mainPanel(
+    h3(textOutput("variable")),
+
+    plotOutput("ITHIMPlot")
+  )
+))

### 2016-06-10 R/shiny/server.R
Updated Shiny effort

@@ -26,31 +26,41 @@ shinyServer(function(input, output) {
 
   # Generate a plot of the requested variable against mpg and only
                                         # include outliers if requested
-
+#browser()
+    
     parameters <- createParameterList(baseline = TRUE)
-    means <- computeMeanMatrices(parameters)
-    quintiles <- getQuintiles(means)
+    means1 <- computeMeanMatrices(parameters)
+    quintiles <- getQuintiles(means1)
 
-    ITHIM.baseline <- list( parameters = parameters, means = means, quintiles = quintiles )
+    ITHIM.baseline <- list( parameters = parameters, means = means1, quintiles = quintiles )
 
+
+
+
+ comparitiveRisk <- reactive({
     parameters <- createParameterList(baseline = FALSE)
-    means <- computeMeanMatrices(parameters)
-    quintiles <- getQuintiles(means)
-
-    ITHIM.scenario <- list( parameters = parameters, means = means, quintiles = quintiles )
-
-#)
-## ITHIM.scenario <- list(
-##   parameters = parameters <- createParameterList(baseline = FALSE),
-##   means = means <- computeMeanMatrices(parameters),
-##   quintiles = quintiles <- getQuintiles(means)
-## )
-    comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-
-
+    parameters <- setParameter(parName="muwt", parValue = input$muwt, parList = parameters)
+    parameters <- setParameter(parName="muct", parValue = input$muct, parList = parameters)
+    means2 <- computeMeanMatrices(parameters)
+    quintiles <- getQuintiles(means2)
+    ITHIM.scenario  <- list( parameters = parameters, means = means2, quintiles = quintiles )
+    compareModels(ITHIM.baseline,ITHIM.scenario)
+})
     output$ITHIMPlot <- renderPlot({
+    #        plot(input$muwt)
+            comparitiveRisk <- comparitiveRisk()
       #  hist(comparitiveRisk$RR.baseline[[as.character(input$variable)]]$M[,1])
       plotRR(comparitiveRisk$RR.baseline[[input$variable]],comparitiveRisk$RR.scenario[[input$variable]]) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle(as.character(input$variable))
 
   })
+
+
+data <- reactive({
+    foo <- comparitiveRisk()
+    data.frame(foo$AF[[input$variable]])
+    })
+    
+    output$values <- renderTable({
+            data()
+              })
 })

### 2016-06-10 R/shiny/ui.R

@@ -11,15 +11,30 @@ shinyUI(pageWithSidebar(
   sidebarPanel(
     selectInput("variable", "Variable:",
                 list("Depression" = "Depression",
-                     "Dementia" = "Dementia"))
+                     "Dementia" = "Dementia")),
+    sliderInput("muwt", "Mean Walking Time (min per week)):",
+                                min=55, max=200, value=107),
+        sliderInput("muct", "Mean Cycling Time (min per week)):",
+                                min=10, max=200, value=39)
+
 
 #    checkboxInput("outliers", "Show outliers", FALSE)
   ),
 
   # Show the caption and plot of the requested variable against mpg
-  mainPanel(
-    h3(textOutput("variable")),
+  ## mainPanel(
+  ##   h3(textOutput("variable")),
 
-    plotOutput("ITHIMPlot")
-  )
+  ##   #plotOutput("ITHIMPlot"),
+  ##   tableOutput("values")
+  ## )
+
+mainPanel(
+        tabsetPanel(
+                  tabPanel("R.R.", plotOutput("ITHIMPlot")),
+#                  tabPanel("Summary", verbatimTextOutput("summary")),
+                  tabPanel("A.F.", tableOutput("values"))
+                )
+      )
+  
 ))

### 2016-06-10 R/shiny/server.R
Deleted shiny directory

@@ -1,66 +1,0 @@
-library(shiny)
-library(datasets)
-library(ITHIM)
-library(ggplot2)
-library(reshape2)
-
-# We tweak the "am" field to have nicer factor labels. Since this doesn't
-# rely on any user inputs we can do this once at startup and then use the
-# value throughout the lifetime of the application
-#mpgData <- mtcars
-#mpgData$am <- factor(mpgData$am, labels = c("Automatic", "Manual"))
-
-# Define server logic required to plot various variables against mpg
-shinyServer(function(input, output) {
-
-  # Compute the forumla text in a reactive expression since it is
-  # shared by the output$caption and output$mpgPlot expressions
-#  formulaText <- reactive({
-#    paste("mpg ~", input$variable)
-#  })
-
-  # Return the formula text for printing as a caption
- # output$caption <- renderText({
- #   formulaText()
-  #})
-
-  # Generate a plot of the requested variable against mpg and only
-                                        # include outliers if requested
-#browser()
-    
-    parameters <- createParameterList(baseline = TRUE)
-    means1 <- computeMeanMatrices(parameters)
-    quintiles <- getQuintiles(means1)
-
-    ITHIM.baseline <- list( parameters = parameters, means = means1, quintiles = quintiles )
-
-
-
-
- comparitiveRisk <- reactive({
-    parameters <- createParameterList(baseline = FALSE)
-    parameters <- setParameter(parName="muwt", parValue = input$muwt, parList = parameters)
-    parameters <- setParameter(parName="muct", parValue = input$muct, parList = parameters)
-    means2 <- computeMeanMatrices(parameters)
-    quintiles <- getQuintiles(means2)
-    ITHIM.scenario  <- list( parameters = parameters, means = means2, quintiles = quintiles )
-    compareModels(ITHIM.baseline,ITHIM.scenario)
-})
-    output$ITHIMPlot <- renderPlot({
-    #        plot(input$muwt)
-            comparitiveRisk <- comparitiveRisk()
-      #  hist(comparitiveRisk$RR.baseline[[as.character(input$variable)]]$M[,1])
-      plotRR(comparitiveRisk$RR.baseline[[input$variable]],comparitiveRisk$RR.scenario[[input$variable]]) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle(as.character(input$variable))
-
-  })
-
-
-data <- reactive({
-    foo <- comparitiveRisk()
[…]

### 2016-06-10 R/shiny/ui.R

@@ -1,40 +1,0 @@
-library(shiny)
-
-# Define UI for miles per gallon application
-shinyUI(pageWithSidebar(
-
-  # Application title
-  headerPanel("ITHIM"),
-
-  # Sidebar with controls to select the variable to plot against mpg
-  # and to specify whether outliers should be included
-  sidebarPanel(
-    selectInput("variable", "Variable:",
-                list("Depression" = "Depression",
-                     "Dementia" = "Dementia")),
-    sliderInput("muwt", "Mean Walking Time (min per week)):",
-                                min=55, max=200, value=107),
-        sliderInput("muct", "Mean Cycling Time (min per week)):",
-                                min=10, max=200, value=39)
-
-
-#    checkboxInput("outliers", "Show outliers", FALSE)
-  ),
-
-  # Show the caption and plot of the requested variable against mpg
-  ## mainPanel(
-  ##   h3(textOutput("variable")),
-
-  ##   #plotOutput("ITHIMPlot"),
-  ##   tableOutput("values")
-  ## )
-
-mainPanel(
-        tabsetPanel(
-                  tabPanel("R.R.", plotOutput("ITHIMPlot")),
-#                  tabPanel("Summary", verbatimTextOutput("summary")),
-                  tabPanel("A.F.", tableOutput("values"))
-                )
-      )
-  
-))

### 2016-06-17 R/ITHIM.Rmd
Created Road Injury Array
Added region argument to getQuintiles

@@
 ```
+
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE}
@@
 install("~/ITHIM/")
-# install_github("syounkin/ITHIM", ref = "devel", force = TRUE)
+# install_github("syounkin/ITHIM", force = TRUE)
 library("ITHIM")
@@
 ITHIM.baseline <- list(
-  parameters = parameters <- createParameterList(baseline = TRUE),
+  parameters = parameters <- createParameterList(vision = "baseline", region = "national"),
   means = means <- computeMeanMatrices(parameters),
-  quintiles = quintiles <- getQuintiles(means)
+  quintiles = quintiles <- getQuintiles(means, region = "national")
 )
 ITHIM.scenario <- list(
-  parameters = parameters <- createParameterList(baseline = FALSE),
+  parameters = parameters <- createParameterList(vision = "scenario", region = "national"),
   means = means <- computeMeanMatrices(parameters),
-  quintiles = quintiles <- getQuintiles(means)
+  quintiles = quintiles <- getQuintiles(means, region = "national")
 )
@@
 dimnames(baselineInjuries) <- list(vehicle,c(vehicle,"NOV"),road,injury)
+
+fatalities <- array(c(0,0,0.6666666667,34,6.3333333333,0,0.3333333333,0,0,0,0,0,4.3333333333,0,0,0.3333333333,0,0,0,0,0.3333333333,0.3333333333,0,0,0.6666666667,0,0.3333333333,0,0,0,48.6666666667,2,0,5.6666666667,0,35.6666666667,0,0,0,2.3333333333,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,2,0,0,0,0,6.3333333333,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,11.3333333333,1,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,18.6666666667,1.3333333333,0,2,0,7.3333333333,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0.3333333333,0,0.3333333333,0.3333333333,20,3,0,0,0,0,0,0.3333333333,0.3333333333,20,3,0,0,0,0,0,0,0,1,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,42.3333333333,4.6666666667,0,2,0,26.3333333333,0,0,0,3.3333333333,0.3333333333,0,0.3333333333,0,1,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,3,0,0,0,6.3333333333,0,0,0.6666666667,0,0),dim=c(9,8,3))
+
+fatalaties <- apply(fatalities,c(1,3),t)
+
+serious <- array(c(0,0,0,6.3333333333,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,40,0.6666666667,0,0.6666666667,0,11.6666666667,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0.3333333333,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,8,0.3333333333,0,0.6666666667,0,3.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.6666666667,0.3333333333,0,0,0,0,0,0,0,0.6666666667,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,29.6666666667,1,0,0,0,11,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0),dim=c(9,8,3))
+
+serious <- apply(serious,c(1,3),t)
+
+baselineInjuries[,,,1] <- fatalaties
+baselineInjuries[,,,2] <- serious
 

### 2016-06-18 R/ITHIM.Rmd
Update to ITHIM.Rmd

@@
 ## R Package _ITHIM_
-This document will eventually serve as a compendium to the R package
+<!-- This document will eventually serve as a compendium to the R package
 _ITHIM_.  It will describe how to use the package to implement the
@@
 As a working document it will serve as a place for SGY to post
-results, questions, interpreations, notes, etc.
+results, questions, interpreations, notes, etc.-->
 
@@
 library("plyr")
+#install("~/ITHIM/")
 ```
@@
 library("devtools")
-install("~/ITHIM/")
-# install_github("syounkin/ITHIM", force = TRUE)
+install_github("syounkin/ITHIM")
 library("ITHIM")
-date()
 ```
 
-### Running in R
+### Bay Area
+
+First we define the baseline and scenario objects.
 
 ```{r instantiate}
-ITHIM.baseline <- list(
-  parameters = parameters <- createParameterList(vision = "baseline", region = "national"),
-  means = means <- computeMeanMatrices(parameters),
-  quintiles = quintiles <- getQuintiles(means, region = "national")
-)
-ITHIM.scenario <- list(
-  parameters = parameters <- createParameterList(vision = "scenario", region = "national"),
-  means = means <- computeMeanMatrices(parameters),
-  quintiles = quintiles <- getQuintiles(means, region = "national")
-)
+ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
+ITHIM.scenario <- createITHIM(vision = "scenario", region = "SFBayArea")
 ```
-Now we can compute the attributable fraction (AF).
+
+Next we compare the scenario to the baseline using compareModels().
 
 ```{r RRs}
-comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
 ```
 
-```{r show}
-with(comparitiveRisk, 100*round(AF$BreastCancer,4))
+#### Deaths
+```{r show, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
 ```
+#### DALY
+```{r show2, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+#### YLD
+```{r show3, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+#### YLL
+```{r show4, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+
+### National
+
+```{r instantiate2}
+ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
+ITHIM.scenario <- createITHIM(vision = "scenario", region = "national")
+```
+
+```{r RRs2}
+comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+```
+
+#### Deaths
+```{r show222, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+#### DALY
+```{r show22, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+#### YLD
+```{r show23, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+#### YLL
+```{r show24, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+
+### Creating Different Visions
+
+```{r vision}
+ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
+comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
+```
+#### YLL
+```{r show233, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+
+### Interrogate Parameters
+
+```{r params}
+names(ITHIM.baseline$parameters)
+ITHIM.baseline$parameters$NonTravelMETs
+ITHIM.baseline$parameters$Rwt
+```
+
+<!--
+
+# Code Graveyard
+
+### Relative Risk Model
+```{r explore}
+names(ITHIM.baseline$parameters)
+```
+```{r RRFigure, echo = FALSE}
+phi <- function(x,k=0.5,phi0=comparativeRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
+plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
+lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
+legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
+```
+
+### Tinkering
 
@@
 AF.sum <- numeric(1e3); i <- 0;
-muwt.vec <- seq(107.1,150,length.out=length(AF.pseudo))
-#muct.vec <- seq(2.7,3.0,length.out=length(AF.pseudo))
+muwt.vec <- seq(107.1,150,length.out=length(AF.sum))
+#muct.vec <- seq(2.7,3.0,length.out=length(AF.sum))
 for(muwt in muwt.vec){
@@
 )
-comparitiveRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-AF.sum[i] <- with(comparitiveRisk, sum(100*AF$BreastCancer))
+comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+AF.sum[i] <- with(comparativeRisk, sum(100*AF$BreastCancer))
 }
@@
 ```{r AF, echo = FALSE}
-with(comparitiveRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
-```
-```{r ap}
-k <- createAirPollutionRRs()
-```
-
-### Relative Risk Model
-```{r explore}
-names(ITHIM.baseline$parameters)
-```
-```{r RRFigure, echo = FALSE}
-phi <- function(x,k=0.5,phi0=comparitiveRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
-plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
-lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
-lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
-lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
-legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
+with(comparativeRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
 ```
@@
 ```{r plot2, echo = FALSE, fig.width = 12}
-plotRR(comparitiveRisk$RR.baseline$Depression,comparitiveRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
+plotRR(comparativeRisk$RR.baseline$Depression,comparativeRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
 ```
@@
 ```{r plot3, echo = FALSE, fig.width = 12, eval = FALSE}
-plotRR(comparitiveRisk$RR.baseline$BreastCancer,comparitiveRisk$RR.scenario$BreastCancer) + coord_cartesian(ylim = c(0.75, 1.05)) + ggtitle("Breast Cancer")
+plotRR(comparativeRisk$RR.baseline$BreastCancer,comparativeRisk$RR.scenario$BreastCancer) + coord_cartesian(ylim = c(0.75, 1.05)) + ggtitle("Breast Cancer")
 ```
@@
 ```{r plot4, echo = FALSE, fig.height = 12}
-D<-melt(comparitiveRisk$AF)
+D<-melt(comparativeRisk$AF)
 names(D)<- c("age","sex","AF","disease")
@@
 ```
-# Code Graveyard
 
@@
 ```
-
+-->
 ### Appendix

### 2016-06-29 R/ITHIM.Rmd
Added Vargo's nonTravel MET code to ITHIM.Rmd

@@
 
+### Example
+
+```{r example}
+example("ITHIM-package")
+```
+
 ### Creating Different Visions
@@
 ```
+
+## Non-Travel METs
+
+```{r vargo, echo = FALSE, eval = TRUE}
+library(ggplot2)
+library(reshape2)
+library(plyr)
+
+summaryFile <- "~/GHI/data/atussum_0311.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_0311.dat
+
+cpsFile <- "~/GHI/data/atuscps_0311.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_0311.dat
+
+stateResultFile <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults.csv
+
+stateResultFile2 <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults_quintile.csv"
+
+regionResultFile <- "~/GHI/data//regionresults_quintile.csv" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/regionresults_quintile.csv
+
+#readLines("~/Box Documents/work/CoBenefits/MaggieShare/AmerTimeUse/atusact_2003.dat", n=10)
+# activity <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusact_2003.dat", header=T, skip=0, sep=",")
+
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
+
+# respondants <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusresp_2003.dat", header=T, skip=0, sep=",")
+
+# roster <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusrost_2003.dat", header=T, skip=0, sep=",")
+
+CPS <- read.table(cpsFile, header=T, skip=0, sep=",")
+
+CPS.2 <- CPS[,c("TUCASEID", "GESTFIPS", "GEREG","GEMETSTA")]
+CPS.2$IDCHAR <- as.character(CPS.2[,"TUCASEID"])
+
+summary.2 <- summary[, c("TUCASEID", "TEAGE",    "TESEX",    "t050203",  "t130101",  "t130102",  "t130103",  "t130104",  "t130106",  "t130108", "t130109", "t130111", "t130113",  "t130115", "t130116",  "t130117", "t130119",  "t130120",  "t130122",  "t130123", "t130124",  "t130126",  "t130127", "t130128", "t130130",  "t130131",  "t130132", "t130133",  "t130134", "t130135", "t130136" , "t130199" , "t139999")]
+summary.2$IDCHAR <- as.character(summary.2[,"TUCASEID"])
+
+summary.2a <- summary.2[,c(1:3, length(summary.2))]
+summary.2b <- summary.2[,c(5:length(summary.2)-1)]
+minutes <- rowSums(summary.2b)
+summary.2a$minutes <- minutes
+
+metkey <- as.matrix(c(3.0,7.3,5.0,6.5,7.5,3.5,5.8,7.8,6.0,8.0,3.8,7.8,8.0,5.3,7.3,9.8,6.3,7.0,7.0,5.0,6.8,3.0,3.5,5.5,6.0,4.0,6.0,3.0,3.0,3.0), nrow = 1, ncol=30)
+summary.2c <- mapply(`*`,summary.2b,metkey)
+mets <- rowSums(summary.2c)
+summary.2a$mets <- mets
+
+summary.2a$agecat <- ifelse(summary.2a$TEAGE <5, 1, 
+                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, 2, 
+                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, 3, 
+                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, 4, 
+                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, 5, 
+                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, 6, 
+                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, 7, 8
+)))))))
+
+summary.2a$agecat2 <- ifelse(summary.2a$TEAGE <5, "age0-4", 
+                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, "age5-14", 
+                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, "age15-29", 
+                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, "age30-44", 
+                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, "age45-59", 
+                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, "age60-69", 
+                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, "age70-79", "age80+"
+)))))))
+
+use <- merge(summary.2a, CPS.2, by.x="TUCASEID", by.y="TUCASEID", all.x=T, sorted=F)
+
+use$sexcat <- ifelse(use$TESEX == 1, "male", "female")
+
+use$regcat <- ifelse(use$GEREG == 1, "northeast",
+              ifelse(use$GEREG == 2, "midwest",
+              ifelse(use$GEREG == 3, "south",  "west")))
+
+use$metcat <- ifelse(use$GEMETSTA == 1, "metro",
+              ifelse(use$GEMETSTA == 2, "non-metro", "unclassified"))
+
+use <- use[,c("minutes","mets","agecat2","GESTFIPS","sexcat","regcat","metcat")]
+
+#head(use)
+
+use.region <- ddply(use, .(sexcat, agecat2, regcat, metcat), summarise, 
+             mean_min = mean(minutes, na.rm=T),
+             mean_mets = mean(mets, na.rm=T),
+             n = length(minutes)
+)
+
+use.state <- ddply(use, .(sexcat, agecat2, GESTFIPS, metcat), summarise, 
+             mean_min = mean(minutes, na.rm=T),
+             mean_mets = mean(mets, na.rm=T),
+             n = length(minutes)
+)
+
+use2 <- subset(use, metcat == "metro")
+
+use.state2 <- ddply(use2, .(sexcat, agecat2, GESTFIPS), summarise, 
+              percentile10 <- quantile(mets, 0.10, na.rm=T), 
+              percentile30 <- quantile(mets, 0.3, na.rm=T),  
+              percentile50 <- quantile(mets, 0.5, na.rm=T),
+              percentile70 <- quantile(mets, 0.7, na.rm=T), 
+              percentile90 <- quantile(mets, 0.9, na.rm=T), 
+              mean <- mean(mets, na.rm=T), 
+              n = length(mets) 
+)
+
+use.region2 <- ddply(use2, .(sexcat, agecat2, regcat), summarise, 
+              percentile10 <- quantile(mets, 0.10, na.rm=T), 
+              percentile30 <- quantile(mets, 0.3, na.rm=T),  
+              percentile50 <- quantile(mets, 0.5, na.rm=T),
+              percentile70 <- quantile(mets, 0.7, na.rm=T), 
+              percentile90 <- quantile(mets, 0.9, na.rm=T), 
+              mean <- mean(mets, na.rm=T), 
+              n = length(mets) 
+)
+
+state.table <- subset(use.state, metcat == "metro")
+
+#write.csv(state.table, stateResultFile, row.names=F)
+
+#write.csv(use.state2, stateResultFile2, row.names=F)
+
+#write.csv(use.region2, regionResultFile, row.names=F)
+
+#ggplot(use.region, aes(x=factor(sexcat), y=mean_mets, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1) +facet_grid(regcat ~ metcat)
+
+#ggplot(subset(use.region, metcat != "unclassified"), aes(x=factor(sexcat), y=mean_mets, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1, position="dodge") +facet_grid(regcat ~ metcat)
+
+#ggplot(subset(use.state, metcat != "unclassified" & GESTFIPS == 17), aes(x=factor(sexcat), y=mean_min, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1, position="dodge") +facet_grid(GESTFIPS ~ metcat)
+```
+
+```{r figures, echo = FALSE, eval = TRUE}
+
+zeroMETs <- use$minutes == 0
+
+ggplot(use[!zeroMETs,], aes(x = mets, ..density.., colour = agecat2)) + geom_freqpoly(binwidth=200) + facet_grid(. ~ sexcat) + coord_cartesian(xlim = c(0, 1500))
+
+```
+
+Note that `r round(100*sum(zeroMETs)/length(zeroMETs),1)`% of the
+entries were zero minutes and the above densities are for non-zero
+METs only.
 

### 2016-06-29 vargo/nontravel_METs_from_ATUS.r

@@ -9,14 +9,16 @@ library(plyr)
 
 # activity <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusact_2003.dat", header=T, skip=0, sep=",")
 
-summary <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_0311.dat", header=T, skip=0, sep=",")
+summaryFile <- "~/GHI/data/atussum_0311.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_0311.dat
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
 
 # respondants <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusresp_2003.dat", header=T, skip=0, sep=",")
 
 # roster <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusrost_2003.dat", header=T, skip=0, sep=",")
 
+cpsFile <- "~/GHI/data/atuscps_0311.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_0311.dat
 
-CPS <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_0311.dat", header=T, skip=0, sep=",")
+CPS <- read.table(cpsFile, header=T, skip=0, sep=",")
 
 CPS.2 <- CPS[,c("TUCASEID", "GESTFIPS", "GEREG","GEMETSTA")]
 CPS.2$IDCHAR <- as.character(CPS.2[,"TUCASEID"])
@@ -94,7 +96,6 @@ use.state2 <- ddply(use2, .(sexcat, agecat2, GESTFIPS), summarise,
               n = length(mets) 
 )
 
-
 use.region2 <- ddply(use2, .(sexcat, agecat2, regcat), summarise, 
               percentile10 <- quantile(mets, 0.10, na.rm=T), 
               percentile30 <- quantile(mets, 0.3, na.rm=T),  
@@ -110,17 +111,24 @@ use.region2 <- ddply(use2, .(sexcat, agecat2, regcat), summarise,
 
 
 state.table <- subset(use.state, metcat == "metro")
[…]

### 2016-06-30 R/ITHIM.Rmd
Added logNormal approximation to nonzero, non-travel MET distribution

@@
 METs only.
+
+```{r nonZeroMETsLogNormal, echo = FALSE, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE}
+parameters <- ddply(use[!zeroMETs,], .(sexcat, agecat2), summarise,
+             mean_min = mean(minutes, na.rm=T),
+             mean_mets = mean(mets, na.rm=T),
+             sd_min = sd(minutes, na.rm=T),
+             sd_mets = sd(mets, na.rm=T)
+)
+cv_mets <- with(parameters, sd_mets/mean_mets)
+cv_min <- with(parameters, sd_min/mean_min)
+cv0_mets<- mean(cv_mets)
+
+mean_mets <- parameters$mean_mets
+sd_mets <- cv0_mets*mean_mets
+
+getLogNormal <- function(mu,sd){
+    dlnorm(seq(0,2000,length.out=1e3), log(mu/sqrt(1+sd^2/mu^2)), sqrt(log(1+sd^2/mu^2)))
+}
+
+foo <- as.data.frame(mapply(getLogNormal, mean_mets, sd_mets))
+names(foo) <- paste0(parameters$sexcat,parameters$agecat2)
+foo <- melt(foo)
+foo <- data.frame(foo, x = seq(0,2000,length.out = 1e3), sex = c(matrix(c("F","M"),nrow= nrow(foo)/2, ncol = 2, byrow = TRUE)), agecat = c(matrix(c("15-29","30-44","45-59","60-69","70-79", "80+"),nrow= nrow(foo)/12, ncol = 12, byrow = TRUE)))
+
+ggplot(foo, aes(x,value, colour = agecat)) + geom_line() + facet_grid( . ~ sex) + coord_cartesian(xlim = c(0, 1500))
+
+```
 

### 2016-06-30 R/ITHIM.Rmd
Update to ITHIM.Rmd

@@
 ggplot(foo, aes(x,value, colour = agecat)) + geom_line() + facet_grid( . ~ sex) + coord_cartesian(xlim = c(0, 1500))
+
+```
+
+```{r matrices}
+nonTravelMETs <- rbind(0,rbind(0,cbind(parameters$mean_mets[7:12], parameters$mean_mets[1:6])))
+nonTravelMETsNormalized <- nonTravelMETs/nonTravelMETs[3,2]
+meanNonTravelMETs <- nonTravelMETs[3,2]
+
+```
+
+```{r plotMatrix}
+D <- melt(nonTravelMETsNormalized)
+ggplot(D, aes(Var2, value)) + geom_bar(aes(fill=Var1), stat = "identity", position = "dodge")
+```
+
+```{r fooFunction}
+
+probActive <- 0.15
+
+METQuintiles <- mapply(getMETQuintiles, nonTravelMETsNormalized*nonTravelMETs, MoreArgs = list(cv = cv0_mets, p = probActive, size = 1e4), SIMPLIFY = FALSE)
 

### 2016-07-11 R/ITHIM.Rmd
First look at MET Key data

--- 2016-06-30 R/ITHIM.Rmd
@@
 
-```{r vargo, echo = FALSE, eval = TRUE}
+```{r vargo, echo = FALSE, eval = FALSE}
 library(ggplot2)
@@
 metkey <- as.matrix(c(3.0,7.3,5.0,6.5,7.5,3.5,5.8,7.8,6.0,8.0,3.8,7.8,8.0,5.3,7.3,9.8,6.3,7.0,7.0,5.0,6.8,3.0,3.5,5.5,6.0,4.0,6.0,3.0,3.0,3.0), nrow = 1, ncol=30)
+
 summary.2c <- mapply(`*`,summary.2b,metkey)
@@
 
-```{r figures, echo = FALSE, eval = TRUE}
+```{r figures, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
 
@@
 
+<!--
      Note that `r round(100*sum(zeroMETs)/length(zeroMETs),1)`% of the
@@
 METs only.
+-->
 
-```{r nonZeroMETsLogNormal, echo = FALSE, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE}
+```{r nonZeroMETsLogNormal, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
 parameters <- ddply(use[!zeroMETs,], .(sexcat, agecat2), summarise,
@@
 
+```
+
+```{r matrices, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
+nonTravelMETs <- rbind(0,rbind(0,cbind(parameters$mean_mets[7:12], parameters$mean_mets[1:6])))
+nonTravelMETsNormalized <- nonTravelMETs/nonTravelMETs[3,2]
+meanNonTravelMETs <- nonTravelMETs[3,2]
+
+```
+
+```{r plotMatrix, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
+D <- melt(nonTravelMETsNormalized)
+ggplot(D, aes(Var2, value)) + geom_bar(aes(fill=Var1), stat = "identity", position = "dodge")
+```
+
+```{r fooFunction, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
+probActive <- 0.15
+METQuintiles <- mapply(getMETQuintiles, nonTravelMETsNormalized*nonTravelMETs, MoreArgs = list(cv = cv0_mets, p = probActive, size = 1e4), SIMPLIFY = FALSE)
+```
+
+```{r newMETs}
+newMETsFile <- "~/GHI/data/ATUS_to_METS_key.csv"
+newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
+dataColumns <- c(1,5)
+D <- newMETs[,dataColumns]
+metaData <- newMETs[,-dataColumns]
+newMETs <- list(Data = D, metaData = metaData)
+```
+
+```{r readnewMETs}
+summaryFile <- "~/GHI/data/atussum_0311.dat"
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
+metaDataColumns <- 1:24
+summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
+activities <- colnames(summaryList$Data)
+metKey <- newMETs$Data$AinsworthMETS
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+names(metKey) <- activityCode
+activities[!(activities %in% names(metKey))]
+names(metKey)[!(names(metKey) %in% activities)]
+```
+
+```{r metKey}
+metKeyVector <- metKey[activities]/60*7
+names(metKeyVector) <- activities
+metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
+METs <- rowSums(summaryList$Data * metKeyMatrix, na.rm = TRUE)
+hist(METs, breaks = 75)
 ```

### 2016-07-11 R/ITHIM.Rmd
mapply with getTotalDistribution
Worked on importing ATUS data

@@
 
-<!--
-     Note that `r round(100*sum(zeroMETs)/length(zeroMETs),1)`% of the
-entries were zero minutes and the above densities are for non-zero
-METs only.
--->
-
 ```{r nonZeroMETsLogNormal, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
@@
 ```{r readnewMETs}
+cpsFile <- "~/GHI/data/atuscps_0311.dat"
+cps <- read.table(cpsFile, header=T, skip=0, sep="," )
+cps <- within(cps, {
+    TUCASEID <- as.character(TUCASEID)
+    TULINENO <- as.character(TULINENO)
+    GEREG <- as.factor(GEREG)
+    })
+cps <- subset(cps, TULINENO == "1")
+
 summaryFile <- "~/GHI/data/atussum_0311.dat"
@@
 metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-METs <- rowSums(summaryList$Data * metKeyMatrix, na.rm = TRUE)
+METs <- data.frame(METs = rowSums(summaryList$Data * metKeyMatrix, na.rm = TRUE), AGE = summaryList$metaData$TEAGE, row.names = summaryList$metaData$TUCASEID)
 hist(METs, breaks = 75)
+```
+
+```{r dists}
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = GEREG)) + geom_freqpoly()
+```
+
+### Total Distribution
+
+```{r total}
+foo <- mapply(getTotalDistribution, muTravel = rep(37,2), cvTravel = rep(1.72,2), muNonTravel = rep(10,2), cvNonTravel = rep(1,2), p = rep(0.5,2), pWalk = rep(54.4/64.2,2), vWalk = rep(2.88,2), size = 1e4, SIMPLIFY = FALSE)
 ```

### 2016-07-12 R/ITHIM.Rmd
Update to MET distribution code.

--- 2016-07-11 R/ITHIM.Rmd
@@
 summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
+
+summaryList$metaData <- within(summaryList$metaData,{
+    AGECAT <- ifelse(TEAGE <5, "age0-4",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "age5-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "age15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "age30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "age45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "age60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "age70-79", "age80+"
+                            )))))))})
+
 activities <- colnames(summaryList$Data)
@@
 names(metKey) <- activityCode
-activities[!(activities %in% names(metKey))]
-names(metKey)[!(names(metKey) %in% activities)]
+rownames(newMETs$metaData) <- activityCode
 ```
@@
 metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-METs <- rowSums(summaryList$Data * metKeyMatrix, na.rm = TRUE)
-hist(METs, breaks = 75)
+
+METs <- data.frame(METs = rowSums(summaryList$Data * metKeyMatrix, na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID)
+
+```
+
+```{r dists}
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly() + facet_grid( SEX ~ GEREG )
+```
+
+### Total Distribution
+
+```{r total, eval = FALSE, echo = FALSE}
+foo <- mapply(getTotalDistribution, muTravel = rep(37,2), cvTravel = rep(1.72,2), muNonTravel = rep(10,2), cvNonTravel = rep(1,2), p = rep(0.5,2), pWalk = rep(54.4/64.2,2), vWalk = rep(2.88,2), size = 1e4, SIMPLIFY = FALSE)
 ```

### 2016-07-12 R/ITHIM.Rmd
Continued to work on MET distribution

@@
 summaryList$metaData <- within(summaryList$metaData,{
+    TESEX <- as.factor(TESEX)
     AGECAT <- ifelse(TEAGE <5, "age0-4",
@@
 ```{r metKey}
-metKeyVector <- metKey[activities]/60*7
+metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
 names(metKeyVector) <- activities
@@
 
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly() + facet_grid( SEX ~ .)
+#ggplot(D, aes(x = METs, ..density.., colour = SEX)) + geom_freqpoly()
 ```
 
-```{r dists}
-D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly() + facet_grid( SEX ~ GEREG )
+```{r paameters}
+parameters <- ddply(D, .(SEX), summarise,
+             mean = mean(METs, na.rm=T),
+             sd = sd(METs, na.rm=T)
+             )
 ```

### 2016-07-18 R/ITHIM.Rmd
Computes nonTravelMET matrix

@@
 ```{r paameters}
-parameters <- ddply(D, .(SEX), summarise,
+parameters <- ddply(D, .(SEX, AGECAT), summarise,
              mean = mean(METs, na.rm=T),
@@
              )
+
+METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
+METmatrix <- METmatrix/METmatrix[3,2]
+
 ```

### 2016-07-20 R/ITHIM.Rmd
Subset activities on travel status

--- 2016-07-12 R/ITHIM.Rmd
@@
 library("plyr")
-#install("~/ITHIM/")
 ```
@@
 library("devtools")
-install_github("syounkin/ITHIM")
+install("~/ITHIM/")
+#install_github("syounkin/ITHIM")
 library("ITHIM")
@@
 
-```{r newMETs}
+```{r newMETs, echo = FALSE}
 newMETsFile <- "~/GHI/data/ATUS_to_METS_key.csv"
@@
 newMETs <- list(Data = D, metaData = metaData)
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+
+rownames(newMETs$Data) <- activityCode
+rownames(newMETs$metaData) <- activityCode
+
+newMETs$metaData <- within(newMETs$metaData, {
+  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
+})
+
+metKey <- newMETs$Data$AinsworthMETS
+names(metKey) <- activityCode
+
 ```
 
-```{r readnewMETs}
+```{r readnewMETs, echo = FALSE}
 cpsFile <- "~/GHI/data/atuscps_0311.dat"
@@
 summaryList$metaData <- within(summaryList$metaData,{
-    AGECAT <- ifelse(TEAGE <5, "age0-4",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "age5-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "age15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "age30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "age45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "age60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "age70-79", "age80+"
+    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
+    AGECAT <- ifelse(TEAGE <5, "00-04",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
                             )))))))})
-
-activities <- colnames(summaryList$Data)
-metKey <- newMETs$Data$AinsworthMETS
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-names(metKey) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-```
-
-```{r metKey}
-metKeyVector <- metKey[activities]/60*7
-names(metKeyVector) <- activities
-metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-
-METs <- data.frame(METs = rowSums(summaryList$Data * metKeyMatrix, na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID)
 
@@
 
-```{r dists}
+We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
+
+```{r metKey, echo = FALSE}
+allActivities <- colnames(summaryList$Data)
+nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
+nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
+
+activities <- nonTravelActivities
+
+metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
+names(metKeyVector) <- activities
+metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
+dimnames(metKeyMatrix) <- list( NULL, activities)
+
+METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
+
 D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly() + facet_grid( SEX ~ GEREG )
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
+
+```
+
+```{r parameters}
+(parameters <- ddply(D, .(SEX, AGECAT), summarise,
+             mean = mean(METs, na.rm=T),
+             sd = sd(METs, na.rm=T)
+             ))
 ```
@@
 ```{r total, eval = FALSE, echo = FALSE}
-foo <- mapply(getTotalDistribution, muTravel = rep(37,2), cvTravel = rep(1.72,2), muNonTravel = rep(10,2), cvNonTravel = rep(1,2), p = rep(0.5,2), pWalk = rep(54.4/64.2,2), vWalk = rep(2.88,2), size = 1e4, SIMPLIFY = FALSE)
+foo <- mapply(getTotalDistribution,
+  muTravel = matrix(35,nrow = 8, ncol = 2),
+  cvTravel = 1.72,
+  muNonTravel = matrix(100,nrow = 8, ncol = 2),
+  cvNonTravel = 1,
+  p = 1,
+  pWalk = rep(54.4/64.2,2),
+  vWalk = rep(2.88,2),
+  size = 1e5, SIMPLIFY = FALSE)
 ```

### 2016-07-20 R/ITHIM.Rmd

@@
 
-```{r parameters}
-(parameters <- ddply(D, .(SEX, AGECAT), summarise,
+```{r paameters}
+parameters <- ddply(D, .(SEX, AGECAT), summarise,
              mean = mean(METs, na.rm=T),
              sd = sd(METs, na.rm=T)
-             ))
+             )
+
+METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
+METmatrix <- METmatrix/METmatrix[3,2]
 ```

### 2016-07-21 R/ITHIM.Rmd
Update to webpage and created Makefile

--- 2016-07-20 R/ITHIM.Rmd
@@
 
-```{r parameters}
+```{r paameters}
 (parameters <- ddply(D, .(SEX, AGECAT), summarise,
@@
              ))
+
+METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
+(METmatrix <- METmatrix/METmatrix[3,2])
 ```

### 2016-07-21 R/ITHIM.Rmd
Install from local package, not GitHub

--- 2016-07-20 R/ITHIM.Rmd
@@
 library("plyr")
+
 ```
@@
 library("devtools")
+#install_github("syounkin/ITHIM", ref="devel")
 install("~/ITHIM/")
-#install_github("syounkin/ITHIM")
 library("ITHIM")
@@
 
-```{r newMETs, echo = FALSE}
+```{r newMETs}
 newMETsFile <- "~/GHI/data/ATUS_to_METS_key.csv"
@@
 newMETs <- list(Data = D, metaData = metaData)
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-
-rownames(newMETs$Data) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-
-newMETs$metaData <- within(newMETs$metaData, {
-  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
-})
-
-metKey <- newMETs$Data$AinsworthMETS
-names(metKey) <- activityCode
-
 ```
 
-```{r readnewMETs, echo = FALSE}
+```{r readnewMETs}
 cpsFile <- "~/GHI/data/atuscps_0311.dat"
@@
 summaryList$metaData <- within(summaryList$metaData,{
-    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
-    AGECAT <- ifelse(TEAGE <5, "00-04",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
+    TESEX <- as.factor(TESEX)
+    AGECAT <- ifelse(TEAGE <5, "age0-4",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "age5-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "age15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "age30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "age45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "age60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "age70-79", "age80+"
                             )))))))})
 
+activities <- colnames(summaryList$Data)
+metKey <- newMETs$Data$AinsworthMETS
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+names(metKey) <- activityCode
+rownames(newMETs$metaData) <- activityCode
 ```
 
-We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
-
-```{r metKey, echo = FALSE}
-allActivities <- colnames(summaryList$Data)
-nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
-nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
-
-activities <- nonTravelActivities
-
+```{r metKey}
 metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
@@
 metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-dimnames(metKeyMatrix) <- list( NULL, activities)
 
-METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
+METs <- data.frame(METs = rowSums(summaryList$Data * metKeyMatrix, na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID)
 
 D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
-
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly() + facet_grid( SEX ~ .)
+#ggplot(D, aes(x = METs, ..density.., colour = SEX)) + geom_freqpoly()
 ```
@@
 METmatrix <- METmatrix/METmatrix[3,2]
+
 ```
@@
 ```{r total, eval = FALSE, echo = FALSE}
-foo <- mapply(getTotalDistribution,
-  muTravel = matrix(35,nrow = 8, ncol = 2),
-  cvTravel = 1.72,
-  muNonTravel = matrix(100,nrow = 8, ncol = 2),
-  cvNonTravel = 1,
-  p = 1,
-  pWalk = rep(54.4/64.2,2),
-  vWalk = rep(2.88,2),
-  size = 1e5, SIMPLIFY = FALSE)
+foo <- mapply(getTotalDistribution, muTravel = rep(37,2), cvTravel = rep(1.72,2), muNonTravel = rep(10,2), cvNonTravel = rep(1,2), p = rep(0.5,2), pWalk = rep(54.4/64.2,2), vWalk = rep(2.88,2), size = 1e4, SIMPLIFY = FALSE)
 ```

### 2016-07-29 R/ITHIM.Rmd
Introduced TODO.org and NHTS data

@@
 
+### NHTS
+
+```{r NHTS}
+foo <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = "-9")
+names(foo) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
+
+foo <- within(foo, {
+    URBRUR <- as.factor(URBRUR)
+    HHSTFIPS <- as.factor(HHSTFIPS)
+    MSACAT <- as.factor(MSACAT)
+    })
+
+```
 <!--

### 2016-08-02 R/ITHIM.Rmd
Included NHTS data in ITHIM.Rmd

--- 2016-07-21 R/ITHIM.Rmd
@@
 library("devtools")
-#install_github("syounkin/ITHIM", ref="devel")
+install_github("smbache/magrittr", ref="master")
+library("magrittr")
 install("~/ITHIM/")
@@
 
+### NHTS
+
+```{r NHTS}
+DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = "-9")
+names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
+
+DAYV2PUB <- within(DAYV2PUB, {
+    URBRUR <- as.factor(URBRUR)
+    HHSTFIPS <- as.factor(HHSTFIPS)
+    MSACAT <- as.factor(MSACAT)
+    TRPTRANS <- as.factor(TRPTRANS)
+    })
+
+HHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/HHV2PUB.CSV", na.strings = "-9")
+names(HHV2PUB) <- c("HOUSEID","VARSTRAT","WTHHFIN","DRVRCNT","CDIVMSAR","CENSUS_D","CENSUS_R","HH_HISP","HH_RACE","HHFAMINC","HHRELATD","HHRESP","HHSIZE","HHSTATE","HHSTFIPS","HHVEHCNT","HOMEOWN","HOMETYPE","MSACAT","MSASIZE","NUMADLT","RAIL","RESP_CNT","SCRESP","TRAVDAY","URBAN","URBANSIZE","URBRUR","WRKCOUNT","TDAYDATE","FLAG100","LIF_CYC","CNTTDHH","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
+
+PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
+names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
+
+VEHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/VEHV2PUB.CSV", na.strings = "-9")
+names(VEHV2PUB) <- c("HOUSEID","WTHHFIN","VEHID","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","CDIVMSAR","CENSUS_D","CENSUS_R","HHSTATE","HHSTFIPS","HYBRID","MAKECODE","MODLCODE","MSACAT","MSASIZE","OD_READ","RAIL","TRAVDAY","URBAN","URBANSIZE","URBRUR","VEHCOMM","VEHOWNMO","VEHYEAR","WHOMAIN","WRKCOUNT","TDAYDATE","VEHAGE","PERSONID","HH_HISP","HH_RACE","HOMEOWN","HOMETYPE","LIF_CYC","ANNMILES","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","BEST_FLG","BESTMILE","BEST_EDT","BEST_OUT","FUELTYPE","GSYRGAL","GSCOST","GSTOTCST","EPATMPG","EPATMPGF","EIADMPG","VEHTYPE","HH_CBSA","HHC_MSA")
+
+NHTS.trips <- data.frame(HOUSEID = DAYV2PUB$HOUSEID,
+                   PERSONID = DAYV2PUB$PERSONID,
+                   TDCASEID = DAYV2PUB$TDCASEID,
+                   AT = ifelse(DAYV2PUB$TRPTRANS == "22", "cycle", ifelse(DAYV2PUB$TRPTRANS == "23","walk",ifelse(!is.na(DAYV2PUB$TRPTRANS),"other",NA))),
+                   TRVL_MIN = as.numeric(DAYV2PUB$TRVL_MIN),
+                   TRVLCMIN = as.numeric(DAYV2PUB$TRVLCMIN)
+                   )
+
+convertToAgeClass <- function(age){
+  age <- as.numeric(age)
+  agecat <- ifelse( age <= 4, "00-04", ifelse( age <= 14, "05-14", ifelse( age <= 29, "15-29", ifelse( age <= 44, "30-44", ifelse( age <= 59, "45-59", ifelse( age <= 69, "60-69", ifelse( age <= 79, "70-79", ifelse( age > 80, "80+", NA))))))))
+  return(agecat)
+}
+
+NHTS.person <- with(PERV2PUB,{
+                    data.frame(HOUSEID = HOUSEID,
+                               PERSONID = PERSONID,
+                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
+                               AGE = as.factor(convertToAgeClass(R_AGE))
+)
+})
+
+NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
+
+NHTS <- subset(NHTS, !is.na(AGE) & !is.na(SEX))
+
+(foo <- ddply(NHTS, .(SEX, AGE, AT), summarise,
+             sum = sum(TRVL_MIN, na.rm=T)
+             ))
+
+ggplot(NHTS, aes(x = TRVL_MIN, ..density.., colour = AGE)) + geom_freqpoly(binwidth = 15) + facet_grid( SEX ~ AT) + labs(x = "Foo", colour = "Age") + coord_cartesian(xlim = c(0, 75))
+
+countPeople <- function(data){
+ nrow(data %>% group_by(HOUSEID, PERSONID) %>% summarize(n()))
+}
+
+```
 <!--

### 2016-08-02 R/ITHIM.Rmd
Minor update to ITHIM.Rmd

@@
 library("reshape2")
-library("plyr")
+library("dplyr")
 ```
@@
 
-convertToAgeClass <- function(age){
-  age <- as.numeric(age)
-  agecat <- ifelse( age <= 4, "00-04", ifelse( age <= 14, "05-14", ifelse( age <= 29, "15-29", ifelse( age <= 44, "30-44", ifelse( age <= 59, "45-59", ifelse( age <= 69, "60-69", ifelse( age <= 79, "70-79", ifelse( age > 80, "80+", NA))))))))
-  return(agecat)
-}
-
 NHTS.person <- with(PERV2PUB,{
@@
 
-(foo <- ddply(NHTS, .(SEX, AGE, AT), summarise,
-             sum = sum(TRVL_MIN, na.rm=T)
-             ))
+ggplot(NHTS, aes(x = TRVL_MIN, ..density.., colour = AGE)) + geom_freqpoly(binwidth = 15) + facet_grid( SEX ~ AT) + labs(x = "Foo", colour = "Age") + coord_cartesian(xlim = c(0, 75))
 
-ggplot(NHTS, aes(x = TRVL_MIN, ..density.., colour = AGE)) + geom_freqpoly(binwidth = 15) + facet_grid( SEX ~ AT) + labs(x = "Foo", colour = "Age") + coord_cartesian(xlim = c(0, 75))
+NHTS %>% group_by(SEX, AGE) %>% summarize(trips = n())
 

### 2016-08-04 R/create_HHS-metro_Burdens.r
Added Vargo's R code for burden values by HHS region (metro areas)

@@ -1,0 +1,210 @@
+
+setwd("~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/")
+
+#######DETAILS ON WONDER QUERY#######
+# ---
+#   Dataset: Underlying Cause of Death, 1999-2014
+# Query Parameters:
+#   Title: ITHIM2014_HHS_urb_age_sex_16years
+# 2013 Urbanization: All
+# Autopsy: All
+# Five-Year Age Groups: All
+# Gender: Female, Male
+# HHS Regions: All
+# Hispanic Origin: All
+# ICD-10 113 Cause List: Malignant neoplasms of colon, rectum and anus (C18-C21), Malignant neoplasms of trachea, bronchus and
+# lung (C33-C34), Malignant neoplasm of breast (C50), #Diabetes mellitus (E10-E14), #Alzheimer's disease (G30), Other forms of
+# chronic ischemic heart disease (I20,I25), Acute and subacute endocarditis (I33), Diseases of pericardium and acute myocarditis
+# (I30-I31,I40), #Essential hypertension and hypertensive renal disease (I10,I12,I15), #Cerebrovascular diseases (I60-I69),
+# #Influenza and pneumonia (J09-J18), Other acute lower respiratory infections (J20-J22,U04), #Chronic lower respiratory diseases
+# (J40-J47), Other diseases of respiratory system (J00-J06,J30- J39,J67,J70-J98), Motor vehicle accidents
+# (V02-V04,V09.0,V09.2,V12-V14,V19.0-V19.2,V19.4-V19.6,V20-V79,V80.3-V80.5,V81.0-V81.1,V82.0-V82.1,V83-V86,V87.0-V87.8,V88.0-V88.8
+#  V89.0,V89.2)
+# Other land transport accidents
+# (V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
+#  V89.1,V89.3,V89.9)
+# 
+# Place of Death: All
+# Race: All
+# Weekday: All
+# Year/Month: All
+# Group By: Five-Year Age Groups, Gender, HHS Region, 2013 Urbanization, ICD-10 113 Cause List
+# Show Totals: Disabled
+# Show Zero Values: True
+# Show Suppressed: True
+# Calculate Rates Per: 100,000
[…]

### 2016-08-04 R/create_HHS-metro_Burdens.r
Checked to make sure Vargo's code runs.  It does.

@@ -1,5 +1,5 @@
-
-setwd("~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/")
+library("plyr")
+#setwd("~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/")
 
 #######DETAILS ON WONDER QUERY#######
 # ---
@@ -23,7 +23,7 @@ setwd("~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/")
 # Other land transport accidents
 # (V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
 #  V89.1,V89.3,V89.9)
-# 
+#
 # Place of Death: All
 # Race: All
 # Weekday: All
@@ -94,8 +94,9 @@ setwd("~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/")
 #################################
 #wonder <- read.table("./ITHIM2014_HHS_urb_age_sex_16years.txt", header=T, sep="\t",colClasses=c("Notes"="null"), strip.white=T)
 
-wonder <- read.csv("./HHSmetro_Age_sex_2010-14.csv", header=T)
-names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")    
+#wonder <- read.csv("./HHSmetro_Age_sex_2010-14.csv", header=T)
+wonder <- read.csv("./../data/burden/HHSmetro_Age_sex_2010-14.csv", header=T)
+names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
 
 foo <- unique(wonder$HHSRegion)
 foo2 <- unique(wonder$HHSRegionCode)
@@ -125,7 +126,7 @@ wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
 wonder <- wonder[, c("HHSRegion","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
 
[…]

### 2016-08-04 R/create_HHS-metro_Burdens.r
Small update to CDC code.

@@ -95,7 +95,7 @@ library("plyr")
 #wonder <- read.table("./ITHIM2014_HHS_urb_age_sex_16years.txt", header=T, sep="\t",colClasses=c("Notes"="null"), strip.white=T)
 
 #wonder <- read.csv("./HHSmetro_Age_sex_2010-14.csv", header=T)
-wonder <- read.csv("./../data/burden/HHSmetro_Age_sex_2010-14.csv", header=T)
+wonder <- read.csv("./../data/CDC/HHSmetro_Age_sex_2010-14.csv", header=T)
 names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
 
 foo <- unique(wonder$HHSRegion)
@@ -151,7 +151,7 @@ wonderSummary <- ddply(wonder2, .(HHSRegion, gender, ITHIMage, ITHIMcause), summ
 
 #read in data for single year population
 #processed in Excel, to combine two WONDER queries, need to go back and script the processing of the raw WONDER output
-pop14 <- read.csv("./../data/burden/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
+pop14 <- read.csv("./../data/CDC/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
 sum(as.numeric(as.character(pop14$Population)),na.rm=T)  #checking US 'Metro' population here 272,667,942
 sum(as.numeric(as.character(pop14$Deaths)),na.rm=T)  #checking US 'Metro' deaths here 2,125,273
 
@@ -180,7 +180,7 @@ sum(popSummary$AllCauseDeaths2014)
 
 HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
 
-USGBD <- read.csv("./../data/burden/USGBD.csv", header=T)
+USGBD <- read.csv("./../data/CDC/USGBD.csv", header=T)
 USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
 USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)
 
@@ -208,4 +208,3 @@ names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly
 burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
 
 write.csv(burdenOut, "./HHSburdens.csv", row.names=F)
-

### 2016-08-04 R/ITHIM.Rmd
Updated NHTS data parsing effort

@@
 library("reshape2")
+library("plyr")
 library("dplyr")
@@
 
+nMatrix <- ddply( subset(NHTS.person, !is.na(AGE) & !is.na(SEX)), .(SEX, AGE), summarise, count = length(PERSONID))
+sum(nMatrix$count) # 305649 number of people in the PERSON table
+
+foo <- group_by(NHTS.trips, HOUSEID,PERSONID)
+foobar <- dplyr::summarise(foo, n = n())
+nrow(foobar) # 262,934 number of people with trip data
+
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
 
-NHTS <- subset(NHTS, !is.na(AGE) & !is.na(SEX))
+NHTS %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarise( n = n()) # 260,576 people in merged data
+
+#NHTS <- subset(NHTS, !is.na(AGE) & !is.na(SEX))
 
@@
 
-NHTS %>% group_by(SEX, AGE) %>% summarize(trips = n())
+NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))
 
 countPeople <- function(data){
- nrow(data %>% group_by(HOUSEID, PERSONID) %>% summarize(n()))
+ nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
 }
+
+foo <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
+counts <- lapply(foo, function(x) lapply(x, countPeople))
+
+```
+
+```{r CDC}
+#setwd("~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/")
+
+#######DETAILS ON WONDER QUERY#######
+# ---
+#   Dataset: Underlying Cause of Death, 1999-2014
+# Query Parameters:
+#   Title: ITHIM2014_HHS_urb_age_sex_16years
+# 2013 Urbanization: All
+# Autopsy: All
+# Five-Year Age Groups: All
+# Gender: Female, Male
+# HHS Regions: All
+# Hispanic Origin: All
+# ICD-10 113 Cause List: Malignant neoplasms of colon, rectum and anus (C18-C21), Malignant neoplasms of trachea, bronchus and
+# lung (C33-C34), Malignant neoplasm of breast (C50), #Diabetes mellitus (E10-E14), #Alzheimer's disease (G30), Other forms of
+# chronic ischemic heart disease (I20,I25), Acute and subacute endocarditis (I33), Diseases of pericardium and acute myocarditis
+# (I30-I31,I40), #Essential hypertension and hypertensive renal disease (I10,I12,I15), #Cerebrovascular diseases (I60-I69),
+# #Influenza and pneumonia (J09-J18), Other acute lower respiratory infections (J20-J22,U04), #Chronic lower respiratory diseases
+# (J40-J47), Other diseases of respiratory system (J00-J06,J30- J39,J67,J70-J98), Motor vehicle accidents
+# (V02-V04,V09.0,V09.2,V12-V14,V19.0-V19.2,V19.4-V19.6,V20-V79,V80.3-V80.5,V81.0-V81.1,V82.0-V82.1,V83-V86,V87.0-V87.8,V88.0-V88.8
+#  V89.0,V89.2)
+# Other land transport accidents
+# (V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
+#  V89.1,V89.3,V89.9)
+#
+# Place of Death: All
+# Race: All
+# Weekday: All
+# Year/Month: All
+# Group By: Five-Year Age Groups, Gender, HHS Region, 2013 Urbanization, ICD-10 113 Cause List
+# Show Totals: Disabled
+# Show Zero Values: True
+# Show Suppressed: True
+# Calculate Rates Per: 100,000
+# Rate Options: Default intercensal populations for years 2001-2009 (except Infant Age Groups)
+# ---
+#   Help: See http://wonder.cdc.gov/wonder/help/ucd.html for more information.
+# ---
+#   Query Date: Jul 19, 2016 4:46:10 PM
+# ---
+#   Suggested Citation: Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
+# 1999-2014 on CDC WONDER Online Database, released 2015. Data are from the Multiple Cause of Death Files, 1999-2014, as compiled
+# from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program. Accessed at
+# http://wonder.cdc.gov/ucd-icd10.html on Jul 19, 2016 4:46:10 PM
+# ---
+#   Messages:
+#   1. Totals and Percent of Total are disabled when data are grouped by 113 or 130 Cause Lists. Check Caveats below for more
+# information.
+# 2. Selections made to any variable in [HHS Regions, 2013 Urbanization] cause equivalent selections in the other variables. More
+# Information: http://wonder.cdc.gov/wonder/help/faq.html#related-variables.
+# 3. Totals are not available for these results due to suppression constraints. More Information:
+#   http://wonder.cdc.gov/wonder/help/faq.html#Privacy.
+# ---
+#   Footnotes:
+#   1. A '#' symbol preceding the label indicates a rankable cause of death. More information.
+# ---
+#   Caveats:
+#   1. Totals and Percent of Total are disabled when data are grouped by a 113 or 130 Cause List because both aggregate and detailed
+# values are displayed in the table. Also be aware that charts and maps containing both aggregate and detail data could be
+# misleading.
+# 2. Population and rates are labeled 'Not Applicable' when they include a subset of ages 85-100+ because populations are not
+# available for those ages. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Ages 85-100.
+# 3. Circumstances in Georgia for the years 2008 and 2009 have resulted in unusually high death counts for the ICD-10 cause of
+# death code R99, "Other ill-defined and unspecified causes of mortality." Caution should be used in interpreting these data.
+# More information: http://wonder.cdc.gov/wonder/help/ucd.html#Georgia-Reporting-Anomalies.
+# 4. Circumstances in New Jersey for the year 2009 have resulted in unusually high death counts for the ICD-10 cause of death code
+# R99, "Other ill-defined and unspecified causes of mortality" and therefore unusually low death counts in other ICD-10 codes,
+# most notably R95, "Sudden Infant Death Syndrome" and X40-X49, "Unintentional poisoning." Caution should be used in
+# interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#New-Jersey-Reporting-Anomalies.
+# 5. Circumstances in California resulted in unusually high death counts for the ICD-10 cause of death code R99, "Other
+# ill-defined and unspecified causes of mortality" for deaths occurring in years 2000 and 2001. Caution should be used in
+# interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#California-Reporting-Anomalies.
+# 6. Data are Suppressed when the data meet the criteria for confidentiality constraints. More information:
+#   http://wonder.cdc.gov/wonder/help/ucd.html#Assurance of Confidentiality.
+# 7. Death rates are flagged as Unreliable when the rate is calculated with a numerator of 20 or less. More information:
+#   http://wonder.cdc.gov/wonder/help/ucd.html#Unreliable.
+# 8. Deaths of persons with Age "Not Stated" are included in "All" counts and rates, but are not distributed among age groups,
+# so are not included in age-specific counts, age-specific rates or in any age-adjusted rates. More information:
+#   http://wonder.cdc.gov/wonder/help/ucd.html#Not Stated.
+# 9. The population figures for year 2014 are bridged-race estimates of the July 1 resident population, from the Vintage 2014
+# postcensal series released by NCHS on June 30, 2015. The population figures for year 2013 are bridged-race estimates of the July
+# 1 resident population, from the Vintage 2013 postcensal series released by NCHS on June 26, 2014. The population figures for
+# year 2012 are bridged-race estimates of the July 1 resident population, from the Vintage 2012 postcensal series released by NCHS
+# on June 13, 2013. The population figures for year 2011 are bridged-race estimates of the July 1 resident population, from the
+# Vintage 2011 postcensal series released by NCHS on July 18, 2012. Population figures for 2010 are April 1 Census counts. The
+# population figures for years 2001 - 2009 are bridged-race estimates of the July 1 resident population, from the revised
+# intercensal county-level 2000 - 2009 series released by NCHS on October 26, 2012. Population figures for 2000 sssssare April 1 Census
+# counts. Population figures for 1999 are from the 1990-1999 intercensal series of July 1 estimates. Population figures for the
+# infant age groups are the number of live births. <br/><b>Note:</b> Rates and population figures for years 2001 - 2009 differ
+# slightly from previously published reports, due to use of the population estimates which were available at the time of release.
+# 10. The population figures used in the calculation of death rates for the age group 'under 1 year' are the estimates of the
+# resident population that is under one year of age. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Age Group.
+#################################
+#wonder <- read.table("./ITHIM2014_HHS_urb_age_sex_16years.txt", header=T, sep="\t",colClasses=c("Notes"="null"), strip.white=T)
+
+#wonder <- read.csv("./HHSmetro_Age_sex_2010-14.csv", header=T)
+wonder <- read.csv("./../data/CDC/HHSmetro_Age_sex_2010-14.csv", header=T)
+names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
+
+foo <- unique(wonder$HHSRegion)
+foo2 <- unique(wonder$HHSRegionCode)
+
+HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow=51,ncol=3)
+
+# remove 'Not stated' Age entries
+wonder <- subset(wonder, WONDERageCode !="NS")
+
+#recode cause of death for ITHIM
+ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD",
+                   "Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
+names(ITHIMcauseKey) <- unique(wonder$ICD10)
+wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
+
+#recode aage for ITHIM
+ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4",
+                 "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8",
+                 "ageClass8", "ageClass8", "ageClass8")
+names(ITHIMageKey) <- unique(wonder$WONDERage)
+wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
+
+wonder <- wonder[, c("HHSRegion","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
+
+#depression was not in the WONDER data, since there are no deaths, have to add it manually
+depress <- expand.grid(HHSRegion=unique(wonder$HHSRegion), gender=unique(wonder$gender),
+                       ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
+
+wonder2 <- rbind(wonder, depress)
+tail(wonder2)
+
+#replace 'Suppressed' values with random number 0-9
+wonder2$newDeaths <- as.numeric(ifelse(wonder2$deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(wonder2$deaths5yr))))
+class(wonder2$newDeaths) = "numeric"
+class(wonder2$population5yr) = "numeric"
+
+#summarize by HHS>Urbanization>Gender<age<cause
+wonderSummary <- ddply(wonder2, .(HHSRegion, gender, ITHIMage, ITHIMcause), summarise,
+                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
+)
+
+#read in data for single year population
+#processed in Excel, to combine two WONDER queries, need to go back and script the processing of the raw WONDER output
+pop14 <- read.csv("./../data/CDC/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
+sum(as.numeric(as.character(pop14$Population)),na.rm=T)  #checking US 'Metro' population here 272,667,942
+sum(as.numeric(as.character(pop14$Deaths)),na.rm=T)  #checking US 'Metro' deaths here 2,125,273
+
+names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
+
+#this is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' request
+ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5",
+                 "ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3",
+                 "ageClass8")
+names(ITHIMageKey2) <- unique(pop14$WONDERage)
+pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
+
+pop14 <-  pop14[, c("HHSRegion","gender","ITHIMage","deaths2014AllCause","population2014")]
+pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))
+
+sum(as.numeric(as.character(pop14$population2014)),na.rm=T)    #checking US 'Metro' population here 272,667,942
+sum(as.numeric(as.character(pop14$deaths2014)),na.rm=T)     #checking US 'Metro' deaths here 2,125,273
+
+popSummary <- ddply(pop14, .(HHSRegion, gender, ITHIMage), summarise,
+                    population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
+                   AllCauseDeaths2014 = sum(as.numeric(as.character(deaths2014AllCause)), na.rm=T)
+)
+
+sum(popSummary$population2014) #checking US population here 272,667,942
+sum(popSummary$AllCauseDeaths2014)
+
+HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
+
+USGBD <- read.csv("./../data/CDC/USGBD.csv", header=T)
+USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
+USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)
+
+burden <- merge(HHSmetroCauseSpecific, USGBD, all.x=T)
+burden$deaths <- burden$numberOfDeaths5yr/5
+burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
+burden$YLL <- burden$ratio * (burden$GBDyll/burden$GBDpop) * burden$population2014
+burden$YLD <- burden$ratio * (burden$GBDyld/burden$GBDpop) * burden$population2014
+burden$DALY <- burden$ratio * (burden$GBDdaly/burden$GBDpop) * burden$population2014
+
+#gah <- subset(burden, ITHIMcause == "RespiratoryDisease")
+
+#ggplot(gah, aes(x=ITHIMage, y=YLD, fill=factor(HHSRegion))) + geom_bar(stat="identity") + geom_point(aes(y=GBDyld)) + facet_grid(gender ~ .)
+#ggplot(gah, aes(x=ITHIMage, y=DALY)) + geom_bar(stat="identity") + geom_point(aes(y=GBDdaly)) + facet_grid(gender ~ HHSRegion)
+
+burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
+levels(burden$sex)  <- c("M","F")
+
+burden$sex
+
+burdenOut <- burden[,c("HHSRegion","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
+names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")
+
+burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
+
+write.csv(burdenOut, "./HHSburdens.csv", row.names=F)
 

### 2016-08-05 R/ITHIM.Rmd
First pass of NHTS data almost complete

@@
     URBRUR <- as.factor(URBRUR)
-    HHSTFIPS <- as.factor(HHSTFIPS)
+    HHSTATE <- as.factor(HHSTATE)
     MSACAT <- as.factor(MSACAT)
@@
 
+convertToHHS <- function(states){
+  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
+  StateToHHS <- HHSRegions[,3]
+  names(StateToHHS) <- HHSRegions[,1]
+  return(StateToHHS[states])
+}
+
+countPeople <- function(data){
+ nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
+}
+
+tripTime <- function(data, mode = "walk"){
+ return(sum(subset(data, AT == mode)$TRVLCMIN, na.rm = TRUE))
+}
+
+reformatData <- function(data){
+foo <- matrix(unlist(data), nrow = 7, ncol = 2, dimnames = list(paste0("ageClass",2:8), c("F","M")))
+return(foo[,c("M","F")])
+}
+
 NHTS.person <- with(PERV2PUB,{
@@
                                SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
-                               AGE = as.factor(convertToAgeClass(R_AGE))
+                               AGE = as.factor(convertToAgeClass(R_AGE)),
+                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
+                               HHS = as.factor(convertToHHS(HHSTATE))
+
 )
@@
 
+NHTS.person <- subset( NHTS.person, MSA == TRUE )
+
+NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
+
+listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
+
+counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
+walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
+cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
+
+countMatrix <- lapply(counts, reformatData)
+walkMatrix <- lapply(walkLength, reformatData)
+cycleMatrix <- lapply(cycleLength, reformatData)
+
+meanWalkList <- mapply(walkMatrix, countMatrix, FUN = "/", SIMPLIFY = FALSE)
+meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = "/", SIMPLIFY = FALSE)
+
+```
+
+```{r NHTSArchive, echo = FALSE, eval = FALSE}
 nMatrix <- ddply( subset(NHTS.person, !is.na(AGE) & !is.na(SEX)), .(SEX, AGE), summarise, count = length(PERSONID))
@@
 nrow(foobar) # 262,934 number of people with trip data
-
-NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
 
@@
 NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))
-
-countPeople <- function(data){
- nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
-}
-
-foo <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
-counts <- lapply(foo, function(x) lapply(x, countPeople))
-
 ```

### 2016-08-08 R/ITHIM.Rmd
Worked on Vargo's CDC code.

--- 2016-08-02 R/ITHIM.Rmd
@@
 ```
+
+## HHS Metro Region Parameter Sets
+
+First we use NHTS to estimate mean walking and cycling times.
+
+### NHTS
+
+#### Read Two of the Four NHTS Data Tables
+
+```{r NHTS}
+DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = "-9")
+names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
+
+DAYV2PUB <- within(DAYV2PUB, {
+    URBRUR <- as.factor(URBRUR)
+    HHSTATE <- as.factor(HHSTATE)
+    MSACAT <- as.factor(MSACAT)
+    TRPTRANS <- as.factor(TRPTRANS)
+})
+
+NHTS.trips <- data.frame(HOUSEID = DAYV2PUB$HOUSEID,
+                   PERSONID = DAYV2PUB$PERSONID,
+                   TDCASEID = DAYV2PUB$TDCASEID,
+                   AT = ifelse(DAYV2PUB$TRPTRANS == "22", "cycle",
+                        ifelse(DAYV2PUB$TRPTRANS == "23", "walk",
+                        ifelse(!is.na(DAYV2PUB$TRPTRANS), "other", NA))),
+                   TRVL_MIN = as.numeric(DAYV2PUB$TRVL_MIN),
+                   TRVLCMIN = as.numeric(DAYV2PUB$TRVLCMIN)
+                   )
+
+PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
+names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
+
+NHTS.person <- with(PERV2PUB,{
+                    data.frame(HOUSEID = HOUSEID,
+                               PERSONID = PERSONID,
+                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
+                               AGE = as.factor(convertToAgeClass(R_AGE)),
+                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
+                               HHS = as.factor(convertToHHS(HHSTATE))
+
+)
+})
+```
+#### Create a handful of functions to help parse data
+
+```{r parsely}
+convertToHHS <- function(states){
+  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
+  StateToHHS <- HHSRegions[,3]
+  names(StateToHHS) <- HHSRegions[,1]
+  return(StateToHHS[states])
+}
+
+countPeople <- function(data){
+ nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
+}
+
+tripTime <- function(data, mode = "walk"){
+ return(sum(subset(data, AT == mode)$TRVLCMIN, na.rm = TRUE))
+}
+
+reformatData <- function(data){
+foo <- matrix(unlist(data), nrow = 7, ncol = 2, dimnames = list(paste0("ageClass",2:8), c("F","M")))
+return(foo[,c("M","F")])
+}
+```
+
+#### Subset and Merge
+
+```{r subset}
+NHTS.person <- subset( NHTS.person, MSA == TRUE )
+
+NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
+```
+
+#### Split into strata by region
+
+```{r split}
+listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
+```
+
+#### Compute Mean Trip Lengths
+
+```{r count}
+counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
+walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
+cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
+
+countMatrix <- lapply(counts, reformatData)
+walkMatrix <- lapply(walkLength, reformatData)
+cycleMatrix <- lapply(cycleLength, reformatData)
+
+(meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE))
+(meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE))
+
+```
+
+### CDC Mortality
+
+Next we parse the output of a CDC Wonder query to estimate mortality rates.
+
+The data files in ~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/.
+
+#### Query Results File
+
+```{r CDC}
+wonder <- read.csv("./../data/CDC/HHSmetro_Age_sex_2010-14.csv", header=T)
+names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
+
+#HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow=51,ncol=3)
+
+wonder <- subset(wonder, WONDERageCode != "NS") # remove 'Not stated' Age entries
+```
+
+The following two recoding chunks should probably be redone.  They
+depend on the ordering of the factors which seems dangerous.
+
+```{r ITHIMCause}
+ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
+names(ITHIMcauseKey) <- unique(wonder$ICD10)
+wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
+```
+```{r recodeAge}
+ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
+names(ITHIMageKey) <- unique(wonder$WONDERage)
+wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
+```
+
+Depression was not in the WONDER data, since there are no deaths, have to add it manually.
+
+```{r wonder}
+wonder <- wonder[, c("HHSRegion","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
+depress <- expand.grid(HHSRegion=unique(wonder$HHSRegion), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
+wonder <- rbind(wonder, depress)
+```
+
+Replace suppressed values with random number 1-9
+
+```{r suppressed}
+wonder <- within( wonder, {
+    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(1:9), as.numeric(as.character(deaths5yr))))
+    })
+```
+
+10 HHS x 2 sex x 8 age x 13 disease = 2,080 strata.
+
+```{r summary}
+wonderSummary <- ddply(wonder2, .(HHSRegion, gender, ITHIMage, ITHIMcause), summarise,
+                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
+)
+```
+
+#### Single Year Data File
+
+```{r singleyear}
+#read in data for single year population
+#processed in Excel, to combine two WONDER queries, need to go back and script the processing of the raw WONDER output
+pop14 <- read.csv("./../data/CDC/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
+sum(as.numeric(as.character(pop14$Population)),na.rm=T)  #checking US 'Metro' population here 272,667,942
+sum(as.numeric(as.character(pop14$Deaths)),na.rm=T)  #checking US 'Metro' deaths here 2,125,273
+
+names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
+
+#this is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' request
+ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5",
+                 "ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3",
+                 "ageClass8")
+names(ITHIMageKey2) <- unique(pop14$WONDERage)
+pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
+
+pop14 <-  pop14[, c("HHSRegion","gender","ITHIMage","deaths2014AllCause","population2014")]
+pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))
+
+sum(as.numeric(as.character(pop14$population2014)),na.rm=T)    #checking US 'Metro' population here 272,667,942
+sum(as.numeric(as.character(pop14$deaths2014)),na.rm=T)     #checking US 'Metro' deaths here 2,125,273
+
+popSummary <- ddply(pop14, .(HHSRegion, gender, ITHIMage), summarise,
+                    population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
+                   AllCauseDeaths2014 = sum(as.numeric(as.character(deaths2014AllCause)), na.rm=T)
+)
+
+sum(popSummary$population2014) #checking US population here 272,667,942
+sum(popSummary$AllCauseDeaths2014)
+
+HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
+
+USGBD <- read.csv("./../data/CDC/USGBD.csv", header=T)
+USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
+USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)
+
+burden <- merge(HHSmetroCauseSpecific, USGBD, all.x=T)
+burden$deaths <- burden$numberOfDeaths5yr/5
+burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
+burden$YLL <- burden$ratio * (burden$GBDyll/burden$GBDpop) * burden$population2014
+burden$YLD <- burden$ratio * (burden$GBDyld/burden$GBDpop) * burden$population2014
+burden$DALY <- burden$ratio * (burden$GBDdaly/burden$GBDpop) * burden$population2014
+
+#gah <- subset(burden, ITHIMcause == "RespiratoryDisease")
+
+#ggplot(gah, aes(x=ITHIMage, y=YLD, fill=factor(HHSRegion))) + geom_bar(stat="identity") + geom_point(aes(y=GBDyld)) + facet_grid(gender ~ .)
+#ggplot(gah, aes(x=ITHIMage, y=DALY)) + geom_bar(stat="identity") + geom_point(aes(y=GBDdaly)) + facet_grid(gender ~ HHSRegion)
+
+burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
+levels(burden$sex)  <- c("M","F")
+
+burden$sex
+
+burdenOut <- burden[,c("HHSRegion","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
+names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")
+
+burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
+
+write.csv(burdenOut, "./HHSburdens.csv", row.names=F)
+
+```
+
+#### DETAILS ON WONDER QUERY
+
+Dataset: Underlying Cause of Death, 1999-2014
+Query Parameters:
+Title: ITHIM2014_HHS_urb_age_sex_16years
+2013 Urbanization: All
+Autopsy: All
+Five-Year Age Groups: All
+Gender: Female, Male
+HHS Regions: All
+Hispanic Origin: All
+ICD-10 113 Cause List: Malignant neoplasms of colon, rectum and anus (C18-C21), Malignant neoplasms of trachea, bronchus and
+lung (C33-C34), Malignant neoplasm of breast (C50), #Diabetes mellitus (E10-E14), #Alzheimer's disease (G30), Other forms of
+chronic ischemic heart disease (I20,I25), Acute and subacute endocarditis (I33), Diseases of pericardium and acute myocarditis
+(I30-I31,I40), #Essential hypertension and hypertensive renal disease
+(I10,I12,I15), #Cerebrovascular diseases (I60-I69), Influenza and pneumonia (J09-J18), Other acute lower respiratory infections (J20-J22,U04), #Chronic lower respiratory diseases
+(J40-J47), Other diseases of respiratory system (J00-J06,J30- J39,J67,J70-J98), Motor vehicle accidents
+(V02-V04,V09.0,V09.2,V12-V14,V19.0-V19.2,V19.4-V19.6,V20-V79,V80.3-V80.5,V81.0-V81.1,V82.0-V82.1,V83-V86,V87.0-V87.8,V88.0-V88.8
+ V89.0,V89.2)
+Other land transport accidents
+(V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
+ V89.1,V89.3,V89.9)
+Place of Death: All
+Race: All
+Weekday: All
+Year/Month: All
+Group By: Five-Year Age Groups, Gender, HHS Region, 2013 Urbanization, ICD-10 113 Cause List
+Show Totals: Disabled
+Show Zero Values: True
+Show Suppressed: True
+Calculate Rates Per: 100,000
+Rate Options: Default intercensal populations for years 2001-2009 (except Infant Age Groups)
+---
+  Help: See http://wonder.cdc.gov/wonder/help/ucd.html for more information.
+---
+  Query Date: Jul 19, 2016 4:46:10 PM
+---
+  Suggested Citation: Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
+1999-2014 on CDC WONDER Online Database, released 2015. Data are from the Multiple Cause of Death Files, 1999-2014, as compiled
+from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program. Accessed at
+http://wonder.cdc.gov/ucd-icd10.html on Jul 19, 2016 4:46:10 PM
+---
+  Messages:
+  1. Totals and Percent of Total are disabled when data are grouped by 113 or 130 Cause Lists. Check Caveats below for more
+information.
+2. Selections made to any variable in [HHS Regions, 2013 Urbanization] cause equivalent selections in the other variables. More
+Information: http://wonder.cdc.gov/wonder/help/faq.html#related-variables.
+3. Totals are not available for these results due to suppression constraints. More Information:
+  http://wonder.cdc.gov/wonder/help/faq.html#Privacy.
+---
+  Footnotes:
+  1. A '#' symbol preceding the label indicates a rankable cause of death. More information.
+---
+  Caveats:
+  1. Totals and Percent of Total are disabled when data are grouped by a 113 or 130 Cause List because both aggregate and detailed
+values are displayed in the table. Also be aware that charts and maps containing both aggregate and detail data could be
+misleading.
+2. Population and rates are labeled 'Not Applicable' when they include a subset of ages 85-100+ because populations are not
+available for those ages. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Ages 85-100.
+3. Circumstances in Georgia for the years 2008 and 2009 have resulted in unusually high death counts for the ICD-10 cause of
+death code R99, "Other ill-defined and unspecified causes of mortality." Caution should be used in interpreting these data.
+More information: http://wonder.cdc.gov/wonder/help/ucd.html#Georgia-Reporting-Anomalies.
+4. Circumstances in New Jersey for the year 2009 have resulted in unusually high death counts for the ICD-10 cause of death code
+R99, "Other ill-defined and unspecified causes of mortality" and therefore unusually low death counts in other ICD-10 codes,
+most notably R95, "Sudden Infant Death Syndrome" and X40-X49, "Unintentional poisoning." Caution should be used in
+interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#New-Jersey-Reporting-Anomalies.
+5. Circumstances in California resulted in unusually high death counts for the ICD-10 cause of death code R99, "Other
+ill-defined and unspecified causes of mortality" for deaths occurring in years 2000 and 2001. Caution should be used in
+interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#California-Reporting-Anomalies.
+6. Data are Suppressed when the data meet the criteria for confidentiality constraints. More information:
+  http://wonder.cdc.gov/wonder/help/ucd.html#Assurance of Confidentiality.
+7. Death rates are flagged as Unreliable when the rate is calculated with a numerator of 20 or less. More information:
+  http://wonder.cdc.gov/wonder/help/ucd.html#Unreliable.
+8. Deaths of persons with Age "Not Stated" are included in "All" counts and rates, but are not distributed among age groups,
+so are not included in age-specific counts, age-specific rates or in any age-adjusted rates. More information:
+  http://wonder.cdc.gov/wonder/help/ucd.html#Not Stated.
+9. The population figures for year 2014 are bridged-race estimates of the July 1 resident population, from the Vintage 2014
+postcensal series released by NCHS on June 30, 2015. The population figures for year 2013 are bridged-race estimates of the July
+1 resident population, from the Vintage 2013 postcensal series released by NCHS on June 26, 2014. The population figures for
+year 2012 are bridged-race estimates of the July 1 resident population, from the Vintage 2012 postcensal series released by NCHS
+on June 13, 2013. The population figures for year 2011 are bridged-race estimates of the July 1 resident population, from the
+Vintage 2011 postcensal series released by NCHS on July 18, 2012. Population figures for 2010 are April 1 Census counts. The
+population figures for years 2001 - 2009 are bridged-race estimates of the July 1 resident population, from the revised
+intercensal county-level 2000 - 2009 series released by NCHS on October 26, 2012. Population figures for 2000 sssssare April 1 Census
+counts. Population figures for 1999 are from the 1990-1999 intercensal series of July 1 estimates. Population figures for the
+infant age groups are the number of live births. <br/><b>Note:</b> Rates and population figures for years 2001 - 2009 differ
+slightly from previously published reports, due to use of the population estimates which were available at the time of release.
+10. The population figures used in the calculation of death rates for the age group 'under 1 year' are the estimates of the
+resident population that is under one year of age. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Age Group.
+
+## Non-Travel METs
+
+```{r newMETs, echo = FALSE}
+newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
+newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
+dataColumns <- c(1,5)
+D <- newMETs[,dataColumns]
+metaData <- newMETs[,-dataColumns]
+newMETs <- list(Data = D, metaData = metaData)
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+
+rownames(newMETs$Data) <- activityCode
+rownames(newMETs$metaData) <- activityCode
+
+newMETs$metaData <- within(newMETs$metaData, {
+  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
+})
+
+metKey <- newMETs$Data$AinsworthMETS
+names(metKey) <- activityCode
+
+```
+
+```{r readnewMETs, echo = FALSE}
+cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
+cps <- read.table(cpsFile, header=T, skip=0, sep="," )
+cps <- within(cps, {
+    TUCASEID <- as.character(TUCASEID)
+    TULINENO <- as.character(TULINENO)
+    GEREG <- as.factor(GEREG)
+    })
+cps <- subset(cps, TULINENO == "1")
+
+summaryFile <- "~/GHI/data/ATUS/atussum_0311.dat"
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
+metaDataColumns <- 1:24
+summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
+
+summaryList$metaData <- within(summaryList$metaData,{
+    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
+    AGECAT <- ifelse(TEAGE <5, "00-04",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
+                            )))))))})
+
+```
+
+We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
+
+```{r metKey, echo = FALSE}
+allActivities <- colnames(summaryList$Data)
+nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
+nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
+
+activities <- nonTravelActivities
+
+metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
+names(metKeyVector) <- activities
+metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
+dimnames(metKeyMatrix) <- list( NULL, activities)
+
+METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
+
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
+
+```
+
+```{r paameters}
+(parameters <- ddply(D, .(SEX, AGECAT), summarise,
+             mean = mean(METs, na.rm=T),
+             sd = sd(METs, na.rm=T)
+             ))
+
+METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
+(METmatrix <- METmatrix/METmatrix[3,2])
+```
+
+## Appendix
 
@@
 
-## Non-Travel METs
+<!--
+
+# Code Graveyard
+
+### Total Distribution
+
+```{r total, eval = FALSE, echo = FALSE}
+foo <- mapply(getTotalDistribution,
+  muTravel = matrix(35,nrow = 8, ncol = 2),
+  cvTravel = 1.72,
+  muNonTravel = matrix(100,nrow = 8, ncol = 2),
+  cvNonTravel = 1,
+  p = 1,
+  pWalk = rep(54.4/64.2,2),
+  vWalk = rep(2.88,2),
+  size = 1e5, SIMPLIFY = FALSE)
+```
 
@@
 
-summaryFile <- "~/GHI/data/atussum_0311.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_0311.dat
+summaryFile <- "~/GHI/data/ATUS/atussum_0311.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_0311.dat
 
-cpsFile <- "~/GHI/data/atuscps_0311.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_0311.dat
+cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_0311.dat
 
@@
 
-```{r newMETs, echo = FALSE}
-newMETsFile <- "~/GHI/data/ATUS_to_METS_key.csv"
-newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
-dataColumns <- c(1,5)
-D <- newMETs[,dataColumns]
-metaData <- newMETs[,-dataColumns]
-newMETs <- list(Data = D, metaData = metaData)
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-
-rownames(newMETs$Data) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-
-newMETs$metaData <- within(newMETs$metaData, {
-  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
-})
-
-metKey <- newMETs$Data$AinsworthMETS
-names(metKey) <- activityCode
-
-```
-
-```{r readnewMETs, echo = FALSE}
-cpsFile <- "~/GHI/data/atuscps_0311.dat"
-cps <- read.table(cpsFile, header=T, skip=0, sep="," )
-cps <- within(cps, {
-    TUCASEID <- as.character(TUCASEID)
-    TULINENO <- as.character(TULINENO)
-    GEREG <- as.factor(GEREG)
-    })
-cps <- subset(cps, TULINENO == "1")
-
-summaryFile <- "~/GHI/data/atussum_0311.dat"
-summary <- read.table(summaryFile, header=T, skip=0, sep=",")
-metaDataColumns <- 1:24
-summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
-
-summaryList$metaData <- within(summaryList$metaData,{
-    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
-    AGECAT <- ifelse(TEAGE <5, "00-04",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
-                            )))))))})
-
-```
-
-We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
-
-```{r metKey, echo = FALSE}
-allActivities <- colnames(summaryList$Data)
-nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
-nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
-
-activities <- nonTravelActivities
-
-metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
-names(metKeyVector) <- activities
-metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-dimnames(metKeyMatrix) <- list( NULL, activities)
-
-METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
-
-D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
-
-```
-
-```{r paameters}
-(parameters <- ddply(D, .(SEX, AGECAT), summarise,
-             mean = mean(METs, na.rm=T),
-             sd = sd(METs, na.rm=T)
-             ))
-
-METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
-(METmatrix <- METmatrix/METmatrix[3,2])
-```
-
-### Total Distribution
-
-```{r total, eval = FALSE, echo = FALSE}
-foo <- mapply(getTotalDistribution,
-  muTravel = matrix(35,nrow = 8, ncol = 2),
-  cvTravel = 1.72,
-  muNonTravel = matrix(100,nrow = 8, ncol = 2),
-  cvNonTravel = 1,
-  p = 1,
-  pWalk = rep(54.4/64.2,2),
-  vWalk = rep(2.88,2),
-  size = 1e5, SIMPLIFY = FALSE)
-```
-
-### NHTS
-
-```{r NHTS}
-DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = "-9")
-names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
-
-DAYV2PUB <- within(DAYV2PUB, {
-    URBRUR <- as.factor(URBRUR)
-    HHSTFIPS <- as.factor(HHSTFIPS)
-    MSACAT <- as.factor(MSACAT)
-    TRPTRANS <- as.factor(TRPTRANS)
-    })
-
+```{r NHTSTables}
 HHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/HHV2PUB.CSV", na.strings = "-9")
@@
 
-PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
-names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
-
 VEHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/VEHV2PUB.CSV", na.strings = "-9")
 names(VEHV2PUB) <- c("HOUSEID","WTHHFIN","VEHID","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","CDIVMSAR","CENSUS_D","CENSUS_R","HHSTATE","HHSTFIPS","HYBRID","MAKECODE","MODLCODE","MSACAT","MSASIZE","OD_READ","RAIL","TRAVDAY","URBAN","URBANSIZE","URBRUR","VEHCOMM","VEHOWNMO","VEHYEAR","WHOMAIN","WRKCOUNT","TDAYDATE","VEHAGE","PERSONID","HH_HISP","HH_RACE","HOMEOWN","HOMETYPE","LIF_CYC","ANNMILES","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","BEST_FLG","BESTMILE","BEST_EDT","BEST_OUT","FUELTYPE","GSYRGAL","GSCOST","GSTOTCST","EPATMPG","EPATMPGF","EIADMPG","VEHTYPE","HH_CBSA","HHC_MSA")
+```
 
-NHTS.trips <- data.frame(HOUSEID = DAYV2PUB$HOUSEID,
-                   PERSONID = DAYV2PUB$PERSONID,
-                   TDCASEID = DAYV2PUB$TDCASEID,
-                   AT = ifelse(DAYV2PUB$TRPTRANS == "22", "cycle", ifelse(DAYV2PUB$TRPTRANS == "23","walk",ifelse(!is.na(DAYV2PUB$TRPTRANS),"other",NA))),
-                   TRVL_MIN = as.numeric(DAYV2PUB$TRVL_MIN),
-                   TRVLCMIN = as.numeric(DAYV2PUB$TRVLCMIN)
-                   )
+```{r NHTSArchive, echo = FALSE, eval = FALSE}
+nMatrix <- ddply( subset(NHTS.person, !is.na(AGE) & !is.na(SEX)), .(SEX, AGE), summarise, count = length(PERSONID))
+sum(nMatrix$count) # 305649 number of people in the PERSON table
 
-NHTS.person <- with(PERV2PUB,{
-                    data.frame(HOUSEID = HOUSEID,
-                               PERSONID = PERSONID,
-                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
-                               AGE = as.factor(convertToAgeClass(R_AGE))
-)
-})
+foo <- group_by(NHTS.trips, HOUSEID,PERSONID)
+foobar <- dplyr::summarise(foo, n = n())
+nrow(foobar) # 262,934 number of people with trip data
 
-NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
+NHTS %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarise( n = n()) # 260,576 people in merged data
 
-NHTS <- subset(NHTS, !is.na(AGE) & !is.na(SEX))
+#NHTS <- subset(NHTS, !is.na(AGE) & !is.na(SEX))
 
@@
 
-NHTS %>% group_by(SEX, AGE) %>% summarize(trips = n())
-
-countPeople <- function(data){
- nrow(data %>% group_by(HOUSEID, PERSONID) %>% summarize(n()))
-}
-
+NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))
 ```
-<!--
-
-# Code Graveyard
 

### 2016-08-08 R/ITHIM.Rmd
Continued working on NHTS and CDC Wonder data sets for regional parameters

@@
 ### NHTS
+
+#### Create a handful of functions to help parse data
+
+```{r parsely}
+convertToHHS <- function(states){
+  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
+  StateToHHS <- HHSRegions[,3]
+  names(StateToHHS) <- HHSRegions[,1]
+  return(StateToHHS[states])
+}
+
+countPeople <- function(data){
+ nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
+}
+
+tripTime <- function(data, mode = "walk"){
+ return(sum(subset(data, AT == mode)$TRVLCMIN, na.rm = TRUE))
+}
+
+reformatData <- function(data){
+foo <- matrix(unlist(data), nrow = 7, ncol = 2, dimnames = list(paste0("ageClass",2:8), c("F","M")))
+return(foo[,c("M","F")])
+}
+```
 
@@
 ```
-#### Create a handful of functions to help parse data
-
-```{r parsely}
-convertToHHS <- function(states){
-  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
-  StateToHHS <- HHSRegions[,3]
-  names(StateToHHS) <- HHSRegions[,1]
-  return(StateToHHS[states])
-}
-
-countPeople <- function(data){
- nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
-}
-
-tripTime <- function(data, mode = "walk"){
- return(sum(subset(data, AT == mode)$TRVLCMIN, na.rm = TRUE))
-}
-
-reformatData <- function(data){
-foo <- matrix(unlist(data), nrow = 7, ncol = 2, dimnames = list(paste0("ageClass",2:8), c("F","M")))
-return(foo[,c("M","F")])
-}
-```
-
 #### Subset and Merge
@@
 NHTS.person <- subset( NHTS.person, MSA == TRUE )
-
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
@@
 cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
-
 countMatrix <- lapply(counts, reformatData)
@@
 cycleMatrix <- lapply(cycleLength, reformatData)
-
-(meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE))
-(meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE))
-
+meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
+meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
+```
+
+```{r showMatrices}
+meanWalkList$HHS1
+meanWalkList$HHS2
+meanCycleList$HHS1
+meanCycleList$HHS2
 ```
@@
 
-#### Query Results File
+#### Query Results File (HHSmetro_Age_sex_2010-14.csv)
 
 ```{r CDC}
-wonder <- read.csv("./../data/CDC/HHSmetro_Age_sex_2010-14.csv", header=T)
+wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
 names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
-
-#HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow=51,ncol=3)
 
@@
 ```{r summary}
-wonderSummary <- ddply(wonder2, .(HHSRegion, gender, ITHIMage, ITHIMcause), summarise,
+wonderSummary <- ddply(wonder, .(HHSRegion, gender, ITHIMage, ITHIMcause), summarise,
                        numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
@@
 
-#### Single Year Data File
+#### Single Year Data File (HHSmetro_Age_sex_2014_AllCause.csv)
+
+Read in data for single year population.  Processed in Excel, to
+combine two WONDER queries, need to go back and script the processing
+of the raw WONDER output
 
 ```{r singleyear}
-#read in data for single year population
-#processed in Excel, to combine two WONDER queries, need to go back and script the processing of the raw WONDER output
-pop14 <- read.csv("./../data/CDC/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
-sum(as.numeric(as.character(pop14$Population)),na.rm=T)  #checking US 'Metro' population here 272,667,942
-sum(as.numeric(as.character(pop14$Deaths)),na.rm=T)  #checking US 'Metro' deaths here 2,125,273
+pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
 
 names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
-
-#this is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' request
-ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5",
-                 "ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3",
-                 "ageClass8")
+```
+This is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' command.
+
+```{r sloppy}
+ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
 names(ITHIMageKey2) <- unique(pop14$WONDERage)
 pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
-
+```
+
+```{r foop}
 pop14 <-  pop14[, c("HHSRegion","gender","ITHIMage","deaths2014AllCause","population2014")]
-pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))
-
-sum(as.numeric(as.character(pop14$population2014)),na.rm=T)    #checking US 'Metro' population here 272,667,942
-sum(as.numeric(as.character(pop14$deaths2014)),na.rm=T)     #checking US 'Metro' deaths here 2,125,273
+pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(1:9), as.numeric(as.character(pop14$deaths2014))))
 
@@
 
-sum(popSummary$population2014) #checking US population here 272,667,942
-sum(popSummary$AllCauseDeaths2014)
-
 HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
-
-USGBD <- read.csv("./../data/CDC/USGBD.csv", header=T)
+```
+#### Read USGBD file (USGBD.csv)
+
+Presumably this comes from an EXCEL file.
+
+```{r USGBD}
+USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
 USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
@@
 
-#gah <- subset(burden, ITHIMcause == "RespiratoryDisease")
-
-#ggplot(gah, aes(x=ITHIMage, y=YLD, fill=factor(HHSRegion))) + geom_bar(stat="identity") + geom_point(aes(y=GBDyld)) + facet_grid(gender ~ .)
-#ggplot(gah, aes(x=ITHIMage, y=DALY)) + geom_bar(stat="identity") + geom_point(aes(y=GBDdaly)) + facet_grid(gender ~ HHSRegion)
-
 burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
@@
 
-burden$sex
-
 burdenOut <- burden[,c("HHSRegion","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
@@
 
-write.csv(burdenOut, "./HHSburdens.csv", row.names=F)
-
-```
-
+#write.csv(burdenOut, "./HHSburdens.csv", row.names=F)
+```
+
+```{r burdenList}
+burdenList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
+burdenList[[10]][[10]][[1]]
+```
+
+<!--
 #### DETAILS ON WONDER QUERY
@@
 
+-->
+
 ## Non-Travel METs
 
-```{r newMETs, echo = FALSE}
+```{r newMETs, echo = TRUE}
 newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
@@
 
-```{r readnewMETs, echo = FALSE}
+```{r readnewMETs, echo = TRUE}
 cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
@@
 
-```{r metKey, echo = FALSE}
+```{r metKey, echo = TRUE}
 allActivities <- colnames(summaryList$Data)
@@
 
-```{r paameters}
+```{r parameters}
 (parameters <- ddply(D, .(SEX, AGECAT), summarise,
@@
 
-## Appendix
-
-### Bay Area
+## Bay Area
 
@@
 
-#### Deaths
+### Deaths
 ```{r show, echo = FALSE}
@@
 ```
-#### DALY
+### DALY
 ```{r show2, echo = FALSE}
@@
 ```
-#### YLD
+### YLD
 ```{r show3, echo = FALSE}
@@
 ```
-#### YLL
+### YLL
 ```{r show4, echo = FALSE}
@@
 
-### National
+## National
 
@@
 
-#### Deaths
+### Deaths
 ```{r show222, echo = FALSE}
@@
 ```
-#### DALY
+### DALY
 ```{r show22, echo = FALSE}
@@
 ```
-#### YLD
+### YLD
 ```{r show23, echo = FALSE}
@@
 
-### Example
+## Example
 
@@
 
-### Interrogate Parameters
+## Interrogate Parameters
 
@@
 ITHIM.baseline$parameters$Rwt
+```
+
+## Appendix
+
+```{r appendix}
+date()
+sessionInfo()
 ```
@@
 -->
-### Appendix
-
-```{r appendix}
-date()
-sessionInfo()
-```

### 2016-08-09 R/ITHIM.Rmd
Now creating regional list of ITHIM objects

@@
 reformatData <- function(data){
-foo <- matrix(unlist(data), nrow = 7, ncol = 2, dimnames = list(paste0("ageClass",2:8), c("F","M")))
+foo <- matrix(unlist(data), nrow = 7, ncol = 2)
+foo <- rbind( c(0.1, 0.1), foo)
+dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
 return(foo[,c("M","F")])
@@
 meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
+RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 0.01, x) )
+RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 0.01, x) )
+RwtList <- lapply( RwtList, function(x) x/x[3,2] )
+RctList <- lapply( RctList, function(x) x/x[3,2] )
+muwtList <- lapply( meanWalkList, function(x) x[3,2] )
+muctList <- lapply( meanCycleList, function(x) x[3,2] )
 ```
@@
 ```{r wonder}
-wonder <- wonder[, c("HHSRegion","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
-depress <- expand.grid(HHSRegion=unique(wonder$HHSRegion), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
+wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
+depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
 wonder <- rbind(wonder, depress)
@@
 wonder <- within( wonder, {
-    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(1:9), as.numeric(as.character(deaths5yr))))
+    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
     })
@@
 ```{r summary}
-wonderSummary <- ddply(wonder, .(HHSRegion, gender, ITHIMage, ITHIMcause), summarise,
+wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
                        numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
@@
 ```{r foop}
-pop14 <-  pop14[, c("HHSRegion","gender","ITHIMage","deaths2014AllCause","population2014")]
-pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(1:9), as.numeric(as.character(pop14$deaths2014))))
-
-popSummary <- ddply(pop14, .(HHSRegion, gender, ITHIMage), summarise,
+pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
+pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))
+
+popSummary <- ddply(pop14, .(HHSRegionCode, gender, ITHIMage), summarise,
                     population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
@@
 
-burdenOut <- burden[,c("HHSRegion","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
+burdenOut <- burden[,c("HHSRegionCode","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
 names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")
@@
 burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
-
-#write.csv(burdenOut, "./HHSburdens.csv", row.names=F)
-```
+burdenOut <- burdenOut[,c("HHS", "disease","sex","dproj","yll","yld","daly")]
+```
+
+Now we create an R list of GBD objects named by HHS region.
 
 ```{r burdenList}
-burdenList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
-burdenList[[10]][[10]][[1]]
+GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
+#saveRDS(GBDList, file = "../data/GBD.rds")
+```
+
+## Create List of ITHIM Objects
+
+```{r ITHIMObjs}
+
+ITHIMList <- list()
+
+ITHIM.default <- createITHIM(vision = "baseline", region = "national")
+
+for (HHS in paste0("HHS",1:10)){
+  ITHIMList[[HHS]] <- ITHIM.default
+  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muwt", muwtList[[HHS]])
+  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muct", muctList[[HHS]])
+  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rwt", RwtList[[HHS]])
+  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rct", RctList[[HHS]])
+  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
+}
+
 ```
@@
 
-```{r newMETs, echo = TRUE}
+```{r newMETs, echo = TRUE, eval = FALSE}
 newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
@@
 
-```{r readnewMETs, echo = TRUE}
+```{r readnewMETs, echo = TRUE, eval = FALSE}
 cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
@@
 
-```{r metKey, echo = TRUE}
+```{r metKey, echo = TRUE, eval = FALSE}
 allActivities <- colnames(summaryList$Data)
@@
 
-```{r foo, eval = FALSE, echo = TRUE}
+```{r foo, eval = FALSE, echo = TRUE, eval = FALSE}
 nonTravelMETs[which.max(nonTravelMETs$x5),]

### 2016-08-10 R/ITHIM.Rmd
Conducted all pairwise comparisons of metro regions

@@
 
-```{r showMatrices}
-meanWalkList$HHS1
-meanWalkList$HHS2
-meanCycleList$HHS1
-meanCycleList$HHS2
-```
-
 ### CDC Mortality
@@
 #saveRDS(GBDList, file = "../data/GBD.rds")
+```
+
+## Non-Travel METs
+
+```{r newMETs, echo = TRUE, eval = TRUE}
+newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
+newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
+dataColumns <- c(1,5)
+D <- newMETs[,dataColumns]
+metaData <- newMETs[,-dataColumns]
+newMETs <- list(Data = D, metaData = metaData)
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+
+rownames(newMETs$Data) <- activityCode
+rownames(newMETs$metaData) <- activityCode
+
+newMETs$metaData <- within(newMETs$metaData, {
+  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
+})
+
+metKey <- newMETs$Data$AinsworthMETS
+names(metKey) <- activityCode
+
+```
+
+```{r readnewMETs, echo = TRUE, eval = TRUE}
+cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
+cps <- read.table(cpsFile, header=T, skip=0, sep="," )
+cps <- within(cps, {
+    TUCASEID <- as.character(TUCASEID)
+    TULINENO <- as.character(TULINENO)
+    GEREG <- as.factor(GEREG)
+    })
+cps <- subset(cps, TULINENO == "1")
+
+summaryFile <- "~/GHI/data/ATUS/atussum_0311.dat"
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
+metaDataColumns <- 1:24
+summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
+
+summaryList$metaData <- within(summaryList$metaData,{
+    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
+    AGECAT <- ifelse(TEAGE <5, "00-04",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
+                            )))))))})
+
+```
+
+We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
+
+```{r metKey, echo = TRUE, eval = TRUE}
+allActivities <- colnames(summaryList$Data)
+nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
+nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
+
+activities <- nonTravelActivities
+
+metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
+names(metKeyVector) <- activities
+metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
+dimnames(metKeyMatrix) <- list( NULL, activities)
+
+METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
+
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
+
+```
+
+```{r parameters, eval = TRUE}
+(parameters <- ddply(D, .(SEX, AGECAT), summarise,
+             mean = mean(METs, na.rm=T),
+             sd = sd(METs, na.rm=T)
+             ))
+
+METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
+(METmatrix <- METmatrix/METmatrix[3,2])
 ```
@@
 
-<!--
-#### DETAILS ON WONDER QUERY
+```{r CRA}
+
+CRA.fn <- function(i, j){
+  foo <- compareModels( ITHIMList[[i]], ITHIMList[[j]])
+  sum(unlist(foo$AF))
+}
+
+CRAMatrix <- matrix(NA, nrow = 10, ncol = 10)
+
+for( i in 1:10 ){
+  CRA <- c()
+  for( j in 1:10 ){
+    CRA <- c( CRA, CRA.fn(i, j))
+  }
+  CRAMatrix[i,] <- CRA
+}
+
+which(CRAMatrix == max(CRAMatrix), arr.ind = TRUE)
+
+```
+
+## DETAILS ON WONDER QUERY
 
@@
 
--->
-
-## Non-Travel METs
-
-```{r newMETs, echo = TRUE, eval = FALSE}
-newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
-newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
-dataColumns <- c(1,5)
-D <- newMETs[,dataColumns]
-metaData <- newMETs[,-dataColumns]
-newMETs <- list(Data = D, metaData = metaData)
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-
-rownames(newMETs$Data) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-
-newMETs$metaData <- within(newMETs$metaData, {
-  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
-})
-
-metKey <- newMETs$Data$AinsworthMETS
-names(metKey) <- activityCode
-
-```
-
-```{r readnewMETs, echo = TRUE, eval = FALSE}
-cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
-cps <- read.table(cpsFile, header=T, skip=0, sep="," )
-cps <- within(cps, {
-    TUCASEID <- as.character(TUCASEID)
-    TULINENO <- as.character(TULINENO)
-    GEREG <- as.factor(GEREG)
-    })
-cps <- subset(cps, TULINENO == "1")
-
-summaryFile <- "~/GHI/data/ATUS/atussum_0311.dat"
-summary <- read.table(summaryFile, header=T, skip=0, sep=",")
-metaDataColumns <- 1:24
-summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
-
-summaryList$metaData <- within(summaryList$metaData,{
-    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
-    AGECAT <- ifelse(TEAGE <5, "00-04",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
-                            )))))))})
-
-```
-
-We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
-
-```{r metKey, echo = TRUE, eval = FALSE}
-allActivities <- colnames(summaryList$Data)
-nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
-nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
-
-activities <- nonTravelActivities
-
-metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
-names(metKeyVector) <- activities
-metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-dimnames(metKeyMatrix) <- list( NULL, activities)
-
-METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
-
-D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
-
-```
-
-```{r parameters}
-(parameters <- ddply(D, .(SEX, AGECAT), summarise,
-             mean = mean(METs, na.rm=T),
-             sd = sd(METs, na.rm=T)
-             ))
-
-METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
-(METmatrix <- METmatrix/METmatrix[3,2])
-```
-
-## Bay Area
+## Example (Bay Area)
 
@@
 
-## National
+## Example (National)
 

### 2016-08-15 R/ITHIM.Rmd
Summary stats for NHTS trip time and distance.

@@
 ```{r NHTS}
-DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = "-9")
+DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
 names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
 
-DAYV2PUB <- within(DAYV2PUB, {
-    URBRUR <- as.factor(URBRUR)
-    HHSTATE <- as.factor(HHSTATE)
-    MSACAT <- as.factor(MSACAT)
-    TRPTRANS <- as.factor(TRPTRANS)
-})
-
-NHTS.trips <- data.frame(HOUSEID = DAYV2PUB$HOUSEID,
-                   PERSONID = DAYV2PUB$PERSONID,
-                   TDCASEID = DAYV2PUB$TDCASEID,
-                   AT = ifelse(DAYV2PUB$TRPTRANS == "22", "cycle",
-                        ifelse(DAYV2PUB$TRPTRANS == "23", "walk",
-                        ifelse(!is.na(DAYV2PUB$TRPTRANS), "other", NA))),
-                   TRVL_MIN = as.numeric(DAYV2PUB$TRVL_MIN),
-                   TRVLCMIN = as.numeric(DAYV2PUB$TRVLCMIN)
-                   )
+## DAYV2PUB <- within(DAYV2PUB, {
+##     URBRUR <- as.factor(URBRUR)
+##     HHSTATE <- as.factor(HHSTATE)
+##     MSACAT <- as.factor(MSACAT)
+##     TRPTRANS <- as.factor(TRPTRANS)
+##     TRPMILES <- as.numeric(TRPMILES)
+## })
+
+NHTS.trips <- with(DAYV2PUB,{
+    data.frame(HOUSEID = HOUSEID,
+                   PERSONID = PERSONID,
+                   TDCASEID = TDCASEID,
+                   AT = ifelse(TRPTRANS == "22", "cycle",
+                        ifelse(TRPTRANS == "23", "walk",
+                        ifelse(!is.na(TRPTRANS), "other", NA))),
+                   TRVL_MIN = as.numeric(TRVL_MIN),
+                   TRVLCMIN = ifelse(as.numeric(TRVLCMIN) != 0, as.numeric(TRVLCMIN), NA),
+                   URBRUR  = as.factor(URBRUR),
+                   HHSTATE = as.factor(HHSTATE),
+                   MSACAT = as.factor(MSACAT),
+                   TRPTRANS = as.factor(TRPTRANS),
+                   TRPMILES = as.numeric(TRPMILES)
+)})
 
@@
 ```
+
+#### Investigate Trip Data
+
+```{r summary}
+summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
+summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
+summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
+summary(with(subset(NHTS.trips, AT == "walk"), TRVLCMIN))
+summary(with(subset(NHTS.trips, AT == "cycle"), TRVLCMIN))
+summary(with(subset(NHTS.trips, AT == "other"), TRVLCMIN))
+
+```
+
 #### Subset and Merge
@@
 
-```{r summary}
+```{r summary2}
 wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,

### 2016-08-16 R/ITHIM.Rmd
Investigated distribution of walk/cycle trip miles and hours

--- 2016-08-09 R/ITHIM.Rmd
@@
 
-```{r NHTS}
-DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = "-9")
+```{r NHTS, eval = FALSE}
+DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
 names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
 
-DAYV2PUB <- within(DAYV2PUB, {
-    URBRUR <- as.factor(URBRUR)
-    HHSTATE <- as.factor(HHSTATE)
-    MSACAT <- as.factor(MSACAT)
-    TRPTRANS <- as.factor(TRPTRANS)
-})
-
-NHTS.trips <- data.frame(HOUSEID = DAYV2PUB$HOUSEID,
-                   PERSONID = DAYV2PUB$PERSONID,
-                   TDCASEID = DAYV2PUB$TDCASEID,
-                   AT = ifelse(DAYV2PUB$TRPTRANS == "22", "cycle",
-                        ifelse(DAYV2PUB$TRPTRANS == "23", "walk",
-                        ifelse(!is.na(DAYV2PUB$TRPTRANS), "other", NA))),
-                   TRVL_MIN = as.numeric(DAYV2PUB$TRVL_MIN),
-                   TRVLCMIN = as.numeric(DAYV2PUB$TRVLCMIN)
-                   )
+## DAYV2PUB <- within(DAYV2PUB, {
+##     URBRUR <- as.factor(URBRUR)
+##     HHSTATE <- as.factor(HHSTATE)
+##     MSACAT <- as.factor(MSACAT)
+##     TRPTRANS <- as.factor(TRPTRANS)
+##     TRPMILES <- as.numeric(TRPMILES)
+## })
+
+NHTS.trips <- with(DAYV2PUB,{
+    data.frame(HOUSEID = HOUSEID,
+                   PERSONID = PERSONID,
+                   TDCASEID = TDCASEID,
+                   AT = ifelse(TRPTRANS == "22", "cycle",
+                        ifelse(TRPTRANS == "23", "walk",
+                        ifelse(!is.na(TRPTRANS), "other", NA))),
+                   TRVL_MIN = as.numeric(TRVL_MIN),
+                   TRVLCMIN = ifelse(as.numeric(TRVLCMIN) != 0, as.numeric(TRVLCMIN), NA),
+                   URBRUR  = as.factor(URBRUR),
+                   HHSTATE = as.factor(HHSTATE),
+                   MSACAT = as.factor(MSACAT),
+                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
+                   TRPTRANS = as.factor(TRPTRANS),
+                   TRPMILES = as.numeric(TRPMILES)
+)})
 
@@
 })
-```
+saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")
+saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
+```
+
+```{r loadData}
+NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
+NHTS.trips <- readRDS(file = "./data/NHTS.trips.rds")
+```
+
+#### Investigate Trip Data
+
+```{r summary, fig.width = 15}
+summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
+summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
+summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
+summary(with(subset(NHTS.trips, AT == "walk"), TRVLCMIN))
+summary(with(subset(NHTS.trips, AT == "cycle"), TRVLCMIN))
+summary(with(subset(NHTS.trips, AT == "other"), TRVLCMIN))
+
+walkHours <- with(subset(NHTS.trips, {
+    AT == "walk" #& TRVLCMIN > 5 & TRPMILES > 2/9
+}), TRVLCMIN/60)
+
+walkMiles <- with(subset(NHTS.trips, {
+    AT == "walk" #& TRVLCMIN > 5 & TRPMILES > 2/9
+}), TRPMILES)
+
+cycleHours <- with(subset(NHTS.trips, {
+      AT == "cycle" #& TRVLCMIN > 5 & TRPMILES > 2/9
+}), TRVLCMIN/60)
+
+cycleMiles <- with(subset(NHTS.trips, {
+      AT == "cycle" #& TRVLCMIN > 5 & TRPMILES > 2/9
+}), TRPMILES)
+
+walkSpeed <- median(walkMiles/walkHours, na.rm = TRUE)
+cycleSpeed <- median(cycleMiles/cycleHours, na.rm = TRUE)
+
+Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, mode = "walk")
+Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, mode = "cycle")
+D <- rbind(Dwalk,Dcycle)
+
+Dspeed <- data.frame(speed = c(1,1), intercept = log10(c(walkSpeed,cycleSpeed)), mode = c("walk","cycle"))
+
+g <- ggplot(D, aes(Hours, Miles, colour = mode)) + geom_point()
+g <- g + geom_abline(aes(slope = speed, intercept = intercept, colour = mode), Dspeed) + facet_grid(. ~ mode)
+g + scale_y_log10() + scale_x_log10() + theme_bw()
+```
+
 #### Subset and Merge
@@
 
-```{r CDC}
+```{r CDC, eval = FALSE}
 wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
@@
 
-```{r ITHIMCause}
+```{r ITHIMCause, eval = FALSE}
 ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
@@
 ```
-```{r recodeAge}
+```{r recodeAge, eval = FALSE}
 ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
@@
 
-```{r wonder}
+```{r wonder, eval = FALSE}
 wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
@@
 
-```{r suppressed}
+```{r suppressed, eval = FALSE}
 wonder <- within( wonder, {
@@
 
-```{r summary}
+```{r summary2, eval = FALSE}
 wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
@@
 
-```{r singleyear}
+```{r singleyear, eval = FALSE}
 pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
@@
 
-```{r sloppy}
+```{r sloppy, eval = FALSE}
 ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
@@
 
-```{r foop}
+```{r foop, eval = FALSE}
 pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
@@
 
-```{r USGBD}
+```{r USGBD, eval = FALSE}
 USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
@@
 
-```{r burdenList}
+```{r burdenList, eval = FALSE}
 GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
@@
 
+## Non-Travel METs
+
+```{r newMETs, echo = TRUE, eval = FALSE}
+newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
+newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
+dataColumns <- c(1,5)
+D <- newMETs[,dataColumns]
+metaData <- newMETs[,-dataColumns]
+newMETs <- list(Data = D, metaData = metaData)
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+
+rownames(newMETs$Data) <- activityCode
+rownames(newMETs$metaData) <- activityCode
+
+newMETs$metaData <- within(newMETs$metaData, {
+  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
+})
+
+metKey <- newMETs$Data$AinsworthMETS
+names(metKey) <- activityCode
+
+```
+
+```{r readnewMETs, echo = TRUE, eval = FALSE}
+cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
+cps <- read.table(cpsFile, header=T, skip=0, sep="," )
+cps <- within(cps, {
+    TUCASEID <- as.character(TUCASEID)
+    TULINENO <- as.character(TULINENO)
+    GEREG <- as.factor(GEREG)
+    })
+cps <- subset(cps, TULINENO == "1")
+
+summaryFile <- "~/GHI/data/ATUS/atussum_0311.dat"
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
+metaDataColumns <- 1:24
+summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
+
+summaryList$metaData <- within(summaryList$metaData,{
+    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
+    AGECAT <- ifelse(TEAGE <5, "00-04",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
+                            )))))))})
+
+```
+
+We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
+
+```{r metKey, echo = TRUE, eval = FALSE}
+allActivities <- colnames(summaryList$Data)
+nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
+nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
+
+activities <- nonTravelActivities
+
+metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
+names(metKeyVector) <- activities
+metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
+dimnames(metKeyMatrix) <- list( NULL, activities)
+
+METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
+
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
+
+```
+
+```{r parameters, eval = FALSE}
+(parameters <- ddply(D, .(SEX, AGECAT), summarise,
+             mean = mean(METs, na.rm=T),
+             sd = sd(METs, na.rm=T)
+             ))
+
+METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
+(METmatrix <- METmatrix/METmatrix[3,2])
+```
+
 ## Create List of ITHIM Objects
 
-```{r ITHIMObjs}
+```{r ITHIMObjs, eval = FALSE}
 
@@
 
-<!--
-#### DETAILS ON WONDER QUERY
+```{r CRA, eval = FALSE}
+
+CRA.fn <- function(i, j){
+  foo <- compareModels( ITHIMList[[i]], ITHIMList[[j]])
+  sum(unlist(foo$AF))
+}
+
+CRAMatrix <- matrix(NA, nrow = 10, ncol = 10)
+
+for( i in 1:10 ){
+  CRA <- c()
+  for( j in 1:10 ){
+    CRA <- c( CRA, CRA.fn(i, j))
+  }
+  CRAMatrix[i,] <- CRA
+}
+
+which(CRAMatrix == max(CRAMatrix), arr.ind = TRUE)
+
+```
+
+## DETAILS ON WONDER QUERY
 
@@
 
-```{r instantiate}
+```{r instantiate, eval = FALSE}
 ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
@@
 
-```{r RRs}
+```{r RRs, eval = FALSE}
 comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
@@
 ### Deaths
-```{r show, echo = FALSE}
+```{r show, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### DALY
-```{r show2, echo = FALSE}
+```{r show2, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLD
-```{r show3, echo = FALSE}
+```{r show3, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLL
-```{r show4, echo = FALSE}
+```{r show4, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-## National
-
-```{r instantiate2}
+## Example (National)
+
+```{r instantiate2, eval = FALSE}
 ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
@@
 
-```{r RRs2}
+```{r RRs2, eval = FALSE}
 comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
@@
 ### Deaths
-```{r show222, echo = FALSE}
+```{r show222, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### DALY
-```{r show22, echo = FALSE}
+```{r show22, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLD
-```{r show23, echo = FALSE}
+```{r show23, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 #### YLL
-```{r show24, echo = FALSE}
+```{r show24, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r example}
+```{r example, eval = FALSE}
 example("ITHIM-package")
@@
 
-```{r vision}
+```{r vision, eval = FALSE}
 ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
@@
 #### YLL
-```{r show233, echo = FALSE}
+```{r show233, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r params}
+```{r params, eval = FALSE}
 names(ITHIM.baseline$parameters)
@@
 
-```{r NHTSTables}
+```{r NHTSTables, eval = FALSE}
 HHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/HHV2PUB.CSV", na.strings = "-9")
@@
 ### Relative Risk Model
-```{r explore}
+```{r explore, eval = FALSE}
 names(ITHIM.baseline$parameters)
 ```
-```{r RRFigure, echo = FALSE}
+```{r RRFigure, echo = FALSE, eval = FALSE}
 phi <- function(x,k=0.5,phi0=comparativeRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
@@
 
-```{r parlist}
+```{r parlist, eval = FALSE}
 AF.sum <- numeric(1e3); i <- 0;
@@
 
-```{r AF, echo = FALSE}
+```{r AF, echo = FALSE, eval = FALSE}
 with(comparativeRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
@@
 
-```{r plot2, echo = FALSE, fig.width = 12}
+```{r plot2, echo = FALSE, fig.width = 12, eval = FALSE}
 plotRR(comparativeRisk$RR.baseline$Depression,comparativeRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
@@
 
-```{r plot4, echo = FALSE, fig.height = 12}
+```{r plot4, echo = FALSE, fig.height = 12, eval = FALSE}
 D<-melt(comparativeRisk$AF)
@@
 ### Road Injury Array
-```{r road}
+```{r road, eval = FALSE}
 baselineInjuries <- array(dim=c(6,7,3,2))

### 2016-08-16 R/ITHIM.Rmd
Investigated whether Speed is logNormal

@@
 
-Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, mode = "walk")
-Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, mode = "cycle")
+Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, Speed = walkMiles/walkHours, mode = "walk")
+Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, Speed = cycleMiles/cycleHours, mode = "cycle")
 D <- rbind(Dwalk,Dcycle)
@@
 g + scale_y_log10() + scale_x_log10() + theme_bw()
+
+g <- ggplot(D, aes(log10(Speed), ..density.., fill = mode)) + geom_histogram(bins = 25) + facet_grid(. ~ mode)
+g
+
 ```

### 2016-08-30 R/ITHIM.Rmd
Visualization of trip time matrices

@@
 library("dplyr")
+library("cowplot")
+library("grid")
+library("scales")
 ```
@@
 
-```{r parsely}
+```{r parsely, eval = TRUE}
 convertToHHS <- function(states){
@@
 
-```{r NHTS, eval = FALSE}
+```{r NHTS, eval = TRUE}
 DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
@@
 
-```{r loadData}
+```{r loadData, eval = TRUE}
 NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
@@
 
-```{r summary, fig.width = 15}
+```{r summary, fig.width = 15, eval = TRUE}
 summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
@@
 
-```{r subset}
+```{r subset, eval = TRUE}
 NHTS.person <- subset( NHTS.person, MSA == TRUE )
@@
 
-```{r split}
+```{r split, eval = TRUE}
 listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
@@
 
-```{r count}
+```{r count, eval = TRUE}
 counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
@@
 
-```{r CDC, eval = FALSE}
+```{r CDC, eval = TRUE}
 wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
@@
 
-```{r ITHIMCause, eval = FALSE}
+```{r ITHIMCause, eval = TRUE}
 ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
@@
 ```
-```{r recodeAge, eval = FALSE}
+```{r recodeAge, eval = TRUE}
 ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
@@
 
-```{r wonder, eval = FALSE}
+```{r wonder, eval = TRUE}
 wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
@@
 
-```{r suppressed, eval = FALSE}
+```{r suppressed, eval = TRUE}
 wonder <- within( wonder, {
@@
 
-```{r summary2, eval = FALSE}
+```{r summary2, eval = TRUE}
 wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
@@
 
-```{r singleyear, eval = FALSE}
+```{r singleyear, eval = TRUE}
 pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
@@
 
-```{r sloppy, eval = FALSE}
+```{r sloppy, eval = TRUE}
 ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
@@
 
-```{r foop, eval = FALSE}
+```{r foop, eval = TRUE}
 pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
@@
 
-```{r USGBD, eval = FALSE}
+```{r USGBD, eval = TRUE}
 USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
@@
 
-```{r burdenList, eval = FALSE}
+```{r burdenList, eval = TRUE}
 GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
@@
 
-```{r ITHIMObjs, eval = FALSE}
+```{r ITHIMObjs, eval = TRUE}
 
@@
   ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
-}
-
-```
-
-```{r CRA, eval = FALSE}
+  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muNonTravel", 15)
+}
+
+saveRDS(ITHIMList, file = "./data/ITHIMList.rds")
+
+```
+
+```{r readRDS, eval = TRUE}
+ITHIMlist <- readRDS(file = "./data/ITHIMList.rds")
+```
+
+```{r fooba}
+sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
+    return(sum(unlist(compareModels(ITHIM.baseline, ITHIM.scenario)$daly.delta)))
+    }
+```
+
+```{r sumDALY}
+ITHIM.national <- createITHIM(vision = "baseline", region = "national")
+ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", 15)
+ITHIM.national <- updateITHIM(ITHIM.national, "muwt", 49.63)
+ITHIM.national <- updateITHIM(ITHIM.national, "muct", 2.50)
+DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
+```
+
+```{r eval = TRUE}
+
+meanMatrix <- function(ITHIM, mode = "walk"){
+   if(mode == "walk"){
+      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
+}else{
+      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
+}
+}
+
+meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "walk")
+meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "cycle")
+
+D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
+
+names(D) <- c("ageClass", "mode", "sex", "time", "HHS")
+
+p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") + facet_grid( mode ~ HHS)
+
+p1
+
+```
+
+```{r CRA, eval = TRUE}
 

### 2016-08-31 R/ITHIM.Rmd
Working on visulization of mean matrices
There's a bug here right now that needs to be fixed.
Visualiztion of Mean Matrices, Trip Speeds and Regional vs. National DALYs

--- 2016-08-16 R/ITHIM.Rmd
@@
 
-```{r parsely}
+```{r parsely, eval = FALSE}
 convertToHHS <- function(states){
@@
 
-```{r loadData}
+```{r loadData, eval = FALSE}
 NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
@@
 
-```{r summary, fig.width = 15}
+```{r summary, fig.width = 20, fig.height = 10, eval = TRUE}
 summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
@@
 walkHours <- with(subset(NHTS.trips, {
-    AT == "walk" #& TRVLCMIN > 5 & TRPMILES > 2/9
+    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVLCMIN*60 < 8 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 0.5
 }), TRVLCMIN/60)
@@
 walkMiles <- with(subset(NHTS.trips, {
-    AT == "walk" #& TRVLCMIN > 5 & TRPMILES > 2/9
+    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVLCMIN*60 < 8 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 0.5
 }), TRPMILES)
@@
 cycleHours <- with(subset(NHTS.trips, {
-      AT == "cycle" #& TRVLCMIN > 5 & TRPMILES > 2/9
+      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVLCMIN*60 < 25 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 1
 }), TRVLCMIN/60)
@@
 cycleMiles <- with(subset(NHTS.trips, {
-      AT == "cycle" #& TRVLCMIN > 5 & TRPMILES > 2/9
+      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVLCMIN*60 < 25 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 1
 }), TRPMILES)
@@
 
-Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, mode = "walk")
-Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, mode = "cycle")
+Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, Speed = walkMiles/walkHours, mode = "walk")
+Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, Speed = cycleMiles/cycleHours, mode = "cycle")
+
 D <- rbind(Dwalk,Dcycle)
@@
 g + scale_y_log10() + scale_x_log10() + theme_bw()
+
+g <- ggplot(D, aes(Speed, ..density.., fill = mode)) + geom_histogram(breaks = seq(0,25,1)) + theme_bw() + scale_x_continuous(name="Speed") + scale_y_continuous(name="Density") + facet_grid(. ~ mode)
+g
+
 ```
@@
 
-```{r subset}
+```{r subset, eval = TRUE, fig.width = 10}
 NHTS.person <- subset( NHTS.person, MSA == TRUE )
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
+barplot(with(NHTS,table(HHS)))
 ```
@@
 
-```{r split}
+```{r split, eval = FALSE}
 listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
@@
 
-```{r count}
+```{r count, eval = FALSE}
 counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
@@
 
-```{r ITHIMObjs, eval = FALSE}
+```{r foobar, eval = TRUE}
+sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
+    return(sum(unlist(compareModels(ITHIM.baseline, ITHIM.scenario)$daly.delta)))
+    }
+```
+
+```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10}
+compute <- FALSE
+
+if(compute){
+
+DALYsTotal <- data.frame()
+
+for( muNonTravel in seq(1,75,length.out = 20) ){
 
   ITHIMList <- list()
-
+  rm(DALYs)
   ITHIM.default <- createITHIM(vision = "baseline", region = "national")
@@
     ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
-}
-
-```
-
-```{r CRA, eval = FALSE}
+    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muNonTravel", muNonTravel)
+  }
+
+  ITHIM.national <- createITHIM(vision = "baseline", region = "national")
+  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+  ITHIM.national <- updateITHIM(ITHIM.national, "muwt", 49.63)
+  ITHIM.national <- updateITHIM(ITHIM.national, "muct", 2.50)
+  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
+  DALYs <- data.frame(HHS = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
+  DALYsTotal <- rbind(DALYsTotal, DALYs)
+}
+}else{
+DALYsTotal <- readRDS(file = "./data/DALYsTotal.rds")
+}
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY, colour = HHS))
+p  + geom_line() + theme_bw() + scale_x_discrete(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
+```
+
+### Visualization of trip time matrices
+
+```{r meanMatrices, eval = TRUE, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE}
+
+ITHIMList <- c(ITHIMList, list("Natl" = ITHIM.national))
+
+meanMatrix <- function(ITHIM, mode = "walk"){
+   if(mode == "walk"){
+      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
+}else{
+      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
+}
+}
+
+meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "walk")
+meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "cycle")
+
+D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
+
+names(D) <- c("ageClass", "mode", "sex", "time", "HHS")
+
+#D <- within(D, HHS <- factor(HHS, levels = c(paste0("HHS",1:10),"Natl")))
+
+p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") + facet_grid( mode ~ HHS)
+
+p1
+
+```
+
+```{r CRA, eval = TRUE}
 

### 2016-09-01 R/ITHIM.Rmd
eval = TRUE for whole script

@@
 
-```{r parsely, eval = FALSE}
+```{r parsely, eval = TRUE}
 convertToHHS <- function(states){
@@
 
-```{r NHTS, eval = FALSE}
+```{r NHTS, eval = TRUE, echo = FALSE}
 DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
@@
 
-```{r loadData, eval = FALSE}
+```{r loadData, eval = FALSE, echo = FALSE}
 NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
@@
 
-```{r summary, fig.width = 20, fig.height = 10, eval = TRUE}
+```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE}
 summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
@@
 
-```{r subset, eval = TRUE, fig.width = 10}
+```{r subset, eval = TRUE, fig.width = 10, echo = FALSE}
 NHTS.person <- subset( NHTS.person, MSA == TRUE )
@@
 
-```{r split, eval = FALSE}
+```{r split, eval = TRUE, echo = FALSE}
 listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
@@
 
-```{r count, eval = FALSE}
+```{r count, eval = TRUE, echo = FALSE}
 counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
@@
 
-```{r CDC, eval = FALSE}
+```{r CDC, eval = TRUE, echo = FALSE}
 wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
@@
 
-```{r ITHIMCause, eval = FALSE}
+```{r ITHIMCause, eval = TRUE, echo = FALSE}
 ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
@@
 ```
-```{r recodeAge, eval = FALSE}
+```{r recodeAge, eval = TRUE, echo = FALSE}
 ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
@@
 
-```{r wonder, eval = FALSE}
+```{r wonder, eval = TRUE, echo = FALSE}
 wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
@@
 
-```{r suppressed, eval = FALSE}
+```{r suppressed, eval = TRUE, echo = FALSE}
 wonder <- within( wonder, {
@@
 
-```{r summary2, eval = FALSE}
+```{r summary2, eval = TRUE, echo = FALSE}
 wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
@@
 
-```{r singleyear, eval = FALSE}
+```{r singleyear, eval = TRUE, echo = FALSE}
 pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
@@
 
-```{r sloppy, eval = FALSE}
+```{r sloppy, eval = TRUE, echo = FALSE}
 ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
@@
 
-```{r foop, eval = FALSE}
+```{r foop, eval = TRUE, echo = FALSE}
 pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
@@
 
-```{r USGBD, eval = FALSE}
+```{r USGBD, eval = TRUE, echo = FALSE}
 USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
@@
 
-```{r burdenList, eval = FALSE}
+```{r burdenList, eval = TRUE, echo = FALSE}
 GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
@@
 
-```{r newMETs, echo = TRUE, eval = FALSE}
+```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
 newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
@@
 
-```{r readnewMETs, echo = TRUE, eval = FALSE}
+```{r readnewMETs, echo = TRUE, eval = TRUE, echo = FALSE}
 cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
@@
 
-```{r metKey, echo = TRUE, eval = FALSE}
+```{r metKey, echo = TRUE, eval = TRUE, echo = FALSE}
 allActivities <- colnames(summaryList$Data)
@@
 
-```{r parameters, eval = FALSE}
+```{r parameters, eval = TRUE, echo = FALSE}
 (parameters <- ddply(D, .(SEX, AGECAT), summarise,
@@
 
-```{r foobar, eval = TRUE}
+```{r foobar, eval = TRUE, echo = FALSE}
 sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
@@
 
-```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10}
-compute <- FALSE
+```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10, echo = FALSE}
+compute <- TRUE
 
@@
 
-```{r meanMatrices, eval = TRUE, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE}
+```{r meanMatrices, eval = TRUE, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 
-```{r CRA, eval = TRUE}
+```{r CRA, eval = TRUE, echo = FALSE}
 
@@
 
+```
 Dataset: Underlying Cause of Death, 1999-2014
@@
 resident population that is under one year of age. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Age Group.
+```
 
@@
 
-```{r instantiate, eval = FALSE}
+```{r instantiate, eval = TRUE, echo = FALSE}
 ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
@@
 
-```{r RRs, eval = FALSE}
+```{r RRs, eval = TRUE, echo = FALSE}
 comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
@@
 ### Deaths
-```{r show, echo = FALSE, eval = FALSE}
+```{r show, echo = FALSE, eval = TRUE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### DALY
-```{r show2, echo = FALSE, eval = FALSE}
+```{r show2, echo = FALSE, eval = TRUE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLD
-```{r show3, echo = FALSE, eval = FALSE}
+```{r show3, echo = FALSE, eval = TRUE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLL
-```{r show4, echo = FALSE, eval = FALSE}
+```{r show4, echo = FALSE, eval = TRUE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r instantiate2, eval = FALSE}
+```{r instantiate2, eval = TRUE, echo = FALSE}
 ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
@@
 
-```{r RRs2, eval = FALSE}
+```{r RRs2, eval = TRUE, echo = FALSE}
 comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
@@
 ### Deaths
-```{r show222, echo = FALSE, eval = FALSE}
+```{r show222, echo = FALSE, eval = TRUE}
 matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### DALY
-```{r show22, echo = FALSE, eval = FALSE}
+```{r show22, echo = FALSE, eval = TRUE}
 matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLD
-```{r show23, echo = FALSE, eval = FALSE}
+```{r show23, echo = FALSE, eval = TRUE}
 matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 #### YLL
-```{r show24, echo = FALSE, eval = FALSE}
+```{r show24, echo = FALSE, eval = TRUE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r example, eval = FALSE}
+```{r example, eval = TRUE, echo = FALSE}
 example("ITHIM-package")
@@
 
-```{r vision, eval = FALSE}
+```{r vision, eval = TRUE, echo = FALSE}
 ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
@@
 #### YLL
-```{r show233, echo = FALSE, eval = FALSE}
+```{r show233, echo = FALSE, eval = TRUE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r params, eval = FALSE}
+```{r params, eval = TRUE, echo = FALSE}
 names(ITHIM.baseline$parameters)

### 2016-09-02 R/ITHIM.Rmd
Fixed bug in HHS mapping

@@
                                MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
-                               HHS = as.factor(convertToHHS(HHSTATE))
+                               HHSTATE = as.character(HHSTATE),
+                               HHS = as.character(convertToHHS(as.character(HHSTATE))), stringsAsFactors = FALSE
 
@@
 burden$deaths <- burden$numberOfDeaths5yr/5
+## neil uses 10, and if deaths<10 sets ratio to 1
 burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
@@
 
-p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") + facet_grid( mode ~ HHS)
+p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") + facet_grid( mode ~ HHS, scales = "free_y")
 

### 2016-09-06 R/ITHIM.Rmd
Added national ITHIM build on NHTS data
Also created future-research.org as a place to organize my thoughts on publications and so on.

@@
 tripTime <- function(data, mode = "walk"){
- return(sum(subset(data, AT == mode)$TRVLCMIN, na.rm = TRUE))
+ return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
 }
@@
 }
+
+reformatData2 <- function(data){
+foo <- matrix(unlist(data), nrow = 7, ncol = 2)
+foo <- rbind( c(0.1, 0.1), foo)
+dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
+return(foo[,c("M","F")])
+}
+
 ```
@@
 summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
-summary(with(subset(NHTS.trips, AT == "walk"), TRVLCMIN))
-summary(with(subset(NHTS.trips, AT == "cycle"), TRVLCMIN))
-summary(with(subset(NHTS.trips, AT == "other"), TRVLCMIN))
+summary(with(subset(NHTS.trips, AT == "walk"), TRVL_MIN))
+summary(with(subset(NHTS.trips, AT == "cycle"), TRVL_MIN))
+summary(with(subset(NHTS.trips, AT == "other"), TRVL_MIN))
 
 walkHours <- with(subset(NHTS.trips, {
-    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVLCMIN*60 < 8 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 0.5
-}), TRVLCMIN/60)
+    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
+}), TRVL_MIN/60)
 
 walkMiles <- with(subset(NHTS.trips, {
-    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVLCMIN*60 < 8 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 0.5
+    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
 }), TRPMILES)
@@
 cycleHours <- with(subset(NHTS.trips, {
-      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVLCMIN*60 < 25 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 1
-}), TRVLCMIN/60)
+      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
+}), TRVL_MIN/60)
 
 cycleMiles <- with(subset(NHTS.trips, {
-      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVLCMIN*60 < 25 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 1
+      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
 }), TRPMILES)
@@
 
-#### Split into strata by region
-
-```{r split, eval = TRUE, echo = FALSE}
+#### Compute Natioal Metro Average
+
+```{r natl}
+listOfLists <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
+counts <- lapply(listOfLists, function(x) lapply(x, countPeople))
+walkLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "walk"))
+cycleLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "cycle"))
+
+countMatrix <- reformatData(counts)
+walkMatrix <- reformatData(walkLength)
+cycleMatrix <- reformatData(cycleLength)
+
+meanWalkList <- walkMatrix/countMatrix*7
+meanCycleList <- cycleMatrix/countMatrix*7
+
+RwtList <- ifelse(meanWalkList == 0, 1e-4, meanWalkList)
+RctList <- ifelse(meanCycleList == 0, 1e-4, meanCycleList)
+
+RwtNational <- RwtList/RwtList[3,2]
+RctNational <- RctList/RctList[3,2]
+muwtNational <- meanWalkList[3,2]
+muctNational <- meanCycleList[3,2]
+
+rm("listOfLists","counts","walkLength","cycleLength","countMatrix","walkMatrix","cycleMatrix","meanWalkList","meanCycleList","RwtList","RctList","muwtList","muctList")
+```
+
+#### Compute Mean Trip Lengths
+
+```{r count, eval = TRUE, echo = FALSE}
 listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
-```
-
-#### Compute Mean Trip Lengths
-
-```{r count, eval = TRUE, echo = FALSE}
 counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
@@
 meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
-RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 0.01, x) )
-RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 0.01, x) )
+RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 1e-4, x) )
+RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 1e-4, x) )
 RwtList <- lapply( RwtList, function(x) x/x[3,2] )
@@
 
-for( muNonTravel in seq(1,75,length.out = 20) ){
+for( muNonTravel in seq(1,75,length.out = 3) ){
 
@@
   ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
-  ITHIM.national <- updateITHIM(ITHIM.national, "muwt", 49.63)
-  ITHIM.national <- updateITHIM(ITHIM.national, "muct", 2.50)
+  ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
+  ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
+  ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
+  ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
   DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)

### 2016-09-06 R/ITHIM.Rmd
Updated ITHIM.Rmd

@@
 reformatData <- function(data){
-foo <- matrix(unlist(data), nrow = 7, ncol = 2)
-foo <- rbind( c(0.1, 0.1), foo)
-dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
-return(foo[,c("M","F")])
-}
-
-reformatData2 <- function(data){
 foo <- matrix(unlist(data), nrow = 7, ncol = 2)
@@
 p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY, colour = HHS))
-p  + geom_line() + theme_bw() + scale_x_discrete(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
+p  + geom_line() + theme_bw() + scale_x_continuous(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
 ```

### 2016-09-08 manuscript/manuscript.tex
Created manuscript.tex
This will be the first ITHIM paper from GHI that introduces the package, the UI and the regional analysis.

\documentclass{bioinfo}
\usepackage{url}
\copyrightyear{2016}
\pubyear{2016}

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\compendium}{\emph{ReadSimCompendium}}
\newcommand{\ithim}{\emph{ITHIM}}
\renewcommand{\bowtie}{\emph{bowtie}}
\newcommand{\ebseq}{\emph{EBSeq}}
\newcommand{\deseq}{\emph{DESeq2}}
\newcommand{\R}{\emph{R}}
\newcommand{\rnaseq}{RNA-seq}
\newcommand{\gi}{gene/isoform}
\newcommand{\gis}{genes/isoforms}
\newcommand{\pest}{\texttt{ParamEstimation}}
\newcommand{\csimu}{\texttt{CountSimu}}
\newcommand{\writeRSEM}{\texttt{writeRSEMFiles}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}

\begin{document}
\firstpage{1}

\title[\package{}]{\package{}: an \R{} package for implemetation of
  the ITHIM model}
\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
  $\ldots$
  \ and Jonathan Patz$^{1}$}
\address{$^{1}$Global Health Institute\\
University of Wisconsin{\textendash}Madison, Madison, WI USA\\
}
\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}

\section{Summary:}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
ullamco laboris nisi ut aliquip ex ea commodo consequat.

\section{Availability:}
The source code and data files for \package{} are publicly available
on GitHub at \url{https://github.com/syounkin/ITHIM}.

\section{Supplementary information:}

\url{http://www.ithim.ghi.wisc.edu}.

\section{Contact:}
\href{patz@wisc.edu}{patz@wisc.edu}
\end{abstract}

\section{Introduction}

ITHIM is a mathematical model that integrates data on travel patterns,
physical activity, fine particulate matter, GHG emissions, and disease
and injuries. Based on population and travel scenarios. The model has
been used to calculate the health impacts of walking and bicycling
short distances usually traveled by car or driving low-emission
automobiles \cite{woodcock2013}.

The model uses comparative risk assessment through which it formulates
a change in the disease burden, resulting from the shift in the
exposure distribution from a baseline scenario to an alternative
scenario.

\begin{methods}
\section{Method}

\begin{itemize}
\item NHTS data
\item GBD data
\item ITHIM model parameters
\end{itemize}

\end{methods}

\section{Regional Example}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
ullamco laboris nisi ut aliquip ex ea commodo consequat.

\begin{figure}[t]
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}\centerline{\includegraphics[width=0.4\textwidth]{./figures/ITHIMObjs-2}}
    \caption{Change in total national DALYs for each regional transportation
estimates.
}\label{dalyFigure}
\end{figure}

\section*{Acknowledgment}

\paragraph{Funding\textcolon} SGY is supported by$\ldots$ , JV, JP, $\ldots$

%\bibliographystyle{natbib}
%\bibliographystyle{bmc_article}
\bibliographystyle{mypapers}
\bibliography{ITHIM}
\end{document}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
% RNA-seq is a powerful technology used to measure genome-wide
% transcription and dozens of statistical methods have been developed
% recently to facilitate its use, including methods for estimating
% expression and identifying differential expression (DE)
% \cite{Oshlack2010}. The ability to evaluate these methods and compare
% their performance in various settings is essential to ensure that
% reliable information is obtained. The Microarray Quality Control
% (MAQC) \cite{shi2006} data sets have proven useful toward this
% end. However, they are limited by small sample sizes and old
% protocol. Evaluation of DE analysis methods can be performed with
% simulated data, and a handful of methods to simulate RNA-seq reads
% have been developed including FluxSimulator and BEERS
% \cite{Griebel2012, Grant2011}.  BEERS is slower than \package{} and,
% as with FluxSimulator, does not allow specification of read count
% means and are therefore not useful for DE-based read simulation.

% We have developed \package{} to simulate reads efficiently under a
% realistic model of DE across two treatment groups, and to use these
% reads to estimate expression levels for downstream analyses, all
% within the \R{} environment. \package{} is built on the software
% package \rsem{} used primarily for transcript quantification
% \cite{li2011}.  To induce differential expression of transcripts we
% adjust the relative transcript abundance across treatment groups in
% the empirical model used by the \rsem{} simulator.  The reads will be
% simulated under the same design as the sample data.  If the sample
% data are paired-end then the simulate reads will also be paired-end.

% Transcript abundances vary substantially across \rnaseq{} experiments.
% Factors such as cell type, organism, and experimental protocol affect
% the expected transcript abundances.  \package{} builds its model with
% sample data from an \rnaseq{} experiment, and uses the empirical read
% count distributions to select a realistic distribution for mean and
% dispersion parameters.  The total expression and fold change
% relationship for the simulated data is also modeled from these data
% and fold change levels are selected to mimic this
% relationship. \package{} also takes advantage of the transcript
% quantification provided by \rsem{} by allowing the user to estimate
% read counts from the simulated reads.  Both the read-based simulation
% and differential expression analysis may be performed within \R{}
% using any of the many \R{} packages available for DE analysis, e.g.,
% DESeq2, EdgeR, EBSeq \cite{love2014, Robinson2010, leng2013}.
% \package{} implements a two stage simulation.  First, it simulates
% read counts per transcript under a negative binomial distribution
% using empirical mean and dispersion estimates.  Second, it calculates
% relative transcript abundance using estimates of transcript length and
% read-count, updates the sample model, and executes the \rsem{} read
% simulator.

% In the first step, read counts for the untreated group are drawn from
% a negative binomial distribution with mean and dispersion parameters
% $\hat{\mu_0}$, $\hat{\varphi_0}$ estimated from the untreated
% group. In the treated group, differentially expressed transcripts are
% simulated using $\mu_1 = \gamma\mu_0$, $\varphi_1=\varphi_0$.  The
% parameter $\gamma$ is chosen to mimic the fold-change vs.\ total
% expression relationship observed empirically, see Figure
% \ref{ma-transcript}.  \sgy{Observed fold-change values decrease with
% expression level.  If fold-change was assigned independently of
% expression level then transcripts with high levels of expression would
% be assigned\ldots}

% Fold change among DE transcripts is chosen by binning the transcripts
% by overall expression into equally sized bins, and using the bin-wise
% standard deviation of the log fold-change to construct an appropriate
% interval for the fold-change, from which a value is sampled randomly.
% The interval is defined as $I_{\textrm{bin}} =
% (k_1\sigma_{\textrm{bin}}, k_2\sigma_{\textrm{bin}})$, where
% $\sigma_{\textrm{bin}}$ is the bin-wise standard deviation of log
% fold-change, and $k_1$ and $k_2$ are chosen by the user.  In the
% following chromosome 22 example, values of $k_1=?$ and $k_2=?$ were
% used with $B=?$ bins.  By tuning parameters $k_1$ and $k_2$ the user
% can control the signal strength while preserving the relationship
% between fold-change and total expression. \sgy{\package{} could
%   easily be adapted to allow user-defined values for $\gamma$, $\mu_i$
%   and $\varphi_i$.}

% Three files are needed for \package{}; a transcript template, the
% read-count estimates for all samples, and a sample model file.  All of
% these are easily created using \rsem{} utilities. A GTF file,
% transcript-to-gene file such as ``knownIsoforms.txt'' from the UCSC
% Genome Browser, and a FASTA reference sequence, or a Multi-FASTA file
% containing transcript sequences, may be used to create the reference
% files.

% In the second stage, the simulated read counts are used to compute the
% relative transcript abundance needed for the read simulation.
% \package{} can return either the gene or isoform level estimated read
% counts from \rsem{}, or simply write the fastq read file.  The
% software package \bowtie{} is used for the alignment
% \cite{langmead2009}.

% We simulated 1,000,000 reads for each of ten samples from two
% treatment groups, and performed DE analysis on all transcripts from
% chromosome 22 using \ebseq{} \cite{leng2013}.  One hundred genes (288
% transcripts) were selected at random to be DE.  In the top of Figure
% \ref{ma-transcript} the log fold-change is plotted versus total
% expression given the simulated counts.  Beneath it is the same figure,
% bottom of Figure \ref{ma-transcript}, but using the estimate counts of
% the simulated reads.  The reads are simulated in such a way as to
% preserve the relationship between fold-change and total expression.

% Using the transcript-wise estimated read counts from the simulated
% data, we performed DE analysis using \ebseq{}.  Of the 288
% differentially expressed transcripts, 179 (62\%) were correctly
% identified by \ebseq{}, with 29 false discoveries.

% %% In this example several of the DE transcripts were assigned a
% %% posterior probability of DE to be zero.  This is due to one of the
% %% treatment groups having zero count.  Also, some of the DE genes are
% %% removed by \ebseq{} because the read count is too low.  Thus some of
% %% the DE transcripts are unidentifiable.  If we remove them we have a
% %% true positive rate of $x$\%.

% Unaligned read data from twenty RNA-seq chips, along with a set of
% chromosome 22 reference files were used to simulate the reads in this
% example.  The example takes about 3 hours to run and may be found in
% the compendium package \compendium{}, available at
% \url{http://www.biostat.wisc.edu/~syounkin/}.

### 2016-09-08 R/ITHIM.Rmd

@@
 
-```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE}
+```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
 library("devtools")
@@
 
-```{r parsely, eval = TRUE}
+```{r parsely, eval = TRUE, echo = FALSE}
 convertToHHS <- function(states){
-  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
+  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
   StateToHHS <- HHSRegions[,3]
@@
 names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
-
-## DAYV2PUB <- within(DAYV2PUB, {
-##     URBRUR <- as.factor(URBRUR)
-##     HHSTATE <- as.factor(HHSTATE)
-##     MSACAT <- as.factor(MSACAT)
-##     TRPTRANS <- as.factor(TRPTRANS)
-##     TRPMILES <- as.numeric(TRPMILES)
-## })
 
@@
 
-```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE}
-summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
-summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
-summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
-summary(with(subset(NHTS.trips, AT == "walk"), TRVL_MIN))
-summary(with(subset(NHTS.trips, AT == "cycle"), TRVL_MIN))
-summary(with(subset(NHTS.trips, AT == "other"), TRVL_MIN))
+```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+#summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
+#summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
+#summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
+#summary(with(subset(NHTS.trips, AT == "walk"), TRVL_MIN))
+#summary(with(subset(NHTS.trips, AT == "cycle"), TRVL_MIN))
+#summary(with(subset(NHTS.trips, AT == "other"), TRVL_MIN))
 
@@
 
-g <- ggplot(D, aes(Speed, ..density.., fill = mode)) + geom_histogram(breaks = seq(0,25,1)) + theme_bw() + scale_x_continuous(name="Speed") + scale_y_continuous(name="Density") + facet_grid(. ~ mode)
+g <- ggplot(D, aes(Speed, ..density.., fill = mode)) + geom_histogram() + theme_bw() + scale_x_continuous(name="Speed") + scale_y_continuous(name="Density") + facet_grid(. ~ mode, scales = "free")
 g
@@
 #### Subset and Merge
+
+How do we merge?  Are there people who were surveyed who had zero
+travel time?  Couldnt have been very many.  But I think there are
+people in the PERSON table that were not interviewed.
 
@@
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
-barplot(with(NHTS,table(HHS)))
 ```
@@
 
-```{r natl}
+```{r natl, eval = TRUE, echo = FALSE, warning = FALSE}
 listOfLists <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
@@
 
-```{r suppressed, eval = TRUE, echo = FALSE}
+```{r suppressed, eval = TRUE, echo = FALSE, warning = FALSE}
 wonder <- within( wonder, {
@@
 ```{r parameters, eval = TRUE, echo = FALSE}
-(parameters <- ddply(D, .(SEX, AGECAT), summarise,
+parameters <- ddply(D, .(SEX, AGECAT), summarise,
              mean = mean(METs, na.rm=T),
              sd = sd(METs, na.rm=T)
-             ))
+             )
 
 METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
-(METmatrix <- METmatrix/METmatrix[3,2])
+METmatrix <- METmatrix/METmatrix[3,2]
 ```
@@
 
-```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10, echo = FALSE}
+```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10, echo = FALSE, warning = FALSE}
 compute <- TRUE
@@
 
-for( muNonTravel in seq(1,75,length.out = 3) ){
+for( muNonTravel in seq(1,50,length.out = 20) ){
 
@@
 
-  for (HHS in paste0("HHS",1:10)){
+  for (HHS in c(paste0("HHS0",1:9),"HHS10")){
     ITHIMList[[HHS]] <- ITHIM.default
@@
 }
+
 }else{
@@
 }
+
+barplot(with(NHTS,table(HHS)))
 p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY, colour = HHS))
-p  + geom_line() + theme_bw() + scale_x_continuous(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
+p  + geom_line(size = 2) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
 ```
@@
 ```
-
-```{r CRA, eval = TRUE, echo = FALSE}
-
-CRA.fn <- function(i, j){
-  foo <- compareModels( ITHIMList[[i]], ITHIMList[[j]])
-  sum(unlist(foo$AF))
-}
-
-CRAMatrix <- matrix(NA, nrow = 10, ncol = 10)
-
-for( i in 1:10 ){
-  CRA <- c()
-  for( j in 1:10 ){
-    CRA <- c( CRA, CRA.fn(i, j))
-  }
-  CRAMatrix[i,] <- CRA
-}
-
-which(CRAMatrix == max(CRAMatrix), arr.ind = TRUE)
-
-```
-
-## DETAILS ON WONDER QUERY
+# Appendix
+
+## Example (Bay Area)
+
+First we define the baseline and scenario objects.
+
+```{r instantiate, eval = TRUE, echo = FALSE}
+ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
+ITHIM.scenario <- createITHIM(vision = "scenario", region = "SFBayArea")
+```
+
+Next we compare the scenario to the baseline using compareModels().
+
+```{r RRs, eval = TRUE, echo = FALSE}
+comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+```
+
+### Deaths
+```{r show, echo = FALSE, eval = TRUE, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+### DALY
+```{r show2, echo = FALSE, eval = TRUE, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+### YLD
+```{r show3, echo = FALSE, eval = TRUE, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+### YLL
+```{r show4, echo = FALSE, eval = TRUE, echo = FALSE}
+matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+
+## Example (National)
+
+```{r instantiate2, eval = TRUE, echo = FALSE}
+ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
+ITHIM.scenario <- createITHIM(vision = "scenario", region = "national")
+```
+
+```{r RRs2, eval = TRUE, echo = FALSE}
+comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
+```
+
+### Deaths
+```{r show222, echo = FALSE, eval = TRUE}
+matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+### DALY
+```{r show22, echo = FALSE, eval = TRUE}
+matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+### YLD
+```{r show23, echo = FALSE, eval = TRUE}
+matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+#### YLL
+```{r show24, echo = FALSE, eval = TRUE}
+matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+
+## Example
+
+```{r example, eval = TRUE, echo = FALSE}
+example("ITHIM-package")
+```
+
+### Creating Different Visions
+
+```{r vision, eval = TRUE, echo = FALSE}
+ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
+comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
+```
+#### YLL
+```{r show233, echo = FALSE, eval = TRUE}
+matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
+```
+
+## Interrogate Parameters
+
+```{r params, eval = TRUE, echo = FALSE}
+names(ITHIM.baseline$parameters)
+ITHIM.baseline$parameters$NonTravelMETs
+ITHIM.baseline$parameters$Rwt
+```
+
+## Appendix
+
+### DETAILS ON WONDER QUERY
 
@@
 
-## Example (Bay Area)
-
-First we define the baseline and scenario objects.
-
-```{r instantiate, eval = TRUE, echo = FALSE}
-ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
-ITHIM.scenario <- createITHIM(vision = "scenario", region = "SFBayArea")
-```
-
-Next we compare the scenario to the baseline using compareModels().
-
-```{r RRs, eval = TRUE, echo = FALSE}
-comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-```
-
-### Deaths
-```{r show, echo = FALSE, eval = TRUE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### DALY
-```{r show2, echo = FALSE, eval = TRUE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### YLD
-```{r show3, echo = FALSE, eval = TRUE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### YLL
-```{r show4, echo = FALSE, eval = TRUE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-
-## Example (National)
-
-```{r instantiate2, eval = TRUE, echo = FALSE}
-ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
-ITHIM.scenario <- createITHIM(vision = "scenario", region = "national")
-```
-
-```{r RRs2, eval = TRUE, echo = FALSE}
-comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-```
-
-### Deaths
-```{r show222, echo = FALSE, eval = TRUE}
-matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### DALY
-```{r show22, echo = FALSE, eval = TRUE}
-matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### YLD
-```{r show23, echo = FALSE, eval = TRUE}
-matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-#### YLL
-```{r show24, echo = FALSE, eval = TRUE}
-matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-
-## Example
-
-```{r example, eval = TRUE, echo = FALSE}
-example("ITHIM-package")
-```
-
-### Creating Different Visions
-
-```{r vision, eval = TRUE, echo = FALSE}
-ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
-comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
-```
-#### YLL
-```{r show233, echo = FALSE, eval = TRUE}
-matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-
-## Interrogate Parameters
-
-```{r params, eval = TRUE, echo = FALSE}
-names(ITHIM.baseline$parameters)
-ITHIM.baseline$parameters$NonTravelMETs
-ITHIM.baseline$parameters$Rwt
-```
-
-## Appendix
-
 ```{r appendix}

### 2016-09-12 manuscript/manuscript.tex
Added symbolic link to manuscript/figures/
Edits to manuscript 1

@@
 \newcommand{\package}{\emph{ITHIM}}
-\newcommand{\compendium}{\emph{ReadSimCompendium}}
 \newcommand{\ithim}{\emph{ITHIM}}
-\renewcommand{\bowtie}{\emph{bowtie}}
-\newcommand{\ebseq}{\emph{EBSeq}}
-\newcommand{\deseq}{\emph{DESeq2}}
 \newcommand{\R}{\emph{R}}
-\newcommand{\rnaseq}{RNA-seq}
-\newcommand{\gi}{gene/isoform}
-\newcommand{\gis}{genes/isoforms}
-\newcommand{\pest}{\texttt{ParamEstimation}}
-\newcommand{\csimu}{\texttt{CountSimu}}
-\newcommand{\writeRSEM}{\texttt{writeRSEMFiles}}
 \newcommand{\sgy}[1]{{\itshape\col{#1}}}
 \newcommand{\XX}{\sgy{XX}}
+\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
 
@@
 \section{Summary:}
+We divided the U.S. metropolitan regions into 10 regions defined by
+the Department of Health and Human Services (HHS1-10).  We performed a
+comparative risk analysis for each of the regions using national
+averages as the baseline.  We investigated the active transport
+component of the ITHIM model by estimating regional travel parameters
+using the National Household Transportation Survey.  The West coast
+and the Northeast of the U.S. have travel patterns that, if mirrored
+by the nation, would save the U.S. thousands to hundreds of thousands
+DALYs.
+
+\section{Availability:}
+\package{} may be explored using the user interface avaiable
+at \webpage{}.  The source code and data files for \package{} are
+publicly available on GitHub at \webpage{}.
+
+\section{Supplementary information:}
+
+See \webpage{} for links to the source code, manual, tutorial and
+links for more information about the ITHIM model.
+
+\section{Contact:}
+\href{javargo@wisc.edu}{javargo@wisc.edu}
+\end{abstract}
+
+\section{Introduction}
+ITHIM is a statistical model that integrates data on travel patterns,
+physical activity, fine particulate matter, GHG emissions, and disease
+and injuries. Based on population and travel scenarios. The model has
+been used to calculate the health impacts of walking and bicycling
+short distances usually traveled by car or driving low-emission
+automobiles \cite{woodcock2013}.
+
+The model uses comparative risk assessment through which it formulates
+a change in the disease burden, resulting from the shift in the
+exposure distribution from a baseline scenario to an alternative
+scenario.
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
+minim veniam, quis nostrud exercitation ullamco laboris nisi ut
+aliquip ex ea commodo consequat. Duis aute irure dolor in
+reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
+pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
+culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
+dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua.
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
+minim veniam, quis nostrud exercitation ullamco laboris nisi ut
+aliquip ex ea commodo consequat. Duis aute irure dolor in
+reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
+pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
+culpa qui officia deserunt mollit anim id est laborum.
+
+\begin{methods}
+\section{Methods}
+\begin{itemize}
+\item The coefficient of variation was fixed at ??? for all models.
+  Earlier ITHIM models estimated this as a function of total active
+  tranpsort time to model the decrease in variance of active transport
+  time that occurs for increasing means, i.e., fitter communities have
+  less variance in active transport time.
+\item The mean non-travel activity for the referent group was modeled
+  as a variable.  The relative means across age-sex strata were fixed,
+  and estimated using ATUS.
+\item The national DALYs were stimated using the attributable fraction
+  provided by the comparitive risk assesment.  The estimates for
+  disease burned came from WHO and their GBD study.
+\item Not all of the regions had the same sample size.  In fact a few
+  of the regions have relativel low sample size.  Maybe we should put
+  confidence intervals on our estimates.
+\end{itemize}
 
@@
 
-\section{Availability:}
-The source code and data files for \package{} are publicly available
-on GitHub at \url{https://github.com/syounkin/ITHIM}.
-
-\section{Supplementary information:}
-
-\url{http://www.ithim.ghi.wisc.edu}.
-
-\section{Contact:}
-\href{patz@wisc.edu}{patz@wisc.edu}
-\end{abstract}
-
-\section{Introduction}
-
-ITHIM is a mathematical model that integrates data on travel patterns,
-physical activity, fine particulate matter, GHG emissions, and disease
-and injuries. Based on population and travel scenarios. The model has
-been used to calculate the health impacts of walking and bicycling
-short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013}.
-
-The model uses comparative risk assessment through which it formulates
-a change in the disease burden, resulting from the shift in the
-exposure distribution from a baseline scenario to an alternative
-scenario.
-
-\begin{methods}
-\section{Method}
-
-\begin{itemize}
-\item NHTS data
-\item GBD data
-\item ITHIM model parameters
-\end{itemize}
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
+minim veniam, quis nostrud exercitation ullamco laboris nisi ut
+aliquip ex ea commodo consequat. Duis aute irure dolor in
+reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
+pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
+culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
+dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
+commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
+velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
+occaecat cupidatat non proident, sunt in culpa qui officia deserunt
+mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
+adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
+magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
+ullamco laboris nisi ut aliquip ex ea commodo consequat.
 
@@
 \section{Regional Example}
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
+minim veniam, quis nostrud exercitation ullamco laboris nisi ut
+aliquip ex ea commodo consequat. Duis aute irure dolor in
+reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
+pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
+culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
+dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
+commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
+velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
+occaecat cupidatat non proident, sunt in culpa qui officia deserunt
+mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
+adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
+magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
+ullamco laboris nisi ut aliquip ex ea commodo consequat.
 
@@
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}\centerline{\includegraphics[width=0.4\textwidth]{./figures/ITHIMObjs-2}}
+    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}\centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
     \caption{Change in total national DALYs for each regional transportation
@@
 
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
+minim veniam, quis nostrud exercitation ullamco laboris nisi ut
+aliquip ex ea commodo consequat. Duis aute irure dolor in
+reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
+pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
+culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
+dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
+commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
+velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
+occaecat cupidatat non proident, sunt in culpa qui officia deserunt
+mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
+adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
+magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
+ullamco laboris nisi ut aliquip ex ea commodo consequat.
+
 \section*{Acknowledgment}
-
 \paragraph{Funding\textcolon} SGY is supported by$\ldots$ , JV, JP, $\ldots$
 
-%\bibliographystyle{natbib}
+\bibliographystyle{natbib}
 %\bibliographystyle{bmc_article}
-\bibliographystyle{mypapers}
+%\bibliographystyle{mypapers}
 \bibliography{ITHIM}

### 2016-09-12 manuscript/manuscript.tex
Edits to manuscript 1

@@
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}\centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
-    \caption{Change in total national DALYs for each regional transportation
-estimates.
-}\label{dalyFigure}
+    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}
+    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
+    \caption{Change in total national DALYs for each regional
+      transportation estimates.}\label{dalyFigure}
 \end{figure}
@@
 \end{document}
-
-%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-%\bibliographystyle{achemnat}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{abbrv}
-%\bibliographystyle{bioinformatics}
-%
-%\bibliographystyle{plain}
-%
-% RNA-seq is a powerful technology used to measure genome-wide
-% transcription and dozens of statistical methods have been developed
-% recently to facilitate its use, including methods for estimating
-% expression and identifying differential expression (DE)
-% \cite{Oshlack2010}. The ability to evaluate these methods and compare
-% their performance in various settings is essential to ensure that
-% reliable information is obtained. The Microarray Quality Control
-% (MAQC) \cite{shi2006} data sets have proven useful toward this
-% end. However, they are limited by small sample sizes and old
-% protocol. Evaluation of DE analysis methods can be performed with
-% simulated data, and a handful of methods to simulate RNA-seq reads
-% have been developed including FluxSimulator and BEERS
-% \cite{Griebel2012, Grant2011}.  BEERS is slower than \package{} and,
-% as with FluxSimulator, does not allow specification of read count
-% means and are therefore not useful for DE-based read simulation.
-
-% We have developed \package{} to simulate reads efficiently under a
-% realistic model of DE across two treatment groups, and to use these
-% reads to estimate expression levels for downstream analyses, all
-% within the \R{} environment. \package{} is built on the software
-% package \rsem{} used primarily for transcript quantification
-% \cite{li2011}.  To induce differential expression of transcripts we
-% adjust the relative transcript abundance across treatment groups in
-% the empirical model used by the \rsem{} simulator.  The reads will be
-% simulated under the same design as the sample data.  If the sample
-% data are paired-end then the simulate reads will also be paired-end.
-
-% Transcript abundances vary substantially across \rnaseq{} experiments.
-% Factors such as cell type, organism, and experimental protocol affect
-% the expected transcript abundances.  \package{} builds its model with
-% sample data from an \rnaseq{} experiment, and uses the empirical read
-% count distributions to select a realistic distribution for mean and
-% dispersion parameters.  The total expression and fold change
-% relationship for the simulated data is also modeled from these data
-% and fold change levels are selected to mimic this
-% relationship. \package{} also takes advantage of the transcript
-% quantification provided by \rsem{} by allowing the user to estimate
-% read counts from the simulated reads.  Both the read-based simulation
-% and differential expression analysis may be performed within \R{}
-% using any of the many \R{} packages available for DE analysis, e.g.,
-% DESeq2, EdgeR, EBSeq \cite{love2014, Robinson2010, leng2013}.
-% \package{} implements a two stage simulation.  First, it simulates
-% read counts per transcript under a negative binomial distribution
-% using empirical mean and dispersion estimates.  Second, it calculates
-% relative transcript abundance using estimates of transcript length and
-% read-count, updates the sample model, and executes the \rsem{} read
-% simulator.
-
-% In the first step, read counts for the untreated group are drawn from
-% a negative binomial distribution with mean and dispersion parameters
-% $\hat{\mu_0}$, $\hat{\varphi_0}$ estimated from the untreated
-% group. In the treated group, differentially expressed transcripts are
-% simulated using $\mu_1 = \gamma\mu_0$, $\varphi_1=\varphi_0$.  The
-% parameter $\gamma$ is chosen to mimic the fold-change vs.\ total
-% expression relationship observed empirically, see Figure
-% \ref{ma-transcript}.  \sgy{Observed fold-change values decrease with
-% expression level.  If fold-change was assigned independently of
-% expression level then transcripts with high levels of expression would
-% be assigned\ldots}
-
-% Fold change among DE transcripts is chosen by binning the transcripts
-% by overall expression into equally sized bins, and using the bin-wise
-% standard deviation of the log fold-change to construct an appropriate
-% interval for the fold-change, from which a value is sampled randomly.
-% The interval is defined as $I_{\textrm{bin}} =
-% (k_1\sigma_{\textrm{bin}}, k_2\sigma_{\textrm{bin}})$, where
-% $\sigma_{\textrm{bin}}$ is the bin-wise standard deviation of log
-% fold-change, and $k_1$ and $k_2$ are chosen by the user.  In the
-% following chromosome 22 example, values of $k_1=?$ and $k_2=?$ were
-% used with $B=?$ bins.  By tuning parameters $k_1$ and $k_2$ the user
-% can control the signal strength while preserving the relationship
-% between fold-change and total expression. \sgy{\package{} could
-%   easily be adapted to allow user-defined values for $\gamma$, $\mu_i$
-%   and $\varphi_i$.}
-
-% Three files are needed for \package{}; a transcript template, the
-% read-count estimates for all samples, and a sample model file.  All of
-% these are easily created using \rsem{} utilities. A GTF file,
-% transcript-to-gene file such as ``knownIsoforms.txt'' from the UCSC
-% Genome Browser, and a FASTA reference sequence, or a Multi-FASTA file
-% containing transcript sequences, may be used to create the reference
-% files.
-
-% In the second stage, the simulated read counts are used to compute the
-% relative transcript abundance needed for the read simulation.
-% \package{} can return either the gene or isoform level estimated read
-% counts from \rsem{}, or simply write the fastq read file.  The
-% software package \bowtie{} is used for the alignment
-% \cite{langmead2009}.
-
-% We simulated 1,000,000 reads for each of ten samples from two
-% treatment groups, and performed DE analysis on all transcripts from
-% chromosome 22 using \ebseq{} \cite{leng2013}.  One hundred genes (288
-% transcripts) were selected at random to be DE.  In the top of Figure
-% \ref{ma-transcript} the log fold-change is plotted versus total
-% expression given the simulated counts.  Beneath it is the same figure,
-% bottom of Figure \ref{ma-transcript}, but using the estimate counts of
-% the simulated reads.  The reads are simulated in such a way as to
-% preserve the relationship between fold-change and total expression.
-
-% Using the transcript-wise estimated read counts from the simulated
-% data, we performed DE analysis using \ebseq{}.  Of the 288
-% differentially expressed transcripts, 179 (62\%) were correctly
-% identified by \ebseq{}, with 29 false discoveries.
-
-% %% In this example several of the DE transcripts were assigned a
-% %% posterior probability of DE to be zero.  This is due to one of the
-% %% treatment groups having zero count.  Also, some of the DE genes are
-% %% removed by \ebseq{} because the read count is too low.  Thus some of
-% %% the DE transcripts are unidentifiable.  If we remove them we have a
-% %% true positive rate of $x$\%.
-
-% Unaligned read data from twenty RNA-seq chips, along with a set of
-% chromosome 22 reference files were used to simulate the reads in this
-% example.  The example takes about 3 hours to run and may be found in
-% the compendium package \compendium{}, available at
-% \url{http://www.biostat.wisc.edu/~syounkin/}.

### 2016-09-13 manuscript/manuscript.tex
Updates to manuscript

--- 2016-09-08 manuscript/manuscript.tex
@@
 short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013}.
-
-The model uses comparative risk assessment through which it formulates
-a change in the disease burden, resulting from the shift in the
-exposure distribution from a baseline scenario to an alternative
-scenario.
+automobiles \cite{woodcock2013,maizlish2013}.
 
@@
 
-\begin{itemize}
-\item NHTS data
-\item GBD data
-\item ITHIM model parameters
-\end{itemize}
+The ITHIM model uses comparative risk assessment through which it
+formulates a change in the disease burden, resulting from the shift in
+the exposure distribution from a baseline scenario to an alternative
+scenario.
 
-\end{methods}
+\subsection{Data Sources}
 
-\section{Regional Example}
+\subsection{U.S. Metro-Regional Analysis}
 
@@
 \end{figure}
-
+\end{methods}
 \section*{Acknowledgment}
@@
 %\bibliographystyle{natbib}
-%\bibliographystyle{bmc_article}
-\bibliographystyle{mypapers}
+\bibliographystyle{bmc_article}
+%\bibliographystyle{mypapers}
 \bibliography{ITHIM}

### 2016-09-13 manuscript/manuscript.tex

--- 2016-09-12 manuscript/manuscript.tex
@@
 culpa qui officia deserunt mollit anim id est laborum.
+
+The ITHIM model uses comparative risk assessment through which it
+formulates a change in the disease burden, resulting from the shift in
+the exposure distribution from a baseline scenario to an alternative
+scenario.
+
+\subsection{Data Sources}
+
+\subsection{U.S. Metro-Regional Analysis}
 
@@
 \section{Regional Example}
+>>>>>>> 72bfec401397d734de4e6e2096ea97c4bd3437ce
 Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
@@
 
+\end{methods}
+\section*{Acknowledgment}
+
 Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do

### 2016-09-13 manuscript/manuscript.tex
Bugs from merge ironed out.

@@
 \begin{methods}
+
 \section{Methods}
+
 \begin{itemize}
@@
 \section{Regional Example}
->>>>>>> 72bfec401397d734de4e6e2096ea97c4bd3437ce
 Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
@@
 
-\end{methods}
-\section*{Acknowledgment}
-
 Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do

### 2016-09-13 manuscript/manuscript.tex
Color coded the text Red for Jason, Blue for Sam

@@
 \section{Summary:}
-We divided the U.S. metropolitan regions into 10 regions defined by
+\col[blue]{We divided the U.S. metropolitan regions into 10 regions defined by
 the Department of Health and Human Services (HHS1-10).  We performed a
@@
 by the nation, would save the U.S. thousands to hundreds of thousands
-DALYs.
+DALYs.}
 
 \section{Availability:}
-\package{} may be explored using the user interface avaiable
+\col[blue]{\package{} may be explored using the user interface avaiable
 at \webpage{}.  The source code and data files for \package{} are
-publicly available on GitHub at \webpage{}.
+publicly available on GitHub at \webpage{}.}
 
@@
 
-See \webpage{} for links to the source code, manual, tutorial and
-links for more information about the ITHIM model.
+\col[blue]{See \webpage{} for links to the source code, manual, tutorial and
+links for more information about the ITHIM model.}
 
 \section{Contact:}
-\href{javargo@wisc.edu}{javargo@wisc.edu}
+\col[blue]{\href{javargo@wisc.edu}{javargo@wisc.edu}}
 \end{abstract}
@@
 \section{Introduction}
-ITHIM is a statistical model that integrates data on travel patterns,
+\col{ITHIM is a statistical model that integrates data on travel patterns,
 physical activity, fine particulate matter, GHG emissions, and disease
@@
 short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013,maizlish2013}.
+automobiles \cite{woodcock2013,maizlish2013}.}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+\col{The ITHIM model uses comparative risk assessment through which it
+formulates a change in the disease burden, resulting from the shift in
+the exposure distribution from a baseline scenario to an alternative
+scenario.}
+
+\col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua.
+incididunt ut labore et dolore magna aliqua.}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+\col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum.
-
-The ITHIM model uses comparative risk assessment through which it
-formulates a change in the disease burden, resulting from the shift in
-the exposure distribution from a baseline scenario to an alternative
-scenario.
-
-\subsection{Data Sources}
-
-\subsection{U.S. Metro-Regional Analysis}
+culpa qui officia deserunt mollit anim id est laborum.}
 
@@
 
-\begin{itemize}
-\item The coefficient of variation was fixed at ??? for all models.
-  Earlier ITHIM models estimated this as a function of total active
-  tranpsort time to model the decrease in variance of active transport
-  time that occurs for increasing means, i.e., fitter communities have
-  less variance in active transport time.
-\item The mean non-travel activity for the referent group was modeled
-  as a variable.  The relative means across age-sex strata were fixed,
-  and estimated using ATUS.
-\item The national DALYs were stimated using the attributable fraction
-  provided by the comparitive risk assesment.  The estimates for
-  disease burned came from WHO and their GBD study.
-\item Not all of the regions had the same sample size.  In fact a few
-  of the regions have relativel low sample size.  Maybe we should put
-  confidence intervals on our estimates.
-\end{itemize}
+\col[blue]{\begin{itemize}
+  \item The coefficient of variation was fixed at ??? for all strata
+    and models.  Earlier ITHIM models estimated this as a function of
+    total active tranpsort time to model the decrease in variance of
+    active transport time that occurs for increasing means, i.e.,
+    fitter communities have less variance in active transport time.
+  \item The mean non-travel activity for the referent group was
+    modeled as a variable.  The relative means across age-sex strata
+    were fixed, and estimated using ATUS.
+  \item The national DALYs were stimated using the attributable
+    fraction provided by the comparitive risk assesment.  The
+    estimates for disease burned came from WHO and their GBD study.
+  \item Not all of the regions had the same sample size.  In fact a
+    few of the regions have relativel low sample size.  Maybe we
+    should put confidence intervals on our estimates.
+  \item The ITHIM model stratifies the population by sex and age using
+    eight age classes.  Within each stratum active transport time is
+    modeled as log-normal with mean $\hat{\mu}_{i,j}$ which is
+    estimated from the National Household Transportation Survey.
+  \item Quintiles, in this case the quantiles for percentages
+    $(0.1,0.3,0.5,0.7,0.9)$, are used to discretize the integral
+    defined in the Comparative Risk Assesment (CRA)
+  \item 
+\end{itemize}}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+\subsection{Data Sources}
+
+\col[red]{\begin{itemize}
+\item National Household Transportation Survey (NHTS)
+\item American time Use Survey
+\item CDC Wonder
+\end{itemize}}
+
+\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing)) elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.
+ullamco laboris nisi ut aliquip ex ea commodo consequat.}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.
+ullamco laboris nisi ut aliquip ex ea commodo consequat.}
 
@@
 
-\section{Regional Example}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+\section{U.S. Metro-Regional Analysis}
+\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.
+ullamco laboris nisi ut aliquip ex ea commodo consequat.}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+\begin{figure}[t]
+    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}
+    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
+    \caption{Change in total national DALYs for each regional
+      transportation estimates.}\label{dalyFigure}
+\end{figure}
+
+\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.
-
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
-    \caption{Change in total national DALYs for each regional
-      transportation estimates.}\label{dalyFigure}
-\end{figure}
-
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
-dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
-occaecat cupidatat non proident, sunt in culpa qui officia deserunt
-mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
-adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
-magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.
+ullamco laboris nisi ut aliquip ex ea commodo consequat.}
 
@@
 
-\bibliographystyle{natbib}
+%\bibliographystyle{natbib}
 %\bibliographystyle{bmc_article}
-%\bibliographystyle{mypapers}
+\bibliographystyle{mypapers}
 \bibliography{ITHIM}

### 2016-09-14 manuscript/manuscript.tex
Updates to manuscript
abstract changes

@@
 \section{Summary:}
-\col[blue]{We divided the U.S. metropolitan regions into 10 regions defined by
-the Department of Health and Human Services (HHS1-10).  We performed a
-comparative risk analysis for each of the regions using national
-averages as the baseline.  We investigated the active transport
-component of the ITHIM model by estimating regional travel parameters
-using the National Household Transportation Survey.  The West coast
+\col[blue]{While increased active transportation, primarily cycling and walk, is often held up as a goal for improved public health — via reduced chronic disease, improved air quality, and reductions in greenhouse gas emissions — obtaining estimates for the quantiative impacts of such shifts remains infrequent, in not altogether absent.
+Using U.S.
+metropolitan counties from  the 10 regions defined by
+the Department of Health and Human Services (HHS1-10) we perform a
+comparative risk analysis between each regions and national
+averages as the baseline.
+We investigated the reduced chronic disease burden of (x,y,z outcomes) from increased physical activity from population shifts in the amount of biking and walking doen for transport.
+Our methods employ the Integrated Transport and Health Impacts Model (ITHIM) informed by regional travel parameters
+using the National Household Transportation Survey, non-travel activity metircs from the American Time Use Survey, and baseline Disease burden from the Centers for Disease Control and Prevention's WONDER database.
+Our findings show that the West coast
 and the Northeast of the U.S. have travel patterns that, if mirrored
@@
 \section{Introduction}
+\col{}
+
 \col{ITHIM is a statistical model that integrates data on travel patterns,
@@
 \end{itemize}}
+
+The ITHIM model uses the comparative risk assesment framework
 

### 2016-09-15 manuscript/manuscript.tex
Worked on figure and changed permissions
I changed the permissins only so it looks nicer when using ls
Updates to ITHIM manuscript 1

@@
 \newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
-
+\newcommand{\af}{\rho}
+\newcommand{\mets}{MET-hrs./week}
 \begin{document}
@@
 \section{Summary:}
-\col[blue]{While increased active transportation, primarily cycling and walk, is often held up as a goal for improved public health — via reduced chronic disease, improved air quality, and reductions in greenhouse gas emissions — obtaining estimates for the quantiative impacts of such shifts remains infrequent, in not altogether absent.
-Using U.S.
-metropolitan counties from  the 10 regions defined by
-the Department of Health and Human Services (HHS1-10) we perform a
-comparative risk analysis between each regions and national
-averages as the baseline.
-We investigated the reduced chronic disease burden of (x,y,z outcomes) from increased physical activity from population shifts in the amount of biking and walking doen for transport.
-Our methods employ the Integrated Transport and Health Impacts Model (ITHIM) informed by regional travel parameters
-using the National Household Transportation Survey, non-travel activity metircs from the American Time Use Survey, and baseline Disease burden from the Centers for Disease Control and Prevention's WONDER database.
-Our findings show that the West coast
-and the Northeast of the U.S. have travel patterns that, if mirrored
-by the nation, would save the U.S. thousands to hundreds of thousands
-DALYs.}
+\col[blue]{While increased active transportation, primarily cycling
+  and walk, is often held up as a goal for improved public health —
+  via reduced chronic disease, improved air quality, and reductions in
+  greenhouse gas emissions --- obtaining estimates for the quantiative
+  impacts of such shifts remains infrequent, in not altogether
+  absent. Using U.S. metropolitan counties from the 10 regions defined
+  by the Department of Health and Human Services (HHS1-10) we perform
+  a comparative risk analysis between each regions and national
+  averages as the baseline.  We investigated the reduced chronic
+  disease burden of (x,y,z outcomes) from increased physical activity
+  from population shifts in the amount of biking and walking doen for
+  transport.  Our methods employ the Integrated Transport and Health
+  Impacts Model (ITHIM) informed by regional travel parameters using
+  the National Household Transportation Survey, non-travel activity
+  metircs from the American Time Use Survey, and baseline Disease
+  burden from the Centers for Disease Control and Prevention's WONDER
+  database.  Our findings show that the West coast and the Northeast
+  of the U.S. have travel patterns that, if mirrored by the nation,
+  would save the U.S. thousands to hundreds of thousands DALYs.}
 
 \section{Availability:}
-\col[blue]{\package{} may be explored using the user interface avaiable
-at \webpage{}.  The source code and data files for \package{} are
-publicly available on GitHub at \webpage{}.}
+\col[blue]{\package{} may be explored using the user interface
+  avaiable at \webpage{}.  The source code and data files for
+  \package{} are publicly available on GitHub at \webpage{}.}
 
@@
 
-\col[blue]{See \webpage{} for links to the source code, manual, tutorial and
-links for more information about the ITHIM model.}
+\col[blue]{See \webpage{} for links to the source code, manual,
+  tutorial and links for more information about the ITHIM model.}
 
@@
 \section{Introduction}
-\col{}
 
-\col{ITHIM is a statistical model that integrates data on travel patterns,
-physical activity, fine particulate matter, GHG emissions, and disease
-and injuries. Based on population and travel scenarios. The model has
-been used to calculate the health impacts of walking and bicycling
-short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013,maizlish2013}.}
+\col{ITHIM is a statistical model that integrates data on travel
+  patterns, physical activity, fine particulate matter, GHG emissions,
+  and disease and injuries. Based on population and travel
+  scenarios. The model has been used to calculate the health impacts
+  of walking and bicycling short distances usually traveled by car or
+  driving low-emission automobiles \cite{woodcock2013,maizlish2013}.}
 
 \col{The ITHIM model uses comparative risk assessment through which it
-formulates a change in the disease burden, resulting from the shift in
-the exposure distribution from a baseline scenario to an alternative
-scenario.}
+  formulates a change in the disease burden, resulting from the shift
+  in the exposure distribution from a baseline scenario to an
+  alternative scenario.}
 
 \col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
+  eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
+  ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
   aliquip ex ea commodo consequat. Duis aute irure dolor in
@@
 \col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
+  eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
+  ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
   aliquip ex ea commodo consequat. Duis aute irure dolor in
@@
 
-\col[blue]{\begin{itemize}
-  \item The coefficient of variation was fixed at ??? for all strata
-    and models.  Earlier ITHIM models estimated this as a function of
-    total active tranpsort time to model the decrease in variance of
-    active transport time that occurs for increasing means, i.e.,
-    fitter communities have less variance in active transport time.
-  \item The mean non-travel activity for the referent group was
-    modeled as a variable.  The relative means across age-sex strata
-    were fixed, and estimated using ATUS.
-  \item The national DALYs were stimated using the attributable
-    fraction provided by the comparitive risk assesment.  The
-    estimates for disease burned came from WHO and their GBD study.
-  \item Not all of the regions had the same sample size.  In fact a
-    few of the regions have relativel low sample size.  Maybe we
-    should put confidence intervals on our estimates.
-  \item The ITHIM model stratifies the population by sex and age using
-    eight age classes.  Within each stratum active transport time is
-    modeled as log-normal with mean $\hat{\mu}_{i,j}$ which is
-    estimated from the National Household Transportation Survey.
-  \item Quintiles, in this case the quantiles for percentages
-    $(0.1,0.3,0.5,0.7,0.9)$, are used to discretize the integral
-    defined in the Comparative Risk Assesment (CRA)
-  \item 
-\end{itemize}}
+The ITHIM model uses the comparative risk assesment framework to
+estimate the population attributable fraction, $\af$.  In our case
+$\af$ represents the proportional change in mortality and morbidity if
+the physical activity levels were adjusted to represent an alternate
+scenario.
 
-The ITHIM model uses the comparative risk assesment framework
+\begin{equation}
+\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
+\end{equation}
+
+Here $R(x)$ represents the risk when the individual has exposure $x$,
+in our case physical activity (\mets).  $P(x)$ and $Q(x)$ denote the
+population density of individuals with exposure $x$, in the baseline
+and alternate scenarios, respectively.  The exposure variable, $x$, is
+modeled as the sum of travel-related exposure and non-travel-related
+exposure.
+
+The time spent walking or cycling, i.e., active transport time, is
+modeled as a log-normal random variable with means defined for each
+age-sex stratum.  The non-travel-related physical activity is also
+modeled as log-normal with stratum-specific means.  Estimates for
+non-travel-related physical activity are diffiuclt.  We used the
+American Time Use Survey data to estimate the relative means for each
+of the age-sex strata.  We then treat the mean of the referent group
+as a variable and compute results over a grid of values for this
+variable.  See Figure \ref{dalyFigure}.
+
+The travel-related exposure is estimated using active transport time
+along with assumptions about how much physical activity is required
+for cycling and walking.  Active transport time and non-travel-related
+physical activity are modeled as independent random variables.  A
+fixed value was used as the coefficient of variation for the
+distributin of active transport time. A different, but also fixed,
+value was used for the non-travel-related physical activity
+distribution.
 
@@
 
-\col[red]{\begin{itemize}
-\item National Household Transportation Survey (NHTS)
-\item American time Use Survey
-\item CDC Wonder
-\end{itemize}}
-
 \col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing)) elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
-dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
-occaecat cupidatat non proident, sunt in culpa qui officia deserunt
-mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
-adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
-magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.}
-
-\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}
+    \centerline{\includegraphics[width=0.4\textwidth]{./figures/HHSmap}}
     \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
@@
 \end{figure}
-
-\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
-dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
-occaecat cupidatat non proident, sunt in culpa qui officia deserunt
-mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
-adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
-magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.}
 
@@
 \end{document}
+
+%   \begin{itemize}
+%   \item The coefficient of variation was fixed at ??? for all strata
+% and models.  Earlier ITHIM models estimated this as a function of
+% total active tranpsort time to model the decrease in variance of
+% active transport time that occurs for increasing means, i.e., fitter
+% communities have less variance in active transport time.
+%   \item The mean non-travel activity for the referent group was
+% modeled as a variable.  The relative means across age-sex strata were
+% fixed, and estimated using ATUS.
+%   \item The national DALYs were stimated using the attributable
+% fraction provided by the comparitive risk assesment.  The estimates
+% for disease burned came from WHO and their GBD study.
+%   \item Not all of the regions had the same sample size.  In fact a
+% few of the regions have relativel low sample size.  Maybe we should
+% put confidence intervals on our estimates.
+%   \item The ITHIM model stratifies the population by sex and age using
+% eight age classes.  Within each stratum active transport time is
+% modeled as log-normal with mean $\hat{\mu}_{i,j}$ which is estimated
+% from the National Household Transportation Survey.
+%   \item Quintiles, in this case the quantiles for percentages
+% $(0.1,0.3,0.5,0.7,0.9)$, are used to discretize the integral defined
+% in the Comparative Risk Assesment (CRA)
+%   \item
+% \end{itemize}
+%  ``the proportional reduction in population
+% disease or mortality that would occur if exposure to a risk factor
+% were reduced to an alternative ideal exposure scenario (eg. no tobacco
+% use).'' \url{http://www.who.int/healthinfo/global_burden_disease/metrics_paf/en/}

### 2016-09-16 R/ITHIM.Rmd
Updated top directory Makefile
Also made a figure smaller so it looks better in the manuscript.
Worked on figure and changed permissions
I changed the permissins only so it looks nicer when using ls
Working on figure for manuscript

@@
 
-```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10, echo = FALSE, warning = FALSE}
+```{r ITHIMObjs, eval = TRUE, fig.width = 8, fig.height = 5, echo = FALSE, warning = FALSE}
 compute <- TRUE
@@
 
-for( muNonTravel in seq(1,50,length.out = 20) ){
+for( muNonTravel in seq(1,10,length.out = 100) ){
 
@@
 barplot(with(NHTS,table(HHS)))
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY, colour = HHS))
-p  + geom_line(size = 2) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = HHS))
+p  + geom_line(size = 2) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean (MET-hrs./week)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, 10)) #+ facet_grid(.
+~ HHS)
 ```

### 2016-09-16 manuscript/manuscript.tex
Update to ITHIM manuscript #1

@@
 \newcommand{\mets}{MET-hrs./week}
+\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
+\newcommand{\vmt}{$\mathrm{VMT}$}
+
 \begin{document}
@@
 
+\subsection{Active Transport}
+
 The time spent walking or cycling, i.e., active transport time, is
@@
 
-\subsection{Data Sources}
+\subsection{Road Injuries}
 
-\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing)) elit, sed do
+\col[red]{\textit{
+We chose to model road injuries in a more simplifed manner than the
+authors of ITHIM.  The original ITHIM requires information about the
+transportational infrastructure that cannot be accurately modeled on a
+regional level, and is more well-suited to say a collection of
+counties as a case study. \cite{maizlish2013}  We use the overall percent
+of travel by walking and road fatalities per U.S. metropolitan area to
+estimate road fatalaties in U.S. metropolitan areas given the percent
+of travel by walking. \cite{UWTechReport}
+}}
+
+\col[red]{\textit{
+We need to estimate the increase/decrease in road injuries/fatalaties
+nation-wide as walking/cycling time increases.  We need a relative
+risk of road fatalaty as a function of active transport time...
+}}
+
+\col[red]{\textit{
+
+}}
+
+\subsection{Air Pollution}
+
+Can we model the decrease in \pEm{2.5} nationally as a function of
+decrease in driving time?  The original plan was to use a baseline
+value for \pEm{2.5} and a linear relationship bewteen change in
+\pEm{2.5} and change in \vmt{}.
+
+\end{methods}
+
+\section{U.S. Metro-Regional Analysis}
+\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 
-\end{methods}
+\subsection{Data Sources}
 
-\section{U.S. Metro-Regional Analysis}
-\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
+\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing)) elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
@@
 %\bibliographystyle{natbib}
-%\bibliographystyle{bmc_article}
-\bibliographystyle{mypapers}
+\bibliographystyle{bmc_article}
+%\bibliographystyle{mypapers}
 \bibliography{ITHIM}
 \end{document}
-
-%   \begin{itemize}
-%   \item The coefficient of variation was fixed at ??? for all strata
-% and models.  Earlier ITHIM models estimated this as a function of
-% total active tranpsort time to model the decrease in variance of
-% active transport time that occurs for increasing means, i.e., fitter
-% communities have less variance in active transport time.
-%   \item The mean non-travel activity for the referent group was
-% modeled as a variable.  The relative means across age-sex strata were
-% fixed, and estimated using ATUS.
-%   \item The national DALYs were stimated using the attributable
-% fraction provided by the comparitive risk assesment.  The estimates
-% for disease burned came from WHO and their GBD study.
-%   \item Not all of the regions had the same sample size.  In fact a
-% few of the regions have relativel low sample size.  Maybe we should
-% put confidence intervals on our estimates.
-%   \item The ITHIM model stratifies the population by sex and age using
-% eight age classes.  Within each stratum active transport time is
-% modeled as log-normal with mean $\hat{\mu}_{i,j}$ which is estimated
-% from the National Household Transportation Survey.
-%   \item Quintiles, in this case the quantiles for percentages
-% $(0.1,0.3,0.5,0.7,0.9)$, are used to discretize the integral defined
-% in the Comparative Risk Assesment (CRA)
-%   \item
-% \end{itemize}
-%  ``the proportional reduction in population
-% disease or mortality that would occur if exposure to a risk factor
-% were reduced to an alternative ideal exposure scenario (eg. no tobacco
-% use).'' \url{http://www.who.int/healthinfo/global_burden_disease/metrics_paf/en/}

### 2016-09-16 manuscript/manuscript.tex
Updates to ITHIM manuscript 1

@@
   averages as the baseline.  We investigated the reduced chronic
-  disease burden of (x,y,z outcomes) from increased physical activity
-  from population shifts in the amount of biking and walking doen for
-  transport.  Our methods employ the Integrated Transport and Health
-  Impacts Model (ITHIM) informed by regional travel parameters using
-  the National Household Transportation Survey, non-travel activity
-  metircs from the American Time Use Survey, and baseline Disease
-  burden from the Centers for Disease Control and Prevention's WONDER
-  database.  Our findings show that the West coast and the Northeast
-  of the U.S. have travel patterns that, if mirrored by the nation,
-  would save the U.S. thousands to hundreds of thousands DALYs.}
+  disease burden of ($x$, $y$, $z$ outcomes) from increased physical
+  activity from population shifts in the amount of biking and walking
+  doen for transport.  Our methods employ the Integrated Transport and
+  Health Impacts Model (ITHIM) informed by regional travel parameters
+  using the National Household Transportation Survey, non-travel
+  activity metircs from the American Time Use Survey, and baseline
+  Disease burden from the Centers for Disease Control and Prevention's
+  WONDER database.  Our findings show that the West coast and the
+  Northeast of the U.S. have travel patterns that, if mirrored by the
+  nation, would save the U.S. thousands to hundreds of thousands
+  DALYs.}
 
@@
 
-The ITHIM model uses the comparative risk assesment framework to
-estimate the population attributable fraction, $\af$.  In our case
-$\af$ represents the proportional change in mortality and morbidity if
-the physical activity levels were adjusted to represent an alternate
-scenario.
+\col[blue]{\textit{
+    The ITHIM model uses the comparative risk
+    assesment framework to estimate the population attributable
+    fraction, $\af$.  In our case $\af$ represents the proportional
+    change in mortality and morbidity if the physical activity levels
+    were adjusted to represent an alternate scenario.
+}}
 
@@
 
+\col[blue]{\textit{
     Here $R(x)$ represents the risk when the individual has exposure $x$,
@@
 exposure.
+}}
 
@@
 
+\col[blue]{\textit{
     The time spent walking or cycling, i.e., active transport time, is
@@
 variable.  See Figure \ref{dalyFigure}.
+}}
 
+\col[blue]{\textit{
 The travel-related exposure is estimated using active transport time
@@
 distribution.
+}}
 
@@
 
-\col[red]{\textit{
+\col[blue]{\textit{
 We chose to model road injuries in a more simplifed manner than the
@@
 
-\col[red]{\textit{
+\col[blue]{\textit{
 We need to estimate the increase/decrease in road injuries/fatalaties
@@
 
-\col[red]{\textit{
-
-}}
-
 \subsection{Air Pollution}
 
+\col[blue]{\textit{
     Can we model the decrease in \pEm{2.5} nationally as a function of
@@
 \pEm{2.5} and change in \vmt{}.
+}}
 

### 2016-09-19 manuscript/manuscript.tex
Updates to ITHIM manuscript #1
I included the figure of mean travel time by mode, region and age-sex stratum.  I also added citations for the data and worked on the Regional Analysis section

@@
 
-\title[\package{}]{\package{}: an \R{} package for implemetation of
-  the ITHIM model}
+\title[\package{}]{A comparative risk assesment of U.S. regional
+  active transport using the ITHIM model}
 \author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
@@
 
-\col[blue]{\textit{
-    The time spent walking or cycling, i.e., active transport time, is
-modeled as a log-normal random variable with means defined for each
-age-sex stratum.  The non-travel-related physical activity is also
-modeled as log-normal with stratum-specific means.  Estimates for
-non-travel-related physical activity are diffiuclt.  We used the
-American Time Use Survey data to estimate the relative means for each
-of the age-sex strata.  We then treat the mean of the referent group
-as a variable and compute results over a grid of values for this
-variable.  See Figure \ref{dalyFigure}.
-}}
+\col[blue]{\textit{The time spent walking or cycling, i.e., active
+    transport time, is modeled as a log-normal random variable with
+    means defined for each age-sex stratum.  The non-travel-related
+    physical activity is also modeled as log-normal with
+    stratum-specific means.}}
 
@@
 
-\col[blue]{\textit{
+\col[red]{\textit{
 We chose to model road injuries in a more simplifed manner than the
@@
 
-\col[blue]{\textit{
+\col[red]{\textit{
 We need to estimate the increase/decrease in road injuries/fatalaties
@@
 
-\col[blue]{\textit{
+\col[red]{\textit{
     Can we model the decrease in \pEm{2.5} nationally as a function of
@@
 \section{U.S. Metro-Regional Analysis}
-\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
-dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
-occaecat cupidatat non proident, sunt in culpa qui officia deserunt
-mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
-adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
-magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.}
 
-\subsection{Data Sources}
+\begin{figure}[t]
+    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig3}}
+    \caption{Foo}\label{meanMatrices}
+\end{figure}
 
-\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing)) elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
-minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
-dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
-commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
-velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
-occaecat cupidatat non proident, sunt in culpa qui officia deserunt
-mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
-adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
-magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
-ullamco laboris nisi ut aliquip ex ea commodo consequat.}
+\col[blue]{\begin{itemize}
+\item We used the National Household Transportation Survey \cite{NHTS} to
+  estimate the mean walk and cycle times for individuals that live in
+  metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
+  corresponds to individuals whose homes are in an MSA or cMSA)
+  \bi\item Data quality issues along with assumptions about travel speed
+created mirky data.  The median values of trip time were used to avoid
+undue influence by outliers.\ei
+\item We then divided these into the 10 regions of the U.S. defined by
+  the Department of Health and Human Services (HH1-10)
+\item Using only the active transport component of the ITHIM model we
+  ran comparative analyses for each of the ten regions, using the
+  national averages as the baseline.
+\item We used data from the American Time Use Survey \cite{ATUS} to
+  estimate the relative means for each of the age-sex strata.  We then
+  treat the mean of the referent group as a variable and compute
+  results over a grid of values for this variable.  See Figure
+  \ref{dalyFigure}.
+\item Values for DALYs were found using the CDC Wonder database \cite{CDCWonder}
+\end{itemize}}
 
@@
 \section*{Acknowledgment}
-\paragraph{Funding\textcolon} SGY is supported by$\ldots$ , JV, JP, $\ldots$
+\paragraph{Funding\textcolon} \col[red]{SGY is supported by$\ldots$ , JV, JP, $\ldots$}
 

### 2016-09-20 R/ITHIM.Rmd
Worked on Figures for Manuscript #1
Also updated the Make file in the top-level directory.  make all runs the whole thing now with dependencies in place.

--- 2016-09-08 R/ITHIM.Rmd
@@
 
-```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10, echo = FALSE, warning = FALSE}
-compute <- TRUE
-
-if(compute){
+```{r ITHIMObjs, eval = TRUE, fig.width = 8, fig.height = 5, echo = FALSE, warning = FALSE}
+upper <- 8
+length <- 100
 
 DALYsTotal <- data.frame()
-
-for( muNonTravel in seq(1,50,length.out = 20) ){
-
 ITHIMList <- list()
-  rm(DALYs)
 ITHIM.default <- createITHIM(vision = "baseline", region = "national")
@@
     ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
-    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muNonTravel", muNonTravel)
   }
@@
 ITHIM.national <- createITHIM(vision = "baseline", region = "national")
-  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
 ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
@@
 ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+
+for( muNonTravel in seq(0.01,upper,length.out = length) ){
+
+  rm(DALYs)
+
+  for (HHS in c(paste0("HHS0",1:9),"HHS10")){
+    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muNonTravel", muNonTravel)
+  }
+
+  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+
   DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
@@
 
-}else{
-  DALYsTotal <- readRDS(file = "./data/DALYsTotal.rds")
-}
-
 barplot(with(NHTS,table(HHS)))
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY, colour = HHS))
-p  + geom_line(size = 2) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
+
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = HHS))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean (MET-hrs./week)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean (MET-hrs./week)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
+
 ```
@@
 
-```{r meanMatrices, eval = TRUE, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE, echo = FALSE}
+```{r meanMatrices, eval = TRUE, fig.width = 10, fig.height = 5, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
-
 names(D) <- c("ageClass", "mode", "sex", "time", "HHS")
 
-#D <- within(D, HHS <- factor(HHS, levels = c(paste0("HHS",1:10),"Natl")))
-
-p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") + facet_grid( mode ~ HHS, scales = "free_y")
-
-p1
+p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") 
+p1 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Minutes per Week") + facet_grid( mode ~ HHS, scales = "free_y")
 

### 2016-09-20 R/create_HHS-metro_Burdens.r

@@ -1,210 +1,0 @@
-library("plyr")
-#setwd("~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/")
-
-#######DETAILS ON WONDER QUERY#######
-# ---
-#   Dataset: Underlying Cause of Death, 1999-2014
-# Query Parameters:
-#   Title: ITHIM2014_HHS_urb_age_sex_16years
-# 2013 Urbanization: All
-# Autopsy: All
-# Five-Year Age Groups: All
-# Gender: Female, Male
-# HHS Regions: All
-# Hispanic Origin: All
-# ICD-10 113 Cause List: Malignant neoplasms of colon, rectum and anus (C18-C21), Malignant neoplasms of trachea, bronchus and
-# lung (C33-C34), Malignant neoplasm of breast (C50), #Diabetes mellitus (E10-E14), #Alzheimer's disease (G30), Other forms of
-# chronic ischemic heart disease (I20,I25), Acute and subacute endocarditis (I33), Diseases of pericardium and acute myocarditis
-# (I30-I31,I40), #Essential hypertension and hypertensive renal disease (I10,I12,I15), #Cerebrovascular diseases (I60-I69),
-# #Influenza and pneumonia (J09-J18), Other acute lower respiratory infections (J20-J22,U04), #Chronic lower respiratory diseases
-# (J40-J47), Other diseases of respiratory system (J00-J06,J30- J39,J67,J70-J98), Motor vehicle accidents
-# (V02-V04,V09.0,V09.2,V12-V14,V19.0-V19.2,V19.4-V19.6,V20-V79,V80.3-V80.5,V81.0-V81.1,V82.0-V82.1,V83-V86,V87.0-V87.8,V88.0-V88.8
-#  V89.0,V89.2)
-# Other land transport accidents
-# (V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
-#  V89.1,V89.3,V89.9)
-#
-# Place of Death: All
-# Race: All
-# Weekday: All
-# Year/Month: All
-# Group By: Five-Year Age Groups, Gender, HHS Region, 2013 Urbanization, ICD-10 113 Cause List
-# Show Totals: Disabled
-# Show Zero Values: True
-# Show Suppressed: True
-# Calculate Rates Per: 100,000
[…]

### 2016-09-20 R/ITHIM.Rmd
Updates to ITHIM.Rmd
Still having problems with AgeClass 1 and HHS9 cycling referent group mean.

@@
 foo <- matrix(unlist(data), nrow = 7, ncol = 2)
-foo <- rbind( c(0.1, 0.1), foo)
+foo <- rbind( c(0, 0), foo)
 dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
@@
 })
-saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")
-saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
-```
-
-```{r loadData, eval = FALSE, echo = FALSE}
-NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
-NHTS.trips <- readRDS(file = "./data/NHTS.trips.rds")
+
 ```
@@
 
-#### Compute Natioal Metro Average
+#### Compute National Metro Average
 
@@
 cycleMatrix <- lapply(cycleLength, reformatData)
-meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
-meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
-RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 1e-4, x) )
-RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 1e-4, x) )
-RwtList <- lapply( RwtList, function(x) x/x[3,2] )
-RctList <- lapply( RctList, function(x) x/x[3,2] )
-muwtList <- lapply( meanWalkList, function(x) x[3,2] )
-muctList <- lapply( meanCycleList, function(x) x[3,2] )
+meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
+meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
+
+#RwtList <- lapply( , function(x) ifelse(x == 0, 0, x) )
+#RctList <- lapply( , function(x) ifelse(x == 0, 0, x) )
+
+zeroReferentCase <- function(x){
+  if(x[3,2]==0){
+    return(matrix(1e-3,nrow=nrow(x),ncol=ncol(x), dimnames = dimnames(x)))
+  }else{
+    return(ifelse(x==0, 1e-3, x/x[3,2]))    
+  }
+}
+
+RwtList <- lapply( meanWalkList, zeroReferentCase)
+RctList <- lapply( meanCycleList, zeroReferentCase)
+muwtList <- lapply( meanWalkList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
+muctList <- lapply( meanCycleList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
 ```
@@
 ```{r ITHIMObjs, eval = TRUE, fig.width = 8, fig.height = 5, echo = FALSE, warning = FALSE}
-upper <- 8
+upper <- 10
 length <- 100
@@
 ITHIM.default <- createITHIM(vision = "baseline", region = "national")
+ITHIM.default <- updateITHIM(ITHIMList[[HHS]], "cvNonTravel", 3)
+ITHIM.default <- updateITHIM(ITHIMList[[HHS]], "cv", 3.0288)
+ITHIM.default <- updateITHIM(ITHIMList[[HHS]], "pWalk", 0.85)
 
@@
 
-ITHIM.national <- createITHIM(vision = "baseline", region = "national")
+ITHIM.national <- ITHIM.default
 ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)

### 2016-09-20 manuscript/manuscript.tex
Worked on Figures for Manuscript #1
Also updated the Make file in the top-level directory.  make all runs the whole thing now with dependencies in place.
Included @vargovargo edits in Abstract Summary

@@
 \col[blue]{While increased active transportation, primarily cycling
-  and walk, is often held up as a goal for improved public health —
+  and walking, is often held up as a goal for improved public health —
   via reduced chronic disease, improved air quality, and reductions in
   greenhouse gas emissions --- obtaining estimates for the quantiative
-  impacts of such shifts remains infrequent, in not altogether
-  absent. Using U.S. metropolitan counties from the 10 regions defined
-  by the Department of Health and Human Services (HHS1-10) we perform
+  impacts of such shifts remains infrequent, if not altogether
+  absent from related policy action. Using U.S. metropolitan counties from the 10 regions defined
+  by the U.S. Department of Health and Human Services (HHS1-10) we perform
   a comparative risk analysis between each regions and national

### 2016-09-21 R/ITHIM.Rmd
Expanded grid for Regional Analysis
Tweaked the figures and removed AgeClass1 from DALY count and figure 1

@@
 library("ITHIM")
+
+fig.width <- 6 # inches
+phi <- (1+sqrt(5))/2 # the golden ratio ;)
 ```
@@
 
-#RwtList <- lapply( , function(x) ifelse(x == 0, 0, x) )
-#RctList <- lapply( , function(x) ifelse(x == 0, 0, x) )
-
 zeroReferentCase <- function(x){
@@
 sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
-    return(sum(unlist(compareModels(ITHIM.baseline, ITHIM.scenario)$daly.delta)))
-    }
-```
-
-```{r ITHIMObjs, eval = TRUE, fig.width = 8, fig.height = 5, echo = FALSE, warning = FALSE}
-upper <- 10
+    return(sum(unlist(data.frame(compareModels(ITHIM.baseline,ITHIM.scenario)$daly.delta)[-1,]))) # AgeClass 1 is removed from totals
+    }
+```
+
+```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
+upper <- 20
 length <- 100
@@
   DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
-  DALYs <- data.frame(HHS = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
+  DALYs <- data.frame(Region = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
   DALYsTotal <- rbind(DALYsTotal, DALYs)
@@
 
-barplot(with(NHTS,table(HHS)))
-
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = HHS))
+barplot(with(NHTS,table(Region)))
+
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = Region))
 
@@
 
-```{r meanMatrices, eval = TRUE, fig.width = 10, fig.height = 5, warning = FALSE, message = FALSE, echo = FALSE}
-
-ITHIMList <- c(ITHIMList, list("Natl" = ITHIM.national))
-
-meanMatrix <- function(ITHIM, mode = "walk"){
-   if(mode == "walk"){
+```{r meanMatrices, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+ITHIMList <- c(ITHIMList, list("National" = ITHIM.national))
+
+meanMatrix <- function(ITHIM, mode = "Walking"){
+   if(mode == "Walking"){
       return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
+   }else if (mode == "Cycling"){
+      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
    }else{
-      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
-}
-}
-
-meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "walk")
-meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "cycle")
+      message("error in meanMatrix function")
+   }
+}
+
+meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "Walking")
+meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "Cycling")
 
 D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
-names(D) <- c("ageClass", "mode", "sex", "time", "HHS")
-
-p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") 
-p1 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Minutes per Week") + facet_grid( mode ~ HHS, scales = "free_y")
+names(D) <- c("ageClass", "mode", "Sex", "time", "Region")
+
+D1 <- subset(D, Region %in% c("HHS01","HHS02","HHS09","HHS10","National") & ageClass != "1" )
+D2 <- subset(D, !(Region %in% c("HHS01","HHS02","HHS09","HHS10")) & ageClass != "1" )
+
+p1 <- ggplot(D1, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
+p1 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")
+
+p2 <- ggplot(D2, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
+p2 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")
 

### 2016-09-21 manuscript/manuscript.tex
Tweaked the figures and removed AgeClass1 from DALY count and figure 1

@@
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig3}}
-    \caption{Foo}\label{meanMatrices}
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3}}
+    \caption{Mean travel time by mode (walking/cycling) for each of
+      the four regions with a decrease in DALYs
+      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
+      national means are displayed on the far right.  These four
+      regions correspond to New England/NewYork and the West
+      Coast.}\label{meanMatrices}
 \end{figure}
@@
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/HHSmap}}
-    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
+    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
+    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
     \caption{Change in total national DALYs for each regional

### 2016-09-21 R/ITHIM.Rmd
Minor updates to manuscript and figures

@@
 
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean (MET-hrs./week)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Non-Travel METs mean (MET-hrs./week)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
 

### 2016-09-22 R/ITHIM.Rmd
Fixed bug in ITHIM.default update

@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
-upper <- 20
-length <- 100
+upper <- 50
+length <- 500
 
@@
 ITHIM.default <- createITHIM(vision = "baseline", region = "national")
-ITHIM.default <- updateITHIM(ITHIMList[[HHS]], "cvNonTravel", 3)
-ITHIM.default <- updateITHIM(ITHIMList[[HHS]], "cv", 3.0288)
-ITHIM.default <- updateITHIM(ITHIMList[[HHS]], "pWalk", 0.85)
+ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
+ITHIM.default <- updateITHIM(ITHIM.default, "cv", 3.0288)
+ITHIM.default <- updateITHIM(ITHIM.default, "pWalk", 0.85)
+ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
 

### 2016-09-27 R/ITHIM.Rmd
Changed upper and length in ITHIMObjs chunk
Small bit of cleanup to ITHIM.Rmd
Switched to stratification by State

@@
                    URBRUR  = as.factor(URBRUR),
-                   HHSTATE = as.factor(HHSTATE),
+                   #HHSTATE = as.factor(HHSTATE),
                    MSACAT = as.factor(MSACAT),
@@
 ```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
-#summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
-#summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
-#summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
-#summary(with(subset(NHTS.trips, AT == "walk"), TRVL_MIN))
-#summary(with(subset(NHTS.trips, AT == "cycle"), TRVL_MIN))
-#summary(with(subset(NHTS.trips, AT == "other"), TRVL_MIN))
-
 walkHours <- with(subset(NHTS.trips, {
@@
 ```{r subset, eval = TRUE, fig.width = 10, echo = FALSE}
-NHTS.person <- subset( NHTS.person, MSA == TRUE )
+NHTS.person <- subset( NHTS.person, MSA == TRUE & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
@@
 ```{r count, eval = TRUE, echo = FALSE}
-listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
+listOfLists <- lapply(split(NHTS, NHTS$HHSTATE),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
 counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
@@
 
-```{r CDC, eval = TRUE, echo = FALSE}
+```{r CDC, eval = FALSE, echo = FALSE}
 wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
@@
 
-```{r ITHIMCause, eval = TRUE, echo = FALSE}
+```{r ITHIMCause, eval = FALSE, echo = FALSE}
 ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
@@
 ```
-```{r recodeAge, eval = TRUE, echo = FALSE}
+```{r recodeAge, eval = FALSE, echo = FALSE}
 ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
@@
 
-```{r wonder, eval = TRUE, echo = FALSE}
+```{r wonder, eval = FALSE, echo = FALSE}
 wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
@@
 
-```{r suppressed, eval = TRUE, echo = FALSE, warning = FALSE}
+```{r suppressed, eval = FALSE, echo = FALSE, warning = FALSE}
 wonder <- within( wonder, {
@@
 
-```{r summary2, eval = TRUE, echo = FALSE}
+```{r summary2, eval = FALSE, echo = FALSE}
 wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
@@
 
-```{r singleyear, eval = TRUE, echo = FALSE}
+```{r singleyear, eval = FALSE, echo = FALSE}
 pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
@@
 
-```{r sloppy, eval = TRUE, echo = FALSE}
+```{r sloppy, eval = FALSE, echo = FALSE}
 ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
@@
 
-```{r foop, eval = TRUE, echo = FALSE}
+```{r foop, eval = FALSE, echo = FALSE}
 pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
@@
 
-```{r USGBD, eval = TRUE, echo = FALSE}
+```{r USGBD, eval = FALSE, echo = FALSE}
 USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
@@
 
-```{r burdenList, eval = TRUE, echo = FALSE}
+```{r burdenList, eval = FALSE, echo = FALSE}
 GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
-upper <- 50
-length <- 500
+upper <- 100
+length <- 100
 
@@
 
-for (HHS in c(paste0("HHS0",1:9),"HHS10")){
-    ITHIMList[[HHS]] <- ITHIM.default
-    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muwt", muwtList[[HHS]])
-    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muct", muctList[[HHS]])
-    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rwt", RwtList[[HHS]])
-    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rct", RctList[[HHS]])
-    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
-  }
-
 ITHIM.national <- ITHIM.default
@@
 
-for( muNonTravel in seq(0.01,upper,length.out = length) ){
+for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+    ITHIMList[[HHSTATE]] <- ITHIM.default
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
+  }
+
+for( muNonTravel in seq(0,upper,length.out = length) ){
 
@@
 
-  for (HHS in c(paste0("HHS0",1:9),"HHS10")){
-    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muNonTravel", muNonTravel)
+  for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
   }
@@
 
-barplot(with(NHTS,table(Region)))
-
 p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = Region))
@@
 
-ITHIMList <- c(ITHIMList, list("National" = ITHIM.national))
+ITHIMList <- c(ITHIMList, list("Natl." = ITHIM.national))
 
@@
 D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
-names(D) <- c("ageClass", "mode", "Sex", "time", "Region")
-
-D1 <- subset(D, Region %in% c("HHS01","HHS02","HHS09","HHS10","National") & ageClass != "1" )
-D2 <- subset(D, !(Region %in% c("HHS01","HHS02","HHS09","HHS10")) & ageClass != "1" )
-
-p1 <- ggplot(D1, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
-p1 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")
-
-p2 <- ggplot(D2, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
-p2 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")
+names(D) <- c("ageClass", "mode", "Sex", "time", "State")
+
+## D1 <- subset(D, Region %in% c("HHS01","HHS02","HHS09","HHS10","National") & ageClass != "1" )
+## D2 <- subset(D, !(Region %in% c("HHS01","HHS02","HHS09","HHS10")) & ageClass != "1" )
+
+p <- ggplot(D, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
+p + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ State, scales = "free_y")
+
+## p2 <- ggplot(D2, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
+## p2 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")
 

### 2016-09-27 R/ITHIM.Rmd
Reordered the State factor so figures looked better
This is sort of a magic number.  I'm not sure how R handles levels to factors when they are missing or whatever.

@@
   DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
-  DALYs <- data.frame(Region = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
+  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
   DALYsTotal <- rbind(DALYsTotal, DALYs)
@@
 
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = Region))
+DALYsTotal <- within(DALYsTotal, {
+  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA"))
+})
+
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
 
@@
 names(D) <- c("ageClass", "mode", "Sex", "time", "State")
+
+D <- within(D,{
+  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
+})
 

### 2016-09-27 R/ITHIM.Rmd
Created figure for California comparison

@@
 
-```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
+```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
 newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
@@
 
-```{r readnewMETs, echo = TRUE, eval = TRUE, echo = FALSE}
+```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
 cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
@@
 
-```{r metKey, echo = TRUE, eval = TRUE, echo = FALSE}
+```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
 allActivities <- colnames(summaryList$Data)
@@
 
-```{r parameters, eval = TRUE, echo = FALSE}
+```{r parameters, eval = FALSE, echo = FALSE}
 parameters <- ddply(D, .(SEX, AGECAT), summarise,
@@
 upper <- 100
-length <- 100
+length <- 10
 
@@
   State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
+  Sex <- factor(Sex, levels = c("F","M"))
 })
@@
 ```
+
+```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+ITHIM.CA <- ITHIMList[["CA"]]
+ITHIM.CA <- updateITHIM(ITHIM.CA, parName = "muNonTravel", 10)
+ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)
+foo <- compareModels(ITHIM.national,ITHIM.CA)
+foobar <- data.frame(cbind(melt(foo$daly.delta),paste0("ageClass",1:8)))
+names(foobar) <- c("delta.DALY","Sex","Disease","ageClass")
+p <- ggplot( foobar, aes(ageClass,-delta.DALY/1e3, fill = Sex) )
+p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( .
+~ Disease) + theme_bw()
+```
+
 # Appendix

### 2016-09-27 manuscript/manuscript.tex
Switched to pdf format for figures
Minor updates to manuscript and figures
Switched to stratification by State
Created figure for California comparison

@@
 
+%% \begin{figure}[t]
+%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
+%%     \caption{Map of HHS Regions and metropolitan areas}\label{HHSmap}
+%% \end{figure}
+
 \section{U.S. Metro-Regional Analysis}
@@
 \begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3}}
-    \caption{Mean travel time by mode (walking/cycling) for each of
-      the four regions with a decrease in DALYs
-      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
-      national means are displayed on the far right.  These four
-      regions correspond to New England/NewYork and the West
-      Coast.}\label{meanMatrices}
+    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
+    \caption{Change in total national DALYs for each regional
+      transportation estimates.}\label{dalyFigure}
 \end{figure}
@@
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{Change in total national DALYs for each regional
-      transportation estimates.}\label{dalyFigure}
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
+    \caption{Mean travel time by mode (walking/cycling) for each of
+      the four regions with a decrease in DALYs
+      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
+      national means are displayed on the far right.  These four
+      regions correspond to New England/NewYork and the West
+      Coast. The age classes are defined as in the original ITHIM
+      implementation using 8 eight age classes defined in terms of
+      years as follows: 1. 0--4, 2. 5--14, 3. 15--29, 4. 30--44,
+      5. 45--59, 6. 60--69, 7. 70--79, and 8. 80+. Age class 1 is
+      omitted from all calclulations in this paper due to lack of
+      data.}\label{meanMatrices}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
+    \caption{California}\label{California}
 \end{figure}

### 2016-09-28 R/ITHIM.Rmd
Worked on state-wise figures

@@
 
+```{r compute, echo = FALSE, eval = TRUE}
+compute <- FALSE
+```
+
 #### Create a handful of functions to help parse data
@@
 
-```{r NHTS, eval = TRUE, echo = FALSE}
+```{r NHTS, eval = compute, echo = FALSE}
 DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
@@
                    URBRUR  = as.factor(URBRUR),
-                   #HHSTATE = as.factor(HHSTATE),
                    MSACAT = as.factor(MSACAT),
@@
 
-#### Investigate Trip Data
-
-```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
+```{r summary, fig.width = 20, fig.height = 10, eval = compute, echo = FALSE, message = FALSE, warning = FALSE}
 walkHours <- with(subset(NHTS.trips, {
@@
 
-How do we merge?  Are there people who were surveyed who had zero
-travel time?  Couldnt have been very many.  But I think there are
-people in the PERSON table that were not interviewed.
-
-```{r subset, eval = TRUE, fig.width = 10, echo = FALSE}
+The PERSON table contains the family members of the person surveyed,
+right?  We want only the peole who were surveyed for their travel
+times.  So, we can\'t use everyone in the PERSON table.  If someone
+had zero trip time they would not exist in the TRIP table.  So, if we
+use just the TRIP table we miss people with zero trip time. So how do
+we merge?  We use an inner join and assume that their is a negligible
+number of people with zero trip time.  Not very many people never stay
+in the house all day.  Please correct me if I am mistaken abiout the
+nature of the PERSON or TRIP tables.
+
+```{r subset, eval = compute, fig.width = 10, echo = FALSE}
 NHTS.person <- subset( NHTS.person, MSA == TRUE & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
+saveRDS(NHTS, file = "./data/NHTS.rds")
+```
+
+```{r loadNHTS, eval = !compute}
+NHTS <- readRDS(file = "./data/NHTS.rds")
 ```
@@
 
+## Create List of ITHIM Objects
+
+```{r foobar, eval = TRUE, echo = FALSE}
+sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
+    return(sum(unlist(data.frame(compareModels(ITHIM.baseline,ITHIM.scenario)$daly.delta)[-1,]))) # AgeClass 1 is removed from totals
+    }
+```
+
+```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+upper <- 25; wd <- 1; length <- upper/wd + 1;
+
+DALYsTotal <- data.frame()
+ITHIMList <- list()
+ITHIM.default <- createITHIM(vision = "baseline", region = "national")
+ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
+ITHIM.default <- updateITHIM(ITHIM.default, "cv", 3.0288)
+ITHIM.default <- updateITHIM(ITHIM.default, "pWalk", 0.85)
+ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
+
+ITHIM.national <- ITHIM.default
+ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+
+for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+    ITHIMList[[HHSTATE]] <- ITHIM.default
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
+  }
+
+for( muNonTravel in seq(0,upper,length.out = length) ){
+
+  rm(DALYs)
+
+  for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
+  }
+
+  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+
+  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
+  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
+  DALYsTotal <- rbind(DALYsTotal, DALYs)
+}
+
+DALYsTotal <- within(DALYsTotal, {
+  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA"))
+})
+
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
+
+```
+
+### Visualization of trip time matrices
+
+```{r meanMatrices, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+ITHIMList <- c(ITHIMList, list("Natl." = ITHIM.national))
+
+meanMatrix <- function(ITHIM, mode = "Walking"){
+   if(mode == "Walking"){
+      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
+   }else if (mode == "Cycling"){
+      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
+   }else{
+      message("error in meanMatrix function")
+   }
+}
+
+meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "Walking")
+meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "Cycling")
+
+D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
+names(D) <- c("ageClass", "mode", "Sex", "time", "State")
+
+D <- within(D,{
+  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
+  Sex <- factor(Sex, levels = c("F","M"))
+})
+
+p <- ggplot(subset(D,ageClass != 1), aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
+p + theme_bw() + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ State, scales = "free_y") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 5))
+
+```
+
+```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)
+
+ITHIM.CA <- ITHIMList[["CA"]]
+ITHIM.CA <- updateITHIM(ITHIM.CA, parName = "muNonTravel", 10)
+foo1 <- compareModels(ITHIM.national,ITHIM.CA)
+foobar1 <- data.frame(cbind(melt(foo1$daly.delta),paste0("ageClass",1:8)),State = "CA")
+
+ITHIM.NY <- ITHIMList[["NY"]]
+ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
+foo2 <- compareModels(ITHIM.national,ITHIM.NY)
+foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
+
+foobar <- rbind(foobar1,foobar2)
+names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
+
+foobar <- within(foobar, {
+  Disease[Disease == "BreastCancer"] <- "Breast"
+  Disease[Disease == "ColonCancer"] <- "Colon"
+  Disease <- factor(Disease, levels = c("Breast", "Colon", "Dementia", "Depression", "Diabetes"))
+})
+
+p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = Sex) )
+p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( State ~ Disease) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
+```
+
+```{r counts, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+D <- melt(counts)
+names(D) <- c("count","AgeClass","Sex", "State")
+D <- within(D,{
+  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
+  Sex <- factor(Sex, levels = c("F","M"))
+})
+p <- ggplot(subset(D,AgeClass != "00-04"), aes(AgeClass, count/1e3, fill = Sex))
+p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Individuals Sampled (Thousands)") + facet_grid( .
+~ State) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6))
+```
+
+# Appendix
+
 ### CDC Mortality
@@
 
-## Create List of ITHIM Objects
-
-```{r foobar, eval = TRUE, echo = FALSE}
-sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
-    return(sum(unlist(data.frame(compareModels(ITHIM.baseline,ITHIM.scenario)$daly.delta)[-1,]))) # AgeClass 1 is removed from totals
-    }
-```
-
-```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
-upper <- 100
-length <- 10
-
-DALYsTotal <- data.frame()
-ITHIMList <- list()
-ITHIM.default <- createITHIM(vision = "baseline", region = "national")
-ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
-ITHIM.default <- updateITHIM(ITHIM.default, "cv", 3.0288)
-ITHIM.default <- updateITHIM(ITHIM.default, "pWalk", 0.85)
-ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
-
-ITHIM.national <- ITHIM.default
-ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
-
-for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
-    ITHIMList[[HHSTATE]] <- ITHIM.default
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
-  }
-
-for( muNonTravel in seq(0,upper,length.out = length) ){
-
-  rm(DALYs)
-
-  for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
-  }
-
-  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
-
-  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
-  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
-  DALYsTotal <- rbind(DALYsTotal, DALYs)
-}
-
-DALYsTotal <- within(DALYsTotal, {
-  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA"))
-})
-
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
-
-```
-
-### Visualization of trip time matrices
-
-```{r meanMatrices, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-
-ITHIMList <- c(ITHIMList, list("Natl." = ITHIM.national))
-
-meanMatrix <- function(ITHIM, mode = "Walking"){
-   if(mode == "Walking"){
-      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
-   }else if (mode == "Cycling"){
-      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
-   }else{
-      message("error in meanMatrix function")
-   }
-}
-
-meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "Walking")
-meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "Cycling")
-
-D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
-names(D) <- c("ageClass", "mode", "Sex", "time", "State")
-
-D <- within(D,{
-  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
-  Sex <- factor(Sex, levels = c("F","M"))
-})
-
-## D1 <- subset(D, Region %in% c("HHS01","HHS02","HHS09","HHS10","National") & ageClass != "1" )
-## D2 <- subset(D, !(Region %in% c("HHS01","HHS02","HHS09","HHS10")) & ageClass != "1" )
-
-p <- ggplot(D, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
-p + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ State, scales = "free_y")
-
-## p2 <- ggplot(D2, aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge") 
-## p2 + theme_bw() + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ Region, scales = "free_y")
-
-```
-
-```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-ITHIM.CA <- ITHIMList[["CA"]]
-ITHIM.CA <- updateITHIM(ITHIM.CA, parName = "muNonTravel", 10)
-ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)
-foo <- compareModels(ITHIM.national,ITHIM.CA)
-foobar <- data.frame(cbind(melt(foo$daly.delta),paste0("ageClass",1:8)))
-names(foobar) <- c("delta.DALY","Sex","Disease","ageClass")
-p <- ggplot( foobar, aes(ageClass,-delta.DALY/1e3, fill = Sex) )
-p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="Age Class", breaks = c(2,4,6,8)) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( .
-~ Disease) + theme_bw()
-```
-
-# Appendix
-
 ## Example (Bay Area)
@@
 
-```{r instantiate, eval = TRUE, echo = FALSE}
+```{r instantiate, eval = FALSE, echo = FALSE}
 ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
@@
 
-```{r RRs, eval = TRUE, echo = FALSE}
+```{r RRs, eval = FALSE, echo = FALSE}
 comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
@@
 ### Deaths
-```{r show, echo = FALSE, eval = TRUE, echo = FALSE}
+```{r show, echo = FALSE, eval = FALSE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### DALY
-```{r show2, echo = FALSE, eval = TRUE, echo = FALSE}
+```{r show2, echo = FALSE, eval = FALSE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLD
-```{r show3, echo = FALSE, eval = TRUE, echo = FALSE}
+```{r show3, echo = FALSE, eval = FALSE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLL
-```{r show4, echo = FALSE, eval = TRUE, echo = FALSE}
+```{r show4, echo = FALSE, eval = FALSE, echo = FALSE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r instantiate2, eval = TRUE, echo = FALSE}
+```{r instantiate2, eval = FALSE, echo = FALSE}
 ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
@@
 
-```{r RRs2, eval = TRUE, echo = FALSE}
+```{r RRs2, eval = FALSE, echo = FALSE}
 comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
@@
 ### Deaths
-```{r show222, echo = FALSE, eval = TRUE}
+```{r show222, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### DALY
-```{r show22, echo = FALSE, eval = TRUE}
+```{r show22, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 ### YLD
-```{r show23, echo = FALSE, eval = TRUE}
+```{r show23, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 #### YLL
-```{r show24, echo = FALSE, eval = TRUE}
+```{r show24, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r example, eval = TRUE, echo = FALSE}
+```{r example, eval = FALSE, echo = FALSE}
 example("ITHIM-package")
@@
 
-```{r vision, eval = TRUE, echo = FALSE}
+```{r vision, eval = FALSE, echo = FALSE}
 ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
@@
 #### YLL
-```{r show233, echo = FALSE, eval = TRUE}
+```{r show233, echo = FALSE, eval = FALSE}
 matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
@@
 
-```{r params, eval = TRUE, echo = FALSE}
+```{r params, eval = FALSE, echo = FALSE}
 names(ITHIM.baseline$parameters)
@@
 NHTS %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarise( n = n()) # 260,576 people in merged data
-
-#NHTS <- subset(NHTS, !is.na(AGE) & !is.na(SEX))
-
 ggplot(NHTS, aes(x = TRVL_MIN, ..density.., colour = AGE)) + geom_freqpoly(binwidth = 15) + facet_grid( SEX ~ AT) + labs(x = "Foo", colour = "Age") + coord_cartesian(xlim = c(0, 75))
-
 NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))

### 2016-09-28 R/ITHIM.Rmd
Removed update of pWalk
pWalk is now calculated using mean times

@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-upper <- 25; wd <- 1; length <- upper/wd + 1;
+upper <- 25; wd <- 10; length <- upper/wd + 1;
 
@@
 ITHIM.default <- updateITHIM(ITHIM.default, "cv", 3.0288)
-ITHIM.default <- updateITHIM(ITHIM.default, "pWalk", 0.85)
 ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)

### 2016-09-30 R/ITHIM.Rmd
Created road fatalaties figure from BikePedDeaths_09-13.csv

@@
 
+```{r roadInjuries, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+roadFatalaties <- read.csv(file = "./../data/BikePedDeaths_09-13.csv", na.strings = c("998","999"))
+roadFatalaties <- within(roadFatalaties, {
+  ageClass <- convertToAgeClass(AGE)
+  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
+  SEX <- factor(SEX, levels = c("F","M"))
+})
+roadFatalaties <- subset(roadFatalaties, !is.na(SEX))
+
+p <- ggplot(roadFatalaties, aes(ageClass, fill = SEX))
+p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + facet_grid( death ~ ., scales = "free_y")
+```
 # Appendix

### 2016-10-03 R/ITHIM.Rmd
Tweaked figures for Maggie K.

@@
 
-```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = Sex) )
-p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( State ~ Disease) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
-```
-
-```{r counts, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( State ~ Disease) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8)) + coord_flip()
+```
+
+```{r counts, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
 D <- melt(counts)
@@
 p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Individuals Sampled (Thousands)") + facet_grid( .
-~ State) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6))
+~ State) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6)) + coord_flip()
 ```
@@
 ```
+
 # Appendix

### 2016-10-03 R/ITHIM.Rmd
Still working on figures for Maggie K.

@@
 foobar <- within(foobar, {
-  Disease[Disease == "BreastCancer"] <- "Breast"
-  Disease[Disease == "ColonCancer"] <- "Colon"
-  Disease <- factor(Disease, levels = c("Breast", "Colon", "Dementia", "Depression", "Diabetes"))
+  Disease[Disease == "BreastCancer"] <- "Breast Cancer"
+  Disease[Disease == "ColonCancer"] <- "Colon Cancer"
+  Disease <- factor(Disease, levels = c("Breast Cancer", "Colon Cancer", "Dementia", "Depression", "Diabetes"))
 })
 
-p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = Sex) )
-p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( State ~ Disease) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8)) + coord_flip()
+p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3) )
+p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + theme_bw() + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
 ```
@@
 })
-p <- ggplot(subset(D,AgeClass != "00-04"), aes(AgeClass, count/1e3, fill = Sex))
-p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Individuals Sampled (Thousands)") + facet_grid( .
-~ State) + theme_bw() + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6)) + coord_flip()
+p <- ggplot(subset(D,AgeClass != "00-04"), aes(AgeClass, count/1e3))
+p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Individuals Sampled (Thousands)") + facet_grid( State ~ .) + theme_bw() + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6))
 ```

### 2016-10-04 R/ITHIM.Rmd
Change to ITHIM.Rmd

@@
 library("scales")
+library("ggthemes")
 ```
@@
 ```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
-
 ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)
@@
 
-p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3) )
-p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + theme_bw() + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
+p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = State) )
+
+upper <- 10
+
+p + geom_bar(stat="identity", position = "dodge") +
+theme(axis.title=element_blank()) +
+facet_grid( Disease ~ ., scales = "free") +
+scale_y_continuous(breaks=seq(1, upper, 1), name="DALYs Averted Nationwide (Thousands)") +
+
+#geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
+scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
+theme_tufte(base_size=12, ticks=T) 
+#
+
+# + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo")) 
+# + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
 ```

### 2016-10-05 manuscript/manuscript.tex
Edits to state-wise manuscript
Still working on figures for Maggie K.

--- 2016-09-21 manuscript/manuscript.tex
@@
   via reduced chronic disease, improved air quality, and reductions in
-  greenhouse gas emissions --- obtaining estimates for the quantiative
-  impacts of such shifts remains infrequent, if not altogether
-  absent from related policy action. Using U.S. metropolitan counties from the 10 regions defined
-  by the U.S. Department of Health and Human Services (HHS1-10) we perform
-  a comparative risk analysis between each regions and national
-  averages as the baseline.  We investigated the reduced chronic
-  disease burden of ($x$, $y$, $z$ outcomes) from increased physical
-  activity from population shifts in the amount of biking and walking
-  doen for transport.  Our methods employ the Integrated Transport and
-  Health Impacts Model (ITHIM) informed by regional travel parameters
-  using the National Household Transportation Survey, non-travel
-  activity metircs from the American Time Use Survey, and baseline
-  Disease burden from the Centers for Disease Control and Prevention's
-  WONDER database.  Our findings show that the West coast and the
-  Northeast of the U.S. have travel patterns that, if mirrored by the
-  nation, would save the U.S. thousands to hundreds of thousands
-  DALYs.}
+  greenhouse gas emissions — obtaining estimates for the quantiative impacts of such shifts remains infrequent, if not altogether absent from related policy action.
+Using U.S.
+metropolitan counties from seven states we perform a comparative risk analysis between each regions and national averages as the baseline.
+We employ a portion of the Integrated Transport and Health Impacts Model (ITHIM) to investigate the reduced chronic disease burden of (x,y,z outcomes) from changing levels of physical activity with population shifts in the amount of biking and walking done for transport.
+The analysis is informed by regional travel parameters from the National Household Transportation Survey, non-travel activity metrics from the American Time Use Survey, and baseline disease burden from the Centers for Disease Control and Prevention's WONDER database and the Global Burden of Disease dataset.
+Our findings show that the California and New York State have travel patterns that, if mirrored by the nation, would save the U.S.
+upwards of fifty thousand disability adjusted life years.}
 
@@
 
+\col{Regular exercise has been shown to provide health benefits in terms of chronic disease prevention \cite{warburton2006}, cognitive function \cite{hillman2008}, and overall well being and satisfaction with life \cite{maher2013}.
+Still only half of Americans obtain the recommended levels of aerobic activity, which is 2.5 hours per week.
+There is considerable variability in the attainment of this goal by location within the country.
+Supportive local built and policy environments are important determinants of population-level activity patterns.
+Particularly, as they facilitate and foster routine moderate physical activity, such as walking, local decissions on design and policy can play a huge role in the Nation's health.}
+
+\col{To better understand the potential for attainable shifts in transportation behavior across the US, we employed a comparative risk assessment methodology to compare national averages for time spent walking and cycling for transport to those of specific states.
+Our estimates can provide an important evidence base to encourage new construction, retrofits, land use policy, and programatic funding to support active transportation.
+Particularly in places with poor attainmnet of the physical activity guidelines, such investments can be importnat pieces of lareger population health interventions that have effective and long-lasting impacts.
+}
+
 \col{ITHIM is a statistical model that integrates data on travel
@@
   alternative scenario.}
-
-\col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-  eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
-  ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-  aliquip ex ea commodo consequat. Duis aute irure dolor in
-  reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-  pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-  culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
-  dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
-  incididunt ut labore et dolore magna aliqua.}
-
-\col{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
-  eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
-  ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
-  aliquip ex ea commodo consequat. Duis aute irure dolor in
-  reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-  pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
-  culpa qui officia deserunt mollit anim id est laborum.}
 
@@
 
+\begin{figure}[t]
+    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
+    \caption{Map of HHS Regions and metropolitan areas}\label{HHSmap}
+\end{figure}
+
 \section{U.S. Metro-Regional Analysis}
@@
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{Change in total national DALYs for each regional
-      transportation estimates.}\label{dalyFigure}
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
+    \caption{Mean travel time by mode (walking/cycling) for each of
+      the four regions with a decrease in DALYs
+      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
+      national means are displayed on the far right.  These four
+      regions correspond to New England/NewYork and the West
+      Coast.}\label{meanMatrices}
 \end{figure}

### 2016-10-05 R/ITHIM.Rmd

@@
 ```{r compute, echo = FALSE, eval = TRUE}
-compute <- FALSE
+compute <- TRUE
 ```
@@
 ```{r subset, eval = compute, fig.width = 10, echo = FALSE}
-NHTS.person <- subset( NHTS.person, MSA == TRUE & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
+NHTS.person <- subset( NHTS.person, MSA == TRUE ) # & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
@@
 ```{r roadInjuries, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-roadFatalaties <- read.csv(file = "./../data/BikePedDeaths_09-13.csv", na.strings = c("998","999"))
+convertState <- function( stateCode ){
+
+stateCode <- as.character(stateCode)
+
+stateCodeMatrix <- matrix(c("Alabama","1","Nebraska","31","Alaska","2","Nevada","32","Arizona","4","New Hampshire","33","Arkansas","5","New Jersey","34","California","6","New Mexico","35","Colorado","8","New York","36","Connecticut","9","North Carolina","37","Delaware","10","North Dakota","38","District of Columbia","11", "Ohio", "39","Florida","12","Oklahoma","40","Georgia","13","Oregon","41","Hawaii","15","Pennsylvania","42","Idaho","16","Puerto Rico","43","Illinois","17","Rhode Island","44","Indiana","18","South Carolina","45","Iowa","19","South Dakota","46","Kansas","20","Tennessee","47","Kentucky","21","Texas","48","Louisiana","22","Utah","49","Maine","23","Vermont","50","Maryland","24","Virgin Islands (since 2004)","52","Massachusetts","25","Virginia","51","Michigan","26","Washington","53","Minnesota","27","West Virginia","54","Mississippi","28","Wisconsin","55","Missouri","29","Wyoming","56","Montana","30"), byrow = TRUE, ncol = 2)
+
+stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
+
+stateNames <- stateCodeMatrix[,1]
+names(stateNames) <- stateCodeMatrix[,2]
+
+stateAbbrevs <- stateAbbrevMatrix[,2]
+names(stateAbbrevs) <- stateAbbrevMatrix[,1]
+
+return(stateAbbrevs[stateNames[stateCode]])
+
+}
+
+roadFatalaties <- read.csv(file = "./../data/FARS/BikePedDeaths99-13.csv", na.strings = c("998","999"))
 roadFatalaties <- within(roadFatalaties, {
@@
   SEX <- factor(SEX, levels = c("F","M"))
+  STATE <- as.factor(convertState(STATE))
 })
-roadFatalaties <- subset(roadFatalaties, !is.na(SEX))
-
-p <- ggplot(roadFatalaties, aes(ageClass, fill = SEX))
-p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + facet_grid( death ~ ., scales = "free_y")
+roadFatalaties <- subset(roadFatalaties, !is.na(SEX) ) #& STATE %in% c("California", "New York"))
+
+p <- ggplot(roadFatalaties, aes(Type, fill = STATE))
+p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5))# + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) #+ facet_grid( Type ~ ., scales = "free_y")
+
+walkTime <- unlist(lapply(meanWalkList, function(x) x[3,2]))
+
+foo <- data.frame(with(roadFatalaties, table(STATE,Type)))
+within(subset(foo, Type == "Ped"), Walk <- walkTime[as.character(STATE)])
+
 ```

### 2016-10-05 manuscript/manuscript.tex

@@
 
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
-    \caption{Map of HHS Regions and metropolitan areas}\label{HHSmap}
-\end{figure}
-
-\section{U.S. Metro-Regional Analysis}
+\section{Seven State Analysis}
 
@@
       regions correspond to New England/NewYork and the West
-      Coast.}\label{meanMatrices}
+      Coast. The age classes are defined as in the original ITHIM
+      implementation using 8 eight age classes defined in terms of
+      years as follows: 1. 0--4, 2. 5--14, 3. 15--29, 4. 30--44,
+      5. 45--59, 6. 60--69, 7. 70--79, and 8. 80+. Age class 1 is
+      omitted from all calclulations in this paper due to lack of
+      data.}\label{meanMatrices}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
+    \caption{California \& New York Averted DALYs Nationwide}\label{California}
+\end{figure}
+
+\begin{figure}[t]
+    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
+    \caption{Counts}\label{counts}
 \end{figure}

### 2016-10-05 manuscript/manuscript.tex
Removed text coloring and a few figures
Rather the code was commented out.  The links to the figures are still there.

@@
 \section{Summary:}
-\col[blue]{While increased active transportation, primarily cycling
-  and walking, is often held up as a goal for improved public health —
-  via reduced chronic disease, improved air quality, and reductions in
-  greenhouse gas emissions — obtaining estimates for the quantiative impacts of such shifts remains infrequent, if not altogether absent from related policy action.
-Using U.S.
-metropolitan counties from seven states we perform a comparative risk analysis between each regions and national averages as the baseline.
-We employ a portion of the Integrated Transport and Health Impacts Model (ITHIM) to investigate the reduced chronic disease burden of (x,y,z outcomes) from changing levels of physical activity with population shifts in the amount of biking and walking done for transport.
-The analysis is informed by regional travel parameters from the National Household Transportation Survey, non-travel activity metrics from the American Time Use Survey, and baseline disease burden from the Centers for Disease Control and Prevention's WONDER database and the Global Burden of Disease dataset.
-Our findings show that the California and New York State have travel patterns that, if mirrored by the nation, would save the U.S.
-upwards of fifty thousand disability adjusted life years.}
+
+While increased active transportation, primarily cycling and walking,
+is often held up as a goal for improved public health --- via reduced
+chronic disease, improved air quality, and reductions in greenhouse
+gas emissions — obtaining estimates for the quantiative impacts of
+such shifts remains infrequent, if not altogether absent from related
+policy action. Using U.S. metropolitan counties from seven states we
+perform a comparative risk analysis between each regions and national
+averages as the baseline.  We employ a portion of the Integrated
+Transport and Health Impacts Model (ITHIM) to investigate the reduced
+chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
+levels of physical activity with population shifts in the amount of
+biking and walking done for transport. The analysis is informed by
+regional travel parameters from the National Household Transportation
+Survey, non-travel activity metrics from the American Time Use Survey,
+and baseline disease burden from the Centers for Disease Control and
+Prevention's WONDER database and the Global Burden of Disease dataset.
+Our findings show that the California and New York State have travel
+patterns that, if mirrored by the nation, would save the U.S. upwards
+of fifty thousand disability adjusted life years.
 
 \section{Availability:}
-\col[blue]{\package{} may be explored using the user interface
-  avaiable at \webpage{}.  The source code and data files for
-  \package{} are publicly available on GitHub at \webpage{}.}
+
+\package{} may be explored using the user interface avaiable
+at \webpage{}.  The source code and data files for \package{} are
+publicly available on GitHub at \webpage{}.
 
@@
 
-\col[blue]{See \webpage{} for links to the source code, manual,
-  tutorial and links for more information about the ITHIM model.}
+See \webpage{} for links to the source code, manual, tutorial and
+links for more information about the ITHIM model.
 
 \section{Contact:}
-\col[blue]{\href{javargo@wisc.edu}{javargo@wisc.edu}}
+\href{javargo@wisc.edu}{javargo@wisc.edu}
+
 \end{abstract}
@@
 
-\col{Regular exercise has been shown to provide health benefits in terms of chronic disease prevention \cite{warburton2006}, cognitive function \cite{hillman2008}, and overall well being and satisfaction with life \cite{maher2013}.
-Still only half of Americans obtain the recommended levels of aerobic activity, which is 2.5 hours per week.
-There is considerable variability in the attainment of this goal by location within the country.
-Supportive local built and policy environments are important determinants of population-level activity patterns.
-Particularly, as they facilitate and foster routine moderate physical activity, such as walking, local decissions on design and policy can play a huge role in the Nation's health.}
+Regular exercise has been shown to provide health benefits in terms of
+chronic disease prevention \cite{warburton2006}, cognitive function
+\cite{hillman2008}, and overall well being and satisfaction with life
+\cite{maher2013}. Still only half of Americans obtain the recommended
+levels of aerobic activity, which is 2.5 hours per week. There is
+considerable variability in the attainment of this goal by location
+within the country. Supportive local built and policy environments are
+important determinants of population-level activity
+patterns. Particularly, as they facilitate and foster routine moderate
+physical activity, such as walking, local decissions on design and
+policy can play a huge role in the Nation's health.
 
-\col{To better understand the potential for attainable shifts in transportation behavior across the US, we employed a comparative risk assessment methodology to compare national averages for time spent walking and cycling for transport to those of specific states.
-Our estimates can provide an important evidence base to encourage new construction, retrofits, land use policy, and programatic funding to support active transportation.
-Particularly in places with poor attainmnet of the physical activity guidelines, such investments can be importnat pieces of lareger population health interventions that have effective and long-lasting impacts.
-}
+To better understand the potential for attainable shifts in
+transportation behavior across the US, we employed a comparative risk
+assessment methodology to compare national averages for time spent
+walking and cycling for transport to those of specific states. Our
+estimates can provide an important evidence base to encourage new
+construction, retrofits, land use policy, and programatic funding to
+support active transportation. Particularly in places with poor
+attainmnet of the physical activity guidelines, such investments can
+be importnat pieces of lareger population health interventions that
+have effective and long-lasting impacts.
 
-\col{ITHIM is a statistical model that integrates data on travel
-  patterns, physical activity, fine particulate matter, GHG emissions,
-  and disease and injuries. Based on population and travel
-  scenarios. The model has been used to calculate the health impacts
-  of walking and bicycling short distances usually traveled by car or
-  driving low-emission automobiles \cite{woodcock2013,maizlish2013}.}
+ITHIM is a statistical model that integrates data on travel patterns,
+physical activity, fine particulate matter, GHG emissions, and disease
+and injuries. Based on population and travel scenarios. The model has
+been used to calculate the health impacts of walking and bicycling
+short distances usually traveled by car or driving low-emission
+automobiles \cite{woodcock2013,maizlish2013}.
 
-\col{The ITHIM model uses comparative risk assessment through which it
-  formulates a change in the disease burden, resulting from the shift
-  in the exposure distribution from a baseline scenario to an
-  alternative scenario.}
+The ITHIM model uses comparative risk assessment through which it
+formulates a change in the disease burden, resulting from the shift in
+the exposure distribution from a baseline scenario to an alternative
+scenario.
 
@@
 
-\col[blue]{\textit{
+\textit{
     The ITHIM model uses the comparative risk
@@
     were adjusted to represent an alternate scenario.
-}}
+}
 
@@
 
-\col[blue]{\textit{
-    Here $R(x)$ represents the risk when the individual has exposure $x$,
-in our case physical activity (\mets).  $P(x)$ and $Q(x)$ denote the
-population density of individuals with exposure $x$, in the baseline
-and alternate scenarios, respectively.  The exposure variable, $x$, is
-modeled as the sum of travel-related exposure and non-travel-related
-exposure.
-}}
+\textit{ Here $R(x)$ represents the risk when the individual has
+  exposure $x$, in our case physical activity (\mets).  $P(x)$ and
+  $Q(x)$ denote the population density of individuals with exposure
+  $x$, in the baseline and alternate scenarios, respectively.  The
+  exposure variable, $x$, is modeled as the sum of travel-related
+  exposure and non-travel-related exposure.}
 
@@
 
-\col[blue]{\textit{The time spent walking or cycling, i.e., active
-    transport time, is modeled as a log-normal random variable with
-    means defined for each age-sex stratum.  The non-travel-related
-    physical activity is also modeled as log-normal with
-    stratum-specific means.}}
+\textit{The time spent walking or cycling, i.e., active transport
+  time, is modeled as a log-normal random variable with means defined
+  for each age-sex stratum.  The non-travel-related physical activity
+  is also modeled as log-normal with stratum-specific means.}
 
-\col[blue]{\textit{
-The travel-related exposure is estimated using active transport time
-along with assumptions about how much physical activity is required
-for cycling and walking.  Active transport time and non-travel-related
-physical activity are modeled as independent random variables.  A
-fixed value was used as the coefficient of variation for the
-distributin of active transport time. A different, but also fixed,
-value was used for the non-travel-related physical activity
-distribution.
-}}
+\textit{ The travel-related exposure is estimated using active
+  transport time along with assumptions about how much physical
+  activity is required for cycling and walking.  Active transport time
+  and non-travel-related physical activity are modeled as independent
+  random variables.  A fixed value was used as the coefficient of
+  variation for the distributin of active transport time. A different,
+  but also fixed, value was used for the non-travel-related physical
+  activity distribution.}
 
-\subsection{Road Injuries}
+%% \subsection{Road Injuries}
 
-\col[red]{\textit{
-We chose to model road injuries in a more simplifed manner than the
-authors of ITHIM.  The original ITHIM requires information about the
-transportational infrastructure that cannot be accurately modeled on a
-regional level, and is more well-suited to say a collection of
-counties as a case study. \cite{maizlish2013}  We use the overall percent
-of travel by walking and road fatalities per U.S. metropolitan area to
-estimate road fatalaties in U.S. metropolitan areas given the percent
-of travel by walking. \cite{UWTechReport}
-}}
+%% \textit{ We chose to model road injuries in a more simplifed manner
+%%   than the authors of ITHIM.  The original ITHIM requires information
+%%   about the transportational infrastructure that cannot be accurately
+%%   modeled on a regional level, and is more well-suited to say a
+%%   collection of counties as a case study. \cite{maizlish2013} We use
+%%   the overall percent of travel by walking and road fatalities per
+%%   U.S. metropolitan area to estimate road fatalaties in
+%%   U.S. metropolitan areas given the percent of travel by
+%%   walking. \cite{UWTechReport}}
 
-\col[red]{\textit{
-We need to estimate the increase/decrease in road injuries/fatalaties
-nation-wide as walking/cycling time increases.  We need a relative
-risk of road fatalaty as a function of active transport time...
-}}
+%% \textit{ We need to estimate the increase/decrease in road
+%%   injuries/fatalaties nation-wide as walking/cycling time increases.
+%%   We need a relative risk of road fatalaty as a function of active
+%%   transport time\ldots}
 
-\subsection{Air Pollution}
+%% \subsection{Air Pollution}
 
-\col[red]{\textit{
-    Can we model the decrease in \pEm{2.5} nationally as a function of
-decrease in driving time?  The original plan was to use a baseline
-value for \pEm{2.5} and a linear relationship bewteen change in
-\pEm{2.5} and change in \vmt{}.
-}}
+%% \textit{ Can we model the decrease in \pEm{2.5} nationally as a
+%%   function of decrease in driving time?  The original plan was to use
+%%   a baseline value for \pEm{2.5} and a linear relationship bewteen
+%%   change in \pEm{2.5} and change in \vmt{}.}
 
@@
 
-\section{Seven State Analysis}
-
 \begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{Change in total national DALYs for each regional
-      transportation estimates.}\label{dalyFigure}
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
+    \caption{California \& New York Averted DALYs Nationwide}\label{California}
 \end{figure}
 
-\col[blue]{\begin{itemize}
+%% \begin{figure}[t]
+%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
+%%     \caption{Counts}\label{counts}
+%% \end{figure}
+
+\section{Seven State Analysis}
+\begin{itemize}
 \item We used the National Household Transportation Survey \cite{NHTS} to
@@
 \item Values for DALYs were found using the CDC Wonder database \cite{CDCWonder}
-\end{itemize}}
+\end{itemize}
+
+%% \begin{figure}[t]
+%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
+%%     \caption{Change in total national DALYs for each regional
+%%       transportation estimates.}\label{dalyFigure}
+%% \end{figure}
 
@@
 
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
-    \caption{California \& New York Averted DALYs Nationwide}\label{California}
-\end{figure}
-
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
-    \caption{Counts}\label{counts}
-\end{figure}
-
 \section*{Acknowledgment}
-\paragraph{Funding\textcolon} \col[red]{SGY is supported by$\ldots$ , JV, JP, $\ldots$}
+\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots
 
 %\bibliographystyle{natbib}
-\bibliographystyle{bmc_article}
-%\bibliographystyle{mypapers}
+%\bibliographystyle{bmc_article}
+\bibliographystyle{mypapers}
 \bibliography{ITHIM}

### 2016-10-05 manuscript/manuscript.tex
Working on edits to Methods section of manuscript.

@@
 \newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
-\newcommand{\af}{\rho}
+\newcommand{\af}{\textrm{PAF}}
 \newcommand{\mets}{MET-hrs./week}
@@
 
-\textit{
-    The ITHIM model uses the comparative risk
-    assesment framework to estimate the population attributable
-    fraction, $\af$.  In our case $\af$ represents the proportional
-    change in mortality and morbidity if the physical activity levels
-    were adjusted to represent an alternate scenario.
-}
+The ITHIM model uses the comparative risk assesment framework to
+estimate the population attributable fraction, $\af$.  In our case the
+$\af$ represents the proportional change in national disease burden if
+age and sex-specific active transport time means were changed.
 
@@
 
-\textit{ Here $R(x)$ represents the risk when the individual has
-  exposure $x$, in our case physical activity (\mets).  $P(x)$ and
-  $Q(x)$ denote the population density of individuals with exposure
-  $x$, in the baseline and alternate scenarios, respectively.  The
-  exposure variable, $x$, is modeled as the sum of travel-related
-  exposure and non-travel-related exposure.}
+Here $R(x)$ represents the risk when the individual has exposure $x$,
+physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
+density of individuals with exposure $x$, in the baseline and
+alternate scenarios, respectively.  The exposure variable, $x$, is
+modeled as the sum of travel-related exposure and non-travel-related
+exposure, which are modeled as two independent random variables.  The
+age-sex specific mean ratios for non-travel-related exposure are
+estimated from the ATUS \cite{ATUS} and the referent group mean is
+treated as a variable.  We see in Figure \ref{dalyFigure} that varying
+it does not greatly affect the estimate for disease burden.
 
@@
 
-\section{Seven State Analysis}
+\section{New York and California}
+
 \begin{itemize}
@@
 
-%% \begin{figure}[t]
-%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-%%     \caption{Change in total national DALYs for each regional
-%%       transportation estimates.}\label{dalyFigure}
-%% \end{figure}
+\begin{figure}[t]
+    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
+    \caption{Change in total national DALYs for each regional
+      transportation estimates.}\label{dalyFigure}
+\end{figure}
 

### 2016-10-05 R/ITHIM.Rmd
Edits to manuscript.

@@
 #geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
-scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
-theme_tufte(base_size=12, ticks=T)
+scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+"))
+# + theme_tufte(base_size=12, ticks=T)
 #
-
 # + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))

### 2016-10-05 manuscript/manuscript.tex
Edits to manuscript.

@@
 \newcommand{\vmt}{$\mathrm{VMT}$}
+\newcommand{\logNormal}{log-normal}
 
@@
 alternate scenarios, respectively.  The exposure variable, $x$, is
-modeled as the sum of travel-related exposure and non-travel-related
-exposure, which are modeled as two independent random variables.  The
-age-sex specific mean ratios for non-travel-related exposure are
-estimated from the ATUS \cite{ATUS} and the referent group mean is
-treated as a variable.  We see in Figure \ref{dalyFigure} that varying
-it does not greatly affect the estimate for disease burden.
+modeled as the sum of two independent random variables, travel-related
+exposure and non-travel-related exposure.  Non-travel-related exposure
+is modeled with a \logNormal{} distribution with the age-sex specific
+mean ratios for non-travel-related exposure are estimated from the
+ATUS \cite{ATUS}.  The referent group mean is treated as a variable
+and the coefficient of variation is treated as constant across age-sex
+strata and estimated with ATUS.  We see in Figure \ref{dalyFigure}
+that varying the mean values for non-travel-related exposure does not
+greatly affect the estimate for disease burden.  The distribution for
+travel-related exposure is estimated from the time spent walking or
+cycling, i.e., active transport time, which as in the original ITHIM
+model has a \logNormal{} distribution with constant coefficient of
+variation across age-sex strata \cite{}.  Travel-related exposure
+is estimated from active transport time using assumptions about how
+much physical activity is required for cycling and walking.
 
-\subsection{Active Transport}
+\end{methods}
 
-\textit{The time spent walking or cycling, i.e., active transport
-  time, is modeled as a log-normal random variable with means defined
-  for each age-sex stratum.  The non-travel-related physical activity
-  is also modeled as log-normal with stratum-specific means.}
+\begin{figure}[t]
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
+    \caption{California \& New York Averted DALYs Nationwide}\label{California}
+\end{figure}
 
-\textit{ The travel-related exposure is estimated using active
-  transport time along with assumptions about how much physical
-  activity is required for cycling and walking.  Active transport time
-  and non-travel-related physical activity are modeled as independent
-  random variables.  A fixed value was used as the coefficient of
-  variation for the distributin of active transport time. A different,
-  but also fixed, value was used for the non-travel-related physical
-  activity distribution.}
+\section{Results}
+
+We used the National Household Transportation Survey \cite{NHTS} to
+estimate the mean walk and cycle times for individuals that live in
+metropolitan areas for each of seven states (CA, TX, NY, FL, VA, NC
+and AZ) which had sufficient sample size for reasonable estimates of
+age-sex specific travel-related walking and cycling means, Figure
+\ref{meanMatrices}.  Of these seven states two were found to have
+active transport behavior that would avert thousands of DALYs if
+adopted nation-wide. Values for national disease-specific DALYs were
+found using the CDC Wonder database \cite{CDCWonder}.
+
+%% \begin{itemize}
+%% \item We used the National Household Transportation Survey \cite{NHTS} to
+%%   estimate the mean walk and cycle times for individuals that live in
+%%   metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
+%%   corresponds to individuals whose homes are in an MSA or cMSA)
+%%   \bi\item Data quality issues along with assumptions about travel speed
+%% created mirky data.  The median values of trip time were used to avoid
+%% undue influence by outliers.\ei
+%% \item We then divided these into the 10 regions of the U.S. defined by
+%%   the Department of Health and Human Services (HH1-10)
+%% \item Using only the active transport component of the ITHIM model we
+%%   ran comparative analyses for each of the ten regions, using the
+%%   national averages as the baseline.
+%% \item We used data from the American Time Use Survey \cite{ATUS} to
+%%   estimate the relative means for each of the age-sex strata.  We then
+%%   treat the mean of the referent group as a variable and compute
+%%   results over a grid of values for this variable.  See Figure
+%%   \ref{dalyFigure}.
+%% \item 
+%% \end{itemize}
+
+\begin{figure}[t]
+    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
+    \caption{Change in total national DALYs for each regional
+      transportation estimates.}\label{dalyFigure}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
+    \caption{Mean travel time by mode (walking/cycling) for each of
+      seven states National means are displayed on the right.}\label{meanMatrices}
+\end{figure}
+
+\section*{Acknowledgment}
+\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots
+
+%\bibliographystyle{natbib}
+\bibliographystyle{bmc_article}
+%\bibliographystyle{mypapers}
+\bibliography{ITHIM}
+\end{document}
 
@@
 %%   change in \pEm{2.5} and change in \vmt{}.}
-
-\end{methods}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
-    \caption{California \& New York Averted DALYs Nationwide}\label{California}
-\end{figure}
-
 %% \begin{figure}[t]
@@
 %% \end{figure}
-
-\section{New York and California}
-
-\begin{itemize}
-\item We used the National Household Transportation Survey \cite{NHTS} to
-  estimate the mean walk and cycle times for individuals that live in
-  metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
-  corresponds to individuals whose homes are in an MSA or cMSA)
-  \bi\item Data quality issues along with assumptions about travel speed
-created mirky data.  The median values of trip time were used to avoid
-undue influence by outliers.\ei
-\item We then divided these into the 10 regions of the U.S. defined by
-  the Department of Health and Human Services (HH1-10)
-\item Using only the active transport component of the ITHIM model we
-  ran comparative analyses for each of the ten regions, using the
-  national averages as the baseline.
-\item We used data from the American Time Use Survey \cite{ATUS} to
-  estimate the relative means for each of the age-sex strata.  We then
-  treat the mean of the referent group as a variable and compute
-  results over a grid of values for this variable.  See Figure
-  \ref{dalyFigure}.
-\item Values for DALYs were found using the CDC Wonder database \cite{CDCWonder}
-\end{itemize}
-
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{Change in total national DALYs for each regional
-      transportation estimates.}\label{dalyFigure}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
-    \caption{Mean travel time by mode (walking/cycling) for each of
-      the four regions with a decrease in DALYs
-      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
-      national means are displayed on the far right.  These four
-      regions correspond to New England/NewYork and the West
-      Coast. The age classes are defined as in the original ITHIM
-      implementation using 8 eight age classes defined in terms of
-      years as follows: 1. 0--4, 2. 5--14, 3. 15--29, 4. 30--44,
-      5. 45--59, 6. 60--69, 7. 70--79, and 8. 80+. Age class 1 is
-      omitted from all calclulations in this paper due to lack of
-      data.}\label{meanMatrices}
-\end{figure}
-
-\section*{Acknowledgment}
-\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots
-
-%\bibliographystyle{natbib}
-%\bibliographystyle{bmc_article}
-\bibliographystyle{mypapers}
-\bibliography{ITHIM}
-\end{document}

### 2016-10-05 manuscript/manuscript.tex
More edits to manuscript.

@@
 \begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
-    \caption{California \& New York Averted DALYs Nationwide}\label{California}
-\end{figure}
-
-\section{Results}
-
-We used the National Household Transportation Survey \cite{NHTS} to
-estimate the mean walk and cycle times for individuals that live in
-metropolitan areas for each of seven states (CA, TX, NY, FL, VA, NC
-and AZ) which had sufficient sample size for reasonable estimates of
-age-sex specific travel-related walking and cycling means, Figure
-\ref{meanMatrices}.  Of these seven states two were found to have
-active transport behavior that would avert thousands of DALYs if
-adopted nation-wide. Values for national disease-specific DALYs were
-found using the CDC Wonder database \cite{CDCWonder}.
-
-%% \begin{itemize}
-%% \item We used the National Household Transportation Survey \cite{NHTS} to
-%%   estimate the mean walk and cycle times for individuals that live in
-%%   metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
-%%   corresponds to individuals whose homes are in an MSA or cMSA)
-%%   \bi\item Data quality issues along with assumptions about travel speed
-%% created mirky data.  The median values of trip time were used to avoid
-%% undue influence by outliers.\ei
-%% \item We then divided these into the 10 regions of the U.S. defined by
-%%   the Department of Health and Human Services (HH1-10)
-%% \item Using only the active transport component of the ITHIM model we
-%%   ran comparative analyses for each of the ten regions, using the
-%%   national averages as the baseline.
-%% \item We used data from the American Time Use Survey \cite{ATUS} to
-%%   estimate the relative means for each of the age-sex strata.  We then
-%%   treat the mean of the referent group as a variable and compute
-%%   results over a grid of values for this variable.  See Figure
-%%   \ref{dalyFigure}.
-%% \item 
-%% \end{itemize}
-
-\begin{figure}[t]
     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{Change in total national DALYs for each regional
-      transportation estimates.}\label{dalyFigure}
+    \caption{}\label{dalyFigure}
 \end{figure}
@@
     \caption{Mean travel time by mode (walking/cycling) for each of
-      seven states National means are displayed on the right.}\label{meanMatrices}
+      seven states.  National means are displayed on the
+      right.}\label{meanMatrices}
 \end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
+    \caption{}\label{California}
+\end{figure}
+
+\section{Results}
+
+We used the National Household Transportation Survey to estimate the
+mean walk and cycle times for individuals that live in metropolitan
+areas for each of seven states (CA, TX, NY, FL, VA, NC and AZ)
+\cite{NHTS}.  These states were selected becaused they had sufficient
+sample size for reasonable estimates of age-sex specific
+travel-related walking and cycling means, Figure \ref{meanMatrices}.
+
+Of these seven states one was found to show active transport behavior
+that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
+Values for national disease-specific DALYs were found using the CDC
+Wonder database \cite{CDCWonder}.  California stood out among the
+estimates for national DALYs averted.  Most of the decrease in DALYs
+is due to prevention of dementia in the oldest age group, $80+$ years
+old.  California exhibits much greater walking means in this age group
+than the national average.  We also see prevention of depression and
+diabetes in the middle age groups due to increased active transport
+time.
 
@@
 %% \subsection{Road Injuries}
-
 %% \textit{ We chose to model road injuries in a more simplifed manner
@@
 %% \end{figure}
+
+%% \begin{itemize}
+%% \item We used the National Household Transportation Survey \cite{NHTS} to
+%%   estimate the mean walk and cycle times for individuals that live in
+%%   metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
+%%   corresponds to individuals whose homes are in an MSA or cMSA)
+%%   \bi\item Data quality issues along with assumptions about travel speed
+%% created mirky data.  The median values of trip time were used to avoid
+%% undue influence by outliers.\ei
+%% \item We then divided these into the 10 regions of the U.S. defined by
+%%   the Department of Health and Human Services (HH1-10)
+%% \item Using only the active transport component of the ITHIM model we
+%%   ran comparative analyses for each of the ten regions, using the
+%%   national averages as the baseline.
+%% \item We used data from the American Time Use Survey \cite{ATUS} to
+%%   estimate the relative means for each of the age-sex strata.  We then
+%%   treat the mean of the referent group as a variable and compute
+%%   results over a grid of values for this variable.  See Figure
+%%   \ref{dalyFigure}.
+%% \item 
+%% \end{itemize}

### 2016-10-05 R/ITHIM.Rmd
Tweaked DALYs averted figure to be just CA

--- 2016-10-05 R/ITHIM.Rmd
@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-upper <- 25; wd <- 10; length <- upper/wd + 1;
+upper <- 25; length <- 10; wd <- upper/length;
 
@@
 
-for( muNonTravel in seq(0,upper,length.out = length) ){
+for( muNonTravel in seq(0, upper, by = wd)){
 
@@
 
-ITHIM.NY <- ITHIMList[["NY"]]
-ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
-foo2 <- compareModels(ITHIM.national,ITHIM.NY)
-foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
-
-foobar <- rbind(foobar1,foobar2)
+foobar <- foobar1
 names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
+
+## ITHIM.NY <- ITHIMList[["NY"]]
+## ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
+## foo2 <- compareModels(ITHIM.national,ITHIM.NY)
+## foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
+## foobar <- rbind(foobar1,foobar2)
+## names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
 
@@
 
-p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = State) )
-
-upper <- 10
+p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = Sex) )
+
+#upper <- 10
 
@@
 facet_grid( Disease ~ ., scales = "free") +
-scale_y_continuous(breaks=seq(1, upper, 1), name="DALYs Averted Nationwide (Thousands)") +
+scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") +
+scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+"))
 
 #geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
-scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
-theme_tufte(base_size=12, ticks=T)
+# + theme_tufte(base_size=12, ticks=T)
 #
-
 # + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))

### 2016-10-05 manuscript/manuscript.tex
Removed commented-out bits
Tweaked DALYs averted figure to be just CA

@@
 
-\section{Availability:}
+%% \section{Availability:}
 
-\package{} may be explored using the user interface avaiable
-at \webpage{}.  The source code and data files for \package{} are
-publicly available on GitHub at \webpage{}.
+%% %% \package{} may be explored using the user interface avaiable
+%% %% at \webpage{}.
+%% The source code and data files for \package{} are publicly available
+%% on GitHub via a link on \webpage{}.
 
@@
 is due to prevention of dementia in the oldest age group, $80+$ years
-old.  California exhibits much greater walking means in this age group
-than the national average.  We also see prevention of depression and
+old.  California exhibits much greater walking and cycling means in this age group
+than the national average. In particular we see the cycling mean We also see prevention of depression and
 diabetes in the middle age groups due to increased active transport
@@
 \end{document}
-
-%% \subsection{Road Injuries}
-%% \textit{ We chose to model road injuries in a more simplifed manner
-%%   than the authors of ITHIM.  The original ITHIM requires information
-%%   about the transportational infrastructure that cannot be accurately
-%%   modeled on a regional level, and is more well-suited to say a
-%%   collection of counties as a case study. \cite{maizlish2013} We use
-%%   the overall percent of travel by walking and road fatalities per
-%%   U.S. metropolitan area to estimate road fatalaties in
-%%   U.S. metropolitan areas given the percent of travel by
-%%   walking. \cite{UWTechReport}}
-
-%% \textit{ We need to estimate the increase/decrease in road
-%%   injuries/fatalaties nation-wide as walking/cycling time increases.
-%%   We need a relative risk of road fatalaty as a function of active
-%%   transport time\ldots}
-
-%% \subsection{Air Pollution}
-
-%% \textit{ Can we model the decrease in \pEm{2.5} nationally as a
-%%   function of decrease in driving time?  The original plan was to use
-%%   a baseline value for \pEm{2.5} and a linear relationship bewteen
-%%   change in \pEm{2.5} and change in \vmt{}.}
-%% \begin{figure}[t]
-%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
-%%     \caption{Counts}\label{counts}
-%% \end{figure}
-
-%% \begin{itemize}
-%% \item We used the National Household Transportation Survey \cite{NHTS} to
-%%   estimate the mean walk and cycle times for individuals that live in
-%%   metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
-%%   corresponds to individuals whose homes are in an MSA or cMSA)
-%%   \bi\item Data quality issues along with assumptions about travel speed
-%% created mirky data.  The median values of trip time were used to avoid
-%% undue influence by outliers.\ei
-%% \item We then divided these into the 10 regions of the U.S. defined by
-%%   the Department of Health and Human Services (HH1-10)
-%% \item Using only the active transport component of the ITHIM model we
-%%   ran comparative analyses for each of the ten regions, using the
-%%   national averages as the baseline.
-%% \item We used data from the American Time Use Survey \cite{ATUS} to
-%%   estimate the relative means for each of the age-sex strata.  We then
-%%   treat the mean of the referent group as a variable and compute
-%%   results over a grid of values for this variable.  See Figure
-%%   \ref{dalyFigure}.
-%% \item 
-%% \end{itemize}

### 2016-10-05 manuscript/manuscript.tex
Cleaned up manuscript a little

@@
 \newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
-\newcommand{\af}{\textrm{PAF}}
+\newcommand{\af}{\mathrm{PAF}}
 \newcommand{\mets}{MET-hrs./week}
@@
 \begin{equation}
-\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
+\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \, \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x}
 \end{equation}
@@
 model has a \logNormal{} distribution with constant coefficient of
-variation across age-sex strata \cite{}.  Travel-related exposure
-is estimated from active transport time using assumptions about how
-much physical activity is required for cycling and walking.
+variation across age-sex strata.  Travel-related exposure is estimated
+from active transport time using assumptions about how much physical
+activity is required for cycling and walking.
 
@@
 is due to prevention of dementia in the oldest age group, $80+$ years
-old.  California exhibits much greater walking and cycling means in this age group
-than the national average. In particular we see the cycling mean We also see prevention of depression and
-diabetes in the middle age groups due to increased active transport
-time.
+old.  California exhibits much greater walking and cycling means in
+this age group than the national average. In particular we see the
+cycling mean We also see prevention of depression and diabetes in the
+middle age groups due to increased active transport time.
 

### 2016-10-10 methodsPaper/tex/authorSummary.tex
Switched to PlosOne template.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur
eget porta erat. Morbi consectetur est vel gravida
pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam
sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus
mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc
pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet
enim ultrices, ultrices elit pulvinar, volutpat risus.

### 2016-10-10 methodsPaper/tex/conclusion.tex
Switched to PlosOne template.

CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.

### 2016-10-10 methodsPaper/tex/discussion.tex
Switched to PlosOne template.

Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat
euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem
et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur
adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id
massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce
fringilla erat porttitor lectus cursus, vel sagittis arcu
lobortis. Aliquam in enim semper, aliquam massa id, cursus
neque. Praesent faucibus semper libero~\cite{bib3}.

### 2016-10-10 methodsPaper/tex/results.tex
Switched to PlosOne template.

--- 2016-10-05 manuscript/manuscript.tex
@@
-\documentclass{bioinfo}
-\usepackage{url}
-\copyrightyear{2016}
-\pubyear{2016}
+% Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
 
-\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
-\newcommand{\bi}{\begin{itemize}}
-\newcommand{\ei}{\end{itemize}}
-\newcommand{\package}{\emph{ITHIM}}
-\newcommand{\ithim}{\emph{ITHIM}}
-\newcommand{\R}{\emph{R}}
-\newcommand{\sgy}[1]{{\itshape\col{#1}}}
-\newcommand{\XX}{\sgy{XX}}
-\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
-\newcommand{\af}{\textrm{PAF}}
-\newcommand{\mets}{MET-hrs./week}
-\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
-\newcommand{\vmt}{$\mathrm{VMT}$}
-\newcommand{\logNormal}{log-normal}
-
-\begin{document}
-\firstpage{1}
-
-\title[\package{}]{A comparative risk assesment of U.S. regional
-  active transport using the ITHIM model}
-\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
-  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
-  $\ldots$
-  \ and Jonathan Patz$^{1}$}
-\address{$^{1}$Global Health Institute\\
-University of Wisconsin{\textendash}Madison, Madison, WI USA\\
-}
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
-
-\editor{Associate Editor: XXXXXXX}
-
-\maketitle
-
-\begin{abstract}
-
-\section{Summary:}
-
-While increased active transportation, primarily cycling and walking,
-is often held up as a goal for improved public health --- via reduced
-chronic disease, improved air quality, and reductions in greenhouse
-gas emissions — obtaining estimates for the quantiative impacts of
-such shifts remains infrequent, if not altogether absent from related
-policy action. Using U.S. metropolitan counties from seven states we
-perform a comparative risk analysis between each regions and national
-averages as the baseline.  We employ a portion of the Integrated
-Transport and Health Impacts Model (ITHIM) to investigate the reduced
-chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
-levels of physical activity with population shifts in the amount of
-biking and walking done for transport. The analysis is informed by
-regional travel parameters from the National Household Transportation
-Survey, non-travel activity metrics from the American Time Use Survey,
-and baseline disease burden from the Centers for Disease Control and
-Prevention's WONDER database and the Global Burden of Disease dataset.
-Our findings show that the California and New York State have travel
-patterns that, if mirrored by the nation, would save the U.S. upwards
-of fifty thousand disability adjusted life years.
-
-%% \section{Availability:}
-
-%% %% \package{} may be explored using the user interface avaiable
-%% %% at \webpage{}.
-%% The source code and data files for \package{} are publicly available
-%% on GitHub via a link on \webpage{}.
-
-\section{Supplementary information:}
-
-See \webpage{} for links to the source code, manual, tutorial and
-links for more information about the ITHIM model.
-
-\section{Contact:}
-\href{javargo@wisc.edu}{javargo@wisc.edu}
-
-\end{abstract}
-
-\section{Introduction}
-
-Regular exercise has been shown to provide health benefits in terms of
-chronic disease prevention \cite{warburton2006}, cognitive function
-\cite{hillman2008}, and overall well being and satisfaction with life
-\cite{maher2013}. Still only half of Americans obtain the recommended
-levels of aerobic activity, which is 2.5 hours per week. There is
-considerable variability in the attainment of this goal by location
-within the country. Supportive local built and policy environments are
-important determinants of population-level activity
-patterns. Particularly, as they facilitate and foster routine moderate
-physical activity, such as walking, local decissions on design and
-policy can play a huge role in the Nation's health.
-
-To better understand the potential for attainable shifts in
-transportation behavior across the US, we employed a comparative risk
-assessment methodology to compare national averages for time spent
-walking and cycling for transport to those of specific states. Our
-estimates can provide an important evidence base to encourage new
-construction, retrofits, land use policy, and programatic funding to
-support active transportation. Particularly in places with poor
-attainmnet of the physical activity guidelines, such investments can
-be importnat pieces of lareger population health interventions that
-have effective and long-lasting impacts.
-
-ITHIM is a statistical model that integrates data on travel patterns,
-physical activity, fine particulate matter, GHG emissions, and disease
-and injuries. Based on population and travel scenarios. The model has
-been used to calculate the health impacts of walking and bicycling
-short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013,maizlish2013}.
-
-The ITHIM model uses comparative risk assessment through which it
-formulates a change in the disease burden, resulting from the shift in
-the exposure distribution from a baseline scenario to an alternative
-scenario.
-
-\begin{methods}
-
-\section{Methods}
-
-The ITHIM model uses the comparative risk assesment framework to
-estimate the population attributable fraction, $\af$.  In our case the
-$\af$ represents the proportional change in national disease burden if
-age and sex-specific active transport time means were changed.
-
-\begin{equation}
-\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
-\end{equation}
-
-Here $R(x)$ represents the risk when the individual has exposure $x$,
-physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
-density of individuals with exposure $x$, in the baseline and
-alternate scenarios, respectively.  The exposure variable, $x$, is
-modeled as the sum of two independent random variables, travel-related
-exposure and non-travel-related exposure.  Non-travel-related exposure
-is modeled with a \logNormal{} distribution with the age-sex specific
-mean ratios for non-travel-related exposure are estimated from the
-ATUS \cite{ATUS}.  The referent group mean is treated as a variable
-and the coefficient of variation is treated as constant across age-sex
-strata and estimated with ATUS.  We see in Figure \ref{dalyFigure}
-that varying the mean values for non-travel-related exposure does not
-greatly affect the estimate for disease burden.  The distribution for
-travel-related exposure is estimated from the time spent walking or
-cycling, i.e., active transport time, which as in the original ITHIM
-model has a \logNormal{} distribution with constant coefficient of
-variation across age-sex strata \cite{}.  Travel-related exposure
-is estimated from active transport time using assumptions about how
-much physical activity is required for cycling and walking.
-
-\end{methods}
-
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{}\label{dalyFigure}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
-    \caption{Mean travel time by mode (walking/cycling) for each of
-      seven states.  National means are displayed on the
-      right.}\label{meanMatrices}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
-    \caption{}\label{California}
-\end{figure}
-
-\section{Results}
+% % Place tables after the first paragraph in which they are cited.
+% \begin{table}[!ht]
+% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+% \centering
+% \caption{
+% {\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
+% \begin{tabular}{|l+l|l|l|l|l|l|l|}
+% \hline
+% \multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
+% $cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
+% $cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
+% $cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
+% \end{tabular}
+% \begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+% \end{flushleft}
+% \label{table1}
+% \end{adjustwidth}
+% \end{table}
 
@@
 is due to prevention of dementia in the oldest age group, $80+$ years
-old.  California exhibits much greater walking and cycling means in this age group
-than the national average. In particular we see the cycling mean We also see prevention of depression and
-diabetes in the middle age groups due to increased active transport
-time.
-
-\section*{Acknowledgment}
-\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots
-
-%\bibliographystyle{natbib}
-\bibliographystyle{bmc_article}
-%\bibliographystyle{mypapers}
-\bibliography{ITHIM}
-\end{document}
+old.  California exhibits much greater walking and cycling means in
+this age group than the national average. In particular we see the
+cycling mean We also see prevention of depression and diabetes in the
+middle age groups due to increased active transport time.

### 2016-10-10 methodsPaperOld/tex/abstract.tex
Switched to PlosOne template.

--- 2016-10-05 manuscript/manuscript.tex
@@
-\documentclass{bioinfo}
-\usepackage{url}
-\copyrightyear{2016}
-\pubyear{2016}
-
-\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
-\newcommand{\bi}{\begin{itemize}}
-\newcommand{\ei}{\end{itemize}}
-\newcommand{\package}{\emph{ITHIM}}
-\newcommand{\ithim}{\emph{ITHIM}}
-\newcommand{\R}{\emph{R}}
-\newcommand{\sgy}[1]{{\itshape\col{#1}}}
-\newcommand{\XX}{\sgy{XX}}
-\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
-\newcommand{\af}{\textrm{PAF}}
-\newcommand{\mets}{MET-hrs./week}
-\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
-\newcommand{\vmt}{$\mathrm{VMT}$}
-\newcommand{\logNormal}{log-normal}
-
-\begin{document}
-\firstpage{1}
-
-\title[\package{}]{A comparative risk assesment of U.S. regional
-  active transport using the ITHIM model}
-\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
-  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
-  $\ldots$
-  \ and Jonathan Patz$^{1}$}
-\address{$^{1}$Global Health Institute\\
-University of Wisconsin{\textendash}Madison, Madison, WI USA\\
-}
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
-
-\editor{Associate Editor: XXXXXXX}
-
-\maketitle
-
-\begin{abstract}
-
 \section{Summary:}
@@
 \href{javargo@wisc.edu}{javargo@wisc.edu}
-
-\end{abstract}
-
-\section{Introduction}
-
-Regular exercise has been shown to provide health benefits in terms of
-chronic disease prevention \cite{warburton2006}, cognitive function
-\cite{hillman2008}, and overall well being and satisfaction with life
-\cite{maher2013}. Still only half of Americans obtain the recommended
-levels of aerobic activity, which is 2.5 hours per week. There is
-considerable variability in the attainment of this goal by location
-within the country. Supportive local built and policy environments are
-important determinants of population-level activity
-patterns. Particularly, as they facilitate and foster routine moderate
-physical activity, such as walking, local decissions on design and
-policy can play a huge role in the Nation's health.
-
-To better understand the potential for attainable shifts in
-transportation behavior across the US, we employed a comparative risk
-assessment methodology to compare national averages for time spent
-walking and cycling for transport to those of specific states. Our
-estimates can provide an important evidence base to encourage new
-construction, retrofits, land use policy, and programatic funding to
-support active transportation. Particularly in places with poor
-attainmnet of the physical activity guidelines, such investments can
-be importnat pieces of lareger population health interventions that
-have effective and long-lasting impacts.
-
-ITHIM is a statistical model that integrates data on travel patterns,
-physical activity, fine particulate matter, GHG emissions, and disease
-and injuries. Based on population and travel scenarios. The model has
-been used to calculate the health impacts of walking and bicycling
-short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013,maizlish2013}.
-
-The ITHIM model uses comparative risk assessment through which it
-formulates a change in the disease burden, resulting from the shift in
-the exposure distribution from a baseline scenario to an alternative
-scenario.
-
-\begin{methods}
-
-\section{Methods}
-
-The ITHIM model uses the comparative risk assesment framework to
-estimate the population attributable fraction, $\af$.  In our case the
-$\af$ represents the proportional change in national disease burden if
-age and sex-specific active transport time means were changed.
-
-\begin{equation}
-\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
-\end{equation}
-
-Here $R(x)$ represents the risk when the individual has exposure $x$,
-physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
-density of individuals with exposure $x$, in the baseline and
-alternate scenarios, respectively.  The exposure variable, $x$, is
-modeled as the sum of two independent random variables, travel-related
-exposure and non-travel-related exposure.  Non-travel-related exposure
-is modeled with a \logNormal{} distribution with the age-sex specific
-mean ratios for non-travel-related exposure are estimated from the
-ATUS \cite{ATUS}.  The referent group mean is treated as a variable
-and the coefficient of variation is treated as constant across age-sex
-strata and estimated with ATUS.  We see in Figure \ref{dalyFigure}
-that varying the mean values for non-travel-related exposure does not
-greatly affect the estimate for disease burden.  The distribution for
-travel-related exposure is estimated from the time spent walking or
-cycling, i.e., active transport time, which as in the original ITHIM
-model has a \logNormal{} distribution with constant coefficient of
-variation across age-sex strata \cite{}.  Travel-related exposure
-is estimated from active transport time using assumptions about how
-much physical activity is required for cycling and walking.
-
-\end{methods}
-
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{}\label{dalyFigure}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
-    \caption{Mean travel time by mode (walking/cycling) for each of
-      seven states.  National means are displayed on the
-      right.}\label{meanMatrices}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
-    \caption{}\label{California}
-\end{figure}
-
-\section{Results}
-
-We used the National Household Transportation Survey to estimate the
-mean walk and cycle times for individuals that live in metropolitan
-areas for each of seven states (CA, TX, NY, FL, VA, NC and AZ)
-\cite{NHTS}.  These states were selected becaused they had sufficient
-sample size for reasonable estimates of age-sex specific
-travel-related walking and cycling means, Figure \ref{meanMatrices}.
-
-Of these seven states one was found to show active transport behavior
-that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
-Values for national disease-specific DALYs were found using the CDC
-Wonder database \cite{CDCWonder}.  California stood out among the
-estimates for national DALYs averted.  Most of the decrease in DALYs
-is due to prevention of dementia in the oldest age group, $80+$ years
-old.  California exhibits much greater walking and cycling means in this age group
-than the national average. In particular we see the cycling mean We also see prevention of depression and
-diabetes in the middle age groups due to increased active transport
-time.
-
-\section*{Acknowledgment}
-\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots
-
-%\bibliographystyle{natbib}
-\bibliographystyle{bmc_article}
-%\bibliographystyle{mypapers}
-\bibliography{ITHIM}
-\end{document}

### 2016-10-10 methodsPaper/tex/materialsAndMethods.tex
Subdivided paper #1 with Plos One template

--- 2016-10-05 manuscript/manuscript.tex
@@
-\documentclass{bioinfo}
-\usepackage{url}
-\copyrightyear{2016}
-\pubyear{2016}
-
-\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
-\newcommand{\bi}{\begin{itemize}}
-\newcommand{\ei}{\end{itemize}}
-\newcommand{\package}{\emph{ITHIM}}
-\newcommand{\ithim}{\emph{ITHIM}}
-\newcommand{\R}{\emph{R}}
-\newcommand{\sgy}[1]{{\itshape\col{#1}}}
-\newcommand{\XX}{\sgy{XX}}
-\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
-\newcommand{\af}{\mathrm{PAF}}
-\newcommand{\mets}{MET-hrs./week}
-\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
-\newcommand{\vmt}{$\mathrm{VMT}$}
-\newcommand{\logNormal}{log-normal}
-
-\begin{document}
-\firstpage{1}
-
-\title[\package{}]{A comparative risk assesment of U.S. regional
-  active transport using the ITHIM model}
-\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
-  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
-  $\ldots$
-  \ and Jonathan Patz$^{1}$}
-\address{$^{1}$Global Health Institute\\
-University of Wisconsin{\textendash}Madison, Madison, WI USA\\
-}
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
-
-\editor{Associate Editor: XXXXXXX}
-
-\maketitle
-
-\begin{abstract}
-
-\section{Summary:}
-
-While increased active transportation, primarily cycling and walking,
-is often held up as a goal for improved public health --- via reduced
-chronic disease, improved air quality, and reductions in greenhouse
-gas emissions — obtaining estimates for the quantiative impacts of
-such shifts remains infrequent, if not altogether absent from related
-policy action. Using U.S. metropolitan counties from seven states we
-perform a comparative risk analysis between each regions and national
-averages as the baseline.  We employ a portion of the Integrated
-Transport and Health Impacts Model (ITHIM) to investigate the reduced
-chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
-levels of physical activity with population shifts in the amount of
-biking and walking done for transport. The analysis is informed by
-regional travel parameters from the National Household Transportation
-Survey, non-travel activity metrics from the American Time Use Survey,
-and baseline disease burden from the Centers for Disease Control and
-Prevention's WONDER database and the Global Burden of Disease dataset.
-Our findings show that the California and New York State have travel
-patterns that, if mirrored by the nation, would save the U.S. upwards
-of fifty thousand disability adjusted life years.
-
-%% \section{Availability:}
-
-%% %% \package{} may be explored using the user interface avaiable
-%% %% at \webpage{}.
-%% The source code and data files for \package{} are publicly available
-%% on GitHub via a link on \webpage{}.
-
-\section{Supplementary information:}
-
-See \webpage{} for links to the source code, manual, tutorial and
-links for more information about the ITHIM model.
-
-\section{Contact:}
-\href{javargo@wisc.edu}{javargo@wisc.edu}
-
-\end{abstract}
-
-\section{Introduction}
-
-Regular exercise has been shown to provide health benefits in terms of
-chronic disease prevention \cite{warburton2006}, cognitive function
-\cite{hillman2008}, and overall well being and satisfaction with life
-\cite{maher2013}. Still only half of Americans obtain the recommended
-levels of aerobic activity, which is 2.5 hours per week. There is
-considerable variability in the attainment of this goal by location
-within the country. Supportive local built and policy environments are
-important determinants of population-level activity
-patterns. Particularly, as they facilitate and foster routine moderate
-physical activity, such as walking, local decissions on design and
-policy can play a huge role in the Nation's health.
-
-To better understand the potential for attainable shifts in
-transportation behavior across the US, we employed a comparative risk
-assessment methodology to compare national averages for time spent
-walking and cycling for transport to those of specific states. Our
-estimates can provide an important evidence base to encourage new
-construction, retrofits, land use policy, and programatic funding to
-support active transportation. Particularly in places with poor
-attainmnet of the physical activity guidelines, such investments can
-be importnat pieces of lareger population health interventions that
-have effective and long-lasting impacts.
-
-ITHIM is a statistical model that integrates data on travel patterns,
-physical activity, fine particulate matter, GHG emissions, and disease
-and injuries. Based on population and travel scenarios. The model has
-been used to calculate the health impacts of walking and bicycling
-short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013,maizlish2013}.
-
-The ITHIM model uses comparative risk assessment through which it
-formulates a change in the disease burden, resulting from the shift in
-the exposure distribution from a baseline scenario to an alternative
-scenario.
-
-\begin{methods}
-
-\section{Methods}
-
 The ITHIM model uses the comparative risk assesment framework to
@@
 activity is required for cycling and walking.
-
-\end{methods}
-
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{}\label{dalyFigure}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
-    \caption{Mean travel time by mode (walking/cycling) for each of
-      seven states.  National means are displayed on the
-      right.}\label{meanMatrices}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
-    \caption{}\label{California}
-\end{figure}
-
-\section{Results}
-
-We used the National Household Transportation Survey to estimate the
-mean walk and cycle times for individuals that live in metropolitan
-areas for each of seven states (CA, TX, NY, FL, VA, NC and AZ)
-\cite{NHTS}.  These states were selected becaused they had sufficient
-sample size for reasonable estimates of age-sex specific
-travel-related walking and cycling means, Figure \ref{meanMatrices}.
-
-Of these seven states one was found to show active transport behavior
-that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
-Values for national disease-specific DALYs were found using the CDC
-Wonder database \cite{CDCWonder}.  California stood out among the
-estimates for national DALYs averted.  Most of the decrease in DALYs
-is due to prevention of dementia in the oldest age group, $80+$ years
-old.  California exhibits much greater walking and cycling means in
-this age group than the national average. In particular we see the
-cycling mean We also see prevention of depression and diabetes in the
-middle age groups due to increased active transport time.
-
-\section*{Acknowledgment}
-\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots
-
-%\bibliographystyle{natbib}
-\bibliographystyle{bmc_article}
-%\bibliographystyle{mypapers}
-\bibliography{ITHIM}
-\end{document}

### 2016-10-11 secondPaper/tex/introduction.tex
Updates to Makefiles and figure links

--- 2016-10-05 manuscript/manuscript.tex
@@
-\documentclass{bioinfo}
-\usepackage{url}
-\copyrightyear{2016}
-\pubyear{2016}
-
-\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
-\newcommand{\bi}{\begin{itemize}}
-\newcommand{\ei}{\end{itemize}}
-\newcommand{\package}{\emph{ITHIM}}
-\newcommand{\ithim}{\emph{ITHIM}}
-\newcommand{\R}{\emph{R}}
-\newcommand{\sgy}[1]{{\itshape\col{#1}}}
-\newcommand{\XX}{\sgy{XX}}
-\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
-\newcommand{\af}{\textrm{PAF}}
-\newcommand{\mets}{MET-hrs./week}
-\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
-\newcommand{\vmt}{$\mathrm{VMT}$}
-\newcommand{\logNormal}{log-normal}
-
-\begin{document}
-\firstpage{1}
-
-\title[\package{}]{A comparative risk assesment of U.S. regional
-  active transport using the ITHIM model}
-\author[Younkin \textit{et~al}]{Samuel G. Younkin$^{1}$,
-  Jason Vargo$^{1}$\footnote{to whom correspondence should be addressed},
-  $\ldots$
-  \ and Jonathan Patz$^{1}$}
-\address{$^{1}$Global Health Institute\\
-University of Wisconsin{\textendash}Madison, Madison, WI USA\\
-}
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
-
-\editor{Associate Editor: XXXXXXX}
-
-\maketitle
-
-\begin{abstract}
-
-\section{Summary:}
-
-While increased active transportation, primarily cycling and walking,
-is often held up as a goal for improved public health --- via reduced
-chronic disease, improved air quality, and reductions in greenhouse
-gas emissions — obtaining estimates for the quantiative impacts of
-such shifts remains infrequent, if not altogether absent from related
-policy action. Using U.S. metropolitan counties from seven states we
-perform a comparative risk analysis between each regions and national
-averages as the baseline.  We employ a portion of the Integrated
-Transport and Health Impacts Model (ITHIM) to investigate the reduced
-chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
-levels of physical activity with population shifts in the amount of
-biking and walking done for transport. The analysis is informed by
-regional travel parameters from the National Household Transportation
-Survey, non-travel activity metrics from the American Time Use Survey,
-and baseline disease burden from the Centers for Disease Control and
-Prevention's WONDER database and the Global Burden of Disease dataset.
-Our findings show that the California and New York State have travel
-patterns that, if mirrored by the nation, would save the U.S. upwards
-of fifty thousand disability adjusted life years.
-
-%% \section{Availability:}
-
-%% %% \package{} may be explored using the user interface avaiable
-%% %% at \webpage{}.
-%% The source code and data files for \package{} are publicly available
-%% on GitHub via a link on \webpage{}.
-
-\section{Supplementary information:}
-
-See \webpage{} for links to the source code, manual, tutorial and
-links for more information about the ITHIM model.
-
-\section{Contact:}
-\href{javargo@wisc.edu}{javargo@wisc.edu}
-
-\end{abstract}
-
-\section{Introduction}
-
 Regular exercise has been shown to provide health benefits in terms of
@@
 scenario.
-
-\begin{methods}
-
-\section{Methods}
-
-The ITHIM model uses the comparative risk assesment framework to
-estimate the population attributable fraction, $\af$.  In our case the
-$\af$ represents the proportional change in national disease burden if
-age and sex-specific active transport time means were changed.
-
-\begin{equation}
-\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
-\end{equation}
-
-Here $R(x)$ represents the risk when the individual has exposure $x$,
-physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
-density of individuals with exposure $x$, in the baseline and
-alternate scenarios, respectively.  The exposure variable, $x$, is
-modeled as the sum of two independent random variables, travel-related
-exposure and non-travel-related exposure.  Non-travel-related exposure
-is modeled with a \logNormal{} distribution with the age-sex specific
-mean ratios for non-travel-related exposure are estimated from the
-ATUS \cite{ATUS}.  The referent group mean is treated as a variable
-and the coefficient of variation is treated as constant across age-sex
-strata and estimated with ATUS.  We see in Figure \ref{dalyFigure}
-that varying the mean values for non-travel-related exposure does not
-greatly affect the estimate for disease burden.  The distribution for
-travel-related exposure is estimated from the time spent walking or
-cycling, i.e., active transport time, which as in the original ITHIM
-model has a \logNormal{} distribution with constant coefficient of
-variation across age-sex strata \cite{}.  Travel-related exposure
-is estimated from active transport time using assumptions about how
-much physical activity is required for cycling and walking.
-
-\end{methods}
-
-\begin{figure}[t]
-    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig2}}
-    \caption{}\label{dalyFigure}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
-    \caption{Mean travel time by mode (walking/cycling) for each of
-      seven states.  National means are displayed on the
-      right.}\label{meanMatrices}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
-    \caption{}\label{California}
-\end{figure}
-
-\section{Results}
-
-We used the National Household Transportation Survey to estimate the
-mean walk and cycle times for individuals that live in metropolitan
-areas for each of seven states (CA, TX, NY, FL, VA, NC and AZ)
-\cite{NHTS}.  These states were selected becaused they had sufficient
-sample size for reasonable estimates of age-sex specific
-travel-related walking and cycling means, Figure \ref{meanMatrices}.
-
-Of these seven states one was found to show active transport behavior
-that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
-Values for national disease-specific DALYs were found using the CDC
-Wonder database \cite{CDCWonder}.  California stood out among the
-estimates for national DALYs averted.  Most of the decrease in DALYs
-is due to prevention of dementia in the oldest age group, $80+$ years
-old.  California exhibits much greater walking and cycling means in this age group
-than the national average. In particular we see the cycling mean We also see prevention of depression and
-diabetes in the middle age groups due to increased active transport
-time.
-
-\section*{Acknowledgment}
-\paragraph{Funding\textcolon} SGY is supported by \ldots,\ JV, JP by \ldots
-
-%\bibliographystyle{natbib}
-\bibliographystyle{bmc_article}
-%\bibliographystyle{mypapers}
-\bibliography{ITHIM}
-\end{document}

### 2016-10-11 R/ITHIM.Rmd
Update to methods paper

@@
 library("magrittr")
-install("~/ITHIM/")
+#install("~/ITHIM/")
+install_github("syounkin/ITHIM", ref="devel")
 library("ITHIM")
@@
 ```{r compute, echo = FALSE, eval = TRUE}
-compute <- TRUE
+compute <- FALSE
 ```
@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-upper <- 25; length <- 10; wd <- upper/length;
+upper <- 10; length <- 5; wd <- upper/length;
 
@@
 p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
+
+```
+
+Do it again, but with quintiles
+
+```{r ITHIMObjsQuintiles, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+DALYsTotal2 <- data.frame()
+ITHIMList <- list()
+ITHIM.default <- createITHIM(vision = "baseline", region = "national")
+ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
+ITHIM.default <- updateITHIM(ITHIM.default, "cv", 3.0288)
+ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )
+
+ITHIM.national <- ITHIM.default
+ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+
+for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+    ITHIMList[[HHSTATE]] <- ITHIM.default
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
+  }
+
+for( muNonTravel in seq(0, upper, by = wd)){
+
+  rm(DALYs)
+
+  for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
+  }
+
+  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+
+  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
+  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
+  DALYsTotal2 <- rbind(DALYsTotal2, DALYs)
+}
+
+DALYsTotal2 <- within(DALYsTotal2, {
+  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA"))
+})
+
+DALYsTotal <- rbind(data.frame(DALYsTotal,quints="Percentiles"), data.frame(DALYsTotal2,quints="Quintiles"))
+
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( .
+~ quints)
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( .
+~ quints)
 
@@
 
-```{r roadInjuries, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r roadInjuries, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
 convertState <- function( stateCode ){

### 2016-10-11 methodsPaper/methodsPaper.tex
Update to methods paper

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}
% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}
% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}
% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}
% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}
% line numbers
\usepackage[right]{lineno}
% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }
% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}
% array package and thick rules for tables
\usepackage{array}
% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}
% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}
% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
\usepackage{setspace} 
\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

% added by sgy
\usepackage{url}
\input{../tex/sgy}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{A Comparative Risk Assessment of Active Transportation in the United
States} % Please use "title case" (capitalize all terms in the title except conjunctions, prepositions, and articles).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Samuel Younkin\textsuperscript{1},\ %,2\Yinyang},
Jason Vargo\textsuperscript{1},\ %\Yinyang},
Neil Maizlish\textsuperscript{2},\ %\Yinyang},
Jonathan Patz\textsuperscript{1}%,3\textcurrency},
%Name4 Surname\textsuperscript{2},
%#Name5 Surname\textsuperscript{2\ddag},
%Name6 Surname\textsuperscript{2\ddag},
%Name7 Surname\textsuperscript{1,2,3*},
%with the Global Health Institute\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Global Health Institute, University of Wisconsin{\textendash}Madison, Madison, WI, USA
\\
\textbf{2} Berkeley, CA, USA
\\
%\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
%\\
%\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
%\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
% \ddag These authors also contributed equally to this work.

% Current address notes
% \textcurrency Current Address: GHI UW{\textendash}Madison, 1300
% University Ave., Madison, WI, 53706, USA
% WI, USA % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
%* correspondingauthor@institute.edu

% Please keep the abstract below 300 words
% Use "Eq" instead of "Equation" for equation citations.

% Results and Discussion can be combined.

\end{flushleft}

\section*{Abstract}
\input{tex/abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
% \section*{Author Summary}
% \input{tex/authorSummary}

\linenumbers

\section*{Introduction}
\input{tex/introduction.tex}

\section*{Materials and Methods}
\input{tex/materialsAndMethods}

\section*{Results}

\begin{figure}[t]
  \centerline{\includegraphics[width=\textwidth]{./figures/fig1}}
    \caption{}\label{dalyFigureQuintiles}
\end{figure}

%% \begin{figure}[t]

%%     \centerline{\includegraphics[width=\textwidth]{./figures/fig2}}
%%     \caption{}\label{dalyFigure}
%% \end{figure}

%% \begin{figure}[t]
%%   \centerline{\includegraphics[width=\textwidth]{./figures/fig4.pdf}}
%%     \caption{}\label{California}
%% \end{figure}

\input{tex/results}

\section*{Discussion}
\input{tex/discussion}

\section*{Conclusion}
\input{tex/conclusion}

%\section*{Supporting Information}

\input{tex/supportingInformation}

\section*{Acknowledgments}

\nolinenumbers

\bibliography{../tex/ITHIM}

\end{document}

% Template for PLoS
% Version 3.3 June 2016
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

% \subsection*{Etiam eget sapien nibh.}

% % For figure citations, please use "Fig" instead of "Figure".
% Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% % Place figure captions after the first paragraph in which they are cited.
% \begin{figure}[!h]
% \caption{{\bf Bold the figure title.}
% Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
% \label{fig1}
% \end{figure}

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here.
% 
% \begin{thebibliography}{10}

% % \bibitem{bib1}
% % Conant GC, Wolfe KH.
% % \newblock {{T}urning a hobby into a job: how duplicated genes find new
% %   functions}.
% % \newblock Nat Rev Genet. 2008 Dec;9(12):938--950.

% % \bibitem{bib2}
% % Ohno S.
% % \newblock Evolution by gene duplication.
% % \newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
% %   Springer-Verlag.; 1970.

% % \bibitem{bib3}
% % Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
% % \newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
% %   infection through a transposon insertion followed by a {D}uplication}.
% % \newblock PLoS Genet. 2011 Oct;7(10):e1002337.
% \input{../tex/ITHIM}
% \end{thebibliography}

%% Manuscript Organization

%% Manuscripts should be organized as follows. Instructions for each element appear below the list.
%% Beginning section

%% The following elements are required, in order:

%% Title page: List title, authors, and affiliations as first page of manuscript
%% Abstract
%% Introduction

%% Middle section

%% The following elements can be renamed as needed and presented in any order:

%% Materials and Methods
%% Results
%% Discussion
%% Conclusions (optional)

%% Ending section

%% The following elements are required, in order:

%% Acknowledgments
%% References
%% Supporting information captions (if applicable)

%% Other elements

%% Figure captions are inserted immediately after the first paragraph in which the figure is cited. Figure files are uploaded separately.
%% Tables are inserted immediately after the first paragraph in which they are cited.
%% Supporting information files are uploaded separately.

### 2016-10-11 methodsPaper/tex/abstract.tex
chmod -R 755 .
Update to methods paper

--- 2016-10-10 methodsPaperOld/tex/abstract.tex
@@
-\section{Summary:}
+%% While increased active transportation, primarily cycling and walking,
+%% is often held up as a goal for improved public health --- via reduced
+%% chronic disease, improved air quality, and reductions in greenhouse
+%% gas emissions — obtaining estimates for the quantiative impacts of
+%% such shifts remains infrequent, if not altogether absent from related
+%% policy action. Using U.S. metropolitan counties from seven states we
+%% perform a comparative risk analysis between each regions and national
+%% averages as the baseline.  We employ a portion of the Integrated
+%% Transport and Health Impacts Model (ITHIM) to investigate the reduced
+%% chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
+%% levels of physical activity with population shifts in the amount of
+%% biking and walking done for transport. The analysis is informed by
+%% regional travel parameters from the National Household Transportation
+%% Survey, non-travel activity metrics from the American Time Use Survey,
+%% and baseline disease burden from the Centers for Disease Control and
+%% Prevention's WONDER database and the Global Burden of Disease dataset.
+%% Our findings show that the California and New York State have travel
+%% patterns that, if mirrored by the nation, would save the U.S. upwards
+%% of fifty thousand disability adjusted life years.
 
-While increased active transportation, primarily cycling and walking,
-is often held up as a goal for improved public health --- via reduced
-chronic disease, improved air quality, and reductions in greenhouse
-gas emissions — obtaining estimates for the quantiative impacts of
-such shifts remains infrequent, if not altogether absent from related
-policy action. Using U.S. metropolitan counties from seven states we
-perform a comparative risk analysis between each regions and national
-averages as the baseline.  We employ a portion of the Integrated
-Transport and Health Impacts Model (ITHIM) to investigate the reduced
-chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
-levels of physical activity with population shifts in the amount of
-biking and walking done for transport. The analysis is informed by
-regional travel parameters from the National Household Transportation
-Survey, non-travel activity metrics from the American Time Use Survey,
-and baseline disease burden from the Centers for Disease Control and
-Prevention's WONDER database and the Global Burden of Disease dataset.
-Our findings show that the California and New York State have travel
-patterns that, if mirrored by the nation, would save the U.S. upwards
-of fifty thousand disability adjusted life years.
-
+%\subsection{Summary:}
 %% \section{Availability:}
@@
 
-\section{Supplementary information:}
+% \section{Supplementary information:}
 
-See \webpage{} for links to the source code, manual, tutorial and
-links for more information about the ITHIM model.
+% See \webpage{} for links to the source code, manual, tutorial and
+% links for more information about the ITHIM model.
 
-\section{Contact:}
-\href{javargo@wisc.edu}{javargo@wisc.edu}
+% \section{Contact:}
+% \href{javargo@wisc.edu}{javargo@wisc.edu}

### 2016-10-12 secondPaper/tex/introduction.tex

@@
 attainmnet of the physical activity guidelines, such investments can
-be importnat pieces of lareger population health interventions that
+be important pieces of larger population health interventions that
 have effective and long-lasting impacts.

### 2016-10-12 R/ITHIM.Rmd
compute <- TRUE
First attempt at new analysis
Changed the way ITHIM.default is treated.

@@
 ```{r subset, eval = compute, fig.width = 10, echo = FALSE}
-NHTS.person <- subset( NHTS.person, MSA == TRUE ) # & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
+NHTS.person <- subset( NHTS.person, MSA == TRUE )
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+stateVec <- c("CA","TX","NY","FL") #c("CA","TX","NY","FL","VA","NC","AZ")
+stateVecOrdered <- c("TX","FL","NY","CA") #c("NC","TX","VA","FL","AZ","NY","CA")
+
 upper <- 10; length <- 5; wd <- upper/length;
@@
 ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
-ITHIM.default <- updateITHIM(ITHIM.default, "cv", 3.0288)
+ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
 ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
@@
 
-for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+for (HHSTATE in stateVec){
     ITHIMList[[HHSTATE]] <- ITHIM.default
@@
 
-  for (HHSTATE in c("CA","TX","NY","FL","VA","NC","AZ")){
+  for (HHSTATE in stateVec){
     ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
@@
 DALYsTotal <- within(DALYsTotal, {
-  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA"))
+  State <- factor(State, levels = stateVecOrdered)
 })
@@
 ITHIMList <- list()
-ITHIM.default <- createITHIM(vision = "baseline", region = "national")
-ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
-ITHIM.default <- updateITHIM(ITHIM.default, "cv", 3.0288)
+
+#ITHIM.default <- createITHIM(vision = "baseline", region = "national")
+#ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
+#ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
+
 ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )
@@
 DALYsTotal2 <- within(DALYsTotal2, {
-  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA"))
+  State <- factor(State, levels = stateVecOrdered)
 })
@@
 
+```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
+
+ITHIM.national <- ITHIM.default
+ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+
+```
+# Appendix
+
 ```{r roadInjuries, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
@@
 ```
-
-# Appendix
 

### 2016-10-12 methodsPaper/tex/introduction.tex
Tested pucher2010 reference in methods paper
Seems like it worked.

--- 2016-10-12 secondPaper/tex/introduction.tex
@@
 
-To better understand the potential for attainable shifts in
-transportation behavior across the US, we employed a comparative risk
-assessment methodology to compare national averages for time spent
-walking and cycling for transport to those of specific states. Our
-estimates can provide an important evidence base to encourage new
-construction, retrofits, land use policy, and programatic funding to
-support active transportation. Particularly in places with poor
-attainmnet of the physical activity guidelines, such investments can
-be important pieces of larger population health interventions that
-have effective and long-lasting impacts.
+%% To better understand the potential for attainable shifts in
+%% transportation behavior across the US, we employed a comparative risk
+%% assessment methodology to compare national averages for time spent
+%% walking and cycling for transport to those of specific states. Our
+%% estimates can provide an important evidence base to encourage new
+%% construction, retrofits, land use policy, and programatic funding to
+%% support active transportation. Particularly in places with poor
+%% attainmnet of the physical activity guidelines, such investments can
+%% be importnat pieces of lareger population health interventions that
+%% have effective and long-lasting impacts.
 
-ITHIM is a statistical model that integrates data on travel patterns,
-physical activity, fine particulate matter, GHG emissions, and disease
-and injuries. Based on population and travel scenarios. The model has
-been used to calculate the health impacts of walking and bicycling
-short distances usually traveled by car or driving low-emission
-automobiles \cite{woodcock2013,maizlish2013}.
+ITHIM is a statistical model that integrates data on active transport,
+physical activity, fine particulate matter and greenhouse gas
+emissions to provide estimates for the proportional change in
+mortality and morbidity for given baseline and alternate travel
+scenarios. The model has been used to calculate the health impacts of
+walking and bicycling short distances usually traveled by car or
+driving low-emission automobiles \cite{woodcock2013,maizlish2013}.
 
-The ITHIM model uses comparative risk assessment through which it
-formulates a change in the disease burden, resulting from the shift in
-the exposure distribution from a baseline scenario to an alternative
-scenario.
+ITHIM uses a comparative risk assessment framework for the active
+transport component of the model.  We improve on the existing
+implementation of the active transport component by including the
+distribution of non-travel-related activity, improving numerical
+precision when computing the population attributable fraction and
+creating a simple user-interface for the custom-built \R{}
+package \package{}.
+
+Dummy ref \cite{pucher2010}.

### 2016-10-12 methodsPaper/methodsPaper.tex
Small updates to manuscripts
Still trying to separate the content into two papers.  I need more material...
added affiliation to me and JP

@@
 
-% Title must be 250 characters or less.
+% Titlemust be 250 characters or less.
 \begin{flushleft}
@@
 Samuel Younkin\textsuperscript{1},\ %,2\Yinyang},
-Jason Vargo\textsuperscript{1},\ %\Yinyang},
-Neil Maizlish\textsuperscript{2},\ %\Yinyang},
-Jonathan Patz\textsuperscript{1}%,3\textcurrency},
+Jason Vargo\textsuperscript{1, 2},\ %\Yinyang},
+Neil Maizlish\textsuperscript{3},\ %\Yinyang},
+Jonathan Patz\textsuperscript{1, 2}%,3\textcurrency},
 %Name4 Surname\textsuperscript{2},
@@
 \\
-\textbf{2} Berkeley, CA, USA
+\textbf{2} Nelson Institute for Environmental Studies Center for Sustainability and the Global Environment, University of Wisconsin{\textendash}Madison, Madison, WI, USA
+\\
+\textbf{3} Berkeley, CA, USA
 \\

### 2016-10-12 methodsPaper/methodsPaper.tex
Update methodsPaper.tex

@@
 Samuel Younkin\textsuperscript{1},\ %,2\Yinyang},
-Jason Vargo\textsuperscript{1, 2},\ %\Yinyang},
+Jason Vargo\textsuperscript{1, 2}\ %\Yinyang},
 Neil Maizlish\textsuperscript{3},\ %\Yinyang},

### 2016-10-12 methodsPaper/tex/introduction.tex
Small updates to manuscripts
Still trying to separate the content into two papers.  I need more material...
added paragraphs

@@
 important determinants of population-level activity
-patterns. Particularly, as they facilitate and foster routine moderate
-physical activity, such as walking, local decissions on design and
-policy can play a huge role in the Nation's health.
+patterns. Particularly, as they facilitate and foster routine and moderate
+physical activity, such as walking, local decisions on design and
+policy can play a huge role in improving the nation's health.
 
-%% To better understand the potential for attainable shifts in
-%% transportation behavior across the US, we employed a comparative risk
-%% assessment methodology to compare national averages for time spent
-%% walking and cycling for transport to those of specific states. Our
-%% estimates can provide an important evidence base to encourage new
-%% construction, retrofits, land use policy, and programatic funding to
-%% support active transportation. Particularly in places with poor
-%% attainmnet of the physical activity guidelines, such investments can
-%% be importnat pieces of lareger population health interventions that
-%% have effective and long-lasting impacts.
+Given the constraints and complexity of local governance and infrastructure investments, land use, and human behavior there is a need for high quality and reliable data and tools for decision-making related to urban transportaiton systems.
+The lack of accurate and availbale estimates of the health impacts of such decisions has limited the effectiveness of efforts to encourage non-motorized transport in the US and in other countries.
+In the last decade new tools, applying comparative risk assessment methodologies, have made progress toward quantifying the multple health impacts of changes to the transportation system and behaviors.
 
-ITHIM is a statistical model that integrates data on active transport,
+ITHIM is one such statistical model that integrates data on active transport,
 physical activity, fine particulate matter and greenhouse gas
@@
 
-Dummy ref \cite{pucher2010}.
+Here we present this implementation using publically-available datasets with coverage of the entire United States, and investigate the sensitivity of the model outputs to variability in the input parameters.
+The results of this work suggest that statistical modeling and tools like the custom-built \R{} package \package{} can improve the application of this useful methodology and increase its utility to both practicioners and researchers.

### 2016-10-12 methodsPaper/tex/materialsAndMethods.tex
Update to methods paper
metro paragraph

@@
 and the coefficient of variation is treated as constant across age-sex
-strata and estimated with ATUS.  We see in Figure \ref{dalyFigure}
-that varying the mean values for non-travel-related exposure does not
-greatly affect the estimate for disease burden.  The distribution for
-travel-related exposure is estimated from the time spent walking or
-cycling, i.e., active transport time, which as in the original ITHIM
-model has a \logNormal{} distribution with constant coefficient of
-variation across age-sex strata.  Travel-related exposure is estimated
-from active transport time using assumptions about how much physical
+strata and estimated with ATUS.  The distribution for travel-related
+exposure is estimated from the time spent walking or cycling, i.e.,
+active transport time, which as in the original ITHIM model has
+a \logNormal{} distribution with constant coefficient of variation
+across age-sex strata.  Travel-related exposure is estimated from
+active transport time using assumptions about how much physical
 activity is required for cycling and walking.
+
+%does this involve using the CDC Wonder data? we could add a bit here about it, but if this is just national estimtes and shifts perhaps we don't need it.
+
+For the purposes of this implementation we limited our analyses to metropolitan regions of the United States.
+To idenfity urban populations we used the National Center for Health Statistics 2013 Urban-Rural Classification Scheme for Counties \cite{ingram2014}.
+We eliminated counties in the US with the classifcation of 'nonmetropolitan' (noncore or micropolitan).
+We used this and similar classification schemes as well as county and state identifiers in other datsets (NHTS, ATUS, CDC WONDER) to ensure consistency in our estimates for input parameters.

### 2016-10-12 R/ITHIM.Rmd
Small updates to manuscripts
Still trying to separate the content into two papers.  I need more material...
Commented-out old text and update Makefile

@@
 ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
+ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
 ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
@@
 ITHIMList <- list()
-
-#ITHIM.default <- createITHIM(vision = "baseline", region = "national")
-#ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
-#ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
 
@@
 
-```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
-
+## New Analysis
+
+```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
 ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
@@
 
-```
+ITHIM.scenario <- ITHIM.national
+ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=2)/10)
+ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=2)/10)
+
+pWalk <- ITHIM.national$means$pWalk[3,2]
+
+DALYs <- c(); DALYs.quints <- c();
+atVec <- seq(45, 120, length.out = 5)
+muNonTravelVec <- c(1,10,25)
+DTotal <- data.frame()
+
+for(at in atVec ){  # minutes active transport/week
+  for( muNonTravel in muNonTravelVec ){
+
+    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muwt", at*pWalk)
+    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muct", at*(1-pWalk))
+    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muNonTravel", muNonTravel)
+
+    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muwt", at*pWalk)
+    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muct", at*(1-pWalk))
+    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muNonTravel", muNonTravel)
+
+    ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+
+    DALYs <- sumDALY(ITHIM.national, ITHIM.scenario)
+    DALYs.quints <- sumDALY(ITHIM.national.quints, ITHIM.scenario.quints)
+
+    D <- data.frame(quants = c("percentiles","quintiles"), DALYs = c(DALYs,DALYs.quints),at = at, muNonTravel = muNonTravel)
+
+    DTotal <- rbind(D,DTotal)
+
+  }
+}
+
+p <- ggplot(DTotal, aes(x = at,y = -DALYs/1e3, colour = quants))
+p + stat_smooth( method = loess) + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)") + facet_grid( .
+~ muNonTravel) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))
+
+#D <- rbind(data.frame(AT = atVec, DALYs = DALYs, quants = "percentiles"),data.frame(AT = atVec, DALYs = DALYs.quints, quants = "quintiles"))
+#+ scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( .
+~ quints)
+#p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( .
+~ quints)
+```
+
 # Appendix
-
-```{r roadInjuries, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-convertState <- function( stateCode ){
-
-stateCode <- as.character(stateCode)
-
-stateCodeMatrix <- matrix(c("Alabama","1","Nebraska","31","Alaska","2","Nevada","32","Arizona","4","New Hampshire","33","Arkansas","5","New Jersey","34","California","6","New Mexico","35","Colorado","8","New York","36","Connecticut","9","North Carolina","37","Delaware","10","North Dakota","38","District of Columbia","11", "Ohio", "39","Florida","12","Oklahoma","40","Georgia","13","Oregon","41","Hawaii","15","Pennsylvania","42","Idaho","16","Puerto Rico","43","Illinois","17","Rhode Island","44","Indiana","18","South Carolina","45","Iowa","19","South Dakota","46","Kansas","20","Tennessee","47","Kentucky","21","Texas","48","Louisiana","22","Utah","49","Maine","23","Vermont","50","Maryland","24","Virgin Islands (since 2004)","52","Massachusetts","25","Virginia","51","Michigan","26","Washington","53","Minnesota","27","West Virginia","54","Mississippi","28","Wisconsin","55","Missouri","29","Wyoming","56","Montana","30"), byrow = TRUE, ncol = 2)
-
-stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
-
-stateNames <- stateCodeMatrix[,1]
-names(stateNames) <- stateCodeMatrix[,2]
-
-stateAbbrevs <- stateAbbrevMatrix[,2]
-names(stateAbbrevs) <- stateAbbrevMatrix[,1]
-
-return(stateAbbrevs[stateNames[stateCode]])
-
-}
-
-roadFatalaties <- read.csv(file = "./../data/FARS/BikePedDeaths99-13.csv", na.strings = c("998","999"))
-roadFatalaties <- within(roadFatalaties, {
-  ageClass <- convertToAgeClass(AGE)
-  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
-  SEX <- factor(SEX, levels = c("F","M"))
-  STATE <- as.factor(convertState(STATE))
-})
-roadFatalaties <- subset(roadFatalaties, !is.na(SEX) ) #& STATE %in% c("California", "New York"))
-
-p <- ggplot(roadFatalaties, aes(Type, fill = STATE))
-p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5))# + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) #+ facet_grid( Type ~ ., scales = "free_y")
-
-walkTime <- unlist(lapply(meanWalkList, function(x) x[3,2]))
-
-foo <- data.frame(with(roadFatalaties, table(STATE,Type)))
-within(subset(foo, Type == "Ped"), Walk <- walkTime[as.character(STATE)])
-
-```
-
-### CDC Mortality
-
-Next we parse the output of a CDC Wonder query to estimate mortality rates.
-
-The data files in ~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/.
-
-#### Query Results File (HHSmetro_Age_sex_2010-14.csv)
-
-```{r CDC, eval = FALSE, echo = FALSE}
-wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
-names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
-
-wonder <- subset(wonder, WONDERageCode != "NS") # remove 'Not stated' Age entries
-```
-
-The following two recoding chunks should probably be redone.  They
-depend on the ordering of the factors which seems dangerous.
-
-```{r ITHIMCause, eval = FALSE, echo = FALSE}
-ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
-names(ITHIMcauseKey) <- unique(wonder$ICD10)
-wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
-```
-```{r recodeAge, eval = FALSE, echo = FALSE}
-ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
-names(ITHIMageKey) <- unique(wonder$WONDERage)
-wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
-```
-
-Depression was not in the WONDER data, since there are no deaths, have to add it manually.
-
-```{r wonder, eval = FALSE, echo = FALSE}
-wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
-depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
-wonder <- rbind(wonder, depress)
-```
-
-Replace suppressed values with random number 1-9
-
-```{r suppressed, eval = FALSE, echo = FALSE, warning = FALSE}
-wonder <- within( wonder, {
-    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
-    })
-```
-
-10 HHS x 2 sex x 8 age x 13 disease = 2,080 strata.
-
-```{r summary2, eval = FALSE, echo = FALSE}
-wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
-                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
-)
-```
-
-#### Single Year Data File (HHSmetro_Age_sex_2014_AllCause.csv)
-
-Read in data for single year population.  Processed in Excel, to
-combine two WONDER queries, need to go back and script the processing
-of the raw WONDER output
-
-```{r singleyear, eval = FALSE, echo = FALSE}
-pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
-
-names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
-```
-This is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' command.
-
-```{r sloppy, eval = FALSE, echo = FALSE}
-ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
-names(ITHIMageKey2) <- unique(pop14$WONDERage)
-pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
-```
-
-```{r foop, eval = FALSE, echo = FALSE}
-pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
-pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))
-
-popSummary <- ddply(pop14, .(HHSRegionCode, gender, ITHIMage), summarise,
-                    population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
-                   AllCauseDeaths2014 = sum(as.numeric(as.character(deaths2014AllCause)), na.rm=T)
-)
-
-HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
-```
-#### Read USGBD file (USGBD.csv)
-
-Presumably this comes from an EXCEL file.
-
-```{r USGBD, eval = FALSE, echo = FALSE}
-USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
-USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
-USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)
-
-burden <- merge(HHSmetroCauseSpecific, USGBD, all.x=T)
-burden$deaths <- burden$numberOfDeaths5yr/5
-## neil uses 10, and if deaths<10 sets ratio to 1
-burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
-burden$YLL <- burden$ratio * (burden$GBDyll/burden$GBDpop) * burden$population2014
-burden$YLD <- burden$ratio * (burden$GBDyld/burden$GBDpop) * burden$population2014
-burden$DALY <- burden$ratio * (burden$GBDdaly/burden$GBDpop) * burden$population2014
-
-burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
-levels(burden$sex)  <- c("M","F")
-
-burdenOut <- burden[,c("HHSRegionCode","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
-names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")
-
-burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
-burdenOut <- burdenOut[,c("HHS", "disease","sex","dproj","yll","yld","daly")]
-```
-
-Now we create an R list of GBD objects named by HHS region.
-
-```{r burdenList, eval = FALSE, echo = FALSE}
-GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
-#saveRDS(GBDList, file = "../data/GBD.rds")
-```
 
@@
 METmatrix <- METmatrix/METmatrix[3,2]
+```
+
+```{r roadInjuries, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+convertState <- function( stateCode ){
+
+stateCode <- as.character(stateCode)
+
+stateCodeMatrix <- matrix(c("Alabama","1","Nebraska","31","Alaska","2","Nevada","32","Arizona","4","New Hampshire","33","Arkansas","5","New Jersey","34","California","6","New Mexico","35","Colorado","8","New York","36","Connecticut","9","North Carolina","37","Delaware","10","North Dakota","38","District of Columbia","11", "Ohio", "39","Florida","12","Oklahoma","40","Georgia","13","Oregon","41","Hawaii","15","Pennsylvania","42","Idaho","16","Puerto Rico","43","Illinois","17","Rhode Island","44","Indiana","18","South Carolina","45","Iowa","19","South Dakota","46","Kansas","20","Tennessee","47","Kentucky","21","Texas","48","Louisiana","22","Utah","49","Maine","23","Vermont","50","Maryland","24","Virgin Islands (since 2004)","52","Massachusetts","25","Virginia","51","Michigan","26","Washington","53","Minnesota","27","West Virginia","54","Mississippi","28","Wisconsin","55","Missouri","29","Wyoming","56","Montana","30"), byrow = TRUE, ncol = 2)
+
+stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
+
+stateNames <- stateCodeMatrix[,1]
+names(stateNames) <- stateCodeMatrix[,2]
+
+stateAbbrevs <- stateAbbrevMatrix[,2]
+names(stateAbbrevs) <- stateAbbrevMatrix[,1]
+
+return(stateAbbrevs[stateNames[stateCode]])
+
+}
+
+roadFatalaties <- read.csv(file = "./../data/FARS/BikePedDeaths99-13.csv", na.strings = c("998","999"))
+roadFatalaties <- within(roadFatalaties, {
+  ageClass <- convertToAgeClass(AGE)
+  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
+  SEX <- factor(SEX, levels = c("F","M"))
+  STATE <- as.factor(convertState(STATE))
+})
+roadFatalaties <- subset(roadFatalaties, !is.na(SEX) ) #& STATE %in% c("California", "New York"))
+
+p <- ggplot(roadFatalaties, aes(Type, fill = STATE))
+p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5))# + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) #+ facet_grid( Type ~ ., scales = "free_y")
+
+walkTime <- unlist(lapply(meanWalkList, function(x) x[3,2]))
+
+foo <- data.frame(with(roadFatalaties, table(STATE,Type)))
+within(subset(foo, Type == "Ped"), Walk <- walkTime[as.character(STATE)])
+
+```
+
+### CDC Mortality
+
+Next we parse the output of a CDC Wonder query to estimate mortality rates.
+
+The data files in ~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/.
+
+#### Query Results File (HHSmetro_Age_sex_2010-14.csv)
+
+```{r CDC, eval = FALSE, echo = FALSE}
+wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
+names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
+
+wonder <- subset(wonder, WONDERageCode != "NS") # remove 'Not stated' Age entries
+```
+
+The following two recoding chunks should probably be redone.  They
+depend on the ordering of the factors which seems dangerous.
+
+```{r ITHIMCause, eval = FALSE, echo = FALSE}
+ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
+names(ITHIMcauseKey) <- unique(wonder$ICD10)
+wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
+```
+```{r recodeAge, eval = FALSE, echo = FALSE}
+ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
+names(ITHIMageKey) <- unique(wonder$WONDERage)
+wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
+```
+
+Depression was not in the WONDER data, since there are no deaths, have to add it manually.
+
+```{r wonder, eval = FALSE, echo = FALSE}
+wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
+depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
+wonder <- rbind(wonder, depress)
+```
+
+Replace suppressed values with random number 1-9
+
+```{r suppressed, eval = FALSE, echo = FALSE, warning = FALSE}
+wonder <- within( wonder, {
+    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
+    })
+```
+
+10 HHS x 2 sex x 8 age x 13 disease = 2,080 strata.
+
+```{r summary2, eval = FALSE, echo = FALSE}
+wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
+                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
+)
+```
+
+#### Single Year Data File (HHSmetro_Age_sex_2014_AllCause.csv)
+
+Read in data for single year population.  Processed in Excel, to
+combine two WONDER queries, need to go back and script the processing
+of the raw WONDER output
+
+```{r singleyear, eval = FALSE, echo = FALSE}
+pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
+
+names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
+```
+This is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' command.
+
+```{r sloppy, eval = FALSE, echo = FALSE}
+ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
+names(ITHIMageKey2) <- unique(pop14$WONDERage)
+pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
+```
+
+```{r foop, eval = FALSE, echo = FALSE}
+pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
+pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))
+
+popSummary <- ddply(pop14, .(HHSRegionCode, gender, ITHIMage), summarise,
+                    population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
+                   AllCauseDeaths2014 = sum(as.numeric(as.character(deaths2014AllCause)), na.rm=T)
+)
+
+HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
+```
+#### Read USGBD file (USGBD.csv)
+
+Presumably this comes from an EXCEL file.
+
+```{r USGBD, eval = FALSE, echo = FALSE}
+USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
+USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
+USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)
+
+burden <- merge(HHSmetroCauseSpecific, USGBD, all.x=T)
+burden$deaths <- burden$numberOfDeaths5yr/5
+## neil uses 10, and if deaths<10 sets ratio to 1
+burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
+burden$YLL <- burden$ratio * (burden$GBDyll/burden$GBDpop) * burden$population2014
+burden$YLD <- burden$ratio * (burden$GBDyld/burden$GBDpop) * burden$population2014
+burden$DALY <- burden$ratio * (burden$GBDdaly/burden$GBDpop) * burden$population2014
+
+burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
+levels(burden$sex)  <- c("M","F")
+
+burdenOut <- burden[,c("HHSRegionCode","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
+names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")
+
+burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
+burdenOut <- burdenOut[,c("HHS", "disease","sex","dproj","yll","yld","daly")]
+```
+
+Now we create an R list of GBD objects named by HHS region.
+
+```{r burdenList, eval = FALSE, echo = FALSE}
+GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
+#saveRDS(GBDList, file = "../data/GBD.rds")
 ```

### 2016-10-12 methodsPaper/tex/results.tex
Subdivided paper #1 with Plos One template
Update to methods paper
Commented-out old text and update Makefile

@@
-% Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
+%% We used the National Household Transportation Survey to estimate the
+%% mean walk and cycle times for individuals that live in metropolitan
+%% areas for each of seven states (CA, TX, NY, FL, VA, NC and AZ)
+%% \cite{NHTS}.  These states were selected becaused they had sufficient
+%% sample size for reasonable estimates of age-sex specific
+%% travel-related walking and cycling means, Figure \ref{meanMatrices}.
 
-% % Place tables after the first paragraph in which they are cited.
-% \begin{table}[!ht]
-% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
-% \centering
-% \caption{
-% {\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
-% \begin{tabular}{|l+l|l|l|l|l|l|l|}
-% \hline
-% \multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
-% $cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
-% $cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
-% $cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
-% \end{tabular}
-% \begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-% \end{flushleft}
-% \label{table1}
-% \end{adjustwidth}
-% \end{table}
+%% \begin{figure}[t]
+%%   \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
+%%     \caption{Mean travel time by mode (walking/cycling) for each of
+%%       seven states.  National means are displayed on the
+%%       right.}\label{meanMatrices}
+%% \end{figure}
 
-We used the National Household Transportation Survey to estimate the
-mean walk and cycle times for individuals that live in metropolitan
-areas for each of seven states (CA, TX, NY, FL, VA, NC and AZ)
-\cite{NHTS}.  These states were selected becaused they had sufficient
-sample size for reasonable estimates of age-sex specific
-travel-related walking and cycling means, Figure \ref{meanMatrices}.
-
-Of these seven states one was found to show active transport behavior
-that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
-Values for national disease-specific DALYs were found using the CDC
-Wonder database \cite{CDCWonder}.  California stood out among the
-estimates for national DALYs averted.  Most of the decrease in DALYs
-is due to prevention of dementia in the oldest age group, $80+$ years
-old.  California exhibits much greater walking and cycling means in
-this age group than the national average. In particular we see the
-cycling mean We also see prevention of depression and diabetes in the
-middle age groups due to increased active transport time.
+%% Of these seven states one was found to show active transport behavior
+%% that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
+%% Values for national disease-specific DALYs were found using the CDC
+%% Wonder database \cite{CDCWonder}.  California stood out among the
+%% estimates for national DALYs averted.  Most of the decrease in DALYs
+%% is due to prevention of dementia in the oldest age group, $80+$ years
+%% old.  California exhibits much greater walking and cycling means in
+%% this age group than the national average. In particular we see the
+%% cycling mean We also see prevention of depression and diabetes in the
+%% middle age groups due to increased active transport time.

### 2016-10-12 methodsPaper/tex/abstract.tex

@@
+Comparative risk assessment exercises to quantify the health potential of active transportation have been performed in the United States on a limited basis.
+Current methodologies, which rely on entensive data collection activities and user expertise, have estimated health benefits for select metropolitan areas under a few alternative scenarios.
+These analyses have largely been conducted using the Intergrated Transport Health Impacts Model (IHTIM), which was developed in the UK and has been adapted for use in the US, primarility California.
+Here we present an implementation of one portion of ITHIM's comparative risk assessment framework to model the health impacts of modified physical activity patterns from biking and walking.
+Using the newly-developed ITHIM package available for the R Software platform publically-available datasets with coverage for the entire US are used to quantify changes in diability adjusted life years.
+We examine the sensitivity of outcomes for the nation to input parameters; active travel time, non-travel physical activity time, specifically.
+The revised R package implemention of ITHIM's physical activity component provides a robust and stable means of examining the potential for changes in transportation behavior in the US, and the distribution of those benefits across age and gender categories.
+
 %% While increased active transportation, primarily cycling and walking,

### 2016-10-17 R/ITHIM.Rmd
Tweaked figure 1
Made a few changes to get running on new iMac

@@
 options(repos='http://cran.rstudio.com/')
+# install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r","devtools"))
 library("ggplot2")
+library("devtools")
 library("reshape2")
@@
 library("dplyr")
-library("cowplot")
-library("grid")
-library("scales")
-library("ggthemes")
+#library("cowplot")
+#library("grid")
+#library("scales")
+#library("ggthemes")
 ```
@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-library("devtools")
-install_github("smbache/magrittr", ref="master")
-library("magrittr")
 #install("~/ITHIM/")
 install_github("syounkin/ITHIM", ref="devel")
+install_github("smbache/magrittr", ref="master")
+
+library("magrittr")
 library("ITHIM")
@@
 atVec <- seq(45, 120, length.out = 5)
-muNonTravelVec <- c(1,10,25)
+muNonTravelVec <- c(1,5,10,25)
 DTotal <- data.frame()

### 2016-10-18 R/ITHIM.Rmd
Changed names of 2003 ATUS data files
Not sure how this happened, unless ATUS changed the name of the files.  I can;t find files with the old name *_0311.dat

@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-#install("~/ITHIM/")
-install_github("syounkin/ITHIM", ref="devel")
+install("~/ITHIM/")
+#install_github("syounkin/ITHIM", ref="devel")
 install_github("smbache/magrittr", ref="master")
@@
 ITHIMList <- list()
-ITHIM.default <- createITHIM(vision = "baseline", region = "national")
+ITHIM.default <- createITHIM()
 ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
@@
 ```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
-cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
+cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat"
 cps <- read.table(cpsFile, header=T, skip=0, sep="," )
@@
 
-summaryFile <- "~/GHI/data/ATUS/atussum_0311.dat"
+summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat"
 summary <- read.table(summaryFile, header=T, skip=0, sep=",")
@@
 
-## Example (Bay Area)
-
-First we define the baseline and scenario objects.
-
-```{r instantiate, eval = FALSE, echo = FALSE}
-ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
-ITHIM.scenario <- createITHIM(vision = "scenario", region = "SFBayArea")
-```
-
-Next we compare the scenario to the baseline using compareModels().
-
-```{r RRs, eval = FALSE, echo = FALSE}
-comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-```
-
-### Deaths
-```{r show, echo = FALSE, eval = FALSE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### DALY
-```{r show2, echo = FALSE, eval = FALSE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### YLD
-```{r show3, echo = FALSE, eval = FALSE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### YLL
-```{r show4, echo = FALSE, eval = FALSE, echo = FALSE}
-matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-
 ## Example (National)
@@
 ```{r instantiate2, eval = FALSE, echo = FALSE}
-ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
-ITHIM.scenario <- createITHIM(vision = "scenario", region = "national")
+ITHIM.baseline <- createITHIM()
+ITHIM.scenario <- createITHIM()
 ```
@@
 
-summaryFile <- "~/GHI/data/ATUS/atussum_0311.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_0311.dat
-
-cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_0311.dat
+summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_2003.dat
+
+cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_2003.dat
 

### 2016-10-18 R/ITHIM.Rmd
Added document() command to ITHIM.Rmd

@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
+document("~/ITHIM/")
 install("~/ITHIM/")

### 2016-10-18 methodsPaper/tex/materialsAndMethods.tex
Commented-out old text and update Makefile
Made a few changes to get running on new iMac
added paragraph on methodology

@@
 
-%does this involve using the CDC Wonder data? we could add a bit here about it, but if this is just national estimtes and shifts perhaps we don't need it.
+% \begin{figure}[t]
+%   \centerline{\includegraphics[width=\textwidth]{./figures/fig2.pdf}}
+%     \caption{}\label{fig1}
+% \end{figure}
 
-For the purposes of this implementation we limited our analyses to metropolitan regions of the United States.
-To idenfity urban populations we used the National Center for Health Statistics 2013 Urban-Rural Classification Scheme for Counties \cite{ingram2014}.
-We eliminated counties in the US with the classifcation of 'nonmetropolitan' (noncore or micropolitan).
-We used this and similar classification schemes as well as county and state identifiers in other datsets (NHTS, ATUS, CDC WONDER) to ensure consistency in our estimates for input parameters.
+% does this involve using the CDC Wonder data? we could add a bit here
+% about it, but if this is just national estimtes and shifts perhaps
+% we don't need it.
+
+For the purposes of this implementation we limited our analyses to
+metropolitan regions of the United States. To idenfity urban
+populations we used the National Center for Health Statistics 2013
+Urban-Rural Classification Scheme for Counties \cite{ingram2014}. We
+eliminated counties in the US with the classifcation of
+'nonmetropolitan' (noncore or micropolitan). We used this and similar
+classification schemes as well as county and state identifiers in
+other datsets (NHTS, ATUS, CDC WONDER) to ensure consistency in our
+estimates for input parameters.
+
+To test the new implementation of ITHIM's active transportation component, we used the UW national metro population estimates for active travel time (walking and cycling) as the baseline.
+We obtained active travel time input parameters for alternate scenarios using states with large sample sizes.
+From these examples we were able to approximate low and high active travel scenarios.
+Finally we applied these active travel times to the US metro population to perform the comparative risk assessment agains the baseline US numbers.
+We then examined the sensitivity of disease burden estimates to input parameters including the active travel time and non-travel activity.

### 2016-10-18 methodsPaper/tex/results.tex
results intro

@@
+Our subset of the US population to only metropolitan counties consisted of 1,090 of 3,141 (34.7\%) counties, or x of y (z\%) of the total US population.
+The age and gender distribution of our subset population was slightly younger (making this up) than the national population.
+Active transport times (walking and cycling) for baseline, low, and high scenarios are shown in \ref{fig:ATtimes}.
+Overall mean walking time for high and low scenarios were X (min/week) higher and Y (min/week) lower than baseline, respectively.
+Baseline non-travel activity time was estimated to be X (min/week) from the ATUS for US metropolitan populations.
+
 %% We used the National Household Transportation Survey to estimate the

### 2016-10-18 methodsPaper/tex/materialsAndMethods.tex
CDC WONDER methods

@@
 We then examined the sensitivity of disease burden estimates to input parameters including the active travel time and non-travel activity.
+
+We report total change (averted or increased) in diability adjusted life years (DALYs) as an estimate of dissease burden change with alternative scenarions.
+To estimate DALYs we use a methodology adopted from previous US implementations of ITHIM  which scale Global Budern of Disease Estimates for the US to smaller populations using mortality rate ratios \cite{Maizlish2013}.
+Age, sex, and casuse specific mortalities were obtained form the CDC WONDER database for US metro counties for the years 2010-2014.
+Cells with less than 10 deaths were imputed using R's random integer generator.

### 2016-10-18 methodsPaper/tex/results.tex
vary NTM paragraph
more results

@@
 
-%% We used the National Household Transportation Survey to estimate the
-%% mean walk and cycle times for individuals that live in metropolitan
-%% areas for each of seven states (CA, TX, NY, FL, VA, NC and AZ)
-%% \cite{NHTS}.  These states were selected becaused they had sufficient
-%% sample size for reasonable estimates of age-sex specific
-%% travel-related walking and cycling means, Figure \ref{meanMatrices}.
-
-%% \begin{figure}[t]
-%%   \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
-%%     \caption{Mean travel time by mode (walking/cycling) for each of
-%%       seven states.  National means are displayed on the
-%%       right.}\label{meanMatrices}
-%% \end{figure}
-
-%% Of these seven states one was found to show active transport behavior
-%% that, if adopted nation-wide, would avert 75-100,000 DALYs overall.
-%% Values for national disease-specific DALYs were found using the CDC
-%% Wonder database \cite{CDCWonder}.  California stood out among the
-%% estimates for national DALYs averted.  Most of the decrease in DALYs
-%% is due to prevention of dementia in the oldest age group, $80+$ years
-%% old.  California exhibits much greater walking and cycling means in
-%% this age group than the national average. In particular we see the
-%% cycling mean We also see prevention of depression and diabetes in the
-%% middle age groups due to increased active transport time.
+Running the model to compare baseline to low and high scenarios produced estimates of X change in the disease burden (1,000s DALYs) for our US population subset.
+We then varied the mean non-travel activity time and active travel time to observe how this chaged results (\ref{fig:currentFig1}).
+We observe the most dramatic responses in disease burden estimates to changes in active travel time when non-travel activity times are lowest.
+This result is consistent with the non-linear dose-response curve for physical activity and several health outcomes.
+The total health benefit of increased active travel time (high scenario) varies from X to Y with changing mean non-travel activity time (0 to 25 minutes).

### 2016-10-19 methodsPaper/tex/results.tex
Update results.tex

@@
-Our subset of the US population to only metropolitan counties consisted of 1,090 of 3,141 (34.7\%) counties, or x of y (z\%) of the total US population.
-The age and gender distribution of our subset population was slightly younger (making this up) than the national population.
+Our subset of the US population to only metropolitan counties consisted of 1,090 of 3,141 (34.7\%) counties, or ???X???
+of ???Y???
+(???Z???\%) of the total US population.
+The age and gender distribution of our subset population was slightly younger (???making this up???) than the national population.
 Active transport times (walking and cycling) for baseline, low, and high scenarios are shown in \ref{fig:ATtimes}.
-Overall mean walking time for high and low scenarios were X (min/week) higher and Y (min/week) lower than baseline, respectively.
-Baseline non-travel activity time was estimated to be X (min/week) from the ATUS for US metropolitan populations.
+Overall mean walking time for high and low scenarios were ???X???
+(min/week) higher and ???Y???
+(min/week) lower than baseline, respectively.
+Baseline non-travel activity time was estimated to be ???X???
+(min/week) from the ATUS for US metropolitan populations.
 
-Running the model to compare baseline to low and high scenarios produced estimates of X change in the disease burden (1,000s DALYs) for our US population subset.
+Comparing baseline to low/high scenarios produced estimates of ???X???
+change in the disease burden (1,000s DALYs) for our US population subset.
 We then varied the mean non-travel activity time and active travel time to observe how this chaged results (\ref{fig:currentFig1}).
@@
 This result is consistent with the non-linear dose-response curve for physical activity and several health outcomes.
-The total health benefit of increased active travel time (high scenario) varies from X to Y with changing mean non-travel activity time (0 to 25 minutes).
+The total health benefit of increased active travel time (high scenario) varies from ???X???
+to ???Y???
+with changing mean non-travel activity time (0 to 25 minutes).
+
+Adapting ITHIM's methods to the R platform enables the rapid and thorough assessment of parameter influence within the model.
+Another important improvement enabled by the shift to a more computationaly robust platform was to examine methodological assumptions of former implementations.
+One critical assumption invovles the approximation of _____ used to calculate the attirbutable fraction (see Figure \ref{fig:overview}).
+Previous implementations used quitiles to approximate this value, essentailly the area under the ___????___ curve.
+We improved upon those methods to use percentiles or 100 points to estimate the same values.
+We find our estimates to be much more stable and reliable at all values of active travel and non-travel activity parameters.
+
+As exhibited in Figure \ref{fig:percentQuint}, the variability in disease burden benfits varies to a much greater degree when using fewer points in the attirbutable fraction approximation.

### 2016-10-19 methodsPaper/tex/materialsAndMethods.tex
Fixed a bug or two and wrapped text (for readability)

@@
 
-To test the new implementation of ITHIM's active transportation component, we used the UW national metro population estimates for active travel time (walking and cycling) as the baseline.
-We obtained active travel time input parameters for alternate scenarios using states with large sample sizes.
-From these examples we were able to approximate low and high active travel scenarios.
-Finally we applied these active travel times to the US metro population to perform the comparative risk assessment agains the baseline US numbers.
-We then examined the sensitivity of disease burden estimates to input parameters including the active travel time and non-travel activity.
+To test the new implementation of ITHIM's active transportation
+component, we used the UW national metro population estimates for
+active travel time (walking and cycling) as the baseline. We obtained
+active travel time input parameters for alternate scenarios using
+states with large sample sizes. From these examples we were able to
+approximate low and high active travel scenarios. Finally we applied
+these active travel times to the US metro population to perform the
+comparative risk assessment agains the baseline US numbers. We then
+examined the sensitivity of disease burden estimates to input
+parameters including the active travel time and non-travel activity.
 
-We report total change (averted or increased) in diability adjusted life years (DALYs) as an estimate of dissease burden change with alternative scenarions.
-To estimate DALYs we use a methodology adopted from previous US implementations of ITHIM  which scale Global Budern of Disease Estimates for the US to smaller populations using mortality rate ratios \cite{Maizlish2013}.
-Age, sex, and casuse specific mortalities were obtained form the CDC WONDER database for US metro counties for the years 2010-2014.
-Cells with less than 10 deaths were imputed using R's random integer generator.
+We report total change (averted or increased) in diability adjusted
+life years (DALYs) as an estimate of dissease burden change with
+alternative scenarions. To estimate DALYs we use a methodology adopted
+from previous US implementations of ITHIM which scale Global Budern of
+Disease Estimates for the US to smaller populations using mortality
+rate ratios \cite{maizlish2013}. Age, sex, and casuse specific
+mortalities were obtained form the CDC WONDER database for US metro
+counties for the years 2010-2014. Cells with less than 10 deaths were
+imputed using R's random integer generator.

### 2016-10-19 methodsPaper/tex/results.tex

@@
-Our subset of the US population to only metropolitan counties consisted of 1,090 of 3,141 (34.7\%) counties, or ???X???
-of ???Y???
-(???Z???\%) of the total US population.
-The age and gender distribution of our subset population was slightly younger (???making this up???) than the national population.
-Active transport times (walking and cycling) for baseline, low, and high scenarios are shown in \ref{fig:ATtimes}.
-Overall mean walking time for high and low scenarios were ???X???
-(min/week) higher and ???Y???
-(min/week) lower than baseline, respectively.
-Baseline non-travel activity time was estimated to be ???X???
-(min/week) from the ATUS for US metropolitan populations.
+Our subset of the US population to only metropolitan counties
+consisted of 1,090 of 3,141 (34.7\%) counties, or ???X??? of ???Y???
+(???Z???\%) of the total US population. The age and gender
+distribution of our subset population was slightly younger (???making
+this up???) than the national population. Active transport times
+(walking and cycling) for baseline, low, and high scenarios are shown
+in \ref{fig:ATtimes}. Overall mean walking time for high and low
+scenarios were ???X??? (min/week) higher and ???Y??? (min/week) lower
+than baseline, respectively. Baseline non-travel activity time was
+estimated to be ???X??? (min/week) from the ATUS for US metropolitan
+populations.
 
 Comparing baseline to low/high scenarios produced estimates of ???X???
-change in the disease burden (1,000s DALYs) for our US population subset.
-We then varied the mean non-travel activity time and active travel time to observe how this chaged results (\ref{fig:currentFig1}).
-We observe the most dramatic responses in disease burden estimates to changes in active travel time when non-travel activity times are lowest.
-This result is consistent with the non-linear dose-response curve for physical activity and several health outcomes.
-The total health benefit of increased active travel time (high scenario) varies from ???X???
-to ???Y???
-with changing mean non-travel activity time (0 to 25 minutes).
+change in the disease burden (1,000s DALYs) for our US population
+subset. We then varied the mean non-travel activity time and active
+travel time to observe how this chaged results
+(\ref{fig:currentFig1}). We observe the most dramatic responses in
+disease burden estimates to changes in active travel time when
+non-travel activity times are lowest. This result is consistent with
+the non-linear dose-response curve for physical activity and several
+health outcomes. The total health benefit of increased active travel
+time (high scenario) varies from ???X??? to ???Y??? with changing mean
+non-travel activity time (0 to 25 minutes).
 
-Adapting ITHIM's methods to the R platform enables the rapid and thorough assessment of parameter influence within the model.
-Another important improvement enabled by the shift to a more computationaly robust platform was to examine methodological assumptions of former implementations.
-One critical assumption invovles the approximation of _____ used to calculate the attirbutable fraction (see Figure \ref{fig:overview}).
-Previous implementations used quitiles to approximate this value, essentailly the area under the ___????___ curve.
-We improved upon those methods to use percentiles or 100 points to estimate the same values.
-We find our estimates to be much more stable and reliable at all values of active travel and non-travel activity parameters.
+Adapting ITHIM's methods to the R platform enables the rapid and
+thorough assessment of parameter influence within the model. Another
+important improvement enabled by the shift to a more computationaly
+robust platform was to examine methodological assumptions of former
+implementations. One critical assumption invovles the approximation of
+??? used to calculate the attirbutable fraction (see
+Figure \ref{fig:overview}). Previous implementations used quitiles to
+approximate this value, essentailly the area under the ???? curve. We
+improved upon those methods to use percentiles or 100 points to
+estimate the same values. We find our estimates to be much more stable
+and reliable at all values of active travel and non-travel activity
+parameters.
 
-As exhibited in Figure \ref{fig:percentQuint}, the variability in disease burden benfits varies to a much greater degree when using fewer points in the attirbutable fraction approximation.
+As exhibited in Figure \ref{fig:percentQuint}, the variability in
+disease burden benfits varies to a much greater degree when using
+fewer points in the attirbutable fraction approximation.

### 2016-10-19 R/ITHIM.Rmd
Tweaked Figure 1 from methods paper
Created two new figures
Also created appendix.Rnw to store old R code.

@@
 library("dplyr")
-#library("cowplot")
-#library("grid")
-#library("scales")
-#library("ggthemes")
 ```
@@
 
-#upper <- 10
-
 p + geom_bar(stat="identity", position = "dodge") +
@@
 
-#geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
-# + theme_tufte(base_size=12, ticks=T)
-#
-# + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))
-# + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
 ```
@@
 
-## New Analysis
+## Figure 1
 
@@
 ITHIM.scenario <- ITHIM.national
-ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=2)/10)
-ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=2)/10)
+ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
+ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)
 
@@
 atVec <- seq(45, 120, length.out = 5)
-muNonTravelVec <- c(1,5,10,25)
+muNonTravelVec <- c(1, 10, 50, 100, 150, 200)
 DTotal <- data.frame()
@@
 
-#D <- rbind(data.frame(AT = atVec, DALYs = DALYs, quants = "percentiles"),data.frame(AT = atVec, DALYs = DALYs.quints, quants = "quintiles"))
-#+ scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( .
-~ quints)
-#p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( .
-~ quints)
 ```
 
-# Appendix
+## Figure 2
 
-## Non-Travel METs
+```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+D <- rbind(getWalkTime(ITHIM.default),getWalkSpeed(ITHIM.default),getCycleTime(ITHIM.default),getNonTravelMETs(ITHIM.default))
+D <- within(D, sex <- factor(sex, levels = c("F","M")))
 
-```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
-newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
-newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
-dataColumns <- c(1,5)
-D <- newMETs[,dataColumns]
-metaData <- newMETs[,-dataColumns]
-newMETs <- list(Data = D, metaData = metaData)
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-
-rownames(newMETs$Data) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-
-newMETs$metaData <- within(newMETs$metaData, {
-  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
-})
-
-metKey <- newMETs$Data$AinsworthMETS
-names(metKey) <- activityCode
-
+p <- ggplot(D, aes(x = ageClass, y = mu, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ ., scales = "free_y")
 ```
 
-```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
-cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat"
-cps <- read.table(cpsFile, header=T, skip=0, sep="," )
-cps <- within(cps, {
-    TUCASEID <- as.character(TUCASEID)
-    TULINENO <- as.character(TULINENO)
-    GEREG <- as.factor(GEREG)
-    })
-cps <- subset(cps, TULINENO == "1")
+```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+D <- melt(ITHIM.default$parameters$GBD)
+names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")
+D <- within(D, sex <- factor(sex, levels = c("F","M")))
 
-summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat"
-summary <- read.table(summaryFile, header=T, skip=0, sep=",")
-metaDataColumns <- 1:24
-summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
-
-summaryList$metaData <- within(summaryList$metaData,{
-    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
-    AGECAT <- ifelse(TEAGE <5, "00-04",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
-                            )))))))})
-
+p <- ggplot(subset(D,disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes")), aes(x = ageClass, y = burden, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ disease, scales = "free_y")
 ```
-
-We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
-
-```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
-allActivities <- colnames(summaryList$Data)
-nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
-nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
-
-activities <- nonTravelActivities
-
-metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
-names(metKeyVector) <- activities
-metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-dimnames(metKeyMatrix) <- list( NULL, activities)
-
-METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
-
-D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
-
-```
-
-```{r parameters, eval = FALSE, echo = FALSE}
-parameters <- ddply(D, .(SEX, AGECAT), summarise,
-             mean = mean(METs, na.rm=T),
-             sd = sd(METs, na.rm=T)
-             )
-
-METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
-METmatrix <- METmatrix/METmatrix[3,2]
-```
-
-```{r roadInjuries, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-convertState <- function( stateCode ){
-
-stateCode <- as.character(stateCode)
-
-stateCodeMatrix <- matrix(c("Alabama","1","Nebraska","31","Alaska","2","Nevada","32","Arizona","4","New Hampshire","33","Arkansas","5","New Jersey","34","California","6","New Mexico","35","Colorado","8","New York","36","Connecticut","9","North Carolina","37","Delaware","10","North Dakota","38","District of Columbia","11", "Ohio", "39","Florida","12","Oklahoma","40","Georgia","13","Oregon","41","Hawaii","15","Pennsylvania","42","Idaho","16","Puerto Rico","43","Illinois","17","Rhode Island","44","Indiana","18","South Carolina","45","Iowa","19","South Dakota","46","Kansas","20","Tennessee","47","Kentucky","21","Texas","48","Louisiana","22","Utah","49","Maine","23","Vermont","50","Maryland","24","Virgin Islands (since 2004)","52","Massachusetts","25","Virginia","51","Michigan","26","Washington","53","Minnesota","27","West Virginia","54","Mississippi","28","Wisconsin","55","Missouri","29","Wyoming","56","Montana","30"), byrow = TRUE, ncol = 2)
-
-stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
-
-stateNames <- stateCodeMatrix[,1]
-names(stateNames) <- stateCodeMatrix[,2]
-
-stateAbbrevs <- stateAbbrevMatrix[,2]
-names(stateAbbrevs) <- stateAbbrevMatrix[,1]
-
-return(stateAbbrevs[stateNames[stateCode]])
-
-}
-
-roadFatalaties <- read.csv(file = "./../data/FARS/BikePedDeaths99-13.csv", na.strings = c("998","999"))
-roadFatalaties <- within(roadFatalaties, {
-  ageClass <- convertToAgeClass(AGE)
-  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
-  SEX <- factor(SEX, levels = c("F","M"))
-  STATE <- as.factor(convertState(STATE))
-})
-roadFatalaties <- subset(roadFatalaties, !is.na(SEX) ) #& STATE %in% c("California", "New York"))
-
-p <- ggplot(roadFatalaties, aes(Type, fill = STATE))
-p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5))# + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) #+ facet_grid( Type ~ ., scales = "free_y")
-
-walkTime <- unlist(lapply(meanWalkList, function(x) x[3,2]))
-
-foo <- data.frame(with(roadFatalaties, table(STATE,Type)))
-within(subset(foo, Type == "Ped"), Walk <- walkTime[as.character(STATE)])
-
-```
-
-### CDC Mortality
-
-Next we parse the output of a CDC Wonder query to estimate mortality rates.
-
-The data files in ~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/.
-
-#### Query Results File (HHSmetro_Age_sex_2010-14.csv)
-
-```{r CDC, eval = FALSE, echo = FALSE}
-wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
-names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")
-
-wonder <- subset(wonder, WONDERageCode != "NS") # remove 'Not stated' Age entries
-```
-
-The following two recoding chunks should probably be redone.  They
-depend on the ordering of the factors which seems dangerous.
-
-```{r ITHIMCause, eval = FALSE, echo = FALSE}
-ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
-names(ITHIMcauseKey) <- unique(wonder$ICD10)
-wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
-```
-```{r recodeAge, eval = FALSE, echo = FALSE}
-ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
-names(ITHIMageKey) <- unique(wonder$WONDERage)
-wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
-```
-
-Depression was not in the WONDER data, since there are no deaths, have to add it manually.
-
-```{r wonder, eval = FALSE, echo = FALSE}
-wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
-depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
-wonder <- rbind(wonder, depress)
-```
-
-Replace suppressed values with random number 1-9
-
-```{r suppressed, eval = FALSE, echo = FALSE, warning = FALSE}
-wonder <- within( wonder, {
-    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
-    })
-```
-
-10 HHS x 2 sex x 8 age x 13 disease = 2,080 strata.
-
-```{r summary2, eval = FALSE, echo = FALSE}
-wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
-                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
-)
-```
-
-#### Single Year Data File (HHSmetro_Age_sex_2014_AllCause.csv)
-
-Read in data for single year population.  Processed in Excel, to
-combine two WONDER queries, need to go back and script the processing
-of the raw WONDER output
-
-```{r singleyear, eval = FALSE, echo = FALSE}
-pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)
-
-names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
-```
-This is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' command.
-
-```{r sloppy, eval = FALSE, echo = FALSE}
-ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
-names(ITHIMageKey2) <- unique(pop14$WONDERage)
-pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
-```
-
-```{r foop, eval = FALSE, echo = FALSE}
-pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
-pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))
-
-popSummary <- ddply(pop14, .(HHSRegionCode, gender, ITHIMage), summarise,
-                    population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
-                   AllCauseDeaths2014 = sum(as.numeric(as.character(deaths2014AllCause)), na.rm=T)
-)
-
-HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
-```
-#### Read USGBD file (USGBD.csv)
-
-Presumably this comes from an EXCEL file.
-
-```{r USGBD, eval = FALSE, echo = FALSE}
-USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
-USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
-USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)
-
-burden <- merge(HHSmetroCauseSpecific, USGBD, all.x=T)
-burden$deaths <- burden$numberOfDeaths5yr/5
-## neil uses 10, and if deaths<10 sets ratio to 1
-burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
-burden$YLL <- burden$ratio * (burden$GBDyll/burden$GBDpop) * burden$population2014
-burden$YLD <- burden$ratio * (burden$GBDyld/burden$GBDpop) * burden$population2014
-burden$DALY <- burden$ratio * (burden$GBDdaly/burden$GBDpop) * burden$population2014
-
-burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
-levels(burden$sex)  <- c("M","F")
-
-burdenOut <- burden[,c("HHSRegionCode","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
-names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")
-
-burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
-burdenOut <- burdenOut[,c("HHS", "disease","sex","dproj","yll","yld","daly")]
-```
-
-Now we create an R list of GBD objects named by HHS region.
-
-```{r burdenList, eval = FALSE, echo = FALSE}
-GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
-#saveRDS(GBDList, file = "../data/GBD.rds")
-```
-
-## Example (National)
-
-```{r instantiate2, eval = FALSE, echo = FALSE}
-ITHIM.baseline <- createITHIM()
-ITHIM.scenario <- createITHIM()
-```
-
-```{r RRs2, eval = FALSE, echo = FALSE}
-comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-```
-
-### Deaths
-```{r show222, echo = FALSE, eval = FALSE}
-matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### DALY
-```{r show22, echo = FALSE, eval = FALSE}
-matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-### YLD
-```{r show23, echo = FALSE, eval = FALSE}
-matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-#### YLL
-```{r show24, echo = FALSE, eval = FALSE}
-matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-
-## Example
-
-```{r example, eval = FALSE, echo = FALSE}
-example("ITHIM-package")
-```
-
-### Creating Different Visions
-
-```{r vision, eval = FALSE, echo = FALSE}
-ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
-comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
-```
-#### YLL
-```{r show233, echo = FALSE, eval = FALSE}
-matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
-```
-
-## Interrogate Parameters
-
-```{r params, eval = FALSE, echo = FALSE}
-names(ITHIM.baseline$parameters)
-ITHIM.baseline$parameters$NonTravelMETs
-ITHIM.baseline$parameters$Rwt
-```
-
-## Appendix
-
-### DETAILS ON WONDER QUERY
-
-```
-Dataset: Underlying Cause of Death, 1999-2014
-Query Parameters:
-Title: ITHIM2014_HHS_urb_age_sex_16years
-2013 Urbanization: All
-Autopsy: All
-Five-Year Age Groups: All
-Gender: Female, Male
-HHS Regions: All
-Hispanic Origin: All
-ICD-10 113 Cause List: Malignant neoplasms of colon, rectum and anus (C18-C21), Malignant neoplasms of trachea, bronchus and
-lung (C33-C34), Malignant neoplasm of breast (C50), #Diabetes mellitus (E10-E14), #Alzheimer's disease (G30), Other forms of
-chronic ischemic heart disease (I20,I25), Acute and subacute endocarditis (I33), Diseases of pericardium and acute myocarditis
-(I30-I31,I40), #Essential hypertension and hypertensive renal disease
-(I10,I12,I15), #Cerebrovascular diseases (I60-I69), Influenza and pneumonia (J09-J18), Other acute lower respiratory infections (J20-J22,U04), #Chronic lower respiratory diseases
-(J40-J47), Other diseases of respiratory system (J00-J06,J30- J39,J67,J70-J98), Motor vehicle accidents
-(V02-V04,V09.0,V09.2,V12-V14,V19.0-V19.2,V19.4-V19.6,V20-V79,V80.3-V80.5,V81.0-V81.1,V82.0-V82.1,V83-V86,V87.0-V87.8,V88.0-V88.8
- V89.0,V89.2)
-Other land transport accidents
-(V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
- V89.1,V89.3,V89.9)
-Place of Death: All
-Race: All
-Weekday: All
-Year/Month: All
-Group By: Five-Year Age Groups, Gender, HHS Region, 2013 Urbanization, ICD-10 113 Cause List
-Show Totals: Disabled
-Show Zero Values: True
-Show Suppressed: True
-Calculate Rates Per: 100,000
-Rate Options: Default intercensal populations for years 2001-2009 (except Infant Age Groups)
----
-  Help: See http://wonder.cdc.gov/wonder/help/ucd.html for more information.
----
-  Query Date: Jul 19, 2016 4:46:10 PM
----
-  Suggested Citation: Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
-1999-2014 on CDC WONDER Online Database, released 2015. Data are from the Multiple Cause of Death Files, 1999-2014, as compiled
-from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program. Accessed at
-http://wonder.cdc.gov/ucd-icd10.html on Jul 19, 2016 4:46:10 PM
----
-  Messages:
-  1. Totals and Percent of Total are disabled when data are grouped by 113 or 130 Cause Lists. Check Caveats below for more
-information.
-2. Selections made to any variable in [HHS Regions, 2013 Urbanization] cause equivalent selections in the other variables. More
-Information: http://wonder.cdc.gov/wonder/help/faq.html#related-variables.
-3. Totals are not available for these results due to suppression constraints. More Information:
-  http://wonder.cdc.gov/wonder/help/faq.html#Privacy.
----
-  Footnotes:
-  1. A '#' symbol preceding the label indicates a rankable cause of death. More information.
----
-  Caveats:
-  1. Totals and Percent of Total are disabled when data are grouped by a 113 or 130 Cause List because both aggregate and detailed
-values are displayed in the table. Also be aware that charts and maps containing both aggregate and detail data could be
-misleading.
-2. Population and rates are labeled 'Not Applicable' when they include a subset of ages 85-100+ because populations are not
-available for those ages. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Ages 85-100.
-3. Circumstances in Georgia for the years 2008 and 2009 have resulted in unusually high death counts for the ICD-10 cause of
-death code R99, "Other ill-defined and unspecified causes of mortality." Caution should be used in interpreting these data.
-More information: http://wonder.cdc.gov/wonder/help/ucd.html#Georgia-Reporting-Anomalies.
-4. Circumstances in New Jersey for the year 2009 have resulted in unusually high death counts for the ICD-10 cause of death code
-R99, "Other ill-defined and unspecified causes of mortality" and therefore unusually low death counts in other ICD-10 codes,
-most notably R95, "Sudden Infant Death Syndrome" and X40-X49, "Unintentional poisoning." Caution should be used in
-interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#New-Jersey-Reporting-Anomalies.
-5. Circumstances in California resulted in unusually high death counts for the ICD-10 cause of death code R99, "Other
-ill-defined and unspecified causes of mortality" for deaths occurring in years 2000 and 2001. Caution should be used in
-interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#California-Reporting-Anomalies.
-6. Data are Suppressed when the data meet the criteria for confidentiality constraints. More information:
-  http://wonder.cdc.gov/wonder/help/ucd.html#Assurance of Confidentiality.
-7. Death rates are flagged as Unreliable when the rate is calculated with a numerator of 20 or less. More information:
-  http://wonder.cdc.gov/wonder/help/ucd.html#Unreliable.
-8. Deaths of persons with Age "Not Stated" are included in "All" counts and rates, but are not distributed among age groups,
-so are not included in age-specific counts, age-specific rates or in any age-adjusted rates. More information:
-  http://wonder.cdc.gov/wonder/help/ucd.html#Not Stated.
-9. The population figures for year 2014 are bridged-race estimates of the July 1 resident population, from the Vintage 2014
-postcensal series released by NCHS on June 30, 2015. The population figures for year 2013 are bridged-race estimates of the July
-1 resident population, from the Vintage 2013 postcensal series released by NCHS on June 26, 2014. The population figures for
-year 2012 are bridged-race estimates of the July 1 resident population, from the Vintage 2012 postcensal series released by NCHS
-on June 13, 2013. The population figures for year 2011 are bridged-race estimates of the July 1 resident population, from the
-Vintage 2011 postcensal series released by NCHS on July 18, 2012. Population figures for 2010 are April 1 Census counts. The
-population figures for years 2001 - 2009 are bridged-race estimates of the July 1 resident population, from the revised
-intercensal county-level 2000 - 2009 series released by NCHS on October 26, 2012. Population figures for 2000 sssssare April 1 Census
-counts. Population figures for 1999 are from the 1990-1999 intercensal series of July 1 estimates. Population figures for the
-infant age groups are the number of live births. <br/><b>Note:</b> Rates and population figures for years 2001 - 2009 differ
-slightly from previously published reports, due to use of the population estimates which were available at the time of release.
-10. The population figures used in the calculation of death rates for the age group 'under 1 year' are the estimates of the
-resident population that is under one year of age. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Age Group.
-```
-
-```{r appendix}
-date()
-sessionInfo()
-```
-
-<!--
-
-# Code Graveyard
-
-### Total Distribution
-
-```{r total, eval = FALSE, echo = FALSE}
-foo <- mapply(getTotalDistribution,
-  muTravel = matrix(35,nrow = 8, ncol = 2),
-  cvTravel = 1.72,
-  muNonTravel = matrix(100,nrow = 8, ncol = 2),
-  cvNonTravel = 1,
-  p = 1,
-  pWalk = rep(54.4/64.2,2),
-  vWalk = rep(2.88,2),
-  size = 1e5, SIMPLIFY = FALSE)
-```
-
-```{r vargo, echo = FALSE, eval = FALSE}
-library(ggplot2)
-library(reshape2)
-library(plyr)
-
-summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_2003.dat
-
-cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_2003.dat
-
-stateResultFile <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults.csv
-
-stateResultFile2 <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults_quintile.csv"
-
-regionResultFile <- "~/GHI/data//regionresults_quintile.csv" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/regionresults_quintile.csv
-
-#readLines("~/Box Documents/work/CoBenefits/MaggieShare/AmerTimeUse/atusact_2003.dat", n=10)
-# activity <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusact_2003.dat", header=T, skip=0, sep=",")
-
-summary <- read.table(summaryFile, header=T, skip=0, sep=",")
-
-# respondants <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusresp_2003.dat", header=T, skip=0, sep=",")
-
-# roster <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusrost_2003.dat", header=T, skip=0, sep=",")
-
-CPS <- read.table(cpsFile, header=T, skip=0, sep=",")
-
-CPS.2 <- CPS[,c("TUCASEID", "GESTFIPS", "GEREG","GEMETSTA")]
-CPS.2$IDCHAR <- as.character(CPS.2[,"TUCASEID"])
-
-summary.2 <- summary[, c("TUCASEID", "TEAGE",    "TESEX",    "t050203",  "t130101",  "t130102",  "t130103",  "t130104",  "t130106",  "t130108", "t130109", "t130111", "t130113",  "t130115", "t130116",  "t130117", "t130119",  "t130120",  "t130122",  "t130123", "t130124",  "t130126",  "t130127", "t130128", "t130130",  "t130131",  "t130132", "t130133",  "t130134", "t130135", "t130136" , "t130199" , "t139999")]
-summary.2$IDCHAR <- as.character(summary.2[,"TUCASEID"])
-
-summary.2a <- summary.2[,c(1:3, length(summary.2))]
-summary.2b <- summary.2[,c(5:length(summary.2)-1)]
-minutes <- rowSums(summary.2b)
-summary.2a$minutes <- minutes
-
-metkey <- as.matrix(c(3.0,7.3,5.0,6.5,7.5,3.5,5.8,7.8,6.0,8.0,3.8,7.8,8.0,5.3,7.3,9.8,6.3,7.0,7.0,5.0,6.8,3.0,3.5,5.5,6.0,4.0,6.0,3.0,3.0,3.0), nrow = 1, ncol=30)
-
-summary.2c <- mapply(`*`,summary.2b,metkey)
-mets <- rowSums(summary.2c)
-summary.2a$mets <- mets
-
-summary.2a$agecat <- ifelse(summary.2a$TEAGE <5, 1,
-                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, 2,
-                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, 3,
-                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, 4,
-                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, 5,
-                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, 6,
-                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, 7, 8
-)))))))
-
-summary.2a$agecat2 <- ifelse(summary.2a$TEAGE <5, "age0-4",
-                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, "age5-14",
-                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, "age15-29",
-                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, "age30-44",
-                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, "age45-59",
-                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, "age60-69",
-                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, "age70-79", "age80+"
-)))))))
-
-use <- merge(summary.2a, CPS.2, by.x="TUCASEID", by.y="TUCASEID", all.x=T, sorted=F)
-
-use$sexcat <- ifelse(use$TESEX == 1, "male", "female")
-
-use$regcat <- ifelse(use$GEREG == 1, "northeast",
-              ifelse(use$GEREG == 2, "midwest",
-              ifelse(use$GEREG == 3, "south",  "west")))
-
-use$metcat <- ifelse(use$GEMETSTA == 1, "metro",
-              ifelse(use$GEMETSTA == 2, "non-metro", "unclassified"))
-
-use <- use[,c("minutes","mets","agecat2","GESTFIPS","sexcat","regcat","metcat")]
-
-#head(use)
-
-use.region <- ddply(use, .(sexcat, agecat2, regcat, metcat), summarise,
-             mean_min = mean(minutes, na.rm=T),
-             mean_mets = mean(mets, na.rm=T),
-             n = length(minutes)
-)
-
-use.state <- ddply(use, .(sexcat, agecat2, GESTFIPS, metcat), summarise,
-             mean_min = mean(minutes, na.rm=T),
-             mean_mets = mean(mets, na.rm=T),
-             n = length(minutes)
-)
-
-use2 <- subset(use, metcat == "metro")
-
-use.state2 <- ddply(use2, .(sexcat, agecat2, GESTFIPS), summarise,
-              percentile10 <- quantile(mets, 0.10, na.rm=T),
-              percentile30 <- quantile(mets, 0.3, na.rm=T),
-              percentile50 <- quantile(mets, 0.5, na.rm=T),
-              percentile70 <- quantile(mets, 0.7, na.rm=T),
-              percentile90 <- quantile(mets, 0.9, na.rm=T),
-              mean <- mean(mets, na.rm=T),
-              n = length(mets)
-)
-
-use.region2 <- ddply(use2, .(sexcat, agecat2, regcat), summarise,
-              percentile10 <- quantile(mets, 0.10, na.rm=T),
-              percentile30 <- quantile(mets, 0.3, na.rm=T),
-              percentile50 <- quantile(mets, 0.5, na.rm=T),
-              percentile70 <- quantile(mets, 0.7, na.rm=T),
-              percentile90 <- quantile(mets, 0.9, na.rm=T),
-              mean <- mean(mets, na.rm=T),
-              n = length(mets)
-)
-
-state.table <- subset(use.state, metcat == "metro")
-
-#write.csv(state.table, stateResultFile, row.names=F)
-
-#write.csv(use.state2, stateResultFile2, row.names=F)
-
-#write.csv(use.region2, regionResultFile, row.names=F)
-
-#ggplot(use.region, aes(x=factor(sexcat), y=mean_mets, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1) +facet_grid(regcat ~ metcat)
-
-#ggplot(subset(use.region, metcat != "unclassified"), aes(x=factor(sexcat), y=mean_mets, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1, position="dodge") +facet_grid(regcat ~ metcat)
-
-#ggplot(subset(use.state, metcat != "unclassified" & GESTFIPS == 17), aes(x=factor(sexcat), y=mean_min, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1, position="dodge") +facet_grid(GESTFIPS ~ metcat)
-```
-
-```{r figures, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
-
-zeroMETs <- use$minutes == 0
-
-ggplot(use[!zeroMETs,], aes(x = mets, ..density.., colour = agecat2)) + geom_freqpoly(binwidth=200) + facet_grid(. ~ sexcat) + coord_cartesian(xlim = c(0, 1500))
-
-```
-
-```{r nonZeroMETsLogNormal, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
-parameters <- ddply(use[!zeroMETs,], .(sexcat, agecat2), summarise,
-             mean_min = mean(minutes, na.rm=T),
-             mean_mets = mean(mets, na.rm=T),
-             sd_min = sd(minutes, na.rm=T),
-             sd_mets = sd(mets, na.rm=T)
-)
-cv_mets <- with(parameters, sd_mets/mean_mets)
-cv_min <- with(parameters, sd_min/mean_min)
-cv0_mets<- mean(cv_mets)
-
-mean_mets <- parameters$mean_mets
-sd_mets <- cv0_mets*mean_mets
-
-getLogNormal <- function(mu,sd){
-    dlnorm(seq(0,2000,length.out=1e3), log(mu/sqrt(1+sd^2/mu^2)), sqrt(log(1+sd^2/mu^2)))
-}
-
-foo <- as.data.frame(mapply(getLogNormal, mean_mets, sd_mets))
-names(foo) <- paste0(parameters$sexcat,parameters$agecat2)
-foo <- melt(foo)
-foo <- data.frame(foo, x = seq(0,2000,length.out = 1e3), sex = c(matrix(c("F","M"),nrow= nrow(foo)/2, ncol = 2, byrow = TRUE)), agecat = c(matrix(c("15-29","30-44","45-59","60-69","70-79", "80+"),nrow= nrow(foo)/12, ncol = 12, byrow = TRUE)))
-
-ggplot(foo, aes(x,value, colour = agecat)) + geom_line() + facet_grid( . ~ sex) + coord_cartesian(xlim = c(0, 1500))
-
-```
-
-```{r matrices, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
-nonTravelMETs <- rbind(0,rbind(0,cbind(parameters$mean_mets[7:12], parameters$mean_mets[1:6])))
-nonTravelMETsNormalized <- nonTravelMETs/nonTravelMETs[3,2]
-meanNonTravelMETs <- nonTravelMETs[3,2]
-
-```
-
-```{r plotMatrix, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
-D <- melt(nonTravelMETsNormalized)
-ggplot(D, aes(Var2, value)) + geom_bar(aes(fill=Var1), stat = "identity", position = "dodge")
-```
-
-```{r fooFunction, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
-probActive <- 0.15
-METQuintiles <- mapply(getMETQuintiles, nonTravelMETsNormalized*nonTravelMETs, MoreArgs = list(cv = cv0_mets, p = probActive, size = 1e4), SIMPLIFY = FALSE)
-```
-
-```{r NHTSTables, eval = FALSE}
-HHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/HHV2PUB.CSV", na.strings = "-9")
-names(HHV2PUB) <- c("HOUSEID","VARSTRAT","WTHHFIN","DRVRCNT","CDIVMSAR","CENSUS_D","CENSUS_R","HH_HISP","HH_RACE","HHFAMINC","HHRELATD","HHRESP","HHSIZE","HHSTATE","HHSTFIPS","HHVEHCNT","HOMEOWN","HOMETYPE","MSACAT","MSASIZE","NUMADLT","RAIL","RESP_CNT","SCRESP","TRAVDAY","URBAN","URBANSIZE","URBRUR","WRKCOUNT","TDAYDATE","FLAG100","LIF_CYC","CNTTDHH","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
-
-VEHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/VEHV2PUB.CSV", na.strings = "-9")
-names(VEHV2PUB) <- c("HOUSEID","WTHHFIN","VEHID","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","CDIVMSAR","CENSUS_D","CENSUS_R","HHSTATE","HHSTFIPS","HYBRID","MAKECODE","MODLCODE","MSACAT","MSASIZE","OD_READ","RAIL","TRAVDAY","URBAN","URBANSIZE","URBRUR","VEHCOMM","VEHOWNMO","VEHYEAR","WHOMAIN","WRKCOUNT","TDAYDATE","VEHAGE","PERSONID","HH_HISP","HH_RACE","HOMEOWN","HOMETYPE","LIF_CYC","ANNMILES","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","BEST_FLG","BESTMILE","BEST_EDT","BEST_OUT","FUELTYPE","GSYRGAL","GSCOST","GSTOTCST","EPATMPG","EPATMPGF","EIADMPG","VEHTYPE","HH_CBSA","HHC_MSA")
-```
-
-```{r NHTSArchive, echo = FALSE, eval = FALSE}
-nMatrix <- ddply( subset(NHTS.person, !is.na(AGE) & !is.na(SEX)), .(SEX, AGE), summarise, count = length(PERSONID))
-sum(nMatrix$count) # 305649 number of people in the PERSON table
-
-foo <- group_by(NHTS.trips, HOUSEID,PERSONID)
-foobar <- dplyr::summarise(foo, n = n())
-nrow(foobar) # 262,934 number of people with trip data
-
-NHTS %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarise( n = n()) # 260,576 people in merged data
-ggplot(NHTS, aes(x = TRVL_MIN, ..density.., colour = AGE)) + geom_freqpoly(binwidth = 15) + facet_grid( SEX ~ AT) + labs(x = "Foo", colour = "Age") + coord_cartesian(xlim = c(0, 75))
-NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))
-```
-
-### Relative Risk Model
-```{r explore, eval = FALSE}
-names(ITHIM.baseline$parameters)
-```
-```{r RRFigure, echo = FALSE, eval = FALSE}
-phi <- function(x,k=0.5,phi0=comparativeRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
-plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
-lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
-lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
-lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
-legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
-```
-
-### Tinkering
-
-Note that some of the AF values will disagree with the EXCEL
-spreadsheet because Diabetes, HHD_Air, CVD_Air and Stroke_Air do not
-combine travel METs with non-travel METs.
-
-```{r parlist, eval = FALSE}
-AF.sum <- numeric(1e3); i <- 0;
-muwt.vec <- seq(107.1,150,length.out=length(AF.sum))
-#muct.vec <- seq(2.7,3.0,length.out=length(AF.sum))
-for(muwt in muwt.vec){
-i <- i+1
-parameters <- createParameterList(baseline = FALSE)
-
-parameters <- setParameter(parName="muwt", parValue = muwt, parList = parameters)
-ITHIM.scenario <- list(
-  parameters = parameters <- parameters,
-  means = means <- computeMeanMatrices(parameters),
-  quintiles = quintiles <- getQuintiles(means)
-)
-comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
-AF.sum[i] <- with(comparativeRisk, sum(100*AF$BreastCancer))
-}
-D <- data.frame(mu.wt = muwt.vec,AF.sum=AF.sum)
-p <- ggplot(D, aes(x = mu.wt, y = AF.sum))
-p + geom_point()
-```
-
-```{r AF, echo = FALSE, eval = FALSE}
-with(comparativeRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
-```
-
-To see the help page for a function use help().
-```{r help, eval = FALSE}
-help(createParameterList)
-```
-
-### Example
-To run the example contained in the _ITHIM_ package run the following.
-
-```{r package, eval = FALSE}
-example("ITHIM-package")
-```
-### Figures
-
-```{r plot2, echo = FALSE, fig.width = 12, eval = FALSE}
-plotRR(comparativeRisk$RR.baseline$Depression,comparativeRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
-```
-
-```{r plot3, echo = FALSE, fig.width = 12, eval = FALSE}
-plotRR(comparativeRisk$RR.baseline$BreastCancer,comparativeRisk$RR.scenario$BreastCancer) + coord_cartesian(ylim = c(0.75, 1.05)) + ggtitle("Breast Cancer")
-```
-
-```{r plot4, echo = FALSE, fig.height = 12, eval = FALSE}
-D<-melt(comparativeRisk$AF)
-names(D)<- c("age","sex","AF","disease")
-p <- ggplot(D, aes(age,AF)) + geom_bar(aes(fill=sex), stat = "identity") + facet_grid( disease ~ .) + coord_cartesian(ylim = c(0, 0.1))
-p
-```
-### Road Injury Array
-```{r road, eval = FALSE}
-baselineInjuries <- array(dim=c(6,7,3,2))
-injury <- c("fatal","serious")
-road <- c("minor road","major road","motorway")
-vehicle <- c("walk","cycle","bus","car","HGV","mbike")
-dimnames(baselineInjuries) <- list(vehicle,c(vehicle,"NOV"),road,injury)
-
-fatalities <- array(c(0,0,0.6666666667,34,6.3333333333,0,0.3333333333,0,0,0,0,0,4.3333333333,0,0,0.3333333333,0,0,0,0,0.3333333333,0.3333333333,0,0,0.6666666667,0,0.3333333333,0,0,0,48.6666666667,2,0,5.6666666667,0,35.6666666667,0,0,0,2.3333333333,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,2,0,0,0,0,6.3333333333,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,11.3333333333,1,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,18.6666666667,1.3333333333,0,2,0,7.3333333333,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0.3333333333,0,0.3333333333,0.3333333333,20,3,0,0,0,0,0,0.3333333333,0.3333333333,20,3,0,0,0,0,0,0,0,1,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,42.3333333333,4.6666666667,0,2,0,26.3333333333,0,0,0,3.3333333333,0.3333333333,0,0.3333333333,0,1,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,3,0,0,0,6.3333333333,0,0,0.6666666667,0,0),dim=c(9,8,3))
-
-fatalaties <- apply(fatalities,c(1,3),t)
-
-serious <- array(c(0,0,0,6.3333333333,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,40,0.6666666667,0,0.6666666667,0,11.6666666667,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0.3333333333,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,8,0.3333333333,0,0.6666666667,0,3.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.6666666667,0.3333333333,0,0,0,0,0,0,0,0.6666666667,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,29.6666666667,1,0,0,0,11,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0),dim=c(9,8,3))
-
-serious <- apply(serious,c(1,3),t)
-
-baselineInjuries[,,,1] <- fatalaties
-baselineInjuries[,,,2] <- serious
-
-Data.RoadType <- data.frame(local=c(0.75,0.53,0.0,0.09,0.29,NA,0.09,NA), arterial=c(0.25,0.47,0.96,0.31,0.42,NA,0.31,NA), highway=c(0.0,0,0.04,0.60,0.29,NA,0.60,NA),row.names=c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike") )
-
-Data.Occupancy <- data.frame(occupancy = c(1.0,1.0,18.9,1.3,1.0,NA,1.0,NA), row.names = c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike"))
-
-```
-
-### Active Transport Estimates (Speed and Distance by Transport Mode)
-```{r plot1, echo = FALSE, eval = FALSE}
-D <- melt(ITHIM.baseline$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
-p <- ggplot(D, aes(age,  activeTransportTime)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
-p + coord_cartesian(ylim = c(0.75, 1.05))
-```
-
-These data were ripped out of the EXCEL workbook and vary per scenario.
-```{r transportData, eval = FALSE, echo = FALSE}
-
-transportMode <- c("walk","cycle","bus","minibus","train","car.driver","car.passenger","mbike","elec.bike.scooter","LGV","HGV.rigid","HGV.articulated")
-
-speed <- c(1.35299774571063,5.39374775150058,5.22298847623309,NA,9.94147658251227,17.36906017290580,14.51012134295660,NA,NA,NA,NA,NA)
-
-dist <- c(2.44072161538462,1.19969769230769,4.37851867307692,NA,5.58545353846154,111.91985351923100,39.12264265384620,0.00000000000000,0.00000000000000,NA,1.05479452054795,NA) # miles per week
-
-Data.Speed <- data.frame(speed,dist, row.names = transportMode)
-```
-
-```{r headDataSpeed, eval = FALSE}
-Data.Speed
-```
-
-```{r transportTime, echo = FALSE, results = "hide", eval = FALSE}
-head(transportTime <- Data.Speed$dist/Data.Speed$speed*60)
-```
-
-### Emissions Data
-
-```{r roadAndCO2Data, echo = FALSE, eval = FALSE, results = "hide"}
-
-CO2 <- c(957.686509527760,221.578783667454,203.624526028894,247.869842164617,333.468424533125,90.000000000000,835.014773125786)
-
-Data.CO2 <- data.frame(CO2, row.names = c("bus","Cars.urban","Cars.mway","LGVs.urban","LGVs.mway","motorbike","HGV"))
-
-Data.RoadType <- data.frame(local=c(0.75,0.53,0.0,0.09,0.29,NA,0.09,NA), arterial=c(0.25,0.47,0.96,0.31,0.42,NA,0.31,NA), highway=c(0.0,0,0.04,0.60,0.29,NA,0.60,NA),row.names=c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike") )
-
-Data.Occupancy <- data.frame(occupancy = c(1.0,1.0,18.9,1.3,1.0,NA,1.0,NA), row.names = c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike"))
-
-```
-
-The amount of carbon emissions varies by transport mode and road type.
-To accomodate this we need the distribution of road type by transport
-mode.  At some point we need the expected number of passengers per
-transport mode and we create it here, arbitrarily perhaps.
-
-```{r showData, echo = TRUE, eval = FALSE, results = "show"}
-Data.CO2
-Data.RoadType
-Data.Occupancy
-```
-
-### Non-(Active Transport) METs
-
-The file NonTravelMETs.csv was created from the tripleWin EXCEL file.
-These data are used in a table look-up for a given county.  The data
-represent the non-travel-related energy expenditures as measured in
-METs.  The data are stratified by age and sex.  Variable names were
-not given for the 7 numerical variables listed.  Here we name them x1
-through x7.
-
-```{r nonTravelMETs, eval = FALSE, echo = FALSE}
-nonTravelMETs <- data.frame(scan(file="./data/NonTravelMETs.csv", sep = ",", what = list("","","","",1,1,1,1,1,1,1), skip = 1))
-names(nonTravelMETs) <- c("ID","sexcat","agecat2","GESTFIPS", paste0("x",1:7))
-```
-
-```{r ggplot, fig.width=12, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE, results = "hide", eval = FALSE}
-D <- melt(nonTravelMETs)
-p <- ggplot(D, aes(x = variable, y = value, colour = agecat2), outlier.colour = "red", outlier.shape = 1 )
-p + geom_boxplot()
-```
-
-Who is the outlier?  Perhaps it is a cohort of laborers, e.g., a county of coal miners?
-
-```{r foo, eval = FALSE, echo = TRUE, eval = FALSE}
-nonTravelMETs[which.max(nonTravelMETs$x5),]
-```
-
-Strange that it is group of middle-aged men.  I see this as a red
-flag.  I remove this datum and redraw the boxplots, but this time only
-for x5, x6 and x7, the others seemed less interesting.
-
-```{r fips, eval = FALSE, echo = FALSE, results = "hide"}
-fips <- read.csv("./data/fips_codes_website.csv")
-any(fips$County.FIPS.Code==38)
-```
-
-```{r Nontravelmets, echo = FALSE, message = FALSE, warning = FALSE, results = "hide", fig.width=12, fig.height=6, eval = FALSE}
-D <- melt(nonTravelMETs[-which.max(nonTravelMETs$x5),-(5:8)])
-p <- ggplot(D, aes(x = variable, y = value, colour = agecat2), outlier.colour = "red", outlier.shape = 1 ) + geom_boxplot()
-
-p + facet_grid(. ~ sexcat)
-```
-
-Thoughts?
-
-### Relative Risk of MET Exposure
-
-In the worksheet _Phy activity RRs_ we see estimates for relative risk
-for various disease outcomes.  My guess is that some transformation is
-being made to convert the exposure from active transport time to METs.
-I don't know why this is needed yet.  I can't make much progress here
-until I understand that better.
-
-### Coefficient of Variation
-
-An estimate for the coefficient of variation is provided without much
-explanation.  The literature cites an empirical result.  The feeling
-is that as mean active transport increases in the population people
-are fitter and there is less variance in active transport time.  I
-guess this has been shown in the same refernce for the coefficient of
-variation estimate.
-
-### Vital Statistics Data
-
-#### 2001 Data with Vargo Script
-
-```{r vitalstats, eval = FALSE, echo = FALSE}
-source("../vargo/process_2001_vital_stats.r")
-saveRDS(mortality2001, file = "./data/mortality2001.rds")
-```
-
-#### 2011 (VS11MORT.DUSMCPUB)
-
-Lets read in the CDC data using the file VS11MORT.DUSMCPUB and see whats up.
-
-```{r VS11, echo = FALSE, eval = FALSE}
-tapePosition <- list(state = c(29:30), skippedVars = c(31:34), county = c(35:37), skippedVars = c(38:68), sex = 69, ageUnit = 70, ageValue = c(71:73), skippedVars = c(74:145), ICD = c(146:149), skippedVars = c(150:444), race = c(445:446)) # Magic number!!!
-colClasses <- c("factor","factor","factor", "factor","integer", "character", "factor")
-fileWidth <- 488 # Magic number!!!
-```
-
-```{r VS11Classes, echo = FALSE, eval = FALSE}
-names(colClasses) <- names(tapePosition)[-which(names(tapePosition)=="skippedVars")]
-colClasses
-```
-
-To include more variables edit the above lines.  The variable listed
-must be contiguous.
-
-```{r convertAge, warning = FALSE, eval = FALSE, echo = FALSE}
-
-varWidths <- unlist(lapply(tapePosition, "length"))
-varWidths <- ifelse(names(tapePosition) == "skippedVars", -varWidths, varWidths)
-widths <- c(-(min(tapePosition[[1]])-1),c(varWidths))
-widths <- c(widths,-(fileWidth-sum(abs(widths))))
-
-Data.Mortality <- read.fwf(file = "~/ITHIM/R/data/VS11MORT.DUSMCPUB.first100", widths = widths, col.names = names(colClasses), header = FALSE, colClasses = colClasses)
-Data.Mortality <- data.frame(Data.Mortality, convertedAge = with(Data.Mortality, convertAge(value = ageValue, unit = ageUnit)))
-
-p <- ggplot(Data.Mortality, aes(x = convertedAge)) + geom_histogram(binwidth = 1)
-p + facet_grid(. ~ sex)
-
-```
-
-```{r ICD, fig.width = 12, echo = FALSE, eval = FALSE}
-Data.IDC.Codes <- read.csv(file="./data/IDC_Codes_key.csv", header = TRUE, colClasses = c("character","character","character"), na.string = "")
-Data.IDC.Codes <- subset(Data.IDC.Codes, Include == "1" & !is.na(ITHIMCause))
-ITHIMCauseOfDeath <- Data.IDC.Codes$ITHIMCause
-names(ITHIMCauseOfDeath) <- Data.IDC.Codes$cause.of.death
-Data.Mortality <- data.frame(Data.Mortality, CauseofDeathITHIM = ITHIMCauseOfDeath[Data.Mortality$ICD])
-Data.Mortality <- subset(Data.Mortality, !is.na(CauseofDeathITHIM) & !is.na(convertedAge))
-D <- data.frame(disease = Data.Mortality$CauseofDeathITHIM)
-p <- ggplot(D, aes(disease))
-p + geom_bar()
-
-```
-
-```{r CoDTable, eval = FALSE}
-with(Data.Mortality, table(CauseofDeathITHIM))
-```
-
-```{r ageClass, fig.width = 12, echo = FALSE, eval = FALSE}
-ageClass <- c(rep("00-04",5),rep("05-14",10),rep("15-29",15),rep("30-44",15),rep("45-59",15),rep("60-69",10),rep("70-79",10),rep("80+",121))
-names(ageClass) <- as.character(0:200)
-Data.Mortality <- data.frame(Data.Mortality, ageClass = as.factor(ageClass[as.character(as.integer(Data.Mortality$convertedAge))]))
-with(Data.Mortality, table(ageClass))
-
-D <- data.frame(ageClass = Data.Mortality$ageClass, disease = Data.Mortality$CauseofDeathITHIM)
-p <- ggplot(D, aes(disease))
-p + geom_bar(aes(fill=ageClass))
-```
-
-```{r histbyDisease, fig.width = 12, echo = FALSE, eval = FALSE}
-D <- data.frame(convertedAge = Data.Mortality$convertedAge, disease = Data.Mortality$CauseofDeathITHIM)
-p <- ggplot(D, aes(convertedAge, ..density.., colour = disease)) + geom_histogram(binwidth=5, show.legend = TRUE)
-p + facet_grid( . ~ disease)
-```
-
-```{r counties, eval = FALSE}
-Data.AgeSex <- read.csv("./data/county_age_sex2002.csv", header=T, col.names = c("county.name","state.name", "MSA","sex","all.ages","age0-4","age5-14","age15-29","age30-44","age45-59","age60-69","age70-79","age80+", "STATE","COUNTY"), colClasses = c("factor","factor","factor","factor","integer","integer","integer","integer","integer","integer","integer","integer","integer", "factor", "factor"))
-
-age.sex.2 <- melt(age.sex[,c(14,15,4,6:13)], id=c("STATE","COUNTY","sex"))
-
-#age.sex.3 <- dcast(age.sex.2, STATE + COUNTY + variable ~ sex, mean)
-
-mort.age <- merge(mort.ITHIM, age.sex.3, by.x=c("state","county", "ITHIM.age"), by.y=c("STATE","COUNTY","variable"))
-
-mort.age$year <- 2001
-
-mort.age$app.pop <- ifelse(mort.age$sex ==1, mort.age$male, mort.age$female)
-
-```
-
-```{r tabulate, eval = FALSE}
-Data.Mortality.Table <- as.data.frame(ftable(with(Data.Mortality, table(sex,ageClass,CauseofDeathITHIM))))
-```
-
-```{r saveMortality, echo = FALSE, eval = FALSE}
-saveRDS(object = Data.Mortality, file = "./data/Data.Mortality.rds")
-saveRDS(object = Data.Mortality.Table, file = "./data/Data.Mortality.Table.rds")
-```
--->

### 2016-10-19 R/appendix.Rmd
Created two new figures
Also created appendix.Rnw to store old R code.

# Appendix

## Non-Travel METs

```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
dataColumns <- c(1,5)
D <- newMETs[,dataColumns]
metaData <- newMETs[,-dataColumns]
newMETs <- list(Data = D, metaData = metaData)
activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))

rownames(newMETs$Data) <- activityCode
rownames(newMETs$metaData) <- activityCode

newMETs$metaData <- within(newMETs$metaData, {
  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
})

metKey <- newMETs$Data$AinsworthMETS
names(metKey) <- activityCode

```

```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat"
cps <- read.table(cpsFile, header=T, skip=0, sep="," )
cps <- within(cps, {
    TUCASEID <- as.character(TUCASEID)
    TULINENO <- as.character(TULINENO)
    GEREG <- as.factor(GEREG)
    })
cps <- subset(cps, TULINENO == "1")

summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat"
summary <- read.table(summaryFile, header=T, skip=0, sep=",")
metaDataColumns <- 1:24
summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])

summaryList$metaData <- within(summaryList$metaData,{
    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
    AGECAT <- ifelse(TEAGE <5, "00-04",
                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
                            )))))))})

```

We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.

```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
allActivities <- colnames(summaryList$Data)
nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]

activities <- nonTravelActivities

metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
names(metKeyVector) <- activities
metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
dimnames(metKeyMatrix) <- list( NULL, activities)

METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )

D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))

```

```{r parameters, eval = FALSE, echo = FALSE}
parameters <- ddply(D, .(SEX, AGECAT), summarise,
             mean = mean(METs, na.rm=T),
             sd = sd(METs, na.rm=T)
             )

METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
METmatrix <- METmatrix/METmatrix[3,2]
```

```{r roadInjuries, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
convertState <- function( stateCode ){

stateCode <- as.character(stateCode)

stateCodeMatrix <- matrix(c("Alabama","1","Nebraska","31","Alaska","2","Nevada","32","Arizona","4","New Hampshire","33","Arkansas","5","New Jersey","34","California","6","New Mexico","35","Colorado","8","New York","36","Connecticut","9","North Carolina","37","Delaware","10","North Dakota","38","District of Columbia","11", "Ohio", "39","Florida","12","Oklahoma","40","Georgia","13","Oregon","41","Hawaii","15","Pennsylvania","42","Idaho","16","Puerto Rico","43","Illinois","17","Rhode Island","44","Indiana","18","South Carolina","45","Iowa","19","South Dakota","46","Kansas","20","Tennessee","47","Kentucky","21","Texas","48","Louisiana","22","Utah","49","Maine","23","Vermont","50","Maryland","24","Virgin Islands (since 2004)","52","Massachusetts","25","Virginia","51","Michigan","26","Washington","53","Minnesota","27","West Virginia","54","Mississippi","28","Wisconsin","55","Missouri","29","Wyoming","56","Montana","30"), byrow = TRUE, ncol = 2)

stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)

stateNames <- stateCodeMatrix[,1]
names(stateNames) <- stateCodeMatrix[,2]

stateAbbrevs <- stateAbbrevMatrix[,2]
names(stateAbbrevs) <- stateAbbrevMatrix[,1]

return(stateAbbrevs[stateNames[stateCode]])

}

roadFatalaties <- read.csv(file = "./../data/FARS/BikePedDeaths99-13.csv", na.strings = c("998","999"))
roadFatalaties <- within(roadFatalaties, {
  ageClass <- convertToAgeClass(AGE)
  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
  SEX <- factor(SEX, levels = c("F","M"))
  STATE <- as.factor(convertState(STATE))
})
roadFatalaties <- subset(roadFatalaties, !is.na(SEX) ) #& STATE %in% c("California", "New York"))

p <- ggplot(roadFatalaties, aes(Type, fill = STATE))
p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5))# + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) #+ facet_grid( Type ~ ., scales = "free_y")

walkTime <- unlist(lapply(meanWalkList, function(x) x[3,2]))

foo <- data.frame(with(roadFatalaties, table(STATE,Type)))
within(subset(foo, Type == "Ped"), Walk <- walkTime[as.character(STATE)])

```

### CDC Mortality

Next we parse the output of a CDC Wonder query to estimate mortality rates.

The data files in ~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/.

#### Query Results File (HHSmetro_Age_sex_2010-14.csv)

```{r CDC, eval = FALSE, echo = FALSE}
wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")

wonder <- subset(wonder, WONDERageCode != "NS") # remove 'Not stated' Age entries
```

The following two recoding chunks should probably be redone.  They
depend on the ordering of the factors which seems dangerous.

```{r ITHIMCause, eval = FALSE, echo = FALSE}
ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
names(ITHIMcauseKey) <- unique(wonder$ICD10)
wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
```
```{r recodeAge, eval = FALSE, echo = FALSE}
ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
names(ITHIMageKey) <- unique(wonder$WONDERage)
wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
```

Depression was not in the WONDER data, since there are no deaths, have to add it manually.

```{r wonder, eval = FALSE, echo = FALSE}
wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
wonder <- rbind(wonder, depress)
```

Replace suppressed values with random number 1-9

```{r suppressed, eval = FALSE, echo = FALSE, warning = FALSE}
wonder <- within( wonder, {
    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
    })
```

10 HHS x 2 sex x 8 age x 13 disease = 2,080 strata.

```{r summary2, eval = FALSE, echo = FALSE}
wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
)
```

#### Single Year Data File (HHSmetro_Age_sex_2014_AllCause.csv)

Read in data for single year population.  Processed in Excel, to
combine two WONDER queries, need to go back and script the processing
of the raw WONDER output

```{r singleyear, eval = FALSE, echo = FALSE}
pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)

names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
```
This is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' command.

```{r sloppy, eval = FALSE, echo = FALSE}
ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
names(ITHIMageKey2) <- unique(pop14$WONDERage)
pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
```

```{r foop, eval = FALSE, echo = FALSE}
pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))

popSummary <- ddply(pop14, .(HHSRegionCode, gender, ITHIMage), summarise,
                    population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
                   AllCauseDeaths2014 = sum(as.numeric(as.character(deaths2014AllCause)), na.rm=T)
)

HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
```
#### Read USGBD file (USGBD.csv)

Presumably this comes from an EXCEL file.

```{r USGBD, eval = FALSE, echo = FALSE}
USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)

burden <- merge(HHSmetroCauseSpecific, USGBD, all.x=T)
burden$deaths <- burden$numberOfDeaths5yr/5
## neil uses 10, and if deaths<10 sets ratio to 1
burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
burden$YLL <- burden$ratio * (burden$GBDyll/burden$GBDpop) * burden$population2014
burden$YLD <- burden$ratio * (burden$GBDyld/burden$GBDpop) * burden$population2014
burden$DALY <- burden$ratio * (burden$GBDdaly/burden$GBDpop) * burden$population2014

burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
levels(burden$sex)  <- c("M","F")

burdenOut <- burden[,c("HHSRegionCode","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")

burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
burdenOut <- burdenOut[,c("HHS", "disease","sex","dproj","yll","yld","daly")]
```

Now we create an R list of GBD objects named by HHS region.

```{r burdenList, eval = FALSE, echo = FALSE}
GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
#saveRDS(GBDList, file = "../data/GBD.rds")
```

## Example (National)

```{r instantiate2, eval = FALSE, echo = FALSE}
ITHIM.baseline <- createITHIM()
ITHIM.scenario <- createITHIM()
```

```{r RRs2, eval = FALSE, echo = FALSE}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

### Deaths
```{r show222, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### DALY
```{r show22, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLD
```{r show23, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLL
```{r show24, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

## Example

```{r example, eval = FALSE, echo = FALSE}
example("ITHIM-package")
```

### Creating Different Visions

```{r vision, eval = FALSE, echo = FALSE}
ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
```
#### YLL
```{r show233, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

## Interrogate Parameters

```{r params, eval = FALSE, echo = FALSE}
names(ITHIM.baseline$parameters)
ITHIM.baseline$parameters$NonTravelMETs
ITHIM.baseline$parameters$Rwt
```

## Appendix

### DETAILS ON WONDER QUERY

```
Dataset: Underlying Cause of Death, 1999-2014
Query Parameters:
Title: ITHIM2014_HHS_urb_age_sex_16years
2013 Urbanization: All
Autopsy: All
Five-Year Age Groups: All
Gender: Female, Male
HHS Regions: All
Hispanic Origin: All
ICD-10 113 Cause List: Malignant neoplasms of colon, rectum and anus (C18-C21), Malignant neoplasms of trachea, bronchus and
lung (C33-C34), Malignant neoplasm of breast (C50), #Diabetes mellitus (E10-E14), #Alzheimer's disease (G30), Other forms of
chronic ischemic heart disease (I20,I25), Acute and subacute endocarditis (I33), Diseases of pericardium and acute myocarditis
(I30-I31,I40), #Essential hypertension and hypertensive renal disease
(I10,I12,I15), #Cerebrovascular diseases (I60-I69), Influenza and pneumonia (J09-J18), Other acute lower respiratory infections (J20-J22,U04), #Chronic lower respiratory diseases
(J40-J47), Other diseases of respiratory system (J00-J06,J30- J39,J67,J70-J98), Motor vehicle accidents
(V02-V04,V09.0,V09.2,V12-V14,V19.0-V19.2,V19.4-V19.6,V20-V79,V80.3-V80.5,V81.0-V81.1,V82.0-V82.1,V83-V86,V87.0-V87.8,V88.0-V88.8
 V89.0,V89.2)
Other land transport accidents
(V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
 V89.1,V89.3,V89.9)
Place of Death: All
Race: All
Weekday: All
Year/Month: All
Group By: Five-Year Age Groups, Gender, HHS Region, 2013 Urbanization, ICD-10 113 Cause List
Show Totals: Disabled
Show Zero Values: True
Show Suppressed: True
Calculate Rates Per: 100,000
Rate Options: Default intercensal populations for years 2001-2009 (except Infant Age Groups)
---
  Help: See http://wonder.cdc.gov/wonder/help/ucd.html for more information.
---
  Query Date: Jul 19, 2016 4:46:10 PM
---
  Suggested Citation: Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
1999-2014 on CDC WONDER Online Database, released 2015. Data are from the Multiple Cause of Death Files, 1999-2014, as compiled
from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program. Accessed at
http://wonder.cdc.gov/ucd-icd10.html on Jul 19, 2016 4:46:10 PM
---
  Messages:
  1. Totals and Percent of Total are disabled when data are grouped by 113 or 130 Cause Lists. Check Caveats below for more
information.
2. Selections made to any variable in [HHS Regions, 2013 Urbanization] cause equivalent selections in the other variables. More
Information: http://wonder.cdc.gov/wonder/help/faq.html#related-variables.
3. Totals are not available for these results due to suppression constraints. More Information:
  http://wonder.cdc.gov/wonder/help/faq.html#Privacy.
---
  Footnotes:
  1. A '#' symbol preceding the label indicates a rankable cause of death. More information.
---
  Caveats:
  1. Totals and Percent of Total are disabled when data are grouped by a 113 or 130 Cause List because both aggregate and detailed
values are displayed in the table. Also be aware that charts and maps containing both aggregate and detail data could be
misleading.
2. Population and rates are labeled 'Not Applicable' when they include a subset of ages 85-100+ because populations are not
available for those ages. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Ages 85-100.
3. Circumstances in Georgia for the years 2008 and 2009 have resulted in unusually high death counts for the ICD-10 cause of
death code R99, "Other ill-defined and unspecified causes of mortality." Caution should be used in interpreting these data.
More information: http://wonder.cdc.gov/wonder/help/ucd.html#Georgia-Reporting-Anomalies.
4. Circumstances in New Jersey for the year 2009 have resulted in unusually high death counts for the ICD-10 cause of death code
R99, "Other ill-defined and unspecified causes of mortality" and therefore unusually low death counts in other ICD-10 codes,
most notably R95, "Sudden Infant Death Syndrome" and X40-X49, "Unintentional poisoning." Caution should be used in
interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#New-Jersey-Reporting-Anomalies.
5. Circumstances in California resulted in unusually high death counts for the ICD-10 cause of death code R99, "Other
ill-defined and unspecified causes of mortality" for deaths occurring in years 2000 and 2001. Caution should be used in
interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#California-Reporting-Anomalies.
6. Data are Suppressed when the data meet the criteria for confidentiality constraints. More information:
  http://wonder.cdc.gov/wonder/help/ucd.html#Assurance of Confidentiality.
7. Death rates are flagged as Unreliable when the rate is calculated with a numerator of 20 or less. More information:
  http://wonder.cdc.gov/wonder/help/ucd.html#Unreliable.
8. Deaths of persons with Age "Not Stated" are included in "All" counts and rates, but are not distributed among age groups,
so are not included in age-specific counts, age-specific rates or in any age-adjusted rates. More information:
  http://wonder.cdc.gov/wonder/help/ucd.html#Not Stated.
9. The population figures for year 2014 are bridged-race estimates of the July 1 resident population, from the Vintage 2014
postcensal series released by NCHS on June 30, 2015. The population figures for year 2013 are bridged-race estimates of the July
1 resident population, from the Vintage 2013 postcensal series released by NCHS on June 26, 2014. The population figures for
year 2012 are bridged-race estimates of the July 1 resident population, from the Vintage 2012 postcensal series released by NCHS
on June 13, 2013. The population figures for year 2011 are bridged-race estimates of the July 1 resident population, from the
Vintage 2011 postcensal series released by NCHS on July 18, 2012. Population figures for 2010 are April 1 Census counts. The
population figures for years 2001 - 2009 are bridged-race estimates of the July 1 resident population, from the revised
intercensal county-level 2000 - 2009 series released by NCHS on October 26, 2012. Population figures for 2000 sssssare April 1 Census
counts. Population figures for 1999 are from the 1990-1999 intercensal series of July 1 estimates. Population figures for the
infant age groups are the number of live births. <br/><b>Note:</b> Rates and population figures for years 2001 - 2009 differ
slightly from previously published reports, due to use of the population estimates which were available at the time of release.
10. The population figures used in the calculation of death rates for the age group 'under 1 year' are the estimates of the
resident population that is under one year of age. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Age Group.
```

```{r appendix}
date()
sessionInfo()
```

<!--

# Code Graveyard

### Total Distribution

```{r total, eval = FALSE, echo = FALSE}
foo <- mapply(getTotalDistribution,
  muTravel = matrix(35,nrow = 8, ncol = 2),
  cvTravel = 1.72,
  muNonTravel = matrix(100,nrow = 8, ncol = 2),
  cvNonTravel = 1,
  p = 1,
  pWalk = rep(54.4/64.2,2),
  vWalk = rep(2.88,2),
  size = 1e5, SIMPLIFY = FALSE)
```

```{r vargo, echo = FALSE, eval = FALSE}
library(ggplot2)
library(reshape2)
library(plyr)

summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_2003.dat

cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_2003.dat

stateResultFile <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults.csv

stateResultFile2 <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults_quintile.csv"

regionResultFile <- "~/GHI/data//regionresults_quintile.csv" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/regionresults_quintile.csv

#readLines("~/Box Documents/work/CoBenefits/MaggieShare/AmerTimeUse/atusact_2003.dat", n=10)
# activity <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusact_2003.dat", header=T, skip=0, sep=",")

summary <- read.table(summaryFile, header=T, skip=0, sep=",")

# respondants <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusresp_2003.dat", header=T, skip=0, sep=",")

# roster <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusrost_2003.dat", header=T, skip=0, sep=",")

CPS <- read.table(cpsFile, header=T, skip=0, sep=",")

CPS.2 <- CPS[,c("TUCASEID", "GESTFIPS", "GEREG","GEMETSTA")]
CPS.2$IDCHAR <- as.character(CPS.2[,"TUCASEID"])

summary.2 <- summary[, c("TUCASEID", "TEAGE",    "TESEX",    "t050203",  "t130101",  "t130102",  "t130103",  "t130104",  "t130106",  "t130108", "t130109", "t130111", "t130113",  "t130115", "t130116",  "t130117", "t130119",  "t130120",  "t130122",  "t130123", "t130124",  "t130126",  "t130127", "t130128", "t130130",  "t130131",  "t130132", "t130133",  "t130134", "t130135", "t130136" , "t130199" , "t139999")]
summary.2$IDCHAR <- as.character(summary.2[,"TUCASEID"])

summary.2a <- summary.2[,c(1:3, length(summary.2))]
summary.2b <- summary.2[,c(5:length(summary.2)-1)]
minutes <- rowSums(summary.2b)
summary.2a$minutes <- minutes

metkey <- as.matrix(c(3.0,7.3,5.0,6.5,7.5,3.5,5.8,7.8,6.0,8.0,3.8,7.8,8.0,5.3,7.3,9.8,6.3,7.0,7.0,5.0,6.8,3.0,3.5,5.5,6.0,4.0,6.0,3.0,3.0,3.0), nrow = 1, ncol=30)

summary.2c <- mapply(`*`,summary.2b,metkey)
mets <- rowSums(summary.2c)
summary.2a$mets <- mets

summary.2a$agecat <- ifelse(summary.2a$TEAGE <5, 1,
                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, 2,
                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, 3,
                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, 4,
                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, 5,
                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, 6,
                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, 7, 8
)))))))

summary.2a$agecat2 <- ifelse(summary.2a$TEAGE <5, "age0-4",
                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, "age5-14",
                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, "age15-29",
                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, "age30-44",
                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, "age45-59",
                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, "age60-69",
                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, "age70-79", "age80+"
)))))))

use <- merge(summary.2a, CPS.2, by.x="TUCASEID", by.y="TUCASEID", all.x=T, sorted=F)

use$sexcat <- ifelse(use$TESEX == 1, "male", "female")

use$regcat <- ifelse(use$GEREG == 1, "northeast",
              ifelse(use$GEREG == 2, "midwest",
              ifelse(use$GEREG == 3, "south",  "west")))

use$metcat <- ifelse(use$GEMETSTA == 1, "metro",
              ifelse(use$GEMETSTA == 2, "non-metro", "unclassified"))

use <- use[,c("minutes","mets","agecat2","GESTFIPS","sexcat","regcat","metcat")]

#head(use)

use.region <- ddply(use, .(sexcat, agecat2, regcat, metcat), summarise,
             mean_min = mean(minutes, na.rm=T),
             mean_mets = mean(mets, na.rm=T),
             n = length(minutes)
)

use.state <- ddply(use, .(sexcat, agecat2, GESTFIPS, metcat), summarise,
             mean_min = mean(minutes, na.rm=T),
             mean_mets = mean(mets, na.rm=T),
             n = length(minutes)
)

use2 <- subset(use, metcat == "metro")

use.state2 <- ddply(use2, .(sexcat, agecat2, GESTFIPS), summarise,
              percentile10 <- quantile(mets, 0.10, na.rm=T),
              percentile30 <- quantile(mets, 0.3, na.rm=T),
              percentile50 <- quantile(mets, 0.5, na.rm=T),
              percentile70 <- quantile(mets, 0.7, na.rm=T),
              percentile90 <- quantile(mets, 0.9, na.rm=T),
              mean <- mean(mets, na.rm=T),
              n = length(mets)
)

use.region2 <- ddply(use2, .(sexcat, agecat2, regcat), summarise,
              percentile10 <- quantile(mets, 0.10, na.rm=T),
              percentile30 <- quantile(mets, 0.3, na.rm=T),
              percentile50 <- quantile(mets, 0.5, na.rm=T),
              percentile70 <- quantile(mets, 0.7, na.rm=T),
              percentile90 <- quantile(mets, 0.9, na.rm=T),
              mean <- mean(mets, na.rm=T),
              n = length(mets)
)

state.table <- subset(use.state, metcat == "metro")

#write.csv(state.table, stateResultFile, row.names=F)

#write.csv(use.state2, stateResultFile2, row.names=F)

#write.csv(use.region2, regionResultFile, row.names=F)

#ggplot(use.region, aes(x=factor(sexcat), y=mean_mets, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1) +facet_grid(regcat ~ metcat)

#ggplot(subset(use.region, metcat != "unclassified"), aes(x=factor(sexcat), y=mean_mets, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1, position="dodge") +facet_grid(regcat ~ metcat)

#ggplot(subset(use.state, metcat != "unclassified" & GESTFIPS == 17), aes(x=factor(sexcat), y=mean_min, fill=factor(agecat2))) + geom_bar(stat="identity", alpha=1, position="dodge") +facet_grid(GESTFIPS ~ metcat)
```

```{r figures, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}

zeroMETs <- use$minutes == 0

ggplot(use[!zeroMETs,], aes(x = mets, ..density.., colour = agecat2)) + geom_freqpoly(binwidth=200) + facet_grid(. ~ sexcat) + coord_cartesian(xlim = c(0, 1500))

```

```{r nonZeroMETsLogNormal, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
parameters <- ddply(use[!zeroMETs,], .(sexcat, agecat2), summarise,
             mean_min = mean(minutes, na.rm=T),
             mean_mets = mean(mets, na.rm=T),
             sd_min = sd(minutes, na.rm=T),
             sd_mets = sd(mets, na.rm=T)
)
cv_mets <- with(parameters, sd_mets/mean_mets)
cv_min <- with(parameters, sd_min/mean_min)
cv0_mets<- mean(cv_mets)

mean_mets <- parameters$mean_mets
sd_mets <- cv0_mets*mean_mets

getLogNormal <- function(mu,sd){
    dlnorm(seq(0,2000,length.out=1e3), log(mu/sqrt(1+sd^2/mu^2)), sqrt(log(1+sd^2/mu^2)))
}

foo <- as.data.frame(mapply(getLogNormal, mean_mets, sd_mets))
names(foo) <- paste0(parameters$sexcat,parameters$agecat2)
foo <- melt(foo)
foo <- data.frame(foo, x = seq(0,2000,length.out = 1e3), sex = c(matrix(c("F","M"),nrow= nrow(foo)/2, ncol = 2, byrow = TRUE)), agecat = c(matrix(c("15-29","30-44","45-59","60-69","70-79", "80+"),nrow= nrow(foo)/12, ncol = 12, byrow = TRUE)))

ggplot(foo, aes(x,value, colour = agecat)) + geom_line() + facet_grid( . ~ sex) + coord_cartesian(xlim = c(0, 1500))

```

```{r matrices, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
nonTravelMETs <- rbind(0,rbind(0,cbind(parameters$mean_mets[7:12], parameters$mean_mets[1:6])))
nonTravelMETsNormalized <- nonTravelMETs/nonTravelMETs[3,2]
meanNonTravelMETs <- nonTravelMETs[3,2]

```

```{r plotMatrix, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
D <- melt(nonTravelMETsNormalized)
ggplot(D, aes(Var2, value)) + geom_bar(aes(fill=Var1), stat = "identity", position = "dodge")
```

```{r fooFunction, echo = FALSE, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE}
probActive <- 0.15
METQuintiles <- mapply(getMETQuintiles, nonTravelMETsNormalized*nonTravelMETs, MoreArgs = list(cv = cv0_mets, p = probActive, size = 1e4), SIMPLIFY = FALSE)
```

```{r NHTSTables, eval = FALSE}
HHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/HHV2PUB.CSV", na.strings = "-9")
names(HHV2PUB) <- c("HOUSEID","VARSTRAT","WTHHFIN","DRVRCNT","CDIVMSAR","CENSUS_D","CENSUS_R","HH_HISP","HH_RACE","HHFAMINC","HHRELATD","HHRESP","HHSIZE","HHSTATE","HHSTFIPS","HHVEHCNT","HOMEOWN","HOMETYPE","MSACAT","MSASIZE","NUMADLT","RAIL","RESP_CNT","SCRESP","TRAVDAY","URBAN","URBANSIZE","URBRUR","WRKCOUNT","TDAYDATE","FLAG100","LIF_CYC","CNTTDHH","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")

VEHV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/VEHV2PUB.CSV", na.strings = "-9")
names(VEHV2PUB) <- c("HOUSEID","WTHHFIN","VEHID","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","CDIVMSAR","CENSUS_D","CENSUS_R","HHSTATE","HHSTFIPS","HYBRID","MAKECODE","MODLCODE","MSACAT","MSASIZE","OD_READ","RAIL","TRAVDAY","URBAN","URBANSIZE","URBRUR","VEHCOMM","VEHOWNMO","VEHYEAR","WHOMAIN","WRKCOUNT","TDAYDATE","VEHAGE","PERSONID","HH_HISP","HH_RACE","HOMEOWN","HOMETYPE","LIF_CYC","ANNMILES","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","BEST_FLG","BESTMILE","BEST_EDT","BEST_OUT","FUELTYPE","GSYRGAL","GSCOST","GSTOTCST","EPATMPG","EPATMPGF","EIADMPG","VEHTYPE","HH_CBSA","HHC_MSA")
```

```{r NHTSArchive, echo = FALSE, eval = FALSE}
nMatrix <- ddply( subset(NHTS.person, !is.na(AGE) & !is.na(SEX)), .(SEX, AGE), summarise, count = length(PERSONID))
sum(nMatrix$count) # 305649 number of people in the PERSON table

foo <- group_by(NHTS.trips, HOUSEID,PERSONID)
foobar <- dplyr::summarise(foo, n = n())
nrow(foobar) # 262,934 number of people with trip data

NHTS %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarise( n = n()) # 260,576 people in merged data
ggplot(NHTS, aes(x = TRVL_MIN, ..density.., colour = AGE)) + geom_freqpoly(binwidth = 15) + facet_grid( SEX ~ AT) + labs(x = "Foo", colour = "Age") + coord_cartesian(xlim = c(0, 75))
NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))
```

### Relative Risk Model
```{r explore, eval = FALSE}
names(ITHIM.baseline$parameters)
```
```{r RRFigure, echo = FALSE, eval = FALSE}
phi <- function(x,k=0.5,phi0=comparativeRisk$RR.baseline$BreastCancer$F[1,1]) phi0^{x^k}
plot(x = xvec <- seq(0,xmax <- 10,length.out=1e3), y = phi(xvec, k = 0.25), main = "Relative Risk Models for MET Exposure", xlab = "METs per week", ylab = "Breast Cancer Relative Risk", type = "l", lwd = 2, col = 1, ylim = c(0.9,1),xlim=c(0,xmax))
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.375), lwd = 2, col = 3)
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 0.5), lwd = 2, col = 2)
lines(x = xvec <- seq(0,xmax,length.out=1e3), y = phi(x = xvec, k = 1), lwd = 2, col = 4)
legend(x="topright",legend=paste0("k=",c("0.25","0.375","0.5","1")), col=1:4,lty=1,lwd=2)
```

### Tinkering

Note that some of the AF values will disagree with the EXCEL
spreadsheet because Diabetes, HHD_Air, CVD_Air and Stroke_Air do not
combine travel METs with non-travel METs.

```{r parlist, eval = FALSE}
AF.sum <- numeric(1e3); i <- 0;
muwt.vec <- seq(107.1,150,length.out=length(AF.sum))
#muct.vec <- seq(2.7,3.0,length.out=length(AF.sum))
for(muwt in muwt.vec){
i <- i+1
parameters <- createParameterList(baseline = FALSE)

parameters <- setParameter(parName="muwt", parValue = muwt, parList = parameters)
ITHIM.scenario <- list(
  parameters = parameters <- parameters,
  means = means <- computeMeanMatrices(parameters),
  quintiles = quintiles <- getQuintiles(means)
)
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
AF.sum[i] <- with(comparativeRisk, sum(100*AF$BreastCancer))
}
D <- data.frame(mu.wt = muwt.vec,AF.sum=AF.sum)
p <- ggplot(D, aes(x = mu.wt, y = AF.sum))
p + geom_point()
```

```{r AF, echo = FALSE, eval = FALSE}
with(comparativeRisk, as.data.frame(lapply(AF[c("BreastCancer","Depression")], function(x) 100*round(x[-(1:2),],4))))
```

To see the help page for a function use help().
```{r help, eval = FALSE}
help(createParameterList)
```

### Example
To run the example contained in the _ITHIM_ package run the following.

```{r package, eval = FALSE}
example("ITHIM-package")
```
### Figures

```{r plot2, echo = FALSE, fig.width = 12, eval = FALSE}
plotRR(comparativeRisk$RR.baseline$Depression,comparativeRisk$RR.scenario$Depression) + coord_cartesian(ylim = c(0.75, 1.05))+ggtitle("Depression")
```

```{r plot3, echo = FALSE, fig.width = 12, eval = FALSE}
plotRR(comparativeRisk$RR.baseline$BreastCancer,comparativeRisk$RR.scenario$BreastCancer) + coord_cartesian(ylim = c(0.75, 1.05)) + ggtitle("Breast Cancer")
```

```{r plot4, echo = FALSE, fig.height = 12, eval = FALSE}
D<-melt(comparativeRisk$AF)
names(D)<- c("age","sex","AF","disease")
p <- ggplot(D, aes(age,AF)) + geom_bar(aes(fill=sex), stat = "identity") + facet_grid( disease ~ .) + coord_cartesian(ylim = c(0, 0.1))
p
```
### Road Injury Array
```{r road, eval = FALSE}
baselineInjuries <- array(dim=c(6,7,3,2))
injury <- c("fatal","serious")
road <- c("minor road","major road","motorway")
vehicle <- c("walk","cycle","bus","car","HGV","mbike")
dimnames(baselineInjuries) <- list(vehicle,c(vehicle,"NOV"),road,injury)

fatalities <- array(c(0,0,0.6666666667,34,6.3333333333,0,0.3333333333,0,0,0,0,0,4.3333333333,0,0,0.3333333333,0,0,0,0,0.3333333333,0.3333333333,0,0,0.6666666667,0,0.3333333333,0,0,0,48.6666666667,2,0,5.6666666667,0,35.6666666667,0,0,0,2.3333333333,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,2,0,0,0,0,6.3333333333,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,11.3333333333,1,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,18.6666666667,1.3333333333,0,2,0,7.3333333333,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0.3333333333,0,0.3333333333,0.3333333333,20,3,0,0,0,0,0,0.3333333333,0.3333333333,20,3,0,0,0,0,0,0,0,1,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,42.3333333333,4.6666666667,0,2,0,26.3333333333,0,0,0,3.3333333333,0.3333333333,0,0.3333333333,0,1,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,3,0,0,0,6.3333333333,0,0,0.6666666667,0,0),dim=c(9,8,3))

fatalaties <- apply(fatalities,c(1,3),t)

serious <- array(c(0,0,0,6.3333333333,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,40,0.6666666667,0,0.6666666667,0,11.6666666667,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0.3333333333,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.6666666667,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,8,0.3333333333,0,0.6666666667,0,3.6666666667,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.6666666667,0.3333333333,0,0,0,0,0,0,0,0.6666666667,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,29.6666666667,1,0,0,0,11,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0.3333333333,0,0,0,0,0,0,0,0,0),dim=c(9,8,3))

serious <- apply(serious,c(1,3),t)

baselineInjuries[,,,1] <- fatalaties
baselineInjuries[,,,2] <- serious

Data.RoadType <- data.frame(local=c(0.75,0.53,0.0,0.09,0.29,NA,0.09,NA), arterial=c(0.25,0.47,0.96,0.31,0.42,NA,0.31,NA), highway=c(0.0,0,0.04,0.60,0.29,NA,0.60,NA),row.names=c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike") )

Data.Occupancy <- data.frame(occupancy = c(1.0,1.0,18.9,1.3,1.0,NA,1.0,NA), row.names = c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike"))

```

### Active Transport Estimates (Speed and Distance by Transport Mode)
```{r plot1, echo = FALSE, eval = FALSE}
D <- melt(ITHIM.baseline$means$meanActiveTransportTime, c("age","sex"), value.name = "activeTransportTime")
p <- ggplot(D, aes(age,  activeTransportTime)) + geom_bar(aes(fill=sex), stat = "identity", position = "dodge")
p + coord_cartesian(ylim = c(0.75, 1.05))
```

These data were ripped out of the EXCEL workbook and vary per scenario.
```{r transportData, eval = FALSE, echo = FALSE}

transportMode <- c("walk","cycle","bus","minibus","train","car.driver","car.passenger","mbike","elec.bike.scooter","LGV","HGV.rigid","HGV.articulated")

speed <- c(1.35299774571063,5.39374775150058,5.22298847623309,NA,9.94147658251227,17.36906017290580,14.51012134295660,NA,NA,NA,NA,NA)

dist <- c(2.44072161538462,1.19969769230769,4.37851867307692,NA,5.58545353846154,111.91985351923100,39.12264265384620,0.00000000000000,0.00000000000000,NA,1.05479452054795,NA) # miles per week

Data.Speed <- data.frame(speed,dist, row.names = transportMode)
```

```{r headDataSpeed, eval = FALSE}
Data.Speed
```

```{r transportTime, echo = FALSE, results = "hide", eval = FALSE}
head(transportTime <- Data.Speed$dist/Data.Speed$speed*60)
```

### Emissions Data

```{r roadAndCO2Data, echo = FALSE, eval = FALSE, results = "hide"}

CO2 <- c(957.686509527760,221.578783667454,203.624526028894,247.869842164617,333.468424533125,90.000000000000,835.014773125786)

Data.CO2 <- data.frame(CO2, row.names = c("bus","Cars.urban","Cars.mway","LGVs.urban","LGVs.mway","motorbike","HGV"))

Data.RoadType <- data.frame(local=c(0.75,0.53,0.0,0.09,0.29,NA,0.09,NA), arterial=c(0.25,0.47,0.96,0.31,0.42,NA,0.31,NA), highway=c(0.0,0,0.04,0.60,0.29,NA,0.60,NA),row.names=c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike") )

Data.Occupancy <- data.frame(occupancy = c(1.0,1.0,18.9,1.3,1.0,NA,1.0,NA), row.names = c("walk","cycle","bus","car","HGV","LGV","mbike","elec.bike"))

```

The amount of carbon emissions varies by transport mode and road type.
To accomodate this we need the distribution of road type by transport
mode.  At some point we need the expected number of passengers per
transport mode and we create it here, arbitrarily perhaps.

```{r showData, echo = TRUE, eval = FALSE, results = "show"}
Data.CO2
Data.RoadType
Data.Occupancy
```

### Non-(Active Transport) METs

The file NonTravelMETs.csv was created from the tripleWin EXCEL file.
These data are used in a table look-up for a given county.  The data
represent the non-travel-related energy expenditures as measured in
METs.  The data are stratified by age and sex.  Variable names were
not given for the 7 numerical variables listed.  Here we name them x1
through x7.

```{r nonTravelMETs, eval = FALSE, echo = FALSE}
nonTravelMETs <- data.frame(scan(file="./data/NonTravelMETs.csv", sep = ",", what = list("","","","",1,1,1,1,1,1,1), skip = 1))
names(nonTravelMETs) <- c("ID","sexcat","agecat2","GESTFIPS", paste0("x",1:7))
```

```{r ggplot, fig.width=12, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE, results = "hide", eval = FALSE}
D <- melt(nonTravelMETs)
p <- ggplot(D, aes(x = variable, y = value, colour = agecat2), outlier.colour = "red", outlier.shape = 1 )
p + geom_boxplot()
```

Who is the outlier?  Perhaps it is a cohort of laborers, e.g., a county of coal miners?

```{r foo, eval = FALSE, echo = TRUE, eval = FALSE}
nonTravelMETs[which.max(nonTravelMETs$x5),]
```

Strange that it is group of middle-aged men.  I see this as a red
flag.  I remove this datum and redraw the boxplots, but this time only
for x5, x6 and x7, the others seemed less interesting.

```{r fips, eval = FALSE, echo = FALSE, results = "hide"}
fips <- read.csv("./data/fips_codes_website.csv")
any(fips$County.FIPS.Code==38)
```

```{r Nontravelmets, echo = FALSE, message = FALSE, warning = FALSE, results = "hide", fig.width=12, fig.height=6, eval = FALSE}
D <- melt(nonTravelMETs[-which.max(nonTravelMETs$x5),-(5:8)])
p <- ggplot(D, aes(x = variable, y = value, colour = agecat2), outlier.colour = "red", outlier.shape = 1 ) + geom_boxplot()

p + facet_grid(. ~ sexcat)
```

Thoughts?

### Relative Risk of MET Exposure

In the worksheet _Phy activity RRs_ we see estimates for relative risk
for various disease outcomes.  My guess is that some transformation is
being made to convert the exposure from active transport time to METs.
I don't know why this is needed yet.  I can't make much progress here
until I understand that better.

### Coefficient of Variation

An estimate for the coefficient of variation is provided without much
explanation.  The literature cites an empirical result.  The feeling
is that as mean active transport increases in the population people
are fitter and there is less variance in active transport time.  I
guess this has been shown in the same refernce for the coefficient of
variation estimate.

### Vital Statistics Data

#### 2001 Data with Vargo Script

```{r vitalstats, eval = FALSE, echo = FALSE}
source("../vargo/process_2001_vital_stats.r")
saveRDS(mortality2001, file = "./data/mortality2001.rds")
```

#### 2011 (VS11MORT.DUSMCPUB)

Lets read in the CDC data using the file VS11MORT.DUSMCPUB and see whats up.

```{r VS11, echo = FALSE, eval = FALSE}
tapePosition <- list(state = c(29:30), skippedVars = c(31:34), county = c(35:37), skippedVars = c(38:68), sex = 69, ageUnit = 70, ageValue = c(71:73), skippedVars = c(74:145), ICD = c(146:149), skippedVars = c(150:444), race = c(445:446)) # Magic number!!!
colClasses <- c("factor","factor","factor", "factor","integer", "character", "factor")
fileWidth <- 488 # Magic number!!!
```

```{r VS11Classes, echo = FALSE, eval = FALSE}
names(colClasses) <- names(tapePosition)[-which(names(tapePosition)=="skippedVars")]
colClasses
```

To include more variables edit the above lines.  The variable listed
must be contiguous.

```{r convertAge, warning = FALSE, eval = FALSE, echo = FALSE}

varWidths <- unlist(lapply(tapePosition, "length"))
varWidths <- ifelse(names(tapePosition) == "skippedVars", -varWidths, varWidths)
widths <- c(-(min(tapePosition[[1]])-1),c(varWidths))
widths <- c(widths,-(fileWidth-sum(abs(widths))))

Data.Mortality <- read.fwf(file = "~/ITHIM/R/data/VS11MORT.DUSMCPUB.first100", widths = widths, col.names = names(colClasses), header = FALSE, colClasses = colClasses)
Data.Mortality <- data.frame(Data.Mortality, convertedAge = with(Data.Mortality, convertAge(value = ageValue, unit = ageUnit)))

p <- ggplot(Data.Mortality, aes(x = convertedAge)) + geom_histogram(binwidth = 1)
p + facet_grid(. ~ sex)

```

```{r ICD, fig.width = 12, echo = FALSE, eval = FALSE}
Data.IDC.Codes <- read.csv(file="./data/IDC_Codes_key.csv", header = TRUE, colClasses = c("character","character","character"), na.string = "")
Data.IDC.Codes <- subset(Data.IDC.Codes, Include == "1" & !is.na(ITHIMCause))
ITHIMCauseOfDeath <- Data.IDC.Codes$ITHIMCause
names(ITHIMCauseOfDeath) <- Data.IDC.Codes$cause.of.death
Data.Mortality <- data.frame(Data.Mortality, CauseofDeathITHIM = ITHIMCauseOfDeath[Data.Mortality$ICD])
Data.Mortality <- subset(Data.Mortality, !is.na(CauseofDeathITHIM) & !is.na(convertedAge))
D <- data.frame(disease = Data.Mortality$CauseofDeathITHIM)
p <- ggplot(D, aes(disease))
p + geom_bar()

```

```{r CoDTable, eval = FALSE}
with(Data.Mortality, table(CauseofDeathITHIM))
```

```{r ageClass, fig.width = 12, echo = FALSE, eval = FALSE}
ageClass <- c(rep("00-04",5),rep("05-14",10),rep("15-29",15),rep("30-44",15),rep("45-59",15),rep("60-69",10),rep("70-79",10),rep("80+",121))
names(ageClass) <- as.character(0:200)
Data.Mortality <- data.frame(Data.Mortality, ageClass = as.factor(ageClass[as.character(as.integer(Data.Mortality$convertedAge))]))
with(Data.Mortality, table(ageClass))

D <- data.frame(ageClass = Data.Mortality$ageClass, disease = Data.Mortality$CauseofDeathITHIM)
p <- ggplot(D, aes(disease))
p + geom_bar(aes(fill=ageClass))
```

```{r histbyDisease, fig.width = 12, echo = FALSE, eval = FALSE}
D <- data.frame(convertedAge = Data.Mortality$convertedAge, disease = Data.Mortality$CauseofDeathITHIM)
p <- ggplot(D, aes(convertedAge, ..density.., colour = disease)) + geom_histogram(binwidth=5, show.legend = TRUE)
p + facet_grid( . ~ disease)
```

```{r counties, eval = FALSE}
Data.AgeSex <- read.csv("./data/county_age_sex2002.csv", header=T, col.names = c("county.name","state.name", "MSA","sex","all.ages","age0-4","age5-14","age15-29","age30-44","age45-59","age60-69","age70-79","age80+", "STATE","COUNTY"), colClasses = c("factor","factor","factor","factor","integer","integer","integer","integer","integer","integer","integer","integer","integer", "factor", "factor"))

age.sex.2 <- melt(age.sex[,c(14,15,4,6:13)], id=c("STATE","COUNTY","sex"))

#age.sex.3 <- dcast(age.sex.2, STATE + COUNTY + variable ~ sex, mean)

mort.age <- merge(mort.ITHIM, age.sex.3, by.x=c("state","county", "ITHIM.age"), by.y=c("STATE","COUNTY","variable"))

mort.age$year <- 2001

mort.age$app.pop <- ifelse(mort.age$sex ==1, mort.age$male, mort.age$female)

```

```{r tabulate, eval = FALSE}
Data.Mortality.Table <- as.data.frame(ftable(with(Data.Mortality, table(sex,ageClass,CauseofDeathITHIM))))
```

```{r saveMortality, echo = FALSE, eval = FALSE}
saveRDS(object = Data.Mortality, file = "./data/Data.Mortality.rds")
saveRDS(object = Data.Mortality.Table, file = "./data/Data.Mortality.Table.rds")
```
-->

### 2016-10-19 methodsPaper/tex/results.tex
Created two new figures
Also created appendix.Rnw to store old R code.

@@
 (walking and cycling) for baseline, low, and high scenarios are shown
-in \ref{fig:ATtimes}. Overall mean walking time for high and low
+in \ref{fig2}. Overall mean walking time for high and low
 scenarios were ???X??? (min/week) higher and ???Y??? (min/week) lower
@@
 subset. We then varied the mean non-travel activity time and active
-travel time to observe how this chaged results
-(\ref{fig:currentFig1}). We observe the most dramatic responses in
-disease burden estimates to changes in active travel time when
-non-travel activity times are lowest. This result is consistent with
-the non-linear dose-response curve for physical activity and several
-health outcomes. The total health benefit of increased active travel
-time (high scenario) varies from ???X??? to ???Y??? with changing mean
-non-travel activity time (0 to 25 minutes).
+travel time to observe how this chaged results (\ref{fig1}). We
+observe the most dramatic responses in disease burden estimates to
+changes in active travel time when non-travel activity times are
+lowest. This result is consistent with the non-linear dose-response
+curve for physical activity and several health outcomes. The total
+health benefit of increased active travel time (high scenario) varies
+from ???X??? to ???Y??? with changing mean non-travel activity time (0
+to 25 minutes).
 
@@
 
-As exhibited in Figure \ref{fig:percentQuint}, the variability in
-disease burden benfits varies to a much greater degree when using
-fewer points in the attirbutable fraction approximation.
+As exhibited in Figure \ref{fig1}, the variability in disease burden
+benfits varies to a much greater degree when using fewer points in the
+attirbutable fraction approximation.

### 2016-10-19 R/ITHIM.Rmd
Updated figures 2 and 3 for methods paper

@@
 atVec <- seq(45, 120, length.out = 5)
-muNonTravelVec <- c(1, 10, 50, 100, 150, 200)
+muNonTravelVec <- c(1, 5, 10)
 DTotal <- data.frame()
@@
 
-```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- rbind(getWalkTime(ITHIM.default),getWalkSpeed(ITHIM.default),getCycleTime(ITHIM.default),getNonTravelMETs(ITHIM.default))
+```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", 5)
+
+D <- rbind(getWalkTime(ITHIM.national),getWalkSpeed(ITHIM.national),getCycleTime(ITHIM.national),getNonTravelMETs(ITHIM.national))
 D <- within(D, sex <- factor(sex, levels = c("F","M")))
@@
 p <- ggplot(D, aes(x = ageClass, y = mu, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ ., scales = "free_y")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( type ~ ., scales = "free")
 ```
 
+## Figure 3
+
 ```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- melt(ITHIM.default$parameters$GBD)
+D <- melt(ITHIM.national$parameters$GBD)
 names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")
@@
 
-p <- ggplot(subset(D,disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes")), aes(x = ageClass, y = burden, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ disease, scales = "free_y")
+p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes")), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ disease, scales = "free_y")
 ```

### 2016-10-20 R/ITHIM.Rmd
Changed leisure activity to 10, instead of 5

@@
 
-ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", 5)
+ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", 10)
 
@@
 
-```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = (fig.width+1), fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
 D <- melt(ITHIM.national$parameters$GBD)
@@
 p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes")), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ disease, scales = "free_y")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( type ~ disease, scales = "free_y")
 ```

### 2016-10-21 R/tutorial.Rmd
Created tutorial.Rmd and tutorial.R

```{r figwidth, eval = TRUE,  warning = FALSE, message = FALSE, echo = FALSE}
fig.width <- 6
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

First load the package from Github.  To do this you need the package
devtools.

```{r, libs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
options(repos='http://cran.rstudio.com/')
install.packages("devtools")
```

Now load the package and install ITHIM.

```{r libs2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
library("devtools")
install_github(repo = "syounkin/ITHIM")
library("ITHIM")
```

Now that ITHIM is installed and loaded the first step is to create the
baseline model.  To do so, using defaults built into the package, use
createITHIM as follows.

```{r create, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
ITHIM.baseline <- createITHIM()
```

There are currently fourteen parameters, many of which are matrices,
in the model.  (Recall that we are only using the physical activity
component.)

```{r parameters, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
names(ITHIM.baseline$parameters)
```

We have means and mean matrices for walk time, cycle time and
non-travel-related activity and coefficients of variation for active
transport time and non-travel-related activity.  nAgeClass has always
been 8 and meanType is always "referent".  GBD is an array of disease
burden estimates and quantiles is the vector of quantiles used for the
approximation of the PAF integral.

Now to create the scenario we need only change one parameter.
Parameters may be changed with updateITHIM.  First we change the mean
walking time muwt to 120 minutes.

```{r scenario, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
ITHIM.scenario <- updateITHIM(ITHIM.baseline, "muwt", 120)
```
Now to run the comparative risk assesment use compareModels.
```{r CRA, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
comparativeRisk <- compareModels(ITHIM.baseline, ITHIM.scenario)
```

```{r AF, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
comparativeRisk$AF$ColonCancer
```

### 2016-10-21 R/tutorial.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("tutorial.Rmd","tutorial.html")

### 2016-10-24 R/ITHIM.Rmd
Now correct for zero trip times when computing means from NHTS data

@@
 
-The PERSON table contains the family members of the person surveyed,
-right?  We want only the peole who were surveyed for their travel
-times.  So, we can\'t use everyone in the PERSON table.  If someone
-had zero trip time they would not exist in the TRIP table.  So, if we
-use just the TRIP table we miss people with zero trip time. So how do
-we merge?  We use an inner join and assume that their is a negligible
-number of people with zero trip time.  Not very many people never stay
-in the house all day.  Please correct me if I am mistaken abiout the
-nature of the PERSON or TRIP tables.
+We now correct for zero trip times in NHTS.
 
@@
 NHTS.person <- subset( NHTS.person, MSA == TRUE )
-NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
+NHTS.trips <- subset(NHTS.trips, !is.na(TRPMILES))
+
+NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
+NHTS <- within(NHTS, TRPMILES <- ifelse(is.na(TRPMILES), 0, TRPMILES))
 saveRDS(NHTS, file = "./data/NHTS.rds")

### 2016-10-24 R/ITHIM.Rmd
TRVLCMIN -> TRVL_MIN

@@
                    TRVL_MIN = as.numeric(TRVL_MIN),
-                   TRVLCMIN = ifelse(as.numeric(TRVLCMIN) != 0, as.numeric(TRVLCMIN), NA),
+                   #TRVLCMIN = as.numeric(TRVLCMIN),
                    URBRUR  = as.factor(URBRUR),
@@
 
-```{r summary, fig.width = 20, fig.height = 10, eval = compute, echo = FALSE, message = FALSE, warning = FALSE}
-walkHours <- with(subset(NHTS.trips, {
-    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
-}), TRVL_MIN/60)
-
-walkMiles <- with(subset(NHTS.trips, {
-    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
-}), TRPMILES)
-
-cycleHours <- with(subset(NHTS.trips, {
-      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
-}), TRVL_MIN/60)
-
-cycleMiles <- with(subset(NHTS.trips, {
-      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
-}), TRPMILES)
-
-walkSpeed <- median(walkMiles/walkHours, na.rm = TRUE)
-cycleSpeed <- median(cycleMiles/cycleHours, na.rm = TRUE)
-
-Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, Speed = walkMiles/walkHours, mode = "walk")
-Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, Speed = cycleMiles/cycleHours, mode = "cycle")
-
-D <- rbind(Dwalk,Dcycle)
-
-Dspeed <- data.frame(speed = c(1,1), intercept = log10(c(walkSpeed,cycleSpeed)), mode = c("walk","cycle"))
-
-g <- ggplot(D, aes(Hours, Miles, colour = mode)) + geom_point()
-g <- g + geom_abline(aes(slope = speed, intercept = intercept, colour = mode), Dspeed) + facet_grid(. ~ mode)
-g + scale_y_log10() + scale_x_log10() + theme_bw()
-
-g <- ggplot(D, aes(Speed, ..density.., fill = mode)) + geom_histogram() + theme_bw() + scale_x_continuous(name="Speed") + scale_y_continuous(name="Density") + facet_grid(. ~ mode, scales = "free")
-g
-
-```
-
 #### Subset and Merge
@@
 NHTS.person <- subset( NHTS.person, MSA == TRUE )
-NHTS.trips <- subset(NHTS.trips, !is.na(TRPMILES))
+NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))
 
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
-NHTS <- within(NHTS, TRPMILES <- ifelse(is.na(TRPMILES), 0, TRPMILES))
+NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
 saveRDS(NHTS, file = "./data/NHTS.rds")

### 2016-10-24 R/appendix.Rmd
TRVLCMIN -> TRVL_MIN

@@
 # Appendix
+
+```{r summary, fig.width = 20, fig.height = 10, eval = compute, echo = FALSE, message = FALSE, warning = FALSE}
+walkHours <- with(subset(NHTS.trips, {
+    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
+}), TRVL_MIN/60)
+
+walkMiles <- with(subset(NHTS.trips, {
+    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
+}), TRPMILES)
+
+cycleHours <- with(subset(NHTS.trips, {
+      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
+}), TRVL_MIN/60)
+
+cycleMiles <- with(subset(NHTS.trips, {
+      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
+}), TRPMILES)
+
+walkSpeed <- median(walkMiles/walkHours, na.rm = TRUE)
+cycleSpeed <- median(cycleMiles/cycleHours, na.rm = TRUE)
+
+Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, Speed = walkMiles/walkHours, mode = "walk")
+Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, Speed = cycleMiles/cycleHours, mode = "cycle")
+
+D <- rbind(Dwalk,Dcycle)
+
+Dspeed <- data.frame(speed = c(1,1), intercept = log10(c(walkSpeed,cycleSpeed)), mode = c("walk","cycle"))
+
+g <- ggplot(D, aes(Hours, Miles, colour = mode)) + geom_point()
+g <- g + geom_abline(aes(slope = speed, intercept = intercept, colour = mode), Dspeed) + facet_grid(. ~ mode)
+g + scale_y_log10() + scale_x_log10() + theme_bw()
+
+g <- ggplot(D, aes(Speed, ..density.., fill = mode)) + geom_histogram() + theme_bw() + scale_x_continuous(name="Speed") + scale_y_continuous(name="Density") + facet_grid(. ~ mode, scales = "free")
+g
+
+```
 

### 2016-10-27 R/ITHIM.Rmd
Included CVD in figure 3 of methods paper

@@
 
-p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes")), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
 p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( type ~ disease, scales = "free_y")

### 2016-10-27 R/ITHIM.Rmd
Transposed facets for figure 3

@@
 p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( type ~ disease, scales = "free_y")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( disease ~ type, scales = "free_y")
 ```

### 2016-10-31 methodsPaper/tex/abstract.tex
Edits to Abstract

--- 2016-10-11 methodsPaper/tex/abstract.tex
@@
+Comparative risk assessment analyses to quantify the health potential
+of active transportation, e.g., walking and cycling, have been
+performed in the United States on a limited basis. Current
+methodologies, which rely on extensive data collection and user
+expertise, have estimated health benefits for select metropolitan
+areas under a few alternative scenarios. These analyses have largely
+been conducted using the Integrated Transport Health Impacts Model
+(ITHIM) which was developed in the UK and has been adapted for use in
+the US, primarily California. Here we present an implementation of the
+active transportation component of ITHIM to model the health impacts
+of modified active transportation levels. Using the newly developed
+software package \package{} written in R, along with publicly
+available national datasets, we quantify the change in disability
+adjusted life years over a wide range of scenarios.  The R package
+implemention of ITHIM's physical activity component provides a robust
+environment for exploring the relationship between active
+transportation and positive health outcomes.  A web-based
+user-interface is available for users who want a first estimate for
+the impact active transportation has on their
+state/county/tract/region.
+
 %% While increased active transportation, primarily cycling and walking,

### 2016-10-31 R/ITHIM.Rmd
Tweaks to GBD figure.

@@
 
-```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = (fig.width+1), fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 1.2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
 D <- melt(ITHIM.national$parameters$GBD)
@@
 
-p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( disease ~ type, scales = "free_y")
+p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( .
+~ disease, scales = "free_y")
 ```

### 2016-10-31 methodsPaper/tex/introduction.tex
publically -> publicly

@@
 important determinants of population-level activity
-patterns. Particularly, as they facilitate and foster routine and moderate
-physical activity, such as walking, local decisions on design and
-policy can play a huge role in improving the nation's health.
+patterns. Particularly, as they facilitate and foster routine and
+moderate physical activity, such as walking, local decisions on design
+and policy can play a huge role in improving the nation's health.
 
-Given the constraints and complexity of local governance and infrastructure investments, land use, and human behavior there is a need for high quality and reliable data and tools for decision-making related to urban transportaiton systems.
-The lack of accurate and availbale estimates of the health impacts of such decisions has limited the effectiveness of efforts to encourage non-motorized transport in the US and in other countries.
-In the last decade new tools, applying comparative risk assessment methodologies, have made progress toward quantifying the multple health impacts of changes to the transportation system and behaviors.
+Given the constraints and complexity of local governance and
+infrastructure investments, land use, and human behavior there is a
+need for high quality and reliable data and tools for decision-making
+related to urban transportaiton systems. The lack of accurate and
+availbale estimates of the health impacts of such decisions has
+limited the effectiveness of efforts to encourage non-motorized
+transport in the US and in other countries. In the last decade new
+tools, applying comparative risk assessment methodologies, have made
+progress toward quantifying the multple health impacts of changes to
+the transportation system and behaviors.
 
-ITHIM is one such statistical model that integrates data on active transport,
-physical activity, fine particulate matter and greenhouse gas
-emissions to provide estimates for the proportional change in
+ITHIM is one such statistical model that integrates data on active
+transport, physical activity, fine particulate matter and greenhouse
+gas emissions to provide estimates for the proportional change in
 mortality and morbidity for given baseline and alternate travel
@@
 precision when computing the population attributable fraction and
-creating a simple user-interface for the custom-built \R{}
-package \package{}.
+creating a simple user-interface for the custom-built \R{} package
+\package{}.
 
-Here we present this implementation using publically-available datasets with coverage of the entire United States, and investigate the sensitivity of the model outputs to variability in the input parameters.
-The results of this work suggest that statistical modeling and tools like the custom-built \R{} package \package{} can improve the application of this useful methodology and increase its utility to both practicioners and researchers.
+Here we present this implementation using publicly-available
+datasets with coverage of the entire United States, and investigate
+the sensitivity of the model outputs to variability in the input
+parameters. The results of this work suggest that statistical modeling
+and tools like the custom-built \R{} package \package{} can improve
+the application of this useful methodology and increase its utility to
+both practicioners and researchers.

### 2016-11-01 methodsPaper/tex/materialsAndMethods.tex
Created two new figures
Also created appendix.Rnw to store old R code.
Added caption to DALYs figure (figure 3)
Added Disease Burden estimate equations.

@@
-The ITHIM model uses the comparative risk assesment framework to
-estimate the population attributable fraction, $\af$.  In our case the
-$\af$ represents the proportional change in national disease burden if
-age and sex-specific active transport time means were changed.
+As with the original ITHIM implementation we employed a comparative
+risk analysis (CRA) across scenarios defined by mean active
+transportation time, i.e., walking and cycling.
+
+We compute the proportional change in national disease burden between
+the baseline model and alternative scenarios, or population
+attributable fraction, $\af$.  The population attributable fraction is
+defined below, and computed for each disease included in the model,
+namely Breast Cancer, Colon Cancer, Depression, Dementia, Diabetes and
+Cardiovascular disease.
 
@@
 
-% \begin{figure}[t]
-%   \centerline{\includegraphics[width=\textwidth]{./figures/fig2.pdf}}
-%     \caption{}\label{fig1}
-% \end{figure}
+\paragraph{Disease Burden Estimation}
+\begin{equation}
+B_1 = B_0(1-\af)\frac{\RR_1}{\sum_j \RR_j}
+\end{equation}
+
+$B_j = \gamma_jB_1$, where $\gamma_i = \frac{\RR_i}{\RR_1}$.  The
+overall burden is then computed as the sum across quintiles,
+$B=\sum B_j$.  Finally, $B_0$ is estimated from CDC data\ldots
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig1}}
+    \caption{}\label{fig1}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig2.pdf}}
+    \caption{}\label{fig2}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
+  \caption{Estimates for disease burden nationwide in terms of thousands of
+    DALYs for each of the six diseases included in the model.
+  }\label{fig3}
+\end{figure}
 

### 2016-11-01 methodsPaper/tex/materialsAndMethods.tex
Added Approximation of Integral formula

@@
 risk analysis (CRA) across scenarios defined by mean active
-transportation time, i.e., walking and cycling.
-
-We compute the proportional change in national disease burden between
-the baseline model and alternative scenarios, or population
-attributable fraction, $\af$.  The population attributable fraction is
-defined below, and computed for each disease included in the model,
-namely Breast Cancer, Colon Cancer, Depression, Dementia, Diabetes and
-Cardiovascular disease.
+transportation time, i.e., walking and cycling.  To do so we
+investigate the proportional change in national disease burden between
+the baseline model and a given alternative scenario, i.e., the
+population attributable fraction, $\af$.  The population attributable
+fraction is defined below and computed for each disease included in
+the model, namely Breast Cancer, Colon Cancer, Depression, Dementia,
+Diabetes and Cardiovascular disease.
 
@@
 activity is required for cycling and walking.
+
+\paragraph{Approximation of $\af$ Integral}
+
+\begin{equation}
+1 - \af \approx \frac{\sum R(x_i)}{\sum R(x_i^\prime)}
+\end{equation}
 

### 2016-11-07 methodsPaper/tex/materialsAndMethods.tex
Edits to Materials & Methods

@@
 the model, namely Breast Cancer, Colon Cancer, Depression, Dementia,
-Diabetes and Cardiovascular disease.
+Diabetes and Cardiovascular disease.  For readability we omit an index
+indicating disease.
 
 \begin{equation}
-\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \, \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x}
+\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \,
+  \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x} \label{paf}
 \end{equation}
@@
 modeled as the sum of two independent random variables, travel-related
-exposure and non-travel-related exposure.  Non-travel-related exposure
-is modeled with a \logNormal{} distribution with the age-sex specific
-mean ratios for non-travel-related exposure are estimated from the
-ATUS \cite{ATUS}.  The referent group mean is treated as a variable
-and the coefficient of variation is treated as constant across age-sex
-strata and estimated with ATUS.  The distribution for travel-related
-exposure is estimated from the time spent walking or cycling, i.e.,
-active transport time, which as in the original ITHIM model has
-a \logNormal{} distribution with constant coefficient of variation
-across age-sex strata.  Travel-related exposure is estimated from
-active transport time using assumptions about how much physical
-activity is required for cycling and walking.
+and non-travel-related physical activity.
 
-\paragraph{Approximation of $\af$ Integral}
+\paragraph{Non-Travel-Related Physical Activity}
 
 \begin{equation}
-1 - \af \approx \frac{\sum R(x_i)}{\sum R(x_i^\prime)}
+X_{ij}^{\mathrm{non-travel}} \sim \logNormalMath\left(\mu_{ij}, \gamma \mu_{ij}\right)
 \end{equation}
 
-\paragraph{Disease Burden Estimation}
+Within each age-sex class the non-travel-related exposure is modeled
+with a \logNormal{} distribution.  The age-sex specific mean ratios,
+$r_{ij} = \frac{\mu_{ij}}{\mu_0}$ and $\gamma$ are estimated from the
+American Time USe Survey data \cite{ATUS}.
+
+\paragraph{Travel-Related Physical Activity}
+
+The distribution for travel-related exposure is estimated from the
+time spent walking or cycling, i.e., active transport time, which as
+in the original ITHIM model is assigned a \logNormal{} distribution
+with constant coefficient of variation across age-sex classes.
+Travel-related exposure is estimated from active transport time using
+assumptions about how much physical activity is required for cycling
+and walking.
+
+\paragraph{Estimation of Disease Burden}
+
+To estimate the disease burden we first approximate the population
+attributable fraction through a discretization of the integral in
+equation \ref{paf}.
+
+With a simulated distribution for total physical activity in hand we
+compute the empirical percentiles in the baseline and alternate
+scenarios, $\mathbf{x}$ and $\mathbf{x}^\prime$. The proportional
+change in disease burden, $\delta$, may be approximated as follows,
+where $R$ is the risk function for total activity.
+
 \begin{equation}
-B_1 = B_0(1-\af)\frac{\RR_1}{\sum_j \RR_j}
+\delta = 1 - \af \approx \frac{\sum R(x_i)}{\sum R(x_i^\prime)}
 \end{equation}
 
-$B_j = \gamma_jB_1$, where $\gamma_i = \frac{\RR_i}{\RR_1}$.  The
-overall burden is then computed as the sum across quintiles,
-$B=\sum B_j$.  Finally, $B_0$ is estimated from CDC data\ldots
+The disease burden, in this case DALYs, is found for each of the
+percentiles using equation \ref{burden}, using $B_0$, the overall nation-wide burden for
+the baseline.  The overall burden is then computed as the sum across
+quintiles, $B=\sum_{j=1}^n B_j$.
 
-\begin{figure}[t]
-  \centerline{\includegraphics[width=\textwidth]{./figures/fig1}}
-    \caption{}\label{fig1}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=\textwidth]{./figures/fig2.pdf}}
-    \caption{}\label{fig2}
-\end{figure}
-
-\begin{figure}[t]
-  \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
-  \caption{Estimates for disease burden nationwide in terms of thousands of
-    DALYs for each of the six diseases included in the model.
-  }\label{fig3}
-\end{figure}
-
-% does this involve using the CDC Wonder data? we could add a bit here
-% about it, but if this is just national estimtes and shifts perhaps
-% we don't need it.
+\begin{equation}
+B_j = \delta B_0\frac{\RR_j}{\sum_{k=1}^n \RR_k}\label{burden}
+\end{equation}
 
@@
 imputed using R's random integer generator.
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig1}}
+    \caption{}\label{fig1}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig2.pdf}}
+    \caption{}\label{fig2}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
+  \caption{Estimates for disease burden nationwide in terms of thousands of
+    DALYs for each of the six diseases included in the model.
+  }\label{fig3}
+\end{figure}

### 2016-11-07 methodsPaper/tex/materialsAndMethods.tex
Edits to Materials & Methods

@@
 \begin{equation}
-\delta = 1 - \af \approx \frac{\sum R(x_i)}{\sum R(x_i^\prime)}
+\delta = 1 - \af \approx \frac{\sum_{i=1}^n R(x_i)}{\sum_{i=1}^n R(x_i^\prime)}
 \end{equation}
@@
 
+\paragraph{Risk Functions}
+
+We assign risk based on dose-response curves for each of the diseases
+included in the model.
+
+\begin{equation}
+\RR(x) = \mathrm{e}^{-r \tilde{x}}
+\end{equation}
+
+where $\tilde{x} = x^k$, $k=\frac{1}{2}$ and $r$ varies by age-sex
+class and disease.  For a table of $r$ values see supplementary
+material.
+
+\paragraph{Nationwide Health Benefits to Active Transport}
+
 For the purposes of this implementation we limited our analyses to
@@
 eliminated counties in the US with the classifcation of
-'nonmetropolitan' (noncore or micropolitan). We used this and similar
+`nonmetropolitan' (noncore or micropolitan). We used this and similar
 classification schemes as well as county and state identifiers in
@@
 these active travel times to the US metro population to perform the
-comparative risk assessment agains the baseline US numbers. We then
+comparative risk assessment against the baseline US numbers. We then
 examined the sensitivity of disease burden estimates to input
 parameters including the active travel time and non-travel activity.
-
-We report total change (averted or increased) in diability adjusted
-life years (DALYs) as an estimate of dissease burden change with
-alternative scenarions. To estimate DALYs we use a methodology adopted
-from previous US implementations of ITHIM which scale Global Budern of
-Disease Estimates for the US to smaller populations using mortality
-rate ratios \cite{maizlish2013}. Age, sex, and casuse specific
-mortalities were obtained form the CDC WONDER database for US metro
-counties for the years 2010-2014. Cells with less than 10 deaths were
-imputed using R's random integer generator.
 

### 2016-11-08 R/ITHIM.Rmd
Edits to Materials & Methods

@@
 ```
+
+## Non-Travel METs
+
+```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
+newMETsFile <- "ATUS_to_METS_key.csv"
+
+newMETsFile <- system.file(newMETsFile, package="ITHIM")
+
+newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
+dataColumns <- c(1,5)
+D <- newMETs[,dataColumns]
+metaData <- newMETs[,-dataColumns]
+newMETs <- list(Data = D, metaData = metaData)
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+
+rownames(newMETs$Data) <- activityCode
+rownames(newMETs$metaData) <- activityCode
+
+newMETs$metaData <- within(newMETs$metaData, {
+  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
+})
+
+metKey <- newMETs$Data$AinsworthMETS
+names(metKey) <- activityCode
+
+```
+
+```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
+cpsFile <- "~/GHI/data/ATUS/atuscps_2015.dat"
+cps <- read.table(cpsFile, header=T, skip=0, sep="," )
+cps <- within(cps, {
+    TUCASEID <- as.character(TUCASEID)
+    TULINENO <- as.character(TULINENO)
+    GEREG <- as.factor(GEREG)
+    })
+cps <- subset(cps, TULINENO == "1")
+
+summaryFile <- "~/GHI/data/ATUS/atussum_2015.dat"
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
+metaDataColumns <- 1:24
+summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
+
+summaryList$metaData <- within(summaryList$metaData,{
+    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
+    AGECAT <- ifelse(TEAGE <5, "00-04",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
+                            )))))))})
+
+```
+
+We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
+
+```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
+allActivities <- colnames(summaryList$Data)
+nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
+nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
+
+activities <- nonTravelActivities
+
+metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
+names(metKeyVector) <- activities
+metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
+dimnames(metKeyMatrix) <- list( NULL, activities)
+
+METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
+
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
+
+```
+
+```{r parameters, eval = FALSE, echo = FALSE}
+parameters <- ddply(D, .(SEX, AGECAT), summarise,
+             mean = mean(METs, na.rm=T),
+             sd = sd(METs, na.rm=T)
+             )
+
+(METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2)))
+
+colnames(METmatrix) <- c(as.character(parameters$SEX[1]),as.character(rev(parameters$SEX)[1]))
+
+METmatrix <- METmatrix[,c("male","female")]
+
+METmatrix <- METmatrix/METmatrix[3,2]
+
+```

### 2016-11-10 R/ITHIM.Rmd
Added roxygen2 to install.packages() and set compute <- TRUE

@@
 options(repos='http://cran.rstudio.com/')
-# install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r","devtools"))
+install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
 library("ggplot2")

### 2016-11-14 R/ITHIM.Rmd
Small change to methods paper figure

@@
 DALYs <- c(); DALYs.quints <- c();
-atVec <- seq(45, 120, length.out = 5)
+atVec <- seq(35, 120, length.out = 10)
 muNonTravelVec <- c(1, 5, 10)
@@
 
-p <- ggplot(DTotal, aes(x = at,y = -DALYs/1e3, colour = quants))
+DTotal <- within(DTotal, muNonTravel <- factor(muNonTravel, levels = c(1,5,10), labels = c("low","medium","high")))
+DTotal <- subset(DTotal, quants == "percentiles")
+
+p <- ggplot(DTotal, aes(x = at,y = -DALYs/1e3))#, colour = quants))
 p + stat_smooth( method = loess) + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)") + facet_grid( .

### 2016-11-15 R/ITHIM.Rmd
Tweak to methods paper figure.

@@
 
-p <- ggplot(DTotal, aes(x = at,y = -DALYs/1e3))#, colour = quants))
-p + stat_smooth( method = loess) + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)") + facet_grid( .
-~ muNonTravel) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))
+p <- ggplot(DTotal, aes(x = at, y = -DALYs/1e3, colour = muNonTravel), colour = muNonTravel)
+p + geom_line() + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) #+ facet_grid( .
+~ muNonTravel)
+
+#p + stat_smooth( method = loess) + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) #+ facet_grid( .
+~ muNonTravel)
 

### 2016-11-16 R/ITHIM.Rmd
Created Dose-Response Figure and tweaked Figure 1 from Methods Paper

--- 2016-11-10 R/ITHIM.Rmd
@@
 options(repos='http://cran.rstudio.com/')
-install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
+#install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
 library("ggplot2")
@@
 DALYs <- c(); DALYs.quints <- c();
-atVec <- seq(45, 120, length.out = 5)
-muNonTravelVec <- c(1, 5, 10)
+atVec <- seq(37.5, 120, length.out = 10)
+muNonTravelVec <- c(1, 10, 25)
 DTotal <- data.frame()
@@
 
-p <- ggplot(DTotal, aes(x = at,y = -DALYs/1e3, colour = quants))
-p + stat_smooth( method = loess) + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)") + facet_grid( .
-~ muNonTravel) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))
+DTotal <- within(DTotal, muNonTravel <- factor(muNonTravel, levels = as.character(muNonTravelVec), labels = c("low","medium","high")))
+DTotal <- subset(DTotal, quants == "percentiles")
+
+p <- ggplot(DTotal, aes(x = at, y = -DALYs/1e3, colour = muNonTravel))
+p + geom_line() + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)")  + scale_colour_discrete(name="Baseline Activity") #+ theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))
 
@@
 ```
+
+```{r doseresponse, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+RR1vec <- unique(unlist(createActiveTransportRRs()))
+lapply(createActiveTransportRRs(),function(x) lapply(x, function(x) unique(c(x))))
+#~~~~~~~~~~~~~~~~~~~~~~~
+# Manual bullshit!
+#~~~~~~~~~~~~~~~~~~~~~~~
+RR1vec <- RR1vec[RR1vec != 1]
+names(RR1vec) <- c("BreastCancer","ColonCancer.M","ColonCancer.F", "CVD", "Dementia", "Depression.young", "Depression.old", "Diabetes")
+
+RRx <- function(x, RR1, k) exp(log(RR1)*x^k)
+x <- seq(0,25,0.1)
+RRmat <- outer(x, RR1vec, RRx, k = 0.5)
+rownames(RRmat) <- x
+D <- melt(RRmat)
+names(D) <- c("x","disease","RR")
+
+p <- ggplot(D, aes(x = x, y = RR, colour = disease))
+p + geom_line()  + theme_bw() + scale_y_continuous(name="Relative Risk") + scale_x_continuous(name="Physical Activity (MET-hrs./week)")
+```

### 2016-11-17 R/appendix.Rmd

@@
 
-```
-
-## Non-Travel METs
-
-```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
-newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
-newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
-dataColumns <- c(1,5)
-D <- newMETs[,dataColumns]
-metaData <- newMETs[,-dataColumns]
-newMETs <- list(Data = D, metaData = metaData)
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-
-rownames(newMETs$Data) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-
-newMETs$metaData <- within(newMETs$metaData, {
-  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
-})
-
-metKey <- newMETs$Data$AinsworthMETS
-names(metKey) <- activityCode
-
-```
-
-```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
-cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat"
-cps <- read.table(cpsFile, header=T, skip=0, sep="," )
-cps <- within(cps, {
-    TUCASEID <- as.character(TUCASEID)
-    TULINENO <- as.character(TULINENO)
-    GEREG <- as.factor(GEREG)
-    })
-cps <- subset(cps, TULINENO == "1")
-
-summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat"
-summary <- read.table(summaryFile, header=T, skip=0, sep=",")
-metaDataColumns <- 1:24
-summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
-
-summaryList$metaData <- within(summaryList$metaData,{
-    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
-    AGECAT <- ifelse(TEAGE <5, "00-04",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
-                            )))))))})
-
-```
-
-We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
-
-```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
-allActivities <- colnames(summaryList$Data)
-nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
-nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
-
-activities <- nonTravelActivities
-
-metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
-names(metKeyVector) <- activities
-metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-dimnames(metKeyMatrix) <- list( NULL, activities)
-
-METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
-
-D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
-
-```
-
-```{r parameters, eval = FALSE, echo = FALSE}
-parameters <- ddply(D, .(SEX, AGECAT), summarise,
-             mean = mean(METs, na.rm=T),
-             sd = sd(METs, na.rm=T)
-             )
-
-METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
-METmatrix <- METmatrix/METmatrix[3,2]
 ```

### 2016-11-17 methodsPaper/tex/results.tex

@@
+We report total change (averted or increased) in diability adjusted
+life years (DALYs) as an estimate of dissease burden change with
+alternative scenarios.
+
+To estimate DALYs we use a methodology adopted from previous US
+implementations of ITHIM which scale Global Budern of Disease
+Estimates for the US to smaller populations using mortality rate
+ratios \cite{maizlish2013}. Age, sex, and casuse specific mortalities
+were obtained form the CDC WONDER database for US metro counties for
+the years 2010-2014. Cells with less than 10 deaths were imputed using
+R's random integer generator.
+
 Our subset of the US population to only metropolitan counties

### 2016-11-19 R/ITHIM.Rmd
Tweaked figures for Jonathan's talk

@@
 
-## ITHIM.NY <- ITHIMList[["NY"]]
-## ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
-## foo2 <- compareModels(ITHIM.national,ITHIM.NY)
-## foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
-## foobar <- rbind(foobar1,foobar2)
-## names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
+ITHIM.NY <- ITHIMList[["NY"]]
+ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
+foo2 <- compareModels(ITHIM.national,ITHIM.NY)
+foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
+foobar <- rbind(foobar1,foobar2)
+names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
 
@@
   Disease[Disease == "ColonCancer"] <- "Colon Cancer"
-  Disease <- factor(Disease, levels = c("Breast Cancer", "Colon Cancer", "Dementia", "Depression", "Diabetes"))
+  Disease <- factor(Disease, levels = c("Breast Cancer", "Colon Cancer", "Dementia", "Depression", "Diabetes", "CVD"))
 })
 
-p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = Sex) )
+p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = State))
 
@@
 theme(axis.title=element_blank()) +
-facet_grid( Disease ~ ., scales = "free") +
+facet_grid( Disease ~ Sex, scales = "free" ) +
 scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") +
@@
 DALYs <- c(); DALYs.quints <- c();
-atVec <- seq(37.5, 120, length.out = 10)
+atVec <- seq(45, 120, length.out = 10)
 muNonTravelVec <- c(1, 10, 25)
@@
 p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Nationwide Burden (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( .
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( .
 ~ disease, scales = "free_y")

### 2016-11-21 R/ITHIM.Rmd
Tweak to California figure

@@
 scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") +
-scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+"))
+scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
+theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6)) +
+scale_fill_manual(values=c("#CC6666", "#9999CC"))
 

### 2016-11-23 R/ITHIM.Rmd
Now methods paper is created with knitr
See files ./lib/*.Rnw for editing content.

@@
 
+## Non-Travel METs
+
+```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
+newMETsFile <- "ATUS_to_METS_key.csv"
+
+newMETsFile <- system.file(newMETsFile, package="ITHIM")
+
+newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
+dataColumns <- c(1,5)
+D <- newMETs[,dataColumns]
+metaData <- newMETs[,-dataColumns]
+newMETs <- list(Data = D, metaData = metaData)
+activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
+
+rownames(newMETs$Data) <- activityCode
+rownames(newMETs$metaData) <- activityCode
+
+newMETs$metaData <- within(newMETs$metaData, {
+  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
+})
+
+metKey <- newMETs$Data$AinsworthMETS
+names(metKey) <- activityCode
+
+```
+
+```{r readnewMETs, echo = TRUE, eval = TRUE, echo = FALSE}
+cpsFile <- "~/GHI/data/ATUS/atuscps_2015.dat"
+cps <- read.table(cpsFile, header=T, skip=0, sep="," )
+cps <- within(cps, {
+    TUCASEID <- as.character(TUCASEID)
+    TULINENO <- as.character(TULINENO)
+    GEREG <- as.factor(GEREG)
+    })
+cps <- subset(cps, TULINENO == "1")
+
+summaryFile <- "~/GHI/data/ATUS/atussum_2015.dat"
+summary <- read.table(summaryFile, header=T, skip=0, sep=",")
+metaDataColumns <- 1:24
+summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
+
+summaryList$metaData <- within(summaryList$metaData,{
+    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
+    AGECAT <- ifelse(TEAGE <5, "00-04",
+                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
+                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
+                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
+                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
+                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
+                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
+                            )))))))})
+
+```
+
+We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
+
+```{r metKey, echo = TRUE, eval = TRUE, echo = FALSE}
+allActivities <- colnames(summaryList$Data)
+nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
+nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
+
+activities <- nonTravelActivities
+
+metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
+names(metKeyVector) <- activities
+metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
+dimnames(metKeyMatrix) <- list( NULL, activities)
+
+METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
+
+D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
+ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
+
+```
+
+```{r parameters, eval = TRUE, echo = FALSE}
+parameters <- ddply(D, .(SEX, AGECAT), summarise,
+             mean = mean(METs, na.rm=T),
+             sd = sd(METs, na.rm=T)
+             )
+
+METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
+
+colnames(METmatrix) <- c(as.character(parameters$SEX[1]),as.character(rev(parameters$SEX)[1]))
+
+METmatrix <- METmatrix[,c("male","female")]
+
+METmatrix <- METmatrix/METmatrix[3,2]
+METmatrix[2,] <- METmatrix[3,]
+rownames(METmatrix) <- paste0("ageClass",1:8)
+colnames(METmatrix) <- c("M","F")
+METmatrix
+```
+
 ## Create List of ITHIM Objects
@@
 ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
+ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravelMatrix", METmatrix)
 ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
@@
 p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
-
-```
-
-Do it again, but with quintiles
-
-```{r ITHIMObjsQuintiles, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-DALYsTotal2 <- data.frame()
-ITHIMList <- list()
-
-ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )
-
-ITHIM.national <- ITHIM.default
-ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
-
-for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- ITHIM.default
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
-  }
-
-for( muNonTravel in seq(0, upper, by = wd)){
-
-  rm(DALYs)
-
-  for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
-  }
-
-  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
-
-  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
-  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
-  DALYsTotal2 <- rbind(DALYsTotal2, DALYs)
-}
-
-DALYsTotal2 <- within(DALYsTotal2, {
-  State <- factor(State, levels = stateVecOrdered)
-})
-
-DALYsTotal <- rbind(data.frame(DALYsTotal,quints="Percentiles"), data.frame(DALYsTotal2,quints="Quintiles"))
-
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( .
-~ quints)
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( .
-~ quints)
 
@@
 p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( .
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( .
 ~ disease, scales = "free_y")
-```
-
-## Non-Travel METs
-
-```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
-newMETsFile <- "ATUS_to_METS_key.csv"
-
-newMETsFile <- system.file(newMETsFile, package="ITHIM")
-
-newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
-dataColumns <- c(1,5)
-D <- newMETs[,dataColumns]
-metaData <- newMETs[,-dataColumns]
-newMETs <- list(Data = D, metaData = metaData)
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-
-rownames(newMETs$Data) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-
-newMETs$metaData <- within(newMETs$metaData, {
-  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
-})
-
-metKey <- newMETs$Data$AinsworthMETS
-names(metKey) <- activityCode
-
-```
-
-```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
-cpsFile <- "~/GHI/data/ATUS/atuscps_2015.dat"
-cps <- read.table(cpsFile, header=T, skip=0, sep="," )
-cps <- within(cps, {
-    TUCASEID <- as.character(TUCASEID)
-    TULINENO <- as.character(TULINENO)
-    GEREG <- as.factor(GEREG)
-    })
-cps <- subset(cps, TULINENO == "1")
-
-summaryFile <- "~/GHI/data/ATUS/atussum_2015.dat"
-summary <- read.table(summaryFile, header=T, skip=0, sep=",")
-metaDataColumns <- 1:24
-summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
-
-summaryList$metaData <- within(summaryList$metaData,{
-    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
-    AGECAT <- ifelse(TEAGE <5, "00-04",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
-                            )))))))})
-
-```
-
-We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
-
-```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
-allActivities <- colnames(summaryList$Data)
-nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
-nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
-
-activities <- nonTravelActivities
-
-metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
-names(metKeyVector) <- activities
-metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-dimnames(metKeyMatrix) <- list( NULL, activities)
-
-METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
-
-D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
-
-```
-
-```{r parameters, eval = FALSE, echo = FALSE}
-parameters <- ddply(D, .(SEX, AGECAT), summarise,
-             mean = mean(METs, na.rm=T),
-             sd = sd(METs, na.rm=T)
-             )
-
-(METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2)))
-
-colnames(METmatrix) <- c(as.character(parameters$SEX[1]),as.character(rev(parameters$SEX)[1]))
-
-METmatrix <- METmatrix[,c("male","female")]
-
-METmatrix <- METmatrix/METmatrix[3,2]
-
 ```
@@
 #~~~~~~~~~~~~~~~~~~~~~~~
-# Manual bullshit!
+# Manual bull$%#@!
 #~~~~~~~~~~~~~~~~~~~~~~~
@@
 ```
+
+### Appendix
+
+Do it again, but with quintiles
+
+```{r ITHIMObjsQuintiles, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+DALYsTotal2 <- data.frame()
+ITHIMList <- list()
+
+ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )
+
+ITHIM.national <- ITHIM.default
+ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+
+for (HHSTATE in stateVec){
+    ITHIMList[[HHSTATE]] <- ITHIM.default
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
+  }
+
+for( muNonTravel in seq(0, upper, by = wd)){
+
+  rm(DALYs)
+
+  for (HHSTATE in stateVec){
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
+  }
+
+  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+
+  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
+  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
+  DALYsTotal2 <- rbind(DALYsTotal2, DALYs)
+}
+
+DALYsTotal2 <- within(DALYsTotal2, {
+  State <- factor(State, levels = stateVecOrdered)
+})
+
+DALYsTotal <- rbind(data.frame(DALYsTotal,quints="Percentiles"), data.frame(DALYsTotal2,quints="Quintiles"))
+
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( .
+~ quints)
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( .
+~ quints)
+
+```

### 2016-11-23 methodsPaper/lib/materialsAndMethods.Rnw
Now methods paper is created with knitr
See files ./lib/*.Rnw for editing content.

--- 2016-11-07 methodsPaper/tex/materialsAndMethods.tex
@@
-As with the original ITHIM implementation we employed a comparative
-risk analysis (CRA) across scenarios defined by mean active
-transportation time, i.e., walking and cycling.  To do so we
-investigate the proportional change in national disease burden between
-the baseline model and a given alternative scenario, i.e., the
-population attributable fraction, $\af$.  The population attributable
-fraction is defined below and computed for each disease included in
-the model, namely Breast Cancer, Colon Cancer, Depression, Dementia,
-Diabetes and Cardiovascular disease.  For readability we omit an index
-indicating disease.
+As with the original ITHIM implementation we employ a comparative risk
+analysis (CRA) across scenarios defined by mean active transportation
+time, i.e., walking and cycling.  To do so we estimate the
+proportional change in disease burden between a baseline model and an
+alternative scenario, namely the population attributable fraction,
+$\af$.  The population attributable fraction is defined below and
+computed for each disease included in the model.
 
 \begin{equation}
-\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \,
-  \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x} \label{paf}
+\af_j = \frac{\int \! R_j(x)P(x) \, \mathrm{d}x  - \int \! R_j(x)Q(x) \,
+  \mathrm{d}x}{\int \!
+R_j(x)P(x) \, \mathrm{d}x} \approx 1 - \frac{{\displaystyle \sum_{i=1}^n} R_j(x_i^\prime)}{{\displaystyle \sum_{i=1}^n} R_j(x_i)}  = 1 - \delta_j % = 1 - \frac{{\displaystyle \sum_{i=1}^n} e^{-\alpha \sqrt{x_i^\prime}}}{{\displaystyle \sum_{i=1}^n} e^{-\alpha \sqrt{x_i}}}
+  \label{paf}
 \end{equation}
@@
 modeled as the sum of two independent random variables, travel-related
-and non-travel-related physical activity.
+and non-travel-related physical activity.  $\mathbf{x}$ and
+$\mathbf{x}^\prime$ represent the quantiles for the exposure
+distribution in baseline and alternative scenarios, respectively.
+$\delta_j$ is the proportional burden for disease $j$ in the alternate
+scenario.
 
@@
 
-\paragraph{Estimation of Disease Burden}
+%To estimate the disease burden we first approximate the population
+%attributable fraction through a discretization of the integral in
+%equation \ref{paf}.
 
-To estimate the disease burden we first approximate the population
-attributable fraction through a discretization of the integral in
-equation \ref{paf}.
+\paragraph{Non-Travel-Related Physical Activity}
 
@@
 
-\begin{equation}
-\delta = 1 - \af \approx \frac{\sum_{i=1}^n R(x_i)}{\sum_{i=1}^n R(x_i^\prime)}
-\end{equation}
+%% \begin{equation}
+%% \delta = 1 - \af \approx \frac{{\displaystyle \sum_{i=1}^{n}} R(x_i)}{{\displaystyle \sum_{i=1}^{n}} R(x_i^\prime)}
+%% \end{equation}
+
+\paragraph{Estimation of Disease Burden}
 
 The disease burden, in this case DALYs, is found for each of the
-percentiles using equation \ref{burden}, using $B_0$, the overall nation-wide burden for
-the baseline.  The overall burden is then computed as the sum across
-quintiles, $B=\sum_{j=1}^n B_j$.
+quantiles using equation \ref{burden} using $B_0$, the overall
+nation-wide burden for the baseline.  The overall burden is then
+computed as the sum across quintiles, $B=\sum_{j=1}^n B_j$.
 
 \begin{equation}
-B_j = \delta B_0\frac{\RR_j}{\sum_{k=1}^n \RR_k}\label{burden}
+B_j = \delta B_0\frac{\RR_j}{\sum_{k=1}^n \RR_j}\label{burden}
 \end{equation}
@@
 \begin{equation}
-\RR(x) = \mathrm{e}^{-r \tilde{x}}
+\RR_j(x) = \mathrm{e}^{-\alpha_j \tilde{x}}
 \end{equation}
@@
 
-\begin{figure}[t]
+\begin{sidewaysfigure}[t]
   \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
@@
   }\label{fig3}
-\end{figure}
+\end{sidewaysfigure}

### 2016-11-23 methodsPaper/tex/materialsAndMethods.Rnw
Now methods paper is created with knitr
See files ./lib/*.Rnw for editing content.

--- 2016-11-23 methodsPaper/lib/materialsAndMethods.Rnw
@@
+\Sexpr{1=1}
 As with the original ITHIM implementation we employ a comparative risk

### 2016-11-29 R/ITHIM.Rmd
Added WI and MI to CA and NY comparison code chunk

@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-stateVec <- c("CA","TX","NY","FL") #c("CA","TX","NY","FL","VA","NC","AZ")
-stateVecOrdered <- c("TX","FL","NY","CA") #c("NC","TX","VA","FL","AZ","NY","CA")
+stateVec <- c("CA","TX","NY","FL","WI","MI") #c("CA","TX","NY","FL","VA","NC","AZ")
+stateVecOrdered <- c("TX","FL","NY","CA","WI","MI") #c("NC","TX","VA","FL","AZ","NY","CA")
 
@@
 
+#~~~~~~~~~~
+
 ITHIM.CA <- ITHIMList[["CA"]]
@@
 names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
+
+#~~~~~~~~~~
 
@@
 })
+
+#~~~~~~~~~~
 
@@
 
-### Appendix
+## Wisconsin & Michigan
 
-Do it again, but with quintiles
+```{r wimi, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
 
-```{r ITHIMObjsQuintiles, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-DALYsTotal2 <- data.frame()
-ITHIMList <- list()
+ITHIM.NY <- ITHIMList[["NY"]]
+ITHIM.CA <- ITHIMList[["CA"]]
+ITHIM.MI <- ITHIMList[["MI"]]
+ITHIM.WI <- ITHIMList[["WI"]]
 
-ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )
+MItoCA <- data.frame(lapply(compareModels(ITHIM.MI,ITHIM.CA)$daly.delta, function(x) sum(x$M + x$F)), baseline = "MI", scenario = "CA")
+WItoCA <- data.frame(lapply(compareModels(ITHIM.WI,ITHIM.CA)$daly.delta, function(x) sum(x$M + x$F)), baseline = "WI", scenario = "CA")
+MItoNY <- data.frame(lapply(compareModels(ITHIM.MI,ITHIM.NY)$daly.delta, function(x) sum(x$M + x$F)), baseline = "MI", scenario = "NY")
+WItoNY <- data.frame(lapply(compareModels(ITHIM.WI,ITHIM.NY)$daly.delta, function(x) sum(x$M + x$F)), baseline = "WI", scenario = "NY")
 
-ITHIM.national <- ITHIM.default
-ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+D <- melt(rbind(MItoCA,WItoCA,MItoNY,WItoNY))
+names(D) <- c("baseline","scenario","disease","DALYs")
 
-for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- ITHIM.default
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
-  }
+p <- ggplot(D, aes(x = disease, y = -DALYs/1e3, fill = scenario)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + scale_y_continuous(name="DALYs Averted (thousands)") + facet_grid( . ~ baseline) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))
 
-for( muNonTravel in seq(0, upper, by = wd)){
-
-  rm(DALYs)
-
-  for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
-  }
-
-  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
-
-  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
-  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
-  DALYsTotal2 <- rbind(DALYsTotal2, DALYs)
-}
-
-DALYsTotal2 <- within(DALYsTotal2, {
-  State <- factor(State, levels = stateVecOrdered)
-})
-
-DALYsTotal <- rbind(data.frame(DALYsTotal,quints="Percentiles"), data.frame(DALYsTotal2,quints="Quintiles"))
-
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( .
-~ quints)
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( .
-~ quints)
-
+#+ scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( .
+~ disease, scales = "free_y")
 ```

### 2016-12-02 R/ITHIM.Rmd
Started working with road fatalities

@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-stateVec <- c("CA","TX","NY","FL","WI","MI") #c("CA","TX","NY","FL","VA","NC","AZ")
-stateVecOrdered <- c("TX","FL","NY","CA","WI","MI") #c("NC","TX","VA","FL","AZ","NY","CA")
+stateVec <- c("CA","TX","NY","FL","WI","MI", "OH") #c("CA","TX","NY","FL","VA","NC","AZ")
+stateVecOrdered <- c("TX","FL","NY","CA","WI","MI", "OH") #c("NC","TX","VA","FL","AZ","NY","CA")
 
@@
 ITHIM.WI <- ITHIMList[["WI"]]
+ITHIM.WI <- ITHIMList[["OH"]]
 
-MItoCA <- data.frame(lapply(compareModels(ITHIM.MI,ITHIM.CA)$daly.delta, function(x) sum(x$M + x$F)), baseline = "MI", scenario = "CA")
-WItoCA <- data.frame(lapply(compareModels(ITHIM.WI,ITHIM.CA)$daly.delta, function(x) sum(x$M + x$F)), baseline = "WI", scenario = "CA")
-MItoNY <- data.frame(lapply(compareModels(ITHIM.MI,ITHIM.NY)$daly.delta, function(x) sum(x$M + x$F)), baseline = "MI", scenario = "NY")
-WItoNY <- data.frame(lapply(compareModels(ITHIM.WI,ITHIM.NY)$daly.delta, function(x) sum(x$M + x$F)), baseline = "WI", scenario = "NY")
+prop <- 1/50
+state1Vec <- c("MI","WI","OH")
+state2Vec <- c("NY","CA")
 
-D <- melt(rbind(MItoCA,WItoCA,MItoNY,WItoNY))
+foobar <- data.frame()
+
+for( state1 in state1Vec ){
+    for( state2 in state2Vec ){
+        foo <- data.frame(lapply(compareModels(ITHIMList[[state1]],ITHIMList[[state2]])$daly.delta, function(x) sum(x$M + x$F)*prop), baseline = state1, scenario = state2)
+        foobar <- rbind(foobar,foo)
+    }
+}
+
+D <- melt(foobar)
 names(D) <- c("baseline","scenario","disease","DALYs")
@@
 p <- ggplot(D, aes(x = disease, y = -DALYs/1e3, fill = scenario)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_y_continuous(name="DALYs Averted (thousands)") + facet_grid( . ~ baseline) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))
+p + theme_bw() + scale_y_continuous(name="DALYs Averted (thousands)") + facet_grid( .
+~ baseline) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + theme(legend.title=element_blank()) + scale_x_discrete(name="Disease")
 
-#+ scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( .
-~ disease, scales = "free_y")
+#+ theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( . ~ disease, scales = "free_y")
 ```
+
+## Road Fatalities
+
+```{r roadFat}
+pedDeath <- readRDS(file = "./data/BikePedDeathModelData.rds")
+pedDeath <- subset(pedDeath, urbSubRur == "Urban")
+pedDeath <- within(pedDeath, pedDeathLogical <- pedDeaths0913 > 0)
+pedDeath <- within(pedDeath,{
+  deathsTotal <- pedDeaths0913 + bikeDeaths0913
+  deathsTotalLogical <- deathsTotal > 0
+  activeTract <- (pctWalkToWork + pctBikeToWork ) > 0
+})
+
+D <- melt(pedDeath)
+
+chisq.test(with(pedDeath, table(deathsTotalLogical, activeTract)))
+
+p <- ggplot(subset(D, variable %in% c("vehMiles","pctWalkToWork","activeTract")), aes(deathsTotalLogical, value, colour = deathsTotalLogical))
+p + geom_boxplot() + facet_grid(variable ~ ., scales = "free_y")
+```

### 2016-12-02 methodsPaper/tex/materialsAndMethods.tex
Edits to Materials & Methods
added ICD-10 codes
need to verify Depression

@@
-As with the original ITHIM implementation we employed a comparative
-risk analysis (CRA) across scenarios defined by mean active
-transportation time, i.e., walking and cycling.  To do so we
+As with the original ITHIM implementation we, employed a comparative
+risk assessment (CRA) across scenarios defined by mean active
+transportation time, i.e., combined walking and cycling.  To do so we
 investigate the proportional change in national disease burden between
@@
 population attributable fraction, $\af$.  The population attributable
-fraction is defined below and computed for each disease included in
-the model, namely Breast Cancer, Colon Cancer, Depression, Dementia,
-Diabetes and Cardiovascular disease.  For readability we omit an index
+fraction is defined below and computed for each disease (ICD-10) included in
+the model, namely Breast Cancer (C50), Colon Cancer (C18-C21), Depression (??), Dementia (G30),
+Diabetes (E10-E14) and Cardiovascular disease (I10, I12, I15, I20-I25, I30-I31, I33, I40).  For readability we omit an index
 indicating disease.
@@
 $r_{ij} = \frac{\mu_{ij}}{\mu_0}$ and $\gamma$ are estimated from the
-American Time USe Survey data \cite{ATUS}.
+American Time Use Survey data \cite{ATUS}.
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Tue Nov  8 09:38:46 2016
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrr}
+  \hline
+ & M & F \\
+  \hline
+0-4 & 0.00 & 0.00 \\
+  5-14 & 0.00 & 0.00 \\
+  15-29 & 0.97 & 1.00 \\
+  30-44 & 1.04 & 1.12 \\
+  45-59 & 0.95 & 0.99 \\
+  60-69 & 0.90 & 0.94 \\
+  70-79 & 0.83 & 0.88 \\
+  80+ & 0.72 & 0.77 \\
+   \hline
+\end{tabular}
+\end{table}
 
@@
 To test the new implementation of ITHIM's active transportation
-component, we used the UW national metro population estimates for
+component, we used the US national metro population estimates for
 active travel time (walking and cycling) as the baseline. We obtained
@@
 states with large sample sizes. From these examples we were able to
-approximate low and high active travel scenarios. Finally we applied
+approximate low and high active travel scenarios. Finally, we applied
 these active travel times to the US metro population to perform the

### 2016-12-02 R/ITHIM.Rmd
Created RoadFatalities.Rmd

@@
 ```
-
-## Road Fatalities
-
-```{r roadFat}
-pedDeath <- readRDS(file = "./data/BikePedDeathModelData.rds")
-pedDeath <- subset(pedDeath, urbSubRur == "Urban")
-pedDeath <- within(pedDeath, pedDeathLogical <- pedDeaths0913 > 0)
-pedDeath <- within(pedDeath,{
-  deathsTotal <- pedDeaths0913 + bikeDeaths0913
-  deathsTotalLogical <- deathsTotal > 0
-  activeTract <- (pctWalkToWork + pctBikeToWork ) > 0
-})
-
-D <- melt(pedDeath)
-
-chisq.test(with(pedDeath, table(deathsTotalLogical, activeTract)))
-
-p <- ggplot(subset(D, variable %in% c("vehMiles","pctWalkToWork","activeTract")), aes(deathsTotalLogical, value, colour = deathsTotalLogical))
-p + geom_boxplot() + facet_grid(variable ~ ., scales = "free_y")
-```

### 2016-12-02 R/RoadFatalities.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("RoadFatalities.Rmd","RoadFatalities.html")

### 2016-12-05 methodsPaper/tex/materialsAndMethods.tex

@@
 fraction is defined below and computed for each disease (ICD-10) included in
-the model, namely Breast Cancer (C50), Colon Cancer (C18-C21), Depression (??), Dementia (G30),
-Diabetes (E10-E14) and Cardiovascular disease (I10, I12, I15, I20-I25, I30-I31, I33, I40).  For readability we omit an index
+the model, namely Breast Cancer (C50), Colon Cancer (C18-C21), Dementia (G30),
+Diabetes (E10-E14) and Cardiovascular disease (I10, I12, I15, I20-I25, I30-I31, I33, I40).
+We also include the disability impacts of depression, though it is not listed as a cause of death in CDC's data.
+For readability we omit an index
 indicating disease.

### 2016-12-05 methodsPaper/tex/results.tex

@@
 We report total change (averted or increased) in diability adjusted
-life years (DALYs) as an estimate of dissease burden change with
+life years (DALYs) as an estimate of disease burden change with
 alternative scenarios.
@@
 were obtained form the CDC WONDER database for US metro counties for
-the years 2010-2014. Cells with less than 10 deaths were imputed using
+the years 2010-2014. Cells with mortality counts less than 10 deaths were suppressed for privacy and imputed (0 to 9) using
 R's random integer generator.
@@
 Our subset of the US population to only metropolitan counties
-consisted of 1,090 of 3,141 (34.7\%) counties, or ???X??? of ???Y???
-(???Z???\%) of the total US population. The age and gender
+consisted of 1,090 of 3,141 (34.7\%) counties, or approximately 272M of 312M
+(87\%) of the total US population. The age and gender
 distribution of our subset population was slightly younger (???making
-this up???) than the national population. Active transport times
-(walking and cycling) for baseline, low, and high scenarios are shown
-in \ref{fig2}. Overall mean walking time for high and low
-scenarios were ???X??? (min/week) higher and ???Y??? (min/week) lower
-than baseline, respectively. Baseline non-travel activity time was
-estimated to be ???X??? (min/week) from the ATUS for US metropolitan
+this up???) than the national population. Baseline active transport times
+(walking and cycling) for the US across age and sex groupings was found to be approximately 43 min/week (???).
+Baseline non-travel activity time was found to be ???X???
+(min/week) from the ATUS for US metropolitan
 populations.
-
-Comparing baseline to low/high scenarios produced estimates of ???X???
-change in the disease burden (1,000s DALYs) for our US population
-subset. We then varied the mean non-travel activity time and active
-travel time to observe how this chaged results (\ref{fig1}). We
-observe the most dramatic responses in disease burden estimates to
-changes in active travel time when non-travel activity times are
-lowest. This result is consistent with the non-linear dose-response
-curve for physical activity and several health outcomes. The total
-health benefit of increased active travel time (high scenario) varies
-from ???X??? to ???Y??? with changing mean non-travel activity time (0
-to 25 minutes).
 
@@
 As exhibited in Figure \ref{fig1}, the variability in disease burden
-benfits varies to a much greater degree when using fewer points in the
-attirbutable fraction approximation.
+benefits varies to a much greater degree when using fewer points in the
+attirbutable fraction approximation. Importantly, we see greater stability when using
+the package and percentiles to approximate attirbutable fraction.
+
+Comparing baseline to scenarios with increased active travel produced estimates of up to
+360,000??? fewer DALYs (with 300\% increase in mean active travel time) for our US population subset. We then varied the mean non-travel activity time and active
+travel time to observe how this chaged results (\ref{fig1}). We
+observe the most dramatic responses in disease burden estimates to
+changes in active travel time when non-travel activity times are
+lowest. This result is consistent with the non-linear dose-response
+curve for physical activity and several health outcomes.

### 2016-12-05 methodsPaper/lib/abstract.Rnw

--- 2016-10-31 methodsPaper/tex/abstract.tex
@@
-Comparative risk assessment analyses to quantify the health potential
+Comparative risk assessments to quantify the health potential
 of active transportation, e.g., walking and cycling, have been
@@
 the impact active transportation has on their
-state/county/tract/region.
-
-%% While increased active transportation, primarily cycling and walking,
-%% is often held up as a goal for improved public health --- via reduced
-%% chronic disease, improved air quality, and reductions in greenhouse
-%% gas emissions — obtaining estimates for the quantiative impacts of
-%% such shifts remains infrequent, if not altogether absent from related
-%% policy action. Using U.S. metropolitan counties from seven states we
-%% perform a comparative risk analysis between each regions and national
-%% averages as the baseline.  We employ a portion of the Integrated
-%% Transport and Health Impacts Model (ITHIM) to investigate the reduced
-%% chronic disease burden of ($x$, $y$, $z$ outcomes) from changing
-%% levels of physical activity with population shifts in the amount of
-%% biking and walking done for transport. The analysis is informed by
-%% regional travel parameters from the National Household Transportation
-%% Survey, non-travel activity metrics from the American Time Use Survey,
-%% and baseline disease burden from the Centers for Disease Control and
-%% Prevention's WONDER database and the Global Burden of Disease dataset.
-%% Our findings show that the California and New York State have travel
-%% patterns that, if mirrored by the nation, would save the U.S. upwards
-%% of fifty thousand disability adjusted life years.
-
-%\subsection{Summary:}
-%% \section{Availability:}
-
-%% %% \package{} may be explored using the user interface avaiable
-%% %% at \webpage{}.
-%% The source code and data files for \package{} are publicly available
-%% on GitHub via a link on \webpage{}.
-
-% \section{Supplementary information:}
-
-% See \webpage{} for links to the source code, manual, tutorial and
-% links for more information about the ITHIM model.
-
-% \section{Contact:}
-% \href{javargo@wisc.edu}{javargo@wisc.edu}
+state/region.

### 2016-12-05 methodsPaper/lib/introduction.Rnw

--- 2016-10-31 methodsPaper/tex/introduction.tex
@@
 \cite{maher2013}. Still only half of Americans obtain the recommended
-levels of aerobic activity, which is 2.5 hours per week. There is
+levels of aerobic activity of 2.5 hours per week. There is
 considerable variability in the attainment of this goal by location
@@
 creating a simple user-interface for the custom-built \R{} package
-\package{}.
-
-Here we present this implementation using publicly-available
-datasets with coverage of the entire United States, and investigate
-the sensitivity of the model outputs to variability in the input
-parameters. The results of this work suggest that statistical modeling
-and tools like the custom-built \R{} package \package{} can improve
-the application of this useful methodology and increase its utility to
-both practicioners and researchers.
+\package{}.  We present this implementation here by using it to
+estimate the relationship between overall, nationwide disease burden
+and active transportation time in the United States.

### 2016-12-05 methodsPaper/tex/materialsAndMethods.Rnw

--- 2016-11-23 methodsPaper/lib/materialsAndMethods.Rnw
@@
-As with the original ITHIM implementation we employ a comparative risk
-analysis (CRA) across scenarios defined by mean active transportation
-time, i.e., walking and cycling.  To do so we estimate the
-proportional change in disease burden between a baseline model and an
-alternative scenario, namely the population attributable fraction,
-$\af$.  The population attributable fraction is defined below and
-computed for each disease included in the model.
+As with the original ITHIM implementation we, employed a comparative
+risk assessment (CRA) across scenarios defined by mean active
+transportation time, i.e., combined walking and cycling.  To do so we
+investigate the proportional change in national disease burden between
+the baseline model and a given alternative scenario, i.e., the
+population attributable fraction, $\af$.  The population attributable
+fraction is defined below and computed for each disease (ICD-10)
+included in the model, namely Breast Cancer (C50), Colon Cancer
+(C18-C21), Dementia (G30), Diabetes (E10-E14) and Cardiovascular
+disease (I10, I12, I15, I20-I25, I30-I31, I33, I40).  We also include
+the disability impacts of depression, though it is not listed as a
+cause of death in CDC's data. For readability we omit an index
+indicating disease.
 
@@
 $r_{ij} = \frac{\mu_{ij}}{\mu_0}$ and $\gamma$ are estimated from the
-American Time USe Survey data \cite{ATUS}.
+American Time Use Survey data \cite{ATUS}.
 

### 2016-12-07 R/ITHIM.Rmd
Created gbdList by state

@@
 METmatrix
+```
+
+## GBD Estimates by State
+
+```{r stateBurden, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+
+stateNameToAbbrev <- function( stateName ){
+  stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
+
+  stateAbbrev <- stateAbbrevMatrix[,2]
+  names(stateAbbrev) <- as.character(stateAbbrevMatrix[,1])
+  return(stateAbbrev[stateName])
+}
+
+gbdParse <- function(gbd){
+  gbdListTemp <- split(gbd,gbd$disease)
+  gbdListTemp[["CVD"]] <- data.frame(disease = "CVD", gbdListTemp$IHD[,c("sex",  "ageClass")], gbdListTemp$IHD[,c("dproj","yll","yld","daly")] + gbdListTemp$InflammatoryHD[,c("dproj","yll","yld","daly")] + gbdListTemp$HHD[,c("dproj","yll","yld","daly")])
+  gbdListTemp2 <- lapply(gbdListTemp,function(x) split(x,as.factor(x$sex)))
+  gbdListTemp2 <- lapply(gbdListTemp2, function(x) list(M=x$M,F=x$F))
+  return(gbdListTemp2)
+}
+
+gbdRDS <-  readRDS(file = "./data/stateMetroBurdens2014.rds")
+gbdRDS <- within(gbdRDS, state <- stateNameToAbbrev(state))
+gbdByState <- split(gbdRDS,gbdRDS$state)
+gbdList <- lapply(gbdByState, gbdParse)
+
 ```

### 2016-12-07 R/ITHIM.Rmd
Created GBD by state figure

@@
 gbdByState <- split(gbdRDS,gbdRDS$state)
+gbdByState <- lapply(gbdByState, function(x) x[,!names(x)=="state"])
 gbdList <- lapply(gbdByState, gbdParse)
@@
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-stateVec <- c("CA","TX","NY","FL","WI","MI", "OH") #c("CA","TX","NY","FL","VA","NC","AZ")
-stateVecOrdered <- c("TX","FL","NY","CA","WI","MI", "OH") #c("NC","TX","VA","FL","AZ","NY","CA")
+stateVec <- c("CA","NY","WI","MI","OH") #c("CA","TX","NY","FL","VA","NC","AZ")
+stateVecOrdered <- stateVec #c("NY","CA","WI","MI","OH") #c("NC","TX","VA","FL","AZ","NY","CA")
 
-upper <- 10; length <- 5; wd <- upper/length;
+upper <- 10; length <- 2; wd <- upper/length;
 
@@
     ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", gbdList[[HHSTATE]])
   }
@@
 
+```
+
+```{r statewiseGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+D <- melt(lapply(ITHIMList, function(ITHIM) return(ITHIM$parameters$GBD)))
+names(D) <- c("disease","sex","ageClass","type", "burden", "sex2", "disease2", "state")
+
+D <- within(D, sex <- factor(sex, levels = c("F","M")))
+D <- within(D, state <- factor(state, levels = c("WI","MI","OH","NY","CA")))
+
+p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( disease ~ state, scales = "free_y")
 ```

### 2016-12-08 R/ITHIM.Rmd
compute <- FALSE to compute <- TRUE
Update to ITHIM.Rmd

@@
 p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( disease ~ state, scales = "free_y")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( disease ~ state, scales = "free_y)"
 ```

### 2016-12-09 R/ITHIM.Rmd
Created Figure for Rust BElt Disease Burden

@@
 
-p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( disease ~ state, scales = "free_y)"
+D <- subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD"))
+D <- subset(D, as.character(state) %in% c("WI","MI","OH"))
+D <- subset(D, type == "daly")
+
+p <- ggplot(D, aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( disease ~ state, scales = "free_y")
+```
+
+```{r rustBelt, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
 ```

### 2016-12-09 R/ITHIM.Rmd
First shot at figure for Jonathan on Dec. 12

@@
 
-```{r rustBelt, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r rustBelt, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+ITHIMListScenario <- lapply(ITHIMList, function(ithim) updateITHIM(ithim, parName = "muwt", parValue = ithim$parameters$muwt + 10*7))
+
+deltaDALYList <- mapply(sumDALY, ITHIMList, ITHIMListScenario)
+
+getDALY <- function( ITHIM ){
+foo <- data.frame(lapply(ITHIM$parameters$GBD, function(x) x$M$daly + x$F$daly))
+sum(foo[,names(foo) %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")])
+}
+
+burdenList <- lapply(ITHIMList, getDALY)
+foo <- data.frame(daly = unlist(burdenList), delta = deltaDALYList)
+foo <- within(foo, {
+  daly <- daly + delta
+  delta <- -delta
+  state <- names(burdenList)
+})
+
+foo <- subset(foo, state %in% c("WI", "MI","OH"))
+
+D <- melt(foo)
+
+cbPalette <- c("#0072B2", "#56B4E9", "#999999", "#E69F00", "#009E73", "#F0E442", "#D55E00", "#CC79A7")
+
+p <- ggplot(D, aes(state, value/1e3, fill = variable))
+p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + coord_cartesian(ylim = c(175, 600)) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
 

### 2016-12-12 R/ITHIM.Rmd
Created more figures for WI/MI/OH

@@
 
-foo <- subset(foo, state %in% c("WI", "MI","OH"))
+foo <- subset(foo, state %in% c("MI","OH"))
 
@@
 p <- ggplot(D, aes(state, value/1e3, fill = variable))
-p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + coord_cartesian(ylim = c(175, 600)) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
+p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip() # + coord_cartesian(ylim = c(175, 600))
+```
 
+```{r rustBelt2, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+D <- rbind(data.frame(getWalkTime(ITHIMList[["MI"]]),state = "MI"), data.frame(getWalkTime(ITHIMList[["WI"]]),state = "WI"), data.frame(getWalkTime(ITHIMList[["OH"]]),state = "OH"))
+
+p <- ggplot(subset(D, ageClass != "ageClass1"), aes(x = ageClass, y = mu, fill = state)) + geom_bar(stat= "identity", position = "dodge")
+p + theme_bw() + facet_grid( .
+~ sex) + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Time Walking per Week (minutes)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + scale_fill_manual(values=cbPalette[-3])
 ```

### 2016-12-13 R/ITHIM.Rmd
Updated pomodoros

@@
 gbdRDS <- within(gbdRDS, state <- stateNameToAbbrev(state))
+
 gbdByState <- split(gbdRDS,gbdRDS$state)

### 2016-12-17 R/ITHIM2.Rmd
Created ITHIM2.Rmd

--- 2016-12-09 R/ITHIM.Rmd
@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-document("~/ITHIM/")
-install("~/ITHIM/")
-#install_github("syounkin/ITHIM", ref="devel")
+#document("~/ITHIM/")
+#install("~/ITHIM/")
+install_github("syounkin/ITHIM", ref="classes")
 install_github("smbache/magrittr", ref="master")
@@
 
-## HHS Metro Region Parameter Sets
-
-First we use NHTS to estimate mean walking and cycling times.
-
-### NHTS
-
-```{r compute, echo = FALSE, eval = TRUE}
-compute <- FALSE
+```{r tilePlot, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+ITHIM <- createITHIM()
+tilePlot(ITHIM, n = 50)
 ```
-
-#### Create a handful of functions to help parse data
-
-```{r parsely, eval = TRUE, echo = FALSE}
-convertToHHS <- function(states){
-  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
-  StateToHHS <- HHSRegions[,3]
-  names(StateToHHS) <- HHSRegions[,1]
-  return(StateToHHS[states])
-}
-
-countPeople <- function(data){
- nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
-}
-
-tripTime <- function(data, mode = "walk"){
- return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
-}
-
-reformatData <- function(data){
-foo <- matrix(unlist(data), nrow = 7, ncol = 2)
-foo <- rbind( c(0, 0), foo)
-dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
-return(foo[,c("M","F")])
-}
-
-```
-
-#### Read Two of the Four NHTS Data Tables
-
-```{r NHTS, eval = compute, echo = FALSE}
-DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
-names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
-
-NHTS.trips <- with(DAYV2PUB,{
-    data.frame(HOUSEID = HOUSEID,
-                   PERSONID = PERSONID,
-                   TDCASEID = TDCASEID,
-                   AT = ifelse(TRPTRANS == "22", "cycle",
-                        ifelse(TRPTRANS == "23", "walk",
-                        ifelse(!is.na(TRPTRANS), "other", NA))),
-                   TRVL_MIN = as.numeric(TRVL_MIN),
-                   #TRVLCMIN = as.numeric(TRVLCMIN),
-                   URBRUR  = as.factor(URBRUR),
-                   MSACAT = as.factor(MSACAT),
-                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
-                   TRPTRANS = as.factor(TRPTRANS),
-                   TRPMILES = as.numeric(TRPMILES)
-)})
-
-PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
-names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
-
-NHTS.person <- with(PERV2PUB,{
-                    data.frame(HOUSEID = HOUSEID,
-                               PERSONID = PERSONID,
-                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
-                               AGE = as.factor(convertToAgeClass(R_AGE)),
-                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
-                               HHSTATE = as.character(HHSTATE),
-                               HHS = as.character(convertToHHS(as.character(HHSTATE))), stringsAsFactors = FALSE
-
-)
-})
-
-```
-
-#### Subset and Merge
-
-We now correct for zero trip times in NHTS.
-
-```{r subset, eval = compute, fig.width = 10, echo = FALSE}
-NHTS.person <- subset( NHTS.person, MSA == TRUE )
-NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))
-
-NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
-NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
-saveRDS(NHTS, file = "./data/NHTS.rds")
-```
-
-```{r loadNHTS, eval = !compute}
-NHTS <- readRDS(file = "./data/NHTS.rds")
-```
-
-#### Compute National Metro Average
-
-```{r natl, eval = TRUE, echo = FALSE, warning = FALSE}
-listOfLists <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
-counts <- lapply(listOfLists, function(x) lapply(x, countPeople))
-walkLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "walk"))
-cycleLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "cycle"))
-
-countMatrix <- reformatData(counts)
-walkMatrix <- reformatData(walkLength)
-cycleMatrix <- reformatData(cycleLength)
-
-meanWalkList <- walkMatrix/countMatrix*7
-meanCycleList <- cycleMatrix/countMatrix*7
-
-RwtList <- ifelse(meanWalkList == 0, 1e-4, meanWalkList)
-RctList <- ifelse(meanCycleList == 0, 1e-4, meanCycleList)
-
-RwtNational <- RwtList/RwtList[3,2]
-RctNational <- RctList/RctList[3,2]
-muwtNational <- meanWalkList[3,2]
-muctNational <- meanCycleList[3,2]
-
-rm("listOfLists","counts","walkLength","cycleLength","countMatrix","walkMatrix","cycleMatrix","meanWalkList","meanCycleList","RwtList","RctList","muwtList","muctList")
-```
-
-#### Compute Mean Trip Lengths
-
-```{r count, eval = TRUE, echo = FALSE}
-listOfLists <- lapply(split(NHTS, NHTS$HHSTATE),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
-counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
-walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
-cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
-countMatrix <- lapply(counts, reformatData)
-walkMatrix <- lapply(walkLength, reformatData)
-cycleMatrix <- lapply(cycleLength, reformatData)
-meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
-meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
-
-zeroReferentCase <- function(x){
-  if(x[3,2]==0){
-    return(matrix(1e-3,nrow=nrow(x),ncol=ncol(x), dimnames = dimnames(x)))
-  }else{
-    return(ifelse(x==0, 1e-3, x/x[3,2]))
-  }
-}
-
-RwtList <- lapply( meanWalkList, zeroReferentCase)
-RctList <- lapply( meanCycleList, zeroReferentCase)
-muwtList <- lapply( meanWalkList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
-muctList <- lapply( meanCycleList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
-```
-
-## Non-Travel METs
-
-```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
-newMETsFile <- "ATUS_to_METS_key.csv"
-
-newMETsFile <- system.file(newMETsFile, package="ITHIM")
-
-newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
-dataColumns <- c(1,5)
-D <- newMETs[,dataColumns]
-metaData <- newMETs[,-dataColumns]
-newMETs <- list(Data = D, metaData = metaData)
-activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))
-
-rownames(newMETs$Data) <- activityCode
-rownames(newMETs$metaData) <- activityCode
-
-newMETs$metaData <- within(newMETs$metaData, {
-  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
-})
-
-metKey <- newMETs$Data$AinsworthMETS
-names(metKey) <- activityCode
-
-```
-
-```{r readnewMETs, echo = TRUE, eval = TRUE, echo = FALSE}
-cpsFile <- "~/GHI/data/ATUS/atuscps_2015.dat"
-cps <- read.table(cpsFile, header=T, skip=0, sep="," )
-cps <- within(cps, {
-    TUCASEID <- as.character(TUCASEID)
-    TULINENO <- as.character(TULINENO)
-    GEREG <- as.factor(GEREG)
-    })
-cps <- subset(cps, TULINENO == "1")
-
-summaryFile <- "~/GHI/data/ATUS/atussum_2015.dat"
-summary <- read.table(summaryFile, header=T, skip=0, sep=",")
-metaDataColumns <- 1:24
-summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])
-
-summaryList$metaData <- within(summaryList$metaData,{
-    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
-    AGECAT <- ifelse(TEAGE <5, "00-04",
-                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
-                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
-                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
-                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
-                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
-                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
-                            )))))))})
-
-```
-
-We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.
-
-```{r metKey, echo = TRUE, eval = TRUE, echo = FALSE}
-allActivities <- colnames(summaryList$Data)
-nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
-nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]
-
-activities <- nonTravelActivities
-
-metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
-names(metKeyVector) <- activities
-metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
-dimnames(metKeyMatrix) <- list( NULL, activities)
-
-METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )
-
-D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
-ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))
-
-```
-
-```{r parameters, eval = TRUE, echo = FALSE}
-parameters <- ddply(D, .(SEX, AGECAT), summarise,
-             mean = mean(METs, na.rm=T),
-             sd = sd(METs, na.rm=T)
-             )
-
-METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
-
-colnames(METmatrix) <- c(as.character(parameters$SEX[1]),as.character(rev(parameters$SEX)[1]))
-
-METmatrix <- METmatrix[,c("male","female")]
-
-METmatrix <- METmatrix/METmatrix[3,2]
-METmatrix[2,] <- METmatrix[3,]
-rownames(METmatrix) <- paste0("ageClass",1:8)
-colnames(METmatrix) <- c("M","F")
-METmatrix
-```
-
-## GBD Estimates by State
-
-```{r stateBurden, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-
-stateNameToAbbrev <- function( stateName ){
-  stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
-
-  stateAbbrev <- stateAbbrevMatrix[,2]
-  names(stateAbbrev) <- as.character(stateAbbrevMatrix[,1])
-  return(stateAbbrev[stateName])
-}
-
-gbdParse <- function(gbd){
-  gbdListTemp <- split(gbd,gbd$disease)
-  gbdListTemp[["CVD"]] <- data.frame(disease = "CVD", gbdListTemp$IHD[,c("sex",  "ageClass")], gbdListTemp$IHD[,c("dproj","yll","yld","daly")] + gbdListTemp$InflammatoryHD[,c("dproj","yll","yld","daly")] + gbdListTemp$HHD[,c("dproj","yll","yld","daly")])
-  gbdListTemp2 <- lapply(gbdListTemp,function(x) split(x,as.factor(x$sex)))
-  gbdListTemp2 <- lapply(gbdListTemp2, function(x) list(M=x$M,F=x$F))
-  return(gbdListTemp2)
-}
-
-gbdRDS <-  readRDS(file = "./data/stateMetroBurdens2014.rds")
-gbdRDS <- within(gbdRDS, state <- stateNameToAbbrev(state))
-gbdByState <- split(gbdRDS,gbdRDS$state)
-gbdByState <- lapply(gbdByState, function(x) x[,!names(x)=="state"])
-gbdList <- lapply(gbdByState, gbdParse)
-
-```
-
-## Create List of ITHIM Objects
-
-```{r foobar, eval = TRUE, echo = FALSE}
-sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
-    return(sum(unlist(data.frame(compareModels(ITHIM.baseline,ITHIM.scenario)$daly.delta)[-1,]))) # AgeClass 1 is removed from totals
-    }
-```
-
-```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-stateVec <- c("CA","NY","WI","MI","OH") #c("CA","TX","NY","FL","VA","NC","AZ")
-stateVecOrdered <- stateVec #c("NY","CA","WI","MI","OH") #c("NC","TX","VA","FL","AZ","NY","CA")
-
-upper <- 10; length <- 2; wd <- upper/length;
-
-DALYsTotal <- data.frame()
-ITHIMList <- list()
-ITHIM.default <- createITHIM()
-ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
-ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
-ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
-ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravelMatrix", METmatrix)
-ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
-
-ITHIM.national <- ITHIM.default
-ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
-
-for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- ITHIM.default
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", gbdList[[HHSTATE]])
-  }
-
-for( muNonTravel in seq(0, upper, by = wd)){
-
-  rm(DALYs)
-
-  for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
-  }
-
-  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
-
-  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
-  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
-  DALYsTotal <- rbind(DALYsTotal, DALYs)
-}
-
-DALYsTotal <- within(DALYsTotal, {
-  State <- factor(State, levels = stateVecOrdered)
-})
-
-p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))
-
-p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
-
-```
-
-```{r statewiseGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- melt(lapply(ITHIMList, function(ITHIM) return(ITHIM$parameters$GBD)))
-names(D) <- c("disease","sex","ageClass","type", "burden", "sex2", "disease2", "state")
-
-D <- within(D, sex <- factor(sex, levels = c("F","M")))
-D <- within(D, state <- factor(state, levels = c("WI","MI","OH","NY","CA")))
-
-D <- subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD"))
-D <- subset(D, as.character(state) %in% c("WI","MI","OH"))
-D <- subset(D, type == "daly")
-
-p <- ggplot(D, aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( disease ~ state, scales = "free_y")
-```
-
-```{r rustBelt, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-
-ITHIMListScenario <- lapply(ITHIMList, function(ithim) updateITHIM(ithim, parName = "muwt", parValue = ithim$parameters$muwt + 10*7))
-
-deltaDALYList <- mapply(sumDALY, ITHIMList, ITHIMListScenario)
-
-getDALY <- function( ITHIM ){
-foo <- data.frame(lapply(ITHIM$parameters$GBD, function(x) x$M$daly + x$F$daly))
-sum(foo[,names(foo) %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")])
-}
-
-burdenList <- lapply(ITHIMList, getDALY)
-foo <- data.frame(daly = unlist(burdenList), delta = deltaDALYList)
-foo <- within(foo, {
-  daly <- daly + delta
-  delta <- -delta
-  state <- names(burdenList)
-})
-
-foo <- subset(foo, state %in% c("WI", "MI","OH"))
-
-D <- melt(foo)
-
-cbPalette <- c("#0072B2", "#56B4E9", "#999999", "#E69F00", "#009E73", "#F0E442", "#D55E00", "#CC79A7")
-
-p <- ggplot(D, aes(state, value/1e3, fill = variable))
-p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + coord_cartesian(ylim = c(175, 600)) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
-
-```
-
-### Visualization of trip time matrices
-
-```{r meanMatrices, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-
-ITHIMList <- c(ITHIMList, list("Natl." = ITHIM.national))
-
-meanMatrix <- function(ITHIM, mode = "Walking"){
-   if(mode == "Walking"){
-      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
-   }else if (mode == "Cycling"){
-      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
-   }else{
-      message("error in meanMatrix function")
-   }
-}
-
-meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "Walking")
-meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "Cycling")
-
-D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
-names(D) <- c("ageClass", "mode", "Sex", "time", "State")
-
-D <- within(D,{
-  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
-  Sex <- factor(Sex, levels = c("F","M"))
-})
-
-p <- ggplot(subset(D,ageClass != 1), aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ State, scales = "free_y") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 5))
-
-```
-
-```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
-ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)
-
-#~~~~~~~~~~
-
-ITHIM.CA <- ITHIMList[["CA"]]
-ITHIM.CA <- updateITHIM(ITHIM.CA, parName = "muNonTravel", 10)
-foo1 <- compareModels(ITHIM.national,ITHIM.CA)
-foobar1 <- data.frame(cbind(melt(foo1$daly.delta),paste0("ageClass",1:8)),State = "CA")
-
-foobar <- foobar1
-names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
-
-#~~~~~~~~~~
-
-ITHIM.NY <- ITHIMList[["NY"]]
-ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
-foo2 <- compareModels(ITHIM.national,ITHIM.NY)
-foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
-foobar <- rbind(foobar1,foobar2)
-names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")
-
-foobar <- within(foobar, {
-  Disease[Disease == "BreastCancer"] <- "Breast Cancer"
-  Disease[Disease == "ColonCancer"] <- "Colon Cancer"
-  Disease <- factor(Disease, levels = c("Breast Cancer", "Colon Cancer", "Dementia", "Depression", "Diabetes", "CVD"))
-})
-
-#~~~~~~~~~~
-
-p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = State))
-
-p + geom_bar(stat="identity", position = "dodge") +
-theme(axis.title=element_blank()) +
-facet_grid( Disease ~ Sex, scales = "free" ) +
-scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") +
-scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
-theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6)) +
-scale_fill_manual(values=c("#CC6666", "#9999CC"))
-
-```
-
-```{r counts, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- melt(counts)
-names(D) <- c("count","AgeClass","Sex", "State")
-D <- within(D,{
-  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
-  Sex <- factor(Sex, levels = c("F","M"))
-})
-p <- ggplot(subset(D,AgeClass != "00-04"), aes(AgeClass, count/1e3))
-p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Individuals Sampled (Thousands)") + facet_grid( State ~ .) + theme_bw() + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6))
-```
-
-## Figure 1
-
-```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
-
-ITHIM.national <- ITHIM.default
-ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
-
-ITHIM.scenario <- ITHIM.national
-ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
-ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)
-
-pWalk <- ITHIM.national$means$pWalk[3,2]
-
-DALYs <- c(); DALYs.quints <- c();
-atVec <- seq(45, 120, length.out = 10)
-muNonTravelVec <- c(1, 10, 25)
-DTotal <- data.frame()
-
-for(at in atVec ){  # minutes active transport/week
-  for( muNonTravel in muNonTravelVec ){
-
-    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muwt", at*pWalk)
-    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muct", at*(1-pWalk))
-    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muNonTravel", muNonTravel)
-
-    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muwt", at*pWalk)
-    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muct", at*(1-pWalk))
-    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muNonTravel", muNonTravel)
-
-    ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
-
-    DALYs <- sumDALY(ITHIM.national, ITHIM.scenario)
-    DALYs.quints <- sumDALY(ITHIM.national.quints, ITHIM.scenario.quints)
-
-    D <- data.frame(quants = c("percentiles","quintiles"), DALYs = c(DALYs,DALYs.quints),at = at, muNonTravel = muNonTravel)
-
-    DTotal <- rbind(D,DTotal)
-
-  }
-}
-
-DTotal <- within(DTotal, muNonTravel <- factor(muNonTravel, levels = as.character(muNonTravelVec), labels = c("low","medium","high")))
-DTotal <- subset(DTotal, quants == "percentiles")
-
-p <- ggplot(DTotal, aes(x = at, y = -DALYs/1e3, colour = muNonTravel))
-p + geom_line() + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)")  + scale_colour_discrete(name="Baseline Activity") #+ theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))
-
-```
-
-## Figure 2
-
-```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
-
-ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", 10)
-
-D <- rbind(getWalkTime(ITHIM.national),getWalkSpeed(ITHIM.national),getCycleTime(ITHIM.national),getNonTravelMETs(ITHIM.national))
-D <- within(D, sex <- factor(sex, levels = c("F","M")))
-
-p <- ggplot(D, aes(x = ageClass, y = mu, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( type ~ ., scales = "free")
-```
-
-## Figure 3
-
-```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 1.2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- melt(ITHIM.national$parameters$GBD)
-names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")
-D <- within(D, sex <- factor(sex, levels = c("F","M")))
-
-p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( .
-~ disease, scales = "free_y")
-```
-
-```{r doseresponse, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-RR1vec <- unique(unlist(createActiveTransportRRs()))
-lapply(createActiveTransportRRs(),function(x) lapply(x, function(x) unique(c(x))))
-#~~~~~~~~~~~~~~~~~~~~~~~
-# Manual bull$%#@!
-#~~~~~~~~~~~~~~~~~~~~~~~
-RR1vec <- RR1vec[RR1vec != 1]
-names(RR1vec) <- c("BreastCancer","ColonCancer.M","ColonCancer.F", "CVD", "Dementia", "Depression.young", "Depression.old", "Diabetes")
-
-RRx <- function(x, RR1, k) exp(log(RR1)*x^k)
-x <- seq(0,25,0.1)
-RRmat <- outer(x, RR1vec, RRx, k = 0.5)
-rownames(RRmat) <- x
-D <- melt(RRmat)
-names(D) <- c("x","disease","RR")
-
-p <- ggplot(D, aes(x = x, y = RR, colour = disease))
-p + geom_line()  + theme_bw() + scale_y_continuous(name="Relative Risk") + scale_x_continuous(name="Physical Activity (MET-hrs./week)")
-```
-
-## Wisconsin & Michigan
-
-```{r wimi, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-
-ITHIM.NY <- ITHIMList[["NY"]]
-ITHIM.CA <- ITHIMList[["CA"]]
-ITHIM.MI <- ITHIMList[["MI"]]
-ITHIM.WI <- ITHIMList[["WI"]]
-ITHIM.WI <- ITHIMList[["OH"]]
-
-prop <- 1/50
-state1Vec <- c("MI","WI","OH")
-state2Vec <- c("NY","CA")
-
-foobar <- data.frame()
-
-for( state1 in state1Vec ){
-    for( state2 in state2Vec ){
-        foo <- data.frame(lapply(compareModels(ITHIMList[[state1]],ITHIMList[[state2]])$daly.delta, function(x) sum(x$M + x$F)*prop), baseline = state1, scenario = state2)
-        foobar <- rbind(foobar,foo)
-    }
-}
-
-D <- melt(foobar)
-names(D) <- c("baseline","scenario","disease","DALYs")
-
-p <- ggplot(D, aes(x = disease, y = -DALYs/1e3, fill = scenario)) + geom_bar(stat= "identity", position = "dodge")
-p + theme_bw() + scale_y_continuous(name="DALYs Averted (thousands)") + facet_grid( .
-~ baseline) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + theme(legend.title=element_blank()) + scale_x_discrete(name="Disease")
-
-#+ theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( . ~ disease, scales = "free_y")
-```

### 2016-12-17 R/ITHIM.R

@@ -1,2 +1,3 @@
 library("knitr")
-knit2html("ITHIM.Rmd","ITHIM.html")
+## knit2html("ITHIM.Rmd","ITHIM.html")
+knit2html("ITHIM2.Rmd","ITHIM2.html")

### 2016-12-19 R/ITHIM.Rmd
Scripts now work with version 0.2

@@
 ITHIM.default <- createITHIM()
-ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
-ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
-ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
-ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravelMatrix", METmatrix)
-ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
+ITHIM.default <- update(ITHIM.default, "cvNonTravel", 3)
+ITHIM.default <- update(ITHIM.default, "cv", 1.65)
+ITHIM.default <- update(ITHIM.default, "muNonTravel", 4)
+ITHIM.default <- update(ITHIM.default, "muNonTravelMatrix", METmatrix)
+ITHIM.default <- update(ITHIM.default, "quantiles", (1:99)/100)
 
 ITHIM.national <- ITHIM.default
-ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+ITHIM.national <- update(ITHIM.national, "muwt", muwtNational)
+ITHIM.national <- update(ITHIM.national, "muct", muctNational)
+ITHIM.national <- update(ITHIM.national, "Rwt", RwtNational)
+ITHIM.national <- update(ITHIM.national, "Rct", RctNational)
 
@@
     ITHIMList[[HHSTATE]] <- ITHIM.default
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", gbdList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "GBD", gbdList[[HHSTATE]])
   }
@@
   for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
+    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
   }
 
-  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+  ITHIM.national <- update(ITHIM.national, "muNonTravel", muNonTravel)
 
@@
 ```{r statewiseGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- melt(lapply(ITHIMList, function(ITHIM) return(ITHIM$parameters$GBD)))
+D <- melt(lapply(ITHIMList, function(ITHIM) return(ITHIM@parameters@GBD)))
 names(D) <- c("disease","sex","ageClass","type", "burden", "sex2", "disease2", "state")
@@
 
-ITHIMListScenario <- lapply(ITHIMList, function(ithim) updateITHIM(ithim, parName = "muwt", parValue = ithim$parameters$muwt + 10*7))
+ITHIMListScenario <- lapply(ITHIMList, function(ithim) update(ithim, parName = "muwt", parValue = ithim@parameters@muwt + 10*7))
 
@@
 getDALY <- function( ITHIM ){
-foo <- data.frame(lapply(ITHIM$parameters$GBD, function(x) x$M$daly + x$F$daly))
+foo <- data.frame(lapply(ITHIM@parameters@GBD, function(x) x$M$daly + x$F$daly))
 sum(foo[,names(foo) %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")])
@@
    if(mode == "Walking"){
-      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
+      return(with(as(ITHIM@parameters,"list"), data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
    }else if (mode == "Cycling"){
-      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
+      return(with(as(ITHIM@parameters,"list"), data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
    }else{
@@
 ```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
-ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)
+ITHIM.national <- update(ITHIM.national, parName = "muNonTravel", 10)
 
@@
 ITHIM.CA <- ITHIMList[["CA"]]
-ITHIM.CA <- updateITHIM(ITHIM.CA, parName = "muNonTravel", 10)
+ITHIM.CA <- update(ITHIM.CA, parName = "muNonTravel", 10)
 foo1 <- compareModels(ITHIM.national,ITHIM.CA)
@@
 ITHIM.NY <- ITHIMList[["NY"]]
-ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
+ITHIM.NY <- update(ITHIM.NY, parName = "muNonTravel", 10)
 foo2 <- compareModels(ITHIM.national,ITHIM.NY)
@@
 ```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)
+ITHIM.default <- update(ITHIM.default, "quantiles", (1:99)/100)
 
 ITHIM.national <- ITHIM.default
-ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+ITHIM.national <- update(ITHIM.national, "muwt", muwtNational)
+ITHIM.national <- update(ITHIM.national, "muct", muctNational)
+ITHIM.national <- update(ITHIM.national, "Rwt", RwtNational)
+ITHIM.national <- update(ITHIM.national, "Rct", RctNational)
 
 ITHIM.scenario <- ITHIM.national
-ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
-ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)
+ITHIM.national.quints <- update(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
+ITHIM.scenario.quints <- update(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)
 
-pWalk <- ITHIM.national$means$pWalk[3,2]
+pWalk <- ITHIM.national@means$pWalk[3,2]
 
 DALYs <- c(); DALYs.quints <- c();
-atVec <- seq(45, 120, length.out = 10)
+atVec <- seq(45, 120, length.out = 3)
 muNonTravelVec <- c(1, 10, 25)
@@
 
-    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muwt", at*pWalk)
-    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muct", at*(1-pWalk))
-    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muNonTravel", muNonTravel)
+    ITHIM.scenario <- update(ITHIM.scenario, "muwt", at*pWalk)
+    ITHIM.scenario <- update(ITHIM.scenario, "muct", at*(1-pWalk))
+    ITHIM.scenario <- update(ITHIM.scenario, "muNonTravel", muNonTravel)
 
-    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muwt", at*pWalk)
-    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muct", at*(1-pWalk))
-    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muNonTravel", muNonTravel)
+    ITHIM.scenario.quints <- update(ITHIM.scenario.quints, "muwt", at*pWalk)
+    ITHIM.scenario.quints <- update(ITHIM.scenario.quints, "muct", at*(1-pWalk))
+    ITHIM.scenario.quints <- update(ITHIM.scenario.quints, "muNonTravel", muNonTravel)
 
-    ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+    ITHIM.national <- update(ITHIM.national, "muNonTravel", muNonTravel)
 
@@
 
-ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", 10)
+ITHIM.national <- update(ITHIM.national, "muNonTravel", 10)
 
@@
 ```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 1.2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- melt(ITHIM.national$parameters$GBD)
+D <- melt(ITHIM.national@parameters@GBD)
 names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")

### 2016-12-19 R/ITHIM2.Rmd
Scripts now work with version 0.2

@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-#document("~/ITHIM/")
-#install("~/ITHIM/")
-install_github("syounkin/ITHIM", ref="classes")
+document("~/ITHIM/")
+install("~/ITHIM/")
+#install_github("syounkin/ITHIM", ref="classes")
 install_github("smbache/magrittr", ref="master")

### 2016-12-19 R/ITHIM.R

@@ -1,3 +1,3 @@
 library("knitr")
-## knit2html("ITHIM.Rmd","ITHIM.html")
+knit2html("ITHIM.Rmd","ITHIM.html")
 knit2html("ITHIM2.Rmd","ITHIM2.html")

### 2016-12-20 R/ITHIM.Rmd
Worked on ITHIMList update

@@
 
-```{r foobar, eval = TRUE, echo = FALSE}
-sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
-    return(sum(unlist(data.frame(compareModels(ITHIM.baseline,ITHIM.scenario)$daly.delta)[-1,]))) # AgeClass 1 is removed from totals
-    }
-```
-
 ```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-stateVec <- c("CA","NY","WI","MI","OH") #c("CA","TX","NY","FL","VA","NC","AZ")
+stateVec <- names(listOfLists) #c("CA","NY","WI","MI","OH") #c("CA","TX","NY","FL","VA","NC","AZ")
+stateVec <- stateVec[stateVec != "ND"]
 stateVecOrdered <- stateVec #c("NY","CA","WI","MI","OH") #c("NC","TX","VA","FL","AZ","NY","CA")
@@
 ITHIM.default <- createITHIM()
-ITHIM.default <- update(ITHIM.default, "cvNonTravel", 3)
-ITHIM.default <- update(ITHIM.default, "cv", 1.65)
-ITHIM.default <- update(ITHIM.default, "muNonTravel", 4)
-ITHIM.default <- update(ITHIM.default, "muNonTravelMatrix", METmatrix)
-ITHIM.default <- update(ITHIM.default, "quantiles", (1:99)/100)
 
-ITHIM.national <- ITHIM.default
-ITHIM.national <- update(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- update(ITHIM.national, "muct", muctNational)
-ITHIM.national <- update(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- update(ITHIM.national, "Rct", RctNational)
+ITHIM.default <- update(ITHIM.default, list( cvNonTravel = 3, cv = 1.65, muNonTravel = 4, muNonTravelMatrix = METmatrix, quantiles = (1:99)/100 ) )
+
+ITHIM.national <- update(ITHIM.default, list(muwt = muwtNational, muct = muctNational, Rwt = RwtNational, Rct = RctNational))
 
@@
     ITHIMList[[HHSTATE]] <- ITHIM.default
-    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
-    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "GBD", gbdList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- update(ITHIM.default, list(muwt = muwtList[[HHSTATE]], muct = muctList[[HHSTATE]], Rwt = RwtList[[HHSTATE]], Rct = RwtList[[HHSTATE]], GBD = gbdList[[HHSTATE]]))
   }
@@
   for (HHSTATE in stateVec){
-    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
+    ITHIMList[[HHSTATE]] <- update(ITHIMList[[HHSTATE]], list(muNonTravel = muNonTravel))
   }

### 2016-12-20 R/ITHIM.Rmd
Version 0.2 has almost made it through ITHIM.Rmd and ITHIM2.Rmd error-free

@@
 
-  ITHIM.national <- update(ITHIM.national, "muNonTravel", muNonTravel)
+  ITHIM.national <- update(ITHIM.national, list(muNonTravel = muNonTravel))
 
@@
 p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()
+
+saveRDS(ITHIMList, file = "./data/ITHIMList.by.state.rds")
 
@@
 
-ITHIMListScenario <- lapply(ITHIMList, function(ithim) update(ithim, parName = "muwt", parValue = ithim@parameters@muwt + 10*7))
+ITHIMListScenario <- lapply(ITHIMList, function(ithim) update(ithim, list(muwt = ithim@parameters@muwt + 10*7)))
 
@@
 ```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
-ITHIM.national <- update(ITHIM.national, parName = "muNonTravel", 10)
+ITHIM.national <- update(ITHIM.national, list(muNonTravel = 10))
 
@@
 ITHIM.CA <- ITHIMList[["CA"]]
-ITHIM.CA <- update(ITHIM.CA, parName = "muNonTravel", 10)
+ITHIM.CA <- update(ITHIM.CA, list(muNonTravel = 10))
 foo1 <- compareModels(ITHIM.national,ITHIM.CA)
@@
 ITHIM.NY <- ITHIMList[["NY"]]
-ITHIM.NY <- update(ITHIM.NY, parName = "muNonTravel", 10)
+ITHIM.NY <- update(ITHIM.NY, list(muNonTravel = 10))
 foo2 <- compareModels(ITHIM.national,ITHIM.NY)
@@
 ```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-ITHIM.default <- update(ITHIM.default, "quantiles", (1:99)/100)
+ITHIM.default <- update(ITHIM.default, list(quantiles = (1:99)/100))
 
 ITHIM.national <- ITHIM.default
-ITHIM.national <- update(ITHIM.national, "muwt", muwtNational)
-ITHIM.national <- update(ITHIM.national, "muct", muctNational)
-ITHIM.national <- update(ITHIM.national, "Rwt", RwtNational)
-ITHIM.national <- update(ITHIM.national, "Rct", RctNational)
+ITHIM.national <- update(ITHIM.national, list(muwt = muwtNational, muct = muctNational,Rwt = RwtNational,Rct = RctNational))
 
 ITHIM.scenario <- ITHIM.national
-ITHIM.national.quints <- update(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
-ITHIM.scenario.quints <- update(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)
+ITHIM.national.quints <- update(ITHIM.national, list(quantiles = seq(1,9,by=2)/10))
+ITHIM.scenario.quints <- update(ITHIM.scenario, list(quantiles = seq(1,9,by=2)/10))
 
@@
 
-    ITHIM.scenario <- update(ITHIM.scenario, "muwt", at*pWalk)
-    ITHIM.scenario <- update(ITHIM.scenario, "muct", at*(1-pWalk))
-    ITHIM.scenario <- update(ITHIM.scenario, "muNonTravel", muNonTravel)
+    ITHIM.scenario <- update(ITHIM.scenario, list(muwt = at*pWalk, muct = at*(1-pWalk), muNonTravel = muNonTravel))
 
-    ITHIM.scenario.quints <- update(ITHIM.scenario.quints, "muwt", at*pWalk)
-    ITHIM.scenario.quints <- update(ITHIM.scenario.quints, "muct", at*(1-pWalk))
-    ITHIM.scenario.quints <- update(ITHIM.scenario.quints, "muNonTravel", muNonTravel)
+    ITHIM.scenario.quints <- update(ITHIM.scenario.quints, list(muwt = at*pWalk, muct = at*(1-pWalk), muNonTravel = muNonTravel))
 
-    ITHIM.national <- update(ITHIM.national, "muNonTravel", muNonTravel)
+    ITHIM.national <- update(ITHIM.national, list(muNonTravel = muNonTravel))
 
@@
 
-ITHIM.national <- update(ITHIM.national, "muNonTravel", 10)
+ITHIM.national <- update(ITHIM.national, list(muNonTravel = 10))
 

### 2016-12-20 R/ITHIM.R

@@ -1,3 +1,3 @@
 library("knitr")
 knit2html("ITHIM.Rmd","ITHIM.html")
-#knit2html("ITHIM2.Rmd","ITHIM2.html")
+knit2html("ITHIM2.Rmd","ITHIM2.html")

### 2016-12-21 R/ITHIM.Rmd
Saves ITHIM List as ITHIMList object
Fixed bug in North Dakota GBD

@@
 stateNameToAbbrev <- function( stateName ){
-  stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
+  stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC","North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)
 
@@
 
+ITHIMList <- as("ITHIMList", ITHIMList)
 saveRDS(ITHIMList, file = "./data/ITHIMList.by.state.rds")

### 2017-01-04 R/tutorial.Rmd
Changes to tutorial and website
Update to tutorial and pomodoros

@@
+## ITHIM Tutorial
+
+### Installing the package
+
 ```{r figwidth, eval = TRUE,  warning = FALSE, message = FALSE, echo = FALSE}
@@
 library("devtools")
-install_github(repo = "syounkin/ITHIM")
+install_github(repo = "syounkin/ITHIM", ref = devel)
 library("ITHIM")
 ```
+
+### Built-in Example
+
+```{r example, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+example("ITHIM-package")
+```
+
+### Step-by-step
 

### 2017-01-04 R/tutorial.Rmd
Update to tutorial

--- 2016-10-21 R/tutorial.Rmd
@@
 library("devtools")
-install_github(repo = "syounkin/ITHIM")
+install_github(repo = "syounkin/ITHIM", ref = "devel")
 library("ITHIM")
 ```
+
+### Step-by-step
 
@@
 ITHIM.baseline <- createITHIM()
+```
+
+```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+
 ```

### 2017-01-17 R/tutorial.Rmd
Update to tutorial

--- 2016-10-21 R/tutorial.Rmd
@@
 phi <- (1+sqrt(5))/2 # the golden ratio ;)
+options(repos='http://cran.rstudio.com/')
 ```
@@
 First load the package from Github.  To do this you need the package
-devtools.
+_devtools_.
 
 ```{r, libs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-options(repos='http://cran.rstudio.com/')
 install.packages("devtools")
+library("devtools")
 ```
 
-Now load the package and install ITHIM.
+Now use _install\_github_ to install _ITHIM_ and be sure to load it with _library_.
 
 ```{r libs2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-library("devtools")
-install_github(repo = "syounkin/ITHIM")
+install_github(repo = "syounkin/ITHIM", ref = "devel")
 library("ITHIM")
@@
 
-Now that ITHIM is installed and loaded the first step is to create the
-baseline model.  To do so, using defaults built into the package, use
+### Step-by-step
+
+Now that ITHIM is installed and loaded the first step is to create an
+ITHIM object that represents the baseline vision.
+
+To do so we may create the default ITHIM object using _createITHIM_.
 createITHIM as follows.
@@
 
+An ITHIM object is an R object that conatins, among other things, a
+set of parameters that define the model.  To view these parameters we
+may use _getParameterSet_.
+
+```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+getParameterSet(ITHIM.baseline)
+```
+
 There are currently fourteen parameters, many of which are matrices,
@@
 component.)
-
-```{r parameters, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-names(ITHIM.baseline$parameters)
-```
 
@@
 Now to create the scenario we need only change one parameter.
-Parameters may be changed with updateITHIM.  First we change the mean
+Parameters may be changed with _update_.  First we change the mean
 walking time muwt to 120 minutes.
@@
 ```{r scenario, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-ITHIM.scenario <- updateITHIM(ITHIM.baseline, "muwt", 120)
+ITHIM.scenario <- update(ITHIM.baseline, list(muwt =120))
 ```

### 2017-01-20 R/ITHIM.Rmd
Save the rds ITHIM object for national run
Small tweak to ITHIM.Rmd

@@
 options(repos='http://cran.rstudio.com/')
-#install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
+install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
 library("ggplot2")
@@
 document("~/ITHIM/")
-install("~/ITHIM/")
-#install_github("syounkin/ITHIM", ref="devel")
+#install("~/ITHIM/")
+install_github("syounkin/ITHIM", ref="devel")
 install_github("smbache/magrittr", ref="master")
@@
 
+saveRDS(ITHIM.national, file = "./data/ITHIM.national.rds")
+
 ITHIM.scenario <- ITHIM.national

### 2017-01-20 R/ITHIM.Rmd
Fixed bug in ITHIM.Rmd

@@
 options(repos='http://cran.rstudio.com/')
-install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
+#install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
 library("ggplot2")
@@
 
-ITHIMList <- as("ITHIMList", ITHIMList)
+ITHIMList <- as(ITHIMList, "ITHIMList")
 saveRDS(ITHIMList, file = "./data/ITHIMList.by.state.rds")

### 2017-01-20 R/ITHIM.R

@@ -1,3 +1,3 @@
 library("knitr")
 knit2html("ITHIM.Rmd","ITHIM.html")
-knit2html("ITHIM2.Rmd","ITHIM2.html")
+#knit2html("ITHIM2.Rmd","ITHIM2.html")

### 2017-01-21 R/tutorial.Rmd
Updates to tutorial and pomodoros

--- 2017-01-04 R/tutorial.Rmd
@@
 library("devtools")
-install_github(repo = "syounkin/ITHIM", ref = "devel")
+#install_github(repo = "syounkin/ITHIM", ref = "devel")
+install("~/ITHIM/")
 library("ITHIM")
@@
 
-### Step-by-step
-
-Now that ITHIM is installed and loaded the first step is to create the
-baseline model.  To do so, using defaults built into the package, use
-createITHIM as follows.
+### Create default ITHIM object
 
@@
 
-```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+### ITHIM Methods
 
+#### getMeans
+
+```{r methods3, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+getMeans(ITHIM.baseline)
 ```
 
-There are currently fourteen parameters, many of which are matrices,
-in the model.  (Recall that we are only using the physical activity
-component.)
+#### update
 
-```{r parameters, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-names(ITHIM.baseline$parameters)
+```{r update, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+ITHIM.scenario <- update(ITHIM.baseline, list(muwt = 2*getMeans(ITHIM.baseline)$walk))
 ```
 
-We have means and mean matrices for walk time, cycle time and
-non-travel-related activity and coefficients of variation for active
-transport time and non-travel-related activity.  nAgeClass has always
-been 8 and meanType is always "referent".  GBD is an array of disease
-burden estimates and quantiles is the vector of quantiles used for the
-approximation of the PAF integral.
+#### compareModels
 
-Now to create the scenario we need only change one parameter.
-Parameters may be changed with updateITHIM.  First we change the mean
-walking time muwt to 120 minutes.
-
-```{r scenario, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-ITHIM.scenario <- updateITHIM(ITHIM.baseline, "muwt", 120)
-```
-Now to run the comparative risk assesment use compareModels.
-```{r CRA, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-comparativeRisk <- compareModels(ITHIM.baseline, ITHIM.scenario)
+```{r compareModels, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+CRA.list <- compareModels(ITHIM.baseline,ITHIM.scenario)
 ```
 
-```{r AF, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-comparativeRisk$AF$ColonCancer
+#### Plot
+
+```{r methods2, eval = FALSE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+plot(ITHIM.baseline)
 ```
+
+#### getMethodsITHIM
+
+```{r getMethods, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+getMethodsITHIM()
+```
+
+### The _ParameterSet_ Object
+
+Each ITHIM object contains a ParameterSet object.
+
+```{r pSet, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+paramSet.baseline <- getParameterSet(ITHIM.baseline)
+paramSet.scenario <- getParameterSet(ITHIM.scenario)
+```
+
+```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+paramSet.baseline
+```
+
+```{r class2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+paramSet.scenario
+```

### 2017-01-23 R/tutorial.Rmd

@@
 First load the package from Github.  To do this you need the package
-devtools.
+_devtools_.
 
 ```{r, libs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-options(repos='http://cran.rstudio.com/')
 install.packages("devtools")
+library("devtools")
 ```
 
-Now load the package and install ITHIM.
+Now use _install\_github_ to install _ITHIM_ and be sure to load it with _library_.
 
@@
 library("devtools")
-#install_github(repo = "syounkin/ITHIM", ref = "devel")
-install("~/ITHIM/")
+install_github(repo = "syounkin/ITHIM", ref = "master")
 library("ITHIM")
@@
 ### Create default ITHIM object
+=======
+Now that ITHIM is installed and loaded the first step is to create an
+ITHIM object that represents the baseline vision.
+
+To do so we may create the default ITHIM object using _createITHIM_.
+createITHIM as follows.
 
@@
 paramSet.scenario <- getParameterSet(ITHIM.scenario)
+=======
+An ITHIM object is an R object that conatins, among other things, a
+set of parameters that define the model.  To view these parameters we
+may use _getParameterSet_.
+
+```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+getParameterSet(ITHIM.baseline)
+```
+
+There are currently fourteen parameters, many of which are matrices,
+in the model.  (Recall that we are only using the physical activity
+component.)
+
+We have means and mean matrices for walk time, cycle time and
+non-travel-related activity and coefficients of variation for active
+transport time and non-travel-related activity.  nAgeClass has always
+been 8 and meanType is always "referent".  GBD is an array of disease
+burden estimates and quantiles is the vector of quantiles used for the
+approximation of the PAF integral.
+
+Now to create the scenario we need only change one parameter.
+Parameters may be changed with _update_.  First we change the mean
+walking time muwt to 120 minutes.
+
+```{r scenario, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+ITHIM.scenario <- update(ITHIM.baseline, list(muwt =120))
+>>>>>>> a14bdd7f9847c86f744bf70f3ffe8d2643d2ffb5
 ```

### 2017-01-25 R/ITHIM.Rmd
Created NHTS sample size figure

@@
 muctList <- lapply( meanCycleList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
+```
+
+```{r sampleSize, eval = TRUE, dev=c('pdf', 'png'), fig.width = 3*fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+foo <- lapply( countMatrix, function(x) c(x) )
+foo <- melt(foo)
+names(foo) <- c("sampleSize", "state")
+p <- ggplot(foo, aes( x = state, y = sampleSize))
+p + geom_bar(stat = "identity") + scale_y_log10() + labs(y = "Sample Size (Referent Group)", x = "State") + theme_bw()
 ```

### 2017-02-06 R/ITHIM.Rmd
Update to cambridge.Rmd

@@
 #install("~/ITHIM/")
-install_github("syounkin/ITHIM", ref="devel")
+install_github("syounkin/ITHIM", ref="cambridge")
 install_github("smbache/magrittr", ref="master")

### 2017-02-06 R/cambridge.Rmd
Update to cambridge.Rmd

# Integrated Transport and Health Impacts Model

## R Package _ITHIM_
<!-- This document will eventually serve as a compendium to the R package
_ITHIM_.  It will describe how to use the package to implement the
ITHIM model, as well as a detailed description of the model itself.

As a working document it will serve as a place for SGY to post
results, questions, interpreations, notes, etc.-->

```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
library("devtools")
#install_github("syounkin/ITHIM", ref="cambridge")
install("~/ITHIM/")
document("~/ITHIM/")
install("~/ITHIM/")
library("ITHIM")
```

```{r misc, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

```{r roadInjuries, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
scenario <- readRDS(file = "~/Downloads/distByRoadTypeScenario4.rds")
base <- readRDS(file = "~/Downloads/distByRoadTypeBaseline.rds")

ITHIM <- createITHIM()
ITHIM.baseline <- update(ITHIM, list(distRoadType = base))
ITHIM.scenario <- update(ITHIM, list(distRoadType = scenario))
ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
```

```{r roadInjuries9, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
ITHIM.baseline <- update(ITHIM.baseline, list(GBD = readGBD(file = "gbdEngland.csv")))
ITHIM.scenario <- update(ITHIM.scenario, list(GBD = readGBD(file = "gbdEngland.csv")))
```

```{r roadInjuries3, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
```

```{r roadInjuries4, message = FALSE, , dev=c('png'), eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}

deltaDeathsVec <- c()

for( sin in sinVec <- seq(0,1,length.out = 25) ){
    ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = sin*matrix(1,nrow = 8, ncol = 2)))
    ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
    RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
    deltaDeathsVec <- c(deltaDeathsVec, deltaDeaths)
}

plot(sinVec, deltaDeathsVec, type = "p", pch = 20)

```
```{r roadInjuries5, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
    ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = 0.5*matrix(1,nrow = 8, ncol = 2)))
deltaDeathsVec <- c()
for( underReporting in underReportingVec <- runif(10,0,1) ){
    cat(underReporting)
    ITHIM.baseline <- update(ITHIM.baseline, list(roadInjuries = lapply(getRoadInjuries(ITHIM.baseline), function(x) x/underReporting)))
    ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
    RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
    deltaDeathsVec <- c(deltaDeathsVec, deltaDeaths)
}
plot(underReportingVec, deltaDeathsVec, type = "p", pch = 20)
```

```{r roadInjuries6, message = FALSE, , dev=c('png'), eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}

for(i in 1:10){

    simRI <- lapply(getRoadInjuries(ITHIM.baseline), function(x){
      lambda <- c(as.matrix(x))
      simCount <- matrix(mapply(function(y){rpois(1,y)},y=lambda), nrow = nrow(x), ncol = ncol(x))
      dimnames(simCount) <- dimnames(x)
      simCount
    })

ITHIM.baseline.sim <- update(ITHIM.baseline, list(roadInjuries = simRI))
    ITHIM.scenario.sim <- updateRoadInjuries(ITHIM.baseline.sim, ITHIM.scenario)
    RIburden <- computeRoadInjuryBurden(ITHIM.baseline.sim, ITHIM.scenario.sim)
    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
cat(c(deltaDeaths, "\n"))

}

```

### 2017-02-06 R/cambridge.Rmd
Update to cambridge.Rmd

@@
 
-```{r roadInjuries4, message = FALSE, , dev=c('png'), eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-
-deltaDeathsVec <- c()
-
-for( sin in sinVec <- seq(0,1,length.out = 25) ){
-    ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = sin*matrix(1,nrow = 8, ncol = 2)))
-    ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
-    RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
-    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
-    deltaDeathsVec <- c(deltaDeathsVec, deltaDeaths)
-}
-
-plot(sinVec, deltaDeathsVec, type = "p", pch = 20)
-
-```
-```{r roadInjuries5, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
-    ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = 0.5*matrix(1,nrow = 8, ncol = 2)))
-deltaDeathsVec <- c()
-for( underReporting in underReportingVec <- runif(10,0,1) ){
-    cat(underReporting)
-    ITHIM.baseline <- update(ITHIM.baseline, list(roadInjuries = lapply(getRoadInjuries(ITHIM.baseline), function(x) x/underReporting)))
-    ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
-    RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
-    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
-    deltaDeathsVec <- c(deltaDeathsVec, deltaDeaths)
-}
-plot(underReportingVec, deltaDeathsVec, type = "p", pch = 20)
-```
-
 ```{r roadInjuries6, message = FALSE, , dev=c('png'), eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-
-for(i in 1:10){
-
 simRI <- lapply(getRoadInjuries(ITHIM.baseline), function(x){
@@
   dimnames(simCount) <- dimnames(x)
-      simCount
+  retrun(simCount)
 })
@@
 cat(c(deltaDeaths, "\n"))
-
-}
-
 ```

### 2017-02-06 R/cambridge.Rmd
Update to cambridge.Rmd

@@
 deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
-cat(c(deltaDeaths, "\n"))
+
 ```

### 2017-02-06 R/cambridge.Rmd
Update to cambridge.Rmd

@@
   dimnames(simCount) <- dimnames(x)
-  retrun(simCount)
+  return(simCount)
 })

### 2017-02-13 R/ICT.Rmd
Cambridge trip files

```{r ICT1, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
library("devtools")
install("~/scenic/")
library("scenic")

prepare_baseline()

```

```{r ICT2, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
# 'inst/extdata/pcycl_baseline2.csv')
pcycl_baseline <- read.csv(system.file("extdata", "pcycl_baseline2.csv", package = "scenic"), header = T)  #cycling probs file
pcycl_baseline <- pcycl_baseline[,-1]
#
# #reads scenarios constants
METcycling <- 6.44
METwalking <- 4.61
METebikes <- 4.50

#Read cycling speeds by age/sex
cyclingspeed <- read.csv(system.file("extdata", "cyclingSpeeds.csv", package = "scenic"),header=T)

# Read Cycling Probabilities into an R Object
probCycling <- read.csv(system.file("extdata", "cycling-probability.csv", package = "scenic"), header = T, as.is = T)

# Convert all probabilities into odss
oddsCycling <- round(probCycling /( 1 - probCycling), 3)

# Convert data.frame into a list
oddsCycling <- unlist(oddsCycling, use.names = F)

#lookup tables for equity=0,1. MUST BE expressed in ODDS!
Pcyc0.eq0 <- oddsCycling[1:4]
Pcyc0.eq1 <- rep(oddsCycling[5], 4)

# Baseline=NTS years 2011-2014 + individuals between 18-84 y.o + not Wales/Scotland
#bl <- readRDS(system.file("extdata", "bl2014_v2.rds", package = "scenic"))

bl <- read.csv(system.file("extdata", "sample_baseline.csv", package = "scenic"), header = T)

# bl = subset(bl, subset = Age_B01ID < 21 & HHoldGOR_B02ID!=10  & HHoldGOR_B02ID!=11)
# bl$Age[bl$Age_B01ID<16] <- '16.59'
# bl$Age[bl$Age_B01ID>=16] <- '60plus'
# bl$Sex[bl$Sex_B01ID==1] <- 'Male'
# bl$Sex[bl$Sex_B01ID==2] <- 'Female'
#
# #bl <- read.csv('bl2012_18_84ag_reduced.csv', header=T, as.is = T)
#
# #IMPORTANT: from database -> ID needs to be deleted, IndividualID renamed to ID.
#
# baseline <- bl
#
# #handle short walks, creating 6x of each
# df <- baseline[baseline$MainMode_B03ID==1,]
# shortwalks <- data.frame()
#
# for (i in 1:6) {
#   shortwalks <- rbind(shortwalks,df)
# }
#
# # Update trip IDs with new IDs
# #shortwalks$TripID <-  c(max(baseline$TripID) + 1:nrow(shortwalks))
#
# baseline <- rbind(baseline,shortwalks)
# baseline <- baseline[order(baseline$ID),]
#
# #add people w/o trips to baseline
# fnotrips  <- readRDS(system.file("extdata", "people_notrips2014_v2.Rds", package = "scenic"))
# #fnotrips  <- read.csv('People_w_NoTrips2012_ENG_v6_anon.csv',header=T,as.is=T)
#
# # Remove 85+ age group + Wales/Scotland
# fnotrips <- subset(fnotrips, subset = Age_B01ID < 21 & HHoldGOR_B02ID!=10 & HHoldGOR_B02ID!=11)
# #fnotrips <- subset(fnotrips, subset = Age_B01ID != 21)
#
# fnotrips$agesex <- ""
#
# fnotrips$Age <- ""
#
# fnotrips$Sex <- ""
#
# fnotrips[(fnotrips$Age_B01ID >= 6 & fnotrips$Age_B01ID <= 15) & fnotrips$Sex_B01ID == 1,]$agesex <- '16.59Male'
#
# fnotrips[(fnotrips$Age_B01ID >= 6 & fnotrips$Age_B01ID <= 15) & fnotrips$Sex_B01ID == 2,]$agesex <- '16.59Female'
#
# fnotrips[fnotrips$Age_B01ID >= 16 & fnotrips$Sex_B01ID == 1,]$agesex <- '60plusMale'
#
# fnotrips[fnotrips$Age_B01ID >= 16 & fnotrips$Sex_B01ID == 2,]$agesex <- '60plusFemale'
#
# fnotrips$Age <- 0
#
# fnotrips[fnotrips$Sex_B01ID == 1,]$Sex <- "Male"
#
# fnotrips[fnotrips$Sex_B01ID == 2,]$Sex <- "Female"
#
#
# fnotrips[(fnotrips$Age_B01ID >= 6 & fnotrips$Age_B01ID <= 15),]$Age <- '16.59'
#
# fnotrips[fnotrips$Age_B01ID >= 16,]$Age <- '60plus'
#
# # Add tripID variable to it
# fnotrips$TripID <- c(max(baseline$TripID) + 1:nrow(fnotrips))
#
# rm(shortwalks, df)
#
# #Sample before running scenarios -
#
# # Removed NAs from the data.frame
# hsematch <- readRDS(system.file("extdata", "hse-nts_match_v2.rds", package = "scenic"))
# #names(hsematch)[c(1,7,8)] <- c('ID', 'health_mmets', 'physical_activity_mmets')
# #hsematch = hsematch[, c(1,7,8)]
#
# #next line commented as people w/o trips already in hse
# #hsematch <- rbind(hsematch, subset(fnotrips, select = c(ID, health_mmets, physical_activity_mmets)))
#
# # Remove health_mmets and physical_activity_mmets from fnotrips
# fnotrips$health_mmets <- NULL
# fnotrips$physical_activity_mmets <- NULL
#
# baseline <- rbind.fill(baseline, fnotrips)
#
# # Read nts age group lookup table
# ag_lookup <- read.csv(system.file("extdata", "nts-adjusted-age-groups.csv", package = "scenic"), header = T, as.is = T)
#
# # Create a new variable 'age_group' for baseline, which converts numeric age categories into age ranges
# baseline$age_group <- ag_lookup$age[match(baseline$Age_B01ID, ag_lookup$nts_group)]
# rm(ag_lookup)
#
#
# baseline[is.na(baseline)] <- 0
# baseline$TripID  <- as.numeric(factor(baseline$TripID))
#
#
# #hsematch <- hsematch[,c(8,9)]  #keep only first and last column > IndivID, mMETs
# hse1 <- setDT(hsematch)[,if(.N<1) .SD else .SD[sample(.N,1,replace=F)],by=ID]
# hse1 = as.data.frame(hse1)
#
# #add times cols. to baseline (used for travel time after cycle switch takes place)
# TripTotalTime1 <- 0
# TripTravelTime1 <- 0
# baseline <- cbind(baseline,TripTotalTime1,TripTravelTime1)
# baseline$TripTotalTime1 <- baseline$TripTotalTime
# baseline$TripTravelTime1 <-baseline$TripTravelTime
#
# #add mmets column to baseline (& save for having total mmets)
# baseline <-inner_join(baseline,hse1,by='ID')
# rm(hse1, hsematch)
#
# #randcycle (used later to calculate if people are cyclists), add col. [prob]
# randcycle <- runif(length(unique(baseline$ID)))
# randcycle <- data.frame(ID=unique(baseline$ID),prob=randcycle)
# baseline <- inner_join(baseline,randcycle,by='ID')
# rm(randcycle)
#
# baseline$prob[baseline$TravDay==0 ] <- 0
#
#
# #keep bl as backup for future scenarios core values
# bl <- baseline
#
# #save PROCESSED baseline in main folder
# #saveRDS(bl,file='bl2014_p_v2.rds')
# rm(bl)
#
# #write.csv(bl,file='bl2012_18_84ag_sw_reduced.csv', row.names=F)
#
#
# ###################################  START CALCULATIONS on BASELINE #############################
#
# #nos. for baseline
# carMiles0 <- sum(baseline[baseline$MainMode_B04ID %in% c(3,4,5,12),'TripDisIncSW'])
# carMiles0 <- round(carMiles0,1)
# METh0 <- round(sum(baseline$METh),1)
# MMETh0 <- round(sum(baseline$MMETh),1)
# # Miles to Kilometres, Grams to metric tonnes, 0.0001
# # Using new Christian's average CO2 value of 0.31 grams
# CO20 <- round(carMiles0 * 1.61 * (3.1 / 1.61) * 1e-1,2)   #(in Kg)
#
# # df <- data.frame()
#
# #i <- c(2,4,8,16,32,64)
#
# # DF which main role is to store info of every case in which Observed > DP. Used in UI to filter out mentioned cases (used in 1_flowgram2-2012)
#
# directProbCasesAboveGivenPerc <- data.frame(MS = numeric(0),
#                                             ebikes = numeric(0),
#                                             equity = numeric(0),
#                                             region = numeric(0))
#
# # directProbs values
#
# directProbs <- c(0.05, 0.1, 0.25, 0.5, 0.75, 1)
#
# # Removing TDR
# # TDR
# # j  <- c(1,0.9,0.8,0.7)
#
# m <- c(0,1)   #ebikes
# n <- c(0,1)   #equity
# num = 1
#
# #keep only used variables
# baseline <-baseline[ , c('Age_B01ID', 'Sex_B01ID', 'HHoldGOR_B02ID', 'CarAccess_B01ID',
#                          'NSSec_B03ID', 'IndIncome2002_B02ID', 'EthGroupTS_B02ID', 'TripID', 'ID',
#                          'MainMode_B03ID', 'MainMode_B04ID', 'MainMode_B11ID',
#                          'TripTotalTime', 'TripTravTime', 'TripDisIncSW',
#                          'agesex', 'Cycled', 'METh', 'MMETh', 'cyclist',
#                          'Age', 'Sex', 'age_group',
#                          'TripTotalTime1', 'TripTravelTime1',
#                          'health_mmets', 'physical_activity_mmets', 'prob')]
#
#
# listOfScenarios <- list()
#
# # for (ebikes in m) {
# #   for (equity in n) {
# #     for  (MS in directProbs) { # all occurences of MS should be replaced
# #       cat(ebikes, equity, MS, "\n")
# #       scenario_name <- paste("MS",MS,"_ebik",ebikes,"_eq" ,equity,sep="")
# #       #assign(scenario_name,flowgram(baseline, MS,ebikes,equity, pcycl_baseline))
# #       tempSc <- flow_gram(baseline, MS,ebikes,equity, pcycl_baseline)
# #       #saveRDS(tempSc, paste0('./temp_data_folder/output/repo_version/', scenario_name, '.rds'))
# #
# #       listOfScenarios[[num]] <- scenario_name
# #       num <- num + 1
# #     }
# #   }
# # }  #j-i-m-n loop
#
# cat('All done! \n')
```

### 2017-02-13 R/cambridge.Rmd
Cambridge trip files

@@
 
-## R Package _ITHIM_
+## ITHIM Package
 <!-- This document will eventually serve as a compendium to the R package
@@
 library("devtools")
-#install_github("syounkin/ITHIM", ref="cambridge")
-install("~/ITHIM/")
-document("~/ITHIM/")
-install("~/ITHIM/")
+install_github("syounkin/ITHIM", ref="devel")
+#install("~/ITHIM/")
+#document("~/ITHIM/")
+#install("~/ITHIM/")
 library("ITHIM")
@@
 
+### Create the ITHIM object
+
+#### Method 1: Using Input Files
+
+We begin by creating the ITHIM object from which both the baseline and
+scenarios are derived.  The necessary input parameters are contained
+in 3 files, listed below.
+
+```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+
+fileNamesList <- list(
+    activeTransportTimeFile = "~/GHI/misc/activeTransportTime.csv",
+    roadInjuriesFile = "~/GHI/misc/stata19.csv",
+#    roadInjuriesFile = "~/GHI/misc/roadInjuries.csv",
+    GBDFile = "~/GHI/misc/gbdEngland_20170209.csv"
+#    GBDFile = "~/GHI/misc/gbd.csv"
+)
+
+ITHIM <- createITHIM(fileNamesList)
+#ITHIM <- createITHIM()
+```
+
+The file activeTtransportTime.csv contains means (absolute) for
+walking and cycling times (min/week), stratified by age and sex.
+Links to sample files can be found below.
+
+http://ithim.ghi.wisc.edu/activeTransportTime.csv
+http://ithim.ghi.wisc.edu/roadInjuries.csv
+http://ithim.ghi.wisc.edu/gbd.csv
+
+Let's verify that our parameters have been set properly.
+
+```{r misc14, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+getWalkTime(ITHIM)
+getCycleTime(ITHIM)
+getRoadInjuries(ITHIM)
+head(getGBD(ITHIM))
+```
+
+## Archive
+
+The ITHIM class, which defines an ITHIM object, contains three slots.
+Two of the slots are used for the physical activity component and the third is simply a container for all of the parameters in the model.
+This container is the ParameterSet class.
+
+```{r roadInjuries11, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+getParameterSet(ITHIM)
+```
+
+## Road Injury Model
+We inspect the road injury counts that are used by default.
+```{r misc4, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+RI <- getRoadInjuries(ITHIM)
+class(RI)
+names(RI)
+RI$FatalLocal
+```
+
+We do not yet have default values for the distance by road type
+matrices.
+
+```{r misc3, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+getDistRoadType(ITHIM)
+```
+
+So, we have to update manually.
+
 ```{r roadInjuries, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
@@
 base <- readRDS(file = "~/Downloads/distByRoadTypeBaseline.rds")
-
-ITHIM <- createITHIM()
 ITHIM.baseline <- update(ITHIM, list(distRoadType = base))
 ITHIM.scenario <- update(ITHIM, list(distRoadType = scenario))
+getDistRoadType(ITHIM.baseline)
+getDistRoadType(ITHIM.scenario)
+```
+
+Now that we have the distances by road type we may compute the
+scenario injury multipliers and compute the scenario road injury
+estimates.
+
+```{r roadInjuries0, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
 ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
@@
 
+The final ingredient is the GBD set of parameters.  We have a file
+stored within the ITHIM package from which we may read in these
+values.  We update the ITHIM objects as follows.
+
 ```{r roadInjuries9, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline <- update(ITHIM.baseline, list(GBD = readGBD(file = "gbdEngland.csv")))
-ITHIM.scenario <- update(ITHIM.scenario, list(GBD = readGBD(file = "gbdEngland.csv")))
+#ITHIM.baseline <- update(ITHIM.baseline, list(GBD = readGBD(file = "~/GHI/misc/gbdEngland.csv")))
+#ITHIM.scenario <- update(ITHIM.scenario, list(GBD = readGBD(file = "~/GHI/misc/gbdEngland.csv")))
+```
+
+```{r roadInjuries10, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+getParameterSet(ITHIM.baseline)
 ```
@@
 RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
-deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
+(deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta))
 ```
 
-```{r roadInjuries6, message = FALSE, , dev=c('png'), eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+## Safety in Numbers
+
+The point of all of this class and method business is to make the code for analyses tidy and human-readable.  Here is an example in which we vary the safety in numbers parameters.
+
+```{r roadInjuries4, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
+
+deltaDeathsVec <- c()
+
+for( sin in sinVec <- seq(0,1,length.out = 10) ){
+    ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = sin*matrix(1,nrow = 8, ncol = 2)))
+    ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
+    RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
+    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta) # Not human-readable enough, should improve
+    deltaDeathsVec <- c(deltaDeathsVec, deltaDeaths)
+}
+
+plot(sinVec, deltaDeathsVec, type = "b", pch = 20)
+
+```
+
+## Uncertainty in Road Injury Estimates
+
+```{r roadInjuries6, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
+
+deltaDeaths.vec <- c()
+for( i in 1:50 ){
+
   simRI <- lapply(getRoadInjuries(ITHIM.baseline), function(x){
@@
     dimnames(simCount) <- dimnames(x)
-  return(simCount)
+    return(as.data.frame(simCount))
   })
-
   ITHIM.baseline.sim <- update(ITHIM.baseline, list(roadInjuries = simRI))
   ITHIM.scenario.sim <- updateRoadInjuries(ITHIM.baseline.sim, ITHIM.scenario)
+
   RIburden <- computeRoadInjuryBurden(ITHIM.baseline.sim, ITHIM.scenario.sim)
   deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
-
+  deltaDeaths.vec <- c(deltaDeaths.vec, deltaDeaths)
+}
+hist(deltaDeaths.vec)
 ```

### 2017-02-13 R/testRI.R

@@ -1,0 +1,21 @@
+rm(list=ls())
+library("devtools")
+install_github("syounkin/ITHIM", ref="devel")
+library("ITHIM")
+fileNamesList <- list(
+    activeTransportTimeFile = "~/GHI/misc/activeTransportTime.csv",
+#    roadInjuriesFile = "~/GHI/misc/stata19.csv",
+    roadInjuriesFile = "~/GHI/misc/roadInjuries.csv",
+    GBDFile = "~/GHI/misc/gbdEngland.csv"
+#    GBDFile = "~/GHI/misc/gbd.csv"
+)
+
+ITHIM <- createITHIM(fileNamesList)
+#ITHIM <- createITHIM()
+scenario <- readRDS(file = "~/Downloads/distByRoadTypeScenario4.rds")
+base <- readRDS(file = "~/Downloads/distByRoadTypeBaseline.rds")
+ITHIM.baseline <- update(ITHIM, list(distRoadType = base))
+ITHIM.scenario <- update(ITHIM, list(distRoadType = base))
+ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
+RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
+(deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta))

### 2017-02-14 R/cambridge.Rmd
Testing original csv files

@@
     activeTransportTimeFile = "~/GHI/misc/activeTransportTime.csv",
-    roadInjuriesFile = "~/GHI/misc/stata19.csv",
-#    roadInjuriesFile = "~/GHI/misc/roadInjuries.csv",
-    GBDFile = "~/GHI/misc/gbdEngland_20170209.csv"
-#    GBDFile = "~/GHI/misc/gbd.csv"
+#    roadInjuriesFile = "~/GHI/misc/stata19.csv",
+    roadInjuriesFile = "~/GHI/misc/roadInjuries.csv",
+#    GBDFile = "~/GHI/misc/gbdEngland_20170209.csv"
+    GBDFile = "~/GHI/misc/gbd.csv"
 )

### 2017-02-14 R/cambridge.Rmd
Updated .gitignore and cambridge.Rmd

@@
 library("devtools")
-install_github("syounkin/ITHIM", ref="devel")
-#install("~/ITHIM/")
-#document("~/ITHIM/")
-#install("~/ITHIM/")
+#install_github("syounkin/ITHIM", ref="cambridge")
+install("~/ITHIM/")
+document("~/ITHIM/")
+install("~/ITHIM/")
 library("ITHIM")

### 2017-02-14 R/ProcessCountyBurden.R
Added ProcessCountyBurden to Makefile
Vargo's markdown script can now be run with make ProcessCountyBurden in the top-level.

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("ProcessCountyBurden.md","ProcessCountyBurden.html")

### 2017-02-15 R/acs_tract_age_and_sex.r
Added acs_tract_age_and_sex.r to repository
Vargo wrote this script to estimate F, the population density, for each census tract.

@@ -1,0 +1,161 @@
+library(jsonlite)
+library(reshape2)
+
+getTractAgeSex <- function(state, county) {
+  varString <-
+    "B01001_003E,B01001_004E,B01001_005E,B01001_006E,B01001_007E,B01001_008E,B01001_009E,B01001_010E,B01001_011E,B01001_012E,B01001_013E,B01001_014E,B01001_015E,B01001_016E,B01001_017E,B01001_018E,B01001_019E,B01001_020E,B01001_021E,B01001_022E,B01001_023E,B01001_024E,B01001_025E,B01001_027E,B01001_028E,B01001_029E,B01001_030E,B01001_031E,B01001_032E,B01001_033E,B01001_034E,B01001_035E,B01001_036E,B01001_037E,B01001_038E,B01001_039E,B01001_040E,B01001_041E,B01001_042E,B01001_043E,B01001_044E,B01001_045E,B01001_046E,B01001_047E,B01001_048E,B01001_049E"
+  
+  ACSpop <-
+    as.data.frame(fromJSON(
+      paste(
+        "http://api.census.gov/data/2015/acs5?get=NAME,",
+        varString,
+        "&for=tract:*&in=state:",
+        state,
+        "+county:",
+        county,
+        "&key=f78d6b6c18608edc379b5a06c55407ceb45e7038",
+        sep = ""
+      )
+    ))
+  ACSpop <- ACSpop[-1,]
+  colnames(ACSpop) <-
+    c(
+      "name",
+      "M_Under 5 years",
+      "M_5 to 9 years",
+      "M_10 to 14 years",
+      "M_15 to 17 years",
+      "M_18 and 19 years",
+      "M_20 years",
+      "M_21 years",
+      "M_22 to 24 years",
+      "M_25 to 29 years",
+      "M_30 to 34 years",
+      "M_35 to 39 years",
+      "M_40 to 44 years",
+      "M_45 to 49 years",
+      "M_50 to 54 years",
+      "M_55 to 59 years",
+      "M_60 and 61 years",
+      "M_62 to 64 years",
+      "M_65 and 66 years",
+      "M_67 to 69 years",
+      "M_70 to 74 years",
+      "M_75 to 79 years",
+      "M_80 to 84 years",
+      "M_85 years and over",
+      "F_Under 5 years",
+      "F_5 to 9 years",
+      "F_10 to 14 years",
+      "F_15 to 17 years",
+      "F_18 and 19 years",
+      "F_20 years",
+      "F_21 years",
[…]

### 2017-02-15 R/acs_tract_age_and_sex.r
Commented out line that writes to Box Sync folder

@@ -157,5 +157,5 @@ DallasWONDERpop <- ddply(DallasCnty, .(state, county, tract, gender, WONDERage),
                                       
                          )
 
-write.csv(DallasWONDERpop, "./Box Sync/work/Louisville/Dallas/dallasWonderPop.csv", row.names=F)
+#write.csv(DallasWONDERpop, "./Box Sync/work/Louisville/Dallas/dallasWonderPop.csv", row.names=F)
 

### 2017-02-15 R/ITHIM.Rmd
sumDALY -> deltaBurden

--- 2017-01-25 R/ITHIM.Rmd
@@
 document("~/ITHIM/")
-#install("~/ITHIM/")
-install_github("syounkin/ITHIM", ref="devel")
+install("~/ITHIM/")
+#install_github("syounkin/ITHIM", ref="cambridge")
 install_github("smbache/magrittr", ref="master")
@@
 
-  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
+  DALYs <- lapply(ITHIMList, deltaBurden, ITHIM.baseline = ITHIM.national)
   DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
@@
 
-deltaDALYList <- mapply(sumDALY, ITHIMList, ITHIMListScenario)
+deltaDALYList <- mapply(deltaBurden, ITHIMList, ITHIMListScenario)
 
@@
 
-    DALYs <- sumDALY(ITHIM.national, ITHIM.scenario)
-    DALYs.quints <- sumDALY(ITHIM.national.quints, ITHIM.scenario.quints)
+    DALYs <- deltaBurden(ITHIM.national, ITHIM.scenario)
+    DALYs.quints <- deltaBurden(ITHIM.national.quints, ITHIM.scenario.quints)
 

### 2017-02-17 R/ITHIM.Rmd
Updated tutorial.Rmd with content from cambridge.Rmd

@@
 ```{r compute, echo = FALSE, eval = TRUE}
+date()
 compute <- FALSE
@@
 
-## Figure 2
-
-```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r displayMeans, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 
-D <- rbind(getWalkTime(ITHIM.national),getWalkSpeed(ITHIM.national),getCycleTime(ITHIM.national),getNonTravelMETs(ITHIM.national))
+D <- rbind(getWalkTime(ITHIM.national),getCycleTime(ITHIM.national),getNonTravelMETs(ITHIM.national))
 D <- within(D, sex <- factor(sex, levels = c("F","M")))
@@
 ```
+
+```{r sessionInfo, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+sessionInfo()
+```

### 2017-02-17 R/cambridge-archive.Rmd
Updated tutorial.Rmd with content from cambridge.Rmd

--- 2017-02-14 R/cambridge.Rmd
@@
-# Integrated Transport and Health Impacts Model
-
-## ITHIM Package
-<!-- This document will eventually serve as a compendium to the R package
-_ITHIM_.  It will describe how to use the package to implement the
-ITHIM model, as well as a detailed description of the model itself.
-
-As a working document it will serve as a place for SGY to post
-results, questions, interpreations, notes, etc.-->
-
-```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
-rm(list=ls())
-library("devtools")
-install_github("syounkin/ITHIM", ref="devel")
-#install("~/ITHIM/")
-#document("~/ITHIM/")
-#install("~/ITHIM/")
-library("ITHIM")
-```
-
-```{r misc, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-fig.width <- 6 # inches
-phi <- (1+sqrt(5))/2 # the golden ratio ;)
-```
-
-### Create the ITHIM object
-
-#### Method 1: Using Input Files
-
-We begin by creating the ITHIM object from which both the baseline and
-scenarios are derived.  The necessary input parameters are contained
-in 3 files, listed below.
-
-```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-
-fileNamesList <- list(
-    activeTransportTimeFile = "~/GHI/misc/activeTransportTime.csv",
-#    roadInjuriesFile = "~/GHI/misc/stata19.csv",
-    roadInjuriesFile = "~/GHI/misc/roadInjuries.csv",
-#    GBDFile = "~/GHI/misc/gbdEngland_20170209.csv"
-    GBDFile = "~/GHI/misc/gbd.csv"
-)
-
-ITHIM <- createITHIM(fileNamesList)
-#ITHIM <- createITHIM()
-```
-
-The file activeTtransportTime.csv contains means (absolute) for
-walking and cycling times (min/week), stratified by age and sex.
-Links to sample files can be found below.
-
-http://ithim.ghi.wisc.edu/activeTransportTime.csv
-http://ithim.ghi.wisc.edu/roadInjuries.csv
-http://ithim.ghi.wisc.edu/gbd.csv
-
 Let's verify that our parameters have been set properly.

### 2017-02-17 R/cambridge.Rmd
Updated tutorial.Rmd with content from cambridge.Rmd

--- 2017-02-14 R/cambridge.Rmd
@@
-# Integrated Transport and Health Impacts Model
+## ITHIM Package
 
-## ITHIM Package
-<!-- This document will eventually serve as a compendium to the R package
-_ITHIM_.  It will describe how to use the package to implement the
-ITHIM model, as well as a detailed description of the model itself.
-
-As a working document it will serve as a place for SGY to post
-results, questions, interpreations, notes, etc.-->
+### Install and Load ITHIM
 
@@
 install_github("syounkin/ITHIM", ref="devel")
-#install("~/ITHIM/")
-#document("~/ITHIM/")
-#install("~/ITHIM/")
 library("ITHIM")
@@
 ```{r misc, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
+#install("~/ITHIM/")
+#document("~/ITHIM/")
+#install("~/ITHIM/")
 fig.width <- 6 # inches
@@
 ```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-
 fileNamesList <- list(
-    activeTransportTimeFile = "~/GHI/misc/activeTransportTime.csv",
-#    roadInjuriesFile = "~/GHI/misc/stata19.csv",
-    roadInjuriesFile = "~/GHI/misc/roadInjuries.csv",
-#    GBDFile = "~/GHI/misc/gbdEngland_20170209.csv"
-    GBDFile = "~/GHI/misc/gbd.csv"
+    activeTransportTimeFile = "~/ITHIM/inst/activeTransportTime.csv",
+    roadInjuriesFile = "~/ITHIM/inst/roadInjuries.csv",
+    GBDFile = "~/ITHIM/inst/gbd.csv"
 )
@@
 ITHIM <- createITHIM(fileNamesList)
-#ITHIM <- createITHIM()
 ```
 
-The file activeTtransportTime.csv contains means (absolute) for
-walking and cycling times (min/week), stratified by age and sex.
 Links to sample files can be found below.
 
-http://ithim.ghi.wisc.edu/activeTransportTime.csv
-http://ithim.ghi.wisc.edu/roadInjuries.csv
-http://ithim.ghi.wisc.edu/gbd.csv
+* [activeTransportTime.csv](http://ithim.ghi.wisc.edu/activeTransportTime.csv)
+* [roadInjuries.csv](http://ithim.ghi.wisc.edu/roadInjuries.csv)
+* [gbd.csv](http://ithim.ghi.wisc.edu/gbd.csv)
 
-Let's verify that our parameters have been set properly.
+These files can also be found in the devel branch of ITHIM repository.
 
-```{r misc14, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-getWalkTime(ITHIM)
-getCycleTime(ITHIM)
-getRoadInjuries(ITHIM)
-head(getGBD(ITHIM))
-```
-
-## Archive
-
-The ITHIM class, which defines an ITHIM object, contains three slots.
-Two of the slots are used for the physical activity component and the third is simply a container for all of the parameters in the model.
-This container is the ParameterSet class.
-
-```{r roadInjuries11, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-getParameterSet(ITHIM)
-```
-
-## Road Injury Model
-We inspect the road injury counts that are used by default.
-```{r misc4, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-RI <- getRoadInjuries(ITHIM)
-class(RI)
-names(RI)
-RI$FatalLocal
-```
-
-We do not yet have default values for the distance by road type
-matrices.
-
-```{r misc3, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-getDistRoadType(ITHIM)
-```
-
-So, we have to update manually.
-
-```{r roadInjuries, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-scenario <- readRDS(file = "~/Downloads/distByRoadTypeScenario4.rds")
-base <- readRDS(file = "~/Downloads/distByRoadTypeBaseline.rds")
-ITHIM.baseline <- update(ITHIM, list(distRoadType = base))
-ITHIM.scenario <- update(ITHIM, list(distRoadType = scenario))
-getDistRoadType(ITHIM.baseline)
-getDistRoadType(ITHIM.scenario)
-```
-
-Now that we have the distances by road type we may compute the
-scenario injury multipliers and compute the scenario road injury
-estimates.
-
-```{r roadInjuries0, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
-```
-
-The final ingredient is the GBD set of parameters.  We have a file
-stored within the ITHIM package from which we may read in these
-values.  We update the ITHIM objects as follows.
-
-```{r roadInjuries9, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-#ITHIM.baseline <- update(ITHIM.baseline, list(GBD = readGBD(file = "~/GHI/misc/gbdEngland.csv")))
-#ITHIM.scenario <- update(ITHIM.scenario, list(GBD = readGBD(file = "~/GHI/misc/gbdEngland.csv")))
-```
-
-```{r roadInjuries10, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-getParameterSet(ITHIM.baseline)
-```
-
-```{r roadInjuries3, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
-(deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta))
-```
-
-## Safety in Numbers
-
-The point of all of this class and method business is to make the code for analyses tidy and human-readable.  Here is an example in which we vary the safety in numbers parameters.
-
-```{r roadInjuries4, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
-
-deltaDeathsVec <- c()
-
-for( sin in sinVec <- seq(0,1,length.out = 10) ){
-    ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = sin*matrix(1,nrow = 8, ncol = 2)))
-    ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
-    RIburden <- computeRoadInjuryBurden(ITHIM.baseline, ITHIM.scenario)
-    deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta) # Not human-readable enough, should improve
-    deltaDeathsVec <- c(deltaDeathsVec, deltaDeaths)
-}
-
-plot(sinVec, deltaDeathsVec, type = "b", pch = 20)
-
-```
-
-## Uncertainty in Road Injury Estimates
-
-```{r roadInjuries6, message = FALSE, , dev=c('png'), eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
-
-deltaDeaths.vec <- c()
-for( i in 1:50 ){
-
-  simRI <- lapply(getRoadInjuries(ITHIM.baseline), function(x){
-    lambda <- c(as.matrix(x))
-    simCount <- matrix(mapply(function(y){rpois(1,y)},y=lambda), nrow = nrow(x), ncol = ncol(x))
-    dimnames(simCount) <- dimnames(x)
-    return(as.data.frame(simCount))
-  })
-  ITHIM.baseline.sim <- update(ITHIM.baseline, list(roadInjuries = simRI))
-  ITHIM.scenario.sim <- updateRoadInjuries(ITHIM.baseline.sim, ITHIM.scenario)
-
-  RIburden <- computeRoadInjuryBurden(ITHIM.baseline.sim, ITHIM.scenario.sim)
-  deltaDeaths <- sum(subset(RIburden, burden == "dproj")$delta)
-  deltaDeaths.vec <- c(deltaDeaths.vec, deltaDeaths)
-}
-hist(deltaDeaths.vec)
-```
+* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/activeTransportTime.csv)
+* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/roadInjuries.csv)
+* [gbd.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/gbd.csv)

### 2017-02-17 R/cambridge.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("cambridge.Rmd","cambridge.html")

### 2017-02-17 R/tutorialScript.R

@@ -1,0 +1,9 @@
+library("devtools")
+install_github("syounkin/ITHIM", ref="devel")
+library("ITHIM")
+fileNamesList <- list(
+    activeTransportTimeFile = "~/ITHIM/inst/activeTransportTime.csv",
+    roadInjuriesFile = "~/ITHIM/inst/roadInjuries.csv",
+    GBDFile = "~/ITHIM/inst/gbd.csv"
+)
+ITHIM <- createITHIM(fileNamesList)

### 2017-02-23 R/RoadFatalities.Rmd
Cleaned up R directory

## Road Fatalities

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
#install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
#document("~/ITHIM/")
#install("~/ITHIM/")
install_github("syounkin/ITHIM", ref="devel")
install_github("smbache/magrittr", ref="master")
#library("magrittr")
library("ITHIM")
fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

```{r roadFat, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
pedDeath <- readRDS(file = "./data/BikePedDeathModelData.rds")
pedDeath <- subset(pedDeath, urbSubRur == "Urban")
pedDeath <- within(pedDeath, pedDeathLogical <- pedDeaths0913 > 0)
pedDeath <- within(pedDeath,{
  deathsTotal <- pedDeaths0913 + bikeDeaths0913
  deathsTotalLogical <- deathsTotal > 0
  pctATToWork <- pctWalkToWork + pctBikeToWork
AT <- ifelse(pctATToWork > 0  & pctATToWork <= 0.10 , "low" ,
ifelse(pctATToWork > 0.10  & pctATToWork <= 0.50 , "medium" ,
ifelse(pctATToWork > 0.50  & pctATToWork <= 1 , "high" , NA)))

})

D <- melt(pedDeath)

chisq.test(with(pedDeath, table(deathsTotalLogical, activeTract)))

p <- ggplot(pedDeath, aes(pedDeaths0913, ..density..,colour = AT))
p #+ geom_freqpoly(binwidth = 10) #+ facet_grid( SEX ~ .)

ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) +

```

AT <- ifelse( ( pctWalkToWork + pctBikeToWork ) > 0  &  ( pctWalkToWork + pctBikeToWork ) <= 0.10 , "low" ,
ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.10  &  ( pctWalkToWork + pctBikeToWork ) <= 0.50 , "medium" ,
ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.50  &  ( pctWalkToWork + pctBikeToWork ) <= 1 , "high" , NA)))

### 2017-02-23 R/tutorial.Rmd
Cleaned up R directory

--- 2017-01-21 R/tutorial.Rmd
@@
-## ITHIM Tutorial
+## ITHIM Package Tutorial
 
-### Installing the package
+### Install and Load ITHIM
 
-```{r figwidth, eval = TRUE,  warning = FALSE, message = FALSE, echo = FALSE}
-fig.width <- 6
-phi <- (1+sqrt(5))/2 # the golden ratio ;)
-```
-
-First load the package from Github.  To do this you need the package
-devtools.
-
-```{r, libs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-options(repos='http://cran.rstudio.com/')
-install.packages("devtools")
-```
-
-Now load the package and install ITHIM.
-
-```{r libs2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
+```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
 library("devtools")
-#install_github(repo = "syounkin/ITHIM", ref = "devel")
-install("~/ITHIM/")
+install_github("syounkin/ITHIM", ref="devel")
 library("ITHIM")
@@
 
-### Create default ITHIM object
+### Create the ITHIM object
 
-```{r create, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-ITHIM.baseline <- createITHIM()
+#### Method 1: Using Input Files
+
+We begin by creating the ITHIM object from which both the baseline and
+scenario are derived.  Currently ITHIM reads three files for the
+necessary input parameters.  Sample files are listed below.
+
+```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+fileNamesList <- list(
+    activeTransportTimeFile = "~/ITHIM/inst/activeTransportTime.csv",
+    roadInjuriesFile = "~/ITHIM/inst/roadInjuries.csv",
+    GBDFile = "~/ITHIM/inst/gbd.csv"
+)
 ```
 
-### ITHIM Methods
+These files can be found in the devel branch of ITHIM repository.
 
-#### getMeans
+* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/activeTransportTime.csv)
+* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/roadInjuries.csv)
+* [gbd.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/gbd.csv)
 
-```{r methods3, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-getMeans(ITHIM.baseline)
+Now create the ITHIM object.
+```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM <- createITHIM(fileNamesList)
 ```
 
-#### update
+##### Active Transport File Format
+```
+ageClass,sex,mode,value
+ageClass1,M,walk,10
+ageClass2,M,walk,9
+...
+ageClass7,F,cycle,15
+ageClass8,F,cycle,13
+```
+The rows must be ordered by ageClass.
+##### Road Injuries File Format
+```
+severity,roadType,strikingMode,victimMode,value
+Fatal,Local,walk,walk,0
+Fatal,Local,cycle,walk,13
+...
+Serious,Highway,LGV,NOV,0
+Serious,Highway,mbike,NOV,0
+```
+ebikes are not included
+##### GBD File Format
+The GBD file format has not been normalized like the others yet.
+```
+disease,sex,ageClass,dproj,yll,yld,daly
+BreastCancer,M,ageClass1,0,0,0,0
+BreastCancer,M,ageClass2,0,0,0,0
+...
+RTIs,F,ageClass7,1486.647682,21930.32348,7921.96966,29852.3031
+RTIs,F,ageClass8,1957.865757,10720.47109,5222.446302,15942.95826
+```
+RTIs stands for road traffic injuries.
 
-```{r update, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-ITHIM.scenario <- update(ITHIM.baseline, list(muwt = 2*getMeans(ITHIM.baseline)$walk))
+```{r misc22, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
+purl("./tutorial.Rmd","./tutorialScript.R", documentation = 0)
 ```
-
-#### compareModels
-
-```{r compareModels, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-CRA.list <- compareModels(ITHIM.baseline,ITHIM.scenario)
-```
-
-#### Plot
-
-```{r methods2, eval = FALSE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-plot(ITHIM.baseline)
-```
-
-#### getMethodsITHIM
-
-```{r getMethods, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-getMethodsITHIM()
-```
-
-### The _ParameterSet_ Object
-
-Each ITHIM object contains a ParameterSet object.
-
-```{r pSet, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-paramSet.baseline <- getParameterSet(ITHIM.baseline)
-paramSet.scenario <- getParameterSet(ITHIM.scenario)
-```
-
-```{r class, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-paramSet.baseline
-```
-
-```{r class2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
-paramSet.scenario
-```

### 2017-02-23 R/tutorial.Rmd
Update to tutorial.Rmd

@@
-## ITHIM Package Tutorial
+# ITHIM Package Tutorial
 
-### Install and Load ITHIM
+```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
+rm(list=ls())
+```
+To begin using ITHIM one first must install it.  The package devtools
+makes this is easy.
 
-```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
+## Install and Load ITHIM
+
+```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
 library("devtools")
-install_github("syounkin/ITHIM", ref="devel")
+install_github("syounkin/ITHIM", ref="master")
 library("ITHIM")
+date()
+packageVersion("ITHIM")
 ```
 
-### Create the ITHIM object
+## Creating an ITHIM object
 
-#### Method 1: Using Input Files
+The ITHIM package uses the ITHIM class to facilitate...
 
-We begin by creating the ITHIM object from which both the baseline and
-scenario are derived.  Currently ITHIM reads three files for the
-necessary input parameters.  Sample files are listed below.
+### Method 1: Using Input Files
+
+We begin by creating an ITHIM object to represent the baseline model.
+Currently ITHIM reads three files for the necessary input parameters.
+Sample files are listed below.
 
@@
 
-These files can be found in the devel branch of ITHIM repository.
+These files can be found in the master branch of ITHIM repository.
 
-* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/activeTransportTime.csv)
-* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/roadInjuries.csv)
-* [gbd.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/gbd.csv)
+* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
+* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
+* [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)
 
-Now create the ITHIM object.
+Using the list of three file names we can create the ITHIM object using `createITHIM`.
+
 ```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM <- createITHIM(fileNamesList)
+ITHIM.baseline <- createITHIM(fileNamesList)
 ```
 
-##### Active Transport File Format
+#### Active Transport File Format
 ```
@@
 The rows must be ordered by ageClass.
-##### Road Injuries File Format
+#### Road Injuries File Format
 ```
@@
 ebikes are not included
-##### GBD File Format
+#### GBD File Format
 The GBD file format has not been normalized like the others yet.
@@
 ```
+
+### Method 2: Using the `update` Method
+
+Suppose we wanted to create a scenario in which the mean cycling time
+and mean walking time are doubled.  We need only to change two means
+in the model and nothing else.  To do so we could avoid creating input
+files by simply updating the baseline object with a few new
+parameters.
+
+First we need to know the name of the parameters we want to update.
+Currently the list of parameters resides in the help page for
+"createParameterList".  We will find a better place for it.
+Alternatively one can inspect the slots directly with slotNames.
+
+```{r misc24, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+help(createParameterList)
+slotNames(getParameterSet(ITHIM.baseline))
+```
+
+Now we need to retrieve the walking and cycling means from the
+baseline model.  To do so we use the method `getMeans`.  To find a
+list of available methods use `getMethodsITHIM` with no arguments.
+
+```{r misc23, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+getMethodsITHIM()
+meanVec <- getMeans(ITHIM.baseline)
+meanWalkingTime <- meanVec$walk
+meanCyclingTime <- meanVec$cycle
+scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
+scenarioParameters
+```
+
+```{r misc250, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM.scenario <- update(ITHIM.baseline, scenarioParameters)
+```
+
+## Physical Activity Component
+
+Once the baseline and scenario ITHIM objects are defined, we may
+compute the change in burden due to physical activity as follows.
+
+```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "total", bur = "daly.delta")
+```

### 2017-02-23 R/ITHIM.Rmd
Updated deltaBurden to work with new version

@@
 
-  DALYs <- lapply(ITHIMList, deltaBurden, ITHIM.baseline = ITHIM.national)
+  DALYs <- lapply(ITHIMList, deltaBurden, baseline = ITHIM.national)
   DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))

### 2017-02-27 R/ITHIM.Rmd
Added line to create sample trip file for Ali

@@
 install("~/ITHIM/")
-#install_github("syounkin/ITHIM", ref="cambridge")
+#install_github("syounkin/ITHIM", ref="devel")
 install_github("smbache/magrittr", ref="master")
@@
 date()
-compute <- FALSE
+compute <- TRUE
 ```
@@
 saveRDS(NHTS, file = "./data/NHTS.rds")
+
+# write.csv(with(NHTS,data.frame(HOUSEID, PERSONID, SEX, AGE, MODE = AT, TRVL_MIN, TRPMILES)), file = "./misc/NHTS.csv", quote = FALSE, row.names = FALSE)
+
 ```

### 2017-02-27 R/RoadFatalities.Rmd
Created US National Road Injury estimates

@@
 library("dplyr")
+library("tidyr")
+#install_github("smbache/magrittr", ref="master")
+#library("magrittr")
 ```
@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-#document("~/ITHIM/")
-#install("~/ITHIM/")
-install_github("syounkin/ITHIM", ref="devel")
-install_github("smbache/magrittr", ref="master")
-#library("magrittr")
+install("~/ITHIM/")
+#install_github("syounkin/ITHIM", ref="devel")
 library("ITHIM")
@@
 
-```{r roadFat, eval = TRUE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+Note that I edited the file 'Countyavg_01-05.csv' by hand a little.
+The changes were simple, e.g., column names and capitilization.
+
+```{r foobar, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+RIbyCounty <- read.csv(file = "~/GHI/misc/Countyavg_01-05.csv", stringsAsFactors = TRUE)
+```
+
+We create a national road injuries estimate by adding up all of the
+county averages.  We do that using the R package `tidyr` and its
+piping mechanism `%>%`.  This is then written to the default file in
+`inst` and read in with `createITHIM` using no arguments (because its
+the default file).
+
+```{r foo, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+RINational <- RIbyCounty %>%
+  gather(strikingMode, count, -(1:5)) %>%
+  complete(severity, roadType, victimMode,strikingMode) %>%
+  group_by(severity,roadType,victimMode,strikingMode) %>%
+  summarise(value = sum(count, na.rm = TRUE)) %>%
+  filter(strikingMode != "nov")
+
+write.csv(RINational, file = "~/ITHIM/inst/roadInjuriesUS.csv", quote = FALSE, row.names = FALSE)
+```
+```{r foo2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+install("~/ITHIM/")
+ITHIM <- createITHIM()
+
+getRoadInjuries(ITHIM)
+```
+Rows are stiking mode and columns are victim mode.
+
+```{r roadFat, eval = FALSE, dev=c('png'), fig.width = fig.width, fig.asp = 1/phi, echo = FALSE, warning = FALSE}
 pedDeath <- readRDS(file = "./data/BikePedDeathModelData.rds")
@@
 
+#AT <- ifelse( ( pctWalkToWork + pctBikeToWork ) > 0  &  ( pctWalkToWork + pctBikeToWork ) <= 0.10 , "low" ,
+#ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.10  &  ( pctWalkToWork + pctBikeToWork ) <= 0.50 , "medium" ,
+#ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.50  &  ( pctWalkToWork + pctBikeToWork ) <= 1 , "high" , NA)))
+
 ```
-
-AT <- ifelse( ( pctWalkToWork + pctBikeToWork ) > 0  &  ( pctWalkToWork + pctBikeToWork ) <= 0.10 , "low" ,
-ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.10  &  ( pctWalkToWork + pctBikeToWork ) <= 0.50 , "medium" ,
-ifelse( ( pctWalkToWork + pctBikeToWork ) > 0.50  &  ( pctWalkToWork + pctBikeToWork ) <= 1 , "high" , NA)))

### 2017-02-28 R/ITHIM2.Rmd

@@
 
-```{r tilePlot, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+```{r tilePlot, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1, echo = TRUE, warning = FALSE}
 ITHIM <- createITHIM()
-tilePlot(ITHIM, n = 50)
+tilePlot(ITHIM, n = 25)
 ```

### 2017-02-28 R/appendix.Rmd

@@
+### Appendix
+
+Do it again, but with quintiles
+
+```{r ITHIMObjsQuintiles, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
+DALYsTotal2 <- data.frame()
+ITHIMList <- list()
+
+ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )
+
+ITHIM.national <- ITHIM.default
+ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
+ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)
+
+for (HHSTATE in stateVec){
+    ITHIMList[[HHSTATE]] <- ITHIM.default
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
+  }
+
+for( muNonTravel in seq(0, upper, by = wd)){
+
+  rm(DALYs)
+
+  for (HHSTATE in stateVec){
+    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
+  }
+
+  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
+
+  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
+  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
+  DALYsTotal2 <- rbind(DALYsTotal2, DALYs)
+}
+
+DALYsTotal2 <- within(DALYsTotal2, {
+  State <- factor(State, levels = stateVecOrdered)
+})
+
+DALYsTotal <- rbind(data.frame(DALYsTotal,quints="Percentiles"), data.frame(DALYsTotal2,quints="Quintiles"))
+
+p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( .
+~ quints)
+
+p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( .
+~ quints)
+
+```
+
 # Appendix

### 2017-02-28 R/tutorial.Rmd

@@
 library("ITHIM")
-date()
-packageVersion("ITHIM")
 ```
@@
 First we need to know the name of the parameters we want to update.
-Currently the list of parameters resides in the help page for
-"createParameterList".  We will find a better place for it.
-Alternatively one can inspect the slots directly with slotNames.
+Currently the list of parameters resides in the help page for the
+ITHIM class.  Alternatively one can inspect the
+slots directly with `slotNames`.
 
-```{r misc24, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-help(createParameterList)
+```{r misc24, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = TRUE}
+help("ITHIM")
+```
+or
+```{r misc124, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
 slotNames(getParameterSet(ITHIM.baseline))
@@
 meanVec <- getMeans(ITHIM.baseline)
-meanWalkingTime <- meanVec$walk
-meanCyclingTime <- meanVec$cycle
+meanVec
+meanWalkingTime <- meanVec$walk; meanCyclingTime <- meanVec$cycle
 scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
@@
 
+Now we use the list of parameters and the baseline ITHIM object to
+create the scenario ITHIM object.
+
 ```{r misc250, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
 ITHIM.scenario <- update(ITHIM.baseline, scenarioParameters)
+```
+
+Let's inspect it to be sure.
+
+```{r misc351, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+getMeans(ITHIM.scenario)
 ```
@@
 ```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "total", bur = "daly.delta")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "total", bur = "deaths")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "BreastCancer", bur = "deaths")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "deaths")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "daly")
 ```
+
+## Appendix
+
+```{r misc350, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM.baseline
+ITHIM.scenario
+```
+### Session Info
+```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
+sessionInfo()
+```

### 2017-03-03 vargo/loopStates.r
Added Vargo's loopStates.r script

@@ -1,0 +1,66 @@
+library(devtools)
+install_github("syounkin/ITHIM", ref="devel", force=TRUE)
+library("ITHIM")
+packageVersion("ITHIM")
+
+######## read in ITHIM objects ####### 
+#available at https://uwmadison.box.com/shared/static/06doqkc3aiaoqdb569r7si1vthlyvgu9.rds
+#states <- readRDS("~/Downloads/ITHIMList.by.state.rds")
+states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
+
+####### Compare States ##################
+
+allStatesResults <- data.frame()
+
+#i=48
+for(i in 1:3){ 
+  results <- data.frame()
+  state <- names(states[i])
+  base <- states[[state]]
+
+  baseWalk <- getMeans(base)$walk
+  baseCycle <-  getMeans(base)$cycle
+  # choose this number which controls the number of runs n=3 > 9 runs of  the CRA
+  n <- 3
+  
+  # mean walking and cyling times are varied from half of the baseline to double the baseline
+  wVec <- seq(baseWalk,baseWalk+n,length.out = n+1)
+  cVec <- seq(baseCycle,baseCycle+n,length.out = n+1)
+  ntVec <- c(2, 5, 10)
+  
+
+  for(muNT in ntVec){
+    for(wlk in 1:(n+1)){
+      for(cyc in 1:(n+1)){
+        scenario <- update(base, list(muwt = wVec[wlk], muct = cVec[cyc], muNonTravel = muNT))
+        activityLevel <- ifelse(muNT == 2, "Low", ifelse(muNT == 5, "Medium", "Low"))
+        comparativeRisk <- data.frame(ST = state,
+                                      cycleTime = getMeans(scenario)$cycle, 
+                                      walkTime= getMeans(scenario)$walk, 
+                                      minIncreaseCycle = cyc-1, 
+                                      minIncreaseWalk = wlk-1, 
+                                      nonTravelActivity = activityLevel,
+                                      pctTotalDALYS = deltaBurden(base, scenario)/getBurden(base),
+                                      pctBreastCancer= deltaBurden(base, scenario, dis = "BreastCancer")/getBurden(base, dis = "BreastCancer"),
+                                      pctColonCancer= deltaBurden(base, scenario, dis = "ColonCancer")/getBurden(base, dis = "ColonCancer"),
[…]

### 2017-03-03 vargo/loopStates.r
Made some changes to Vargo's loopStates.r

@@ -2,6 +2,7 @@ library(devtools)
 install_github("syounkin/ITHIM", ref="devel", force=TRUE)
 library("ITHIM")
 packageVersion("ITHIM")
+rm(list=ls())
 
 ######## read in ITHIM objects ####### 
 #available at https://uwmadison.box.com/shared/static/06doqkc3aiaoqdb569r7si1vthlyvgu9.rds
@@ -12,55 +13,74 @@ states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
 
 allStatesResults <- data.frame()
 
-#i=48
-for(i in 1:3){ 
-  results <- data.frame()
-  state <- names(states[i])
-  base <- states[[state]]
+for(i in 1:length(states)){
 
-  baseWalk <- getMeans(base)$walk
-  baseCycle <-  getMeans(base)$cycle
-  # choose this number which controls the number of runs n=3 > 9 runs of  the CRA
-  n <- 3
-  
-  # mean walking and cyling times are varied from half of the baseline to double the baseline
-  wVec <- seq(baseWalk,baseWalk+n,length.out = n+1)
-  cVec <- seq(baseCycle,baseCycle+n,length.out = n+1)
-  ntVec <- c(2, 5, 10)
-  
+    results <- data.frame()
+    state <- names(states[i])
+    base <- states[[state]]
 
-  for(muNT in ntVec){
-    for(wlk in 1:(n+1)){
-      for(cyc in 1:(n+1)){
-        scenario <- update(base, list(muwt = wVec[wlk], muct = cVec[cyc], muNonTravel = muNT))
-        activityLevel <- ifelse(muNT == 2, "Low", ifelse(muNT == 5, "Medium", "Low"))
-        comparativeRisk <- data.frame(ST = state,
-                                      cycleTime = getMeans(scenario)$cycle, 
-                                      walkTime= getMeans(scenario)$walk, 
-                                      minIncreaseCycle = cyc-1, 
-                                      minIncreaseWalk = wlk-1, 
-                                      nonTravelActivity = activityLevel,
-                                      pctTotalDALYS = deltaBurden(base, scenario)/getBurden(base),
-                                      pctBreastCancer= deltaBurden(base, scenario, dis = "BreastCancer")/getBurden(base, dis = "BreastCancer"),
-                                      pctColonCancer= deltaBurden(base, scenario, dis = "ColonCancer")/getBurden(base, dis = "ColonCancer"),
[…]

### 2017-03-03 R/ITHIM.Rmd
Removed commented-out lines from ITHIM.Rmd

@@
 ## R Package _ITHIM_
-<!-- This document will eventually serve as a compendium to the R package
-_ITHIM_.  It will describe how to use the package to implement the
-ITHIM model, as well as a detailed description of the model itself.
-
-As a working document it will serve as a place for SGY to post
-results, questions, interpreations, notes, etc.-->
 

### 2017-03-03 R/tutorial.Rmd
Update to tutorial

@@
 library("devtools")
-install_github("syounkin/ITHIM", ref="master")
+#install_github("syounkin/ITHIM", ref="master")
+install_github("syounkin/ITHIM", ref="devel")
 library("ITHIM")
@@
 ```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "total", bur = "deaths")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "all", bur = "deaths")
 deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "BreastCancer", bur = "deaths")

### 2017-03-03 vargo/loopStates.r
Updated loopStates.r

@@ -4,7 +4,7 @@ library("ITHIM")
 packageVersion("ITHIM")
 rm(list=ls())
 
-######## read in ITHIM objects ####### 
+######## read in ITHIM objects #######
 #available at https://uwmadison.box.com/shared/static/06doqkc3aiaoqdb569r7si1vthlyvgu9.rds
 #states <- readRDS("~/Downloads/ITHIMList.by.state.rds")
 states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
@@ -13,6 +13,7 @@ states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
 
 allStatesResults <- data.frame()
 
+#for(i in 1:2){
 for(i in 1:length(states)){
 
     results <- data.frame()
@@ -39,15 +40,16 @@ for(i in 1:length(states)){
     ntVec <- c(2, 5, 10)
 
     for(muNT in ntVec){
-        for(wlk in 1:(n+1)){
-            for(cyc in 1:(n+1)){
-                scenario <- update(base, list(muwt = wVec[wlk], muct = cVec[cyc], muNonTravel = muNT))
-                activityLevel <- ifelse(muNT == 2, "Low", ifelse(muNT == 5, "Medium", "Low"))
+        base <- update(base, list(muNonTravel = muNT))
+        for(wlk in wVec){
+            for(cyc in cVec){
+                scenario <- update(base, list(muwt = wlk, muct = cyc, muNonTravel = muNT))
+                activityLevel <- ifelse(muNT == ntVec[1], "Low", ifelse(muNT == ntVec[2], "Medium", ifelse(muNT == ntVec[3], "High", NA)))
                 comparativeRisk <- data.frame(ST = state,
-                                              cycleTime = getMeans(scenario)$cycle, 
-                                              walkTime= getMeans(scenario)$walk, 
-                                              minIncreaseCycle = cyc-1, 
-                                              minIncreaseWalk = wlk-1, 
+                                              cycleTime = getMeans(scenario)$cycle,
+                                              walkTime= getMeans(scenario)$walk,
+                                              minIncreaseCycle = cyc-baseCycle,
+                                              minIncreaseWalk = wlk-baseWalk,
                                               nonTravelActivity = activityLevel)
 
                 deltaDALYs <- data.frame(deltaBurden(base, scenario, dis = "all"),
@@ -59,28 +61,40 @@ for(i in 1:length(states)){
[…]

### 2017-03-04 vargo/loopStates.r
Fixed bug in loopStates.r and ran dos2unix on the file too

@@ -1,6 +1,7 @@
 library(devtools)
 install_github("syounkin/ITHIM", ref="devel", force=TRUE)
 library("ITHIM")
+library("tidyr")
 packageVersion("ITHIM")
 rm(list=ls())
 

### 2017-03-06 vargo/loopStates.r
Edits to state loop code
Changed walk and cycle vectors to range from 0 to 20 and returns -deltaBurden instead of total + deltaBurden.

@@ -14,6 +14,8 @@ states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
 
 allStatesResults <- data.frame()
 
+n <- 5
+
 #for(i in 1:2){
 for(i in 1:length(states)){
 
@@ -35,9 +37,8 @@ for(i in 1:length(states)){
     baseWalk <- getMeans(base)$walk
     baseCycle <-  getMeans(base)$cycle
 
-    n <- 4
-    wVec <- seq(baseWalk,baseWalk+n,length.out = n+1)
-    cVec <- seq(baseCycle,baseCycle+n,length.out = n+1)
+    wVec <- baseWalk + seq(0, 20, length.out = n)
+    cVec <- baseCycle + seq(0, 20, length.out = n)
     ntVec <- c(2, 5, 10)
 
     for(muNT in ntVec){
@@ -62,12 +63,15 @@ for(i in 1:length(states)){
                                          deltaBurden(base, scenario, dis = "Diabetes")
                                          )
 
+                deltaDALYs <- data.frame(ifelse(t(unlist(deltaDALYs)) > 0, 0, deltaDALYs))
+
+
                 names(deltaDALYs) <- c("all","BreastCancer","ColonCancer","CVD","Dementia","Depression","Diabetes")
+
 
                 foo <- cbind(comparativeRisk, deltaDALYs)
 
                 foo <- foo %>% gather("disease","value",7:13)
-#                comparativeRisk <-  cbind(comparativeRisk, deltaDALYs, t(totalDALYs))
 
                 results <- rbind(foo, results)
             }
@@ -82,7 +86,7 @@ allStatesResults <- data.frame(allStatesResults, total = t(totalDALYs))
 
 
 
-allStatesResults <- allStatesResults %>% dplyr::mutate( burdenScenario = total + value, pctReduction = -100*value/total)
+allStatesResults <- allStatesResults %>% dplyr::mutate( deltaBurden = -value, pctReduction = -100*value/total)
 
 allStatesResults <- melt(allStatesResults[,!names(allStatesResults) %in% c("value","total")], id =c("ST","cycleTime","walkTime","minIncreaseCycle","minIncreaseWalk","nonTravelActivity","disease"))
 

### 2017-03-06 vargo/loopStates.r
A few minor edits to state loop

@@ -10,17 +10,15 @@ rm(list=ls())
 #states <- readRDS("~/Downloads/ITHIMList.by.state.rds")
 states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
 
-####### Compare States ##################
-
 allStatesResults <- data.frame()
 
 n <- 5
 
 #for(i in 1:2){
-for(i in 1:length(states)){
+for(state in names(states)){
 
     results <- data.frame()
-    state <- names(states[i])
+
     base <- states[[state]]
 
     totalDALYs <- data.frame(getBurden(base, dis = "all"),
@@ -39,7 +37,7 @@ for(i in 1:length(states)){
 
     wVec <- baseWalk + seq(0, 20, length.out = n)
     cVec <- baseCycle + seq(0, 20, length.out = n)
-    ntVec <- c(2, 5, 10)
+    ntVec <- c(2, 5, 10) # must be ordered and have length 3
 
     for(muNT in ntVec){
         base <- update(base, list(muNonTravel = muNT))

### 2017-03-06 vargo/loopStates.r
Updated quantiles in baseline object
This may be the source of the bug we had been experiencing.

@@ -13,13 +13,14 @@ states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
 allStatesResults <- data.frame()
 
 n <- 5
+ntVec <- c(2, 5, 10) # must be ordered and have length 3
 
-#for(i in 1:2){
 for(state in names(states)){
 
     results <- data.frame()
 
     base <- states[[state]]
+    base <- update(base, list(quantiles = seq(0.1,0.9,by=0.1)))
 
     totalDALYs <- data.frame(getBurden(base, dis = "all"),
                     getBurden(base, dis = "BreastCancer"),
@@ -37,7 +38,6 @@ for(state in names(states)){
 
     wVec <- baseWalk + seq(0, 20, length.out = n)
     cVec <- baseCycle + seq(0, 20, length.out = n)
-    ntVec <- c(2, 5, 10) # must be ordered and have length 3
 
     for(muNT in ntVec){
         base <- update(base, list(muNonTravel = muNT))

### 2017-03-06 vargo/loopStates.Rmd
Wrapped loopStates in a markdown file and added to Makefile

Foo.
```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE}
library(devtools)
install_github("syounkin/ITHIM", ref="devel", force=TRUE)
library("ITHIM")
library("tidyr")
packageVersion("ITHIM")
rm(list=ls())

######## read in ITHIM objects #######
#available at https://uwmadison.box.com/shared/static/06doqkc3aiaoqdb569r7si1vthlyvgu9.rds
#states <- readRDS("~/Downloads/ITHIMList.by.state.rds")
states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")

allStatesResults <- data.frame()

n <- 5
ntVec <- c(2, 5, 10) # must be ordered and have length 3

for(state in names(states)){

    results <- data.frame()

    base <- states[[state]]
    base <- update(base, list(quantiles = seq(0.1,0.9,by=0.1)))

    totalDALYs <- data.frame(getBurden(base, dis = "all"),
                    getBurden(base, dis = "BreastCancer"),
                    getBurden(base, dis = "ColonCancer"),
                    getBurden(base, dis = "CVD"),
                    getBurden(base, dis = "Dementia"),
                    getBurden(base, dis = "Depression"),
                    getBurden(base, dis = "Diabetes")
                    )

    names(totalDALYs) <- c("DALY.all","DALY.BreastCancer","DALY.ColonCancer","DALY.CVD","DALY.Dementia","DALY.Depression","DALY.Diabetes")

    baseWalk <- getMeans(base)$walk
    baseCycle <-  getMeans(base)$cycle

    wVec <- baseWalk + seq(0, 20, length.out = n)
    cVec <- baseCycle + seq(0, 20, length.out = n)

    for(muNT in ntVec){
        base <- update(base, list(muNonTravel = muNT))
        for(wlk in wVec){
            for(cyc in cVec){
                scenario <- update(base, list(muwt = wlk, muct = cyc, muNonTravel = muNT))
                activityLevel <- ifelse(muNT == ntVec[1], "Low", ifelse(muNT == ntVec[2], "Medium", ifelse(muNT == ntVec[3], "High", NA)))
                comparativeRisk <- data.frame(ST = state,
                                              cycleTime = getMeans(scenario)$cycle,
                                              walkTime= getMeans(scenario)$walk,
                                              minIncreaseCycle = cyc-baseCycle,
                                              minIncreaseWalk = wlk-baseWalk,
                                              nonTravelActivity = activityLevel)

                deltaDALYs <- data.frame(deltaBurden(base, scenario, dis = "all"),
                                         deltaBurden(base, scenario, dis = "BreastCancer"),
                                         deltaBurden(base, scenario, dis = "ColonCancer"),
                                         deltaBurden(base, scenario, dis = "CVD"),
                                         deltaBurden(base, scenario, dis = "Dementia"),
                                         deltaBurden(base, scenario, dis = "Depression"),
                                         deltaBurden(base, scenario, dis = "Diabetes")
                                         )

                deltaDALYs <- data.frame(ifelse(t(unlist(deltaDALYs)) > 0, 0, deltaDALYs))

                names(deltaDALYs) <- c("all","BreastCancer","ColonCancer","CVD","Dementia","Depression","Diabetes")

                foo <- cbind(comparativeRisk, deltaDALYs)

                foo <- foo %>% gather("disease","value",7:13)

                results <- rbind(foo, results)
            }
        }
    }

    allStatesResults <- rbind(results, allStatesResults)

}

allStatesResults <- data.frame(allStatesResults, total = t(totalDALYs))

allStatesResults <- allStatesResults %>% dplyr::mutate( deltaBurden = -value, pctReduction = -100*value/total)

allStatesResults <- melt(allStatesResults[,!names(allStatesResults) %in% c("value","total")], id =c("ST","cycleTime","walkTime","minIncreaseCycle","minIncreaseWalk","nonTravelActivity","disease"))

## allStatesResults <- within(allStatesResults,{
##                            pctTotalDALYS = deltaDALY.all/DALY.all
##                            pctBreastCancer = deltaDALY.BreastCancer/DALY.BreastCancer
##                            pctColonCancer = deltaDALY.ColonCancer/DALY.ColonCancer
##                            pctCVD = deltaDALY.CVD/DALY.CVD
##                            pctDementia = deltaDALY.Dementia/DALY.Dementia
##                            pctDepression = deltaDALY.Depression/DALY.Depression
##                            pctDiabetes = deltaDALY.Diabetes/DALY.Diabetes
## })

write.csv(allStatesResults, file = "~/GHI/data/stateResults.csv", quote = FALSE, row.names = FALSE)
```

### 2017-03-06 vargo/loopStates.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("loopStates.Rmd","loopStates.html")

### 2017-03-06 vargo/loopStates.Rmd
Increased upper limits of walk and cycle vectors to 60 (from 20)

@@
 Foo.
-```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE}
+```{r states, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE}
+
 library(devtools)
@@
 
-######## read in ITHIM objects #######
-#available at https://uwmadison.box.com/shared/static/06doqkc3aiaoqdb569r7si1vthlyvgu9.rds
-#states <- readRDS("~/Downloads/ITHIMList.by.state.rds")
 states <- readRDS("~/GHI/R/data/ITHIMList.by.state.rds")
@@
 
-n <- 5
+n <- 5 # length of walk and cycle vectors
 ntVec <- c(2, 5, 10) # must be ordered and have length 3
@@
     base <- states[[state]]
-    base <- update(base, list(quantiles = seq(0.1,0.9,by=0.1)))
+    base <- update(base, list(quantiles = seq(0.1, 0.9, by = 0.1)))
 
@@
 
-    wVec <- baseWalk + seq(0, 20, length.out = n)
-    cVec <- baseCycle + seq(0, 20, length.out = n)
+    wVec <- baseWalk + seq(0, 60, length.out = n)
+    cVec <- baseCycle + seq(0, 60, length.out = n)
 
@@
 
-## allStatesResults <- within(allStatesResults,{
-##                            pctTotalDALYS = deltaDALY.all/DALY.all
-##                            pctBreastCancer = deltaDALY.BreastCancer/DALY.BreastCancer
-##                            pctColonCancer = deltaDALY.ColonCancer/DALY.ColonCancer
-##                            pctCVD = deltaDALY.CVD/DALY.CVD
-##                            pctDementia = deltaDALY.Dementia/DALY.Dementia
-##                            pctDepression = deltaDALY.Depression/DALY.Depression
-##                            pctDiabetes = deltaDALY.Diabetes/DALY.Diabetes
-## })
-
 write.csv(allStatesResults, file = "~/GHI/data/stateResults.csv", quote = FALSE, row.names = FALSE)

### 2017-03-06 vargo/loopStates.Rmd
Fixed bug in state loop

@@
 
-for(state in names(states)){
+stateVec <- names(states)
+
+for(state in stateVec){
 
@@
                                               minIncreaseWalk = wlk-baseWalk,
-                                              nonTravelActivity = activityLevel)
+                                              nonTravelActivity = activityLevel, stringsAsFactors = FALSE)
 
@@
 
+                foo <- data.frame(foo, total = t(totalDALYs))
+
                 results <- rbind(foo, results)
@@
 
-allStatesResults <- data.frame(allStatesResults, total = t(totalDALYs))
+summary(allStatesResults)
 
@@
 
+summary(allStatesResults)
+
 allStatesResults <- melt(allStatesResults[,!names(allStatesResults) %in% c("value","total")], id =c("ST","cycleTime","walkTime","minIncreaseCycle","minIncreaseWalk","nonTravelActivity","disease"))
+
+summary(allStatesResults)
 

### 2017-03-09 tomek/roadTest.R
Added Tomek's road injury test script Made a few changes to it.

@@ -1,0 +1,29 @@
+rm(list=ls())
+library("ITHIM") # Make sure you're on the devel branch
+
+ITHIM <- createITHIM()
+
+# DistByRoadType
+
+ITHIM.baseline <- readDistByRoadType(ITHIM, system.file("distByRoadTypeBaseline.csv", package = "ITHIM"))
+ITHIM.scenario <- readDistByRoadType(ITHIM, system.file("distByRoadTypeScenario.csv", package = "ITHIM"))
+
+# SafetyInNumbers
+
+ITHIM.baseline <- readSafetyInNumbers(ITHIM.baseline, system.file("SiN.csv", package = "ITHIM"))
+ITHIM.scenario <- readSafetyInNumbers(ITHIM.scenario, system.file("SiN.csv", package = "ITHIM"))
+
+# RoadInjuries for baseline
+
+ITHIM.baseline <- readRoadInjuries(ITHIM.baseline, system.file("roadInjuriesBaseline.csv", package = "ITHIM"))
+
+# RoadInjuries for scenario using scenario multiplier and baseline RoadInjuries
+
+ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
+
+##################################################################
+# refactored functions - individual call
+
+testComputeMultiplier <- computeMultiplier(baseline = ITHIM.baseline, scenario = ITHIM.scenario, safetyInNumbers = getSiN(ITHIM.baseline))
+
+testMultiplyInjuries <- multiplyInjuries(ITHIM.baseline = ITHIM.baseline, ITHIM.scenario = ITHIM.scenario)

### 2017-03-09 R/tutorial.Rmd
Updated tutorial to use new createITHIM method

@@
 
-## Creating an ITHIM object
-
-The ITHIM package uses the ITHIM class to facilitate...
-
-### Method 1: Using Input Files
+## Creating a Baseline ITHIM object
 
@@
 ```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-fileNamesList <- list(
-    activeTransportTimeFile = "~/ITHIM/inst/activeTransportTime.csv",
-    roadInjuriesFile = "~/ITHIM/inst/roadInjuries.csv",
-    GBDFile = "~/ITHIM/inst/gbd.csv"
-)
+activeTransportFile <- "~/ITHIM/inst/activeTransportTime.csv"
+GBDFile <- "~/ITHIM/inst/gbd.csv"
 ```
@@
 * [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
-* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
 * [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)
@@
 ```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline <- createITHIM(fileNamesList)
+ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile, GBDFile = GBDFile)
 ```
@@
 The rows must be ordered by ageClass.
-#### Road Injuries File Format
-```
-severity,roadType,strikingMode,victimMode,value
-Fatal,Local,walk,walk,0
-Fatal,Local,cycle,walk,13
-...
-Serious,Highway,LGV,NOV,0
-Serious,Highway,mbike,NOV,0
-```
-ebikes are not included
 #### GBD File Format
@@
 
-### Method 2: Using the `update` Method
+### Road Injury Data
+
+If road injury data is present it can be included in the same manner.
+
+```{r misc123, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+roadInjuriesFile <- "~/ITHIM/inst/roadInjuries.csv"
+```
+
+```{r misc220, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile, GBDFile = GBDFile, roadInjuriesFile = roadInjuriesFile)
+```
+
+#### Road Injury File Format
+* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
+
+#### Road Injuries File Format
+```
+severity,roadType,strikingMode,victimMode,value
+Fatal,Local,walk,walk,0
+Fatal,Local,cycle,walk,13
+...
+Serious,Highway,LGV,NOV,0
+Serious,Highway,mbike,NOV,0
+```
+ebikes are not included
+
+## Creating Scenarios Using the `update` Method
 
@@
 Currently the list of parameters resides in the help page for the
-ITHIM class.  Alternatively one can inspect the
-slots directly with `slotNames`.
+ITHIM class (`help("ITHIM")`).  Alternatively one can use
+`getParameterNames`.
 
-```{r misc24, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = TRUE}
-help("ITHIM")
-```
-or
 ```{r misc124, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-slotNames(getParameterSet(ITHIM.baseline))
+getParameterNames(ITHIM.baseline)
 ```
@@
 
-```{r misc350, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline
-ITHIM.scenario
-```
-### Session Info
+### Session Info & Package Test
 ```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
+library("testthat")
+test_package("ITHIM")
 sessionInfo()

### 2017-03-16 R/check.R
Created check.R
This script documents, installs and tests the package.

@@ -1,0 +1,9 @@
+library("devtools")
+document("~/ITHIM/")
+install("~/ITHIM/")
+
+library("ITHIM")
+example("ITHIM")
+
+library("testthat")
+test_package("ITHIM")

### 2017-03-16 R/tutorial.Rmd
Update to tutorial

@@
 ```
-To begin using ITHIM one first must install it.  The package devtools
-makes this is easy.
+To begin using ITHIM one first must install it.  The package
+_devtools_ makes this is easy.  Install _devtools_ using
+`install.packages("devtools")` or some other way in _RStudio_.
 
@@
 
+We use the function `install_github` to install the _master_ branch of
+the _ITHIM_ package repository.
+
 ```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
 library("devtools")
-#install_github("syounkin/ITHIM", ref="master")
-install_github("syounkin/ITHIM", ref="devel")
+install_github("syounkin/ITHIM", ref="master")
 library("ITHIM")
@@
 We begin by creating an ITHIM object to represent the baseline model.
-Currently ITHIM reads three files for the necessary input parameters.
-Sample files are listed below.
+Currently _ITHIM_ can read up to three files containing input
+parameters, but it is not necessary to provide all three.  Two sample
+files are listed below that contain model parameters for the active
+transport component of _ITHIM_.
 
@@
 
-These files can be found in the master branch of ITHIM repository.
+For a description of the file format see the help page for
+`createITHIM` using `help("createITHIM")`.
+
+```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
+                              GBDFile = GBDFile)
+```
 
@@
 
-Using the list of three file names we can create the ITHIM object using `createITHIM`.
+### File Formats
+#### Active Transport File Format
 
-```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile, GBDFile = GBDFile)
+The active transport file contains mean walking and cycling times
+stratified by sex and _ITHIM_ age class.  The default file is located
+in _inst/_.
+
+```
+ageClass,sex,mode,value
+ageClass1,M,walk,21.47
+ageClass2,M,walk,53.11
+ageClass3,M,walk,50.13
+...
+ageClass6,F,cycle,1.53
+ageClass7,F,cycle,0.53
+ageClass8,F,cycle,0.23
 ```
 
-#### Active Transport File Format
+Notes about file format:
+* The rows must be ordered by ageClass.
+* The mean values are in minutes per week
+
+#### Disease Burden File Format
+
+The disease burden file contains estimates for disease burden
+stratified by _ITHIM_ age class and sex.
+
 ```
-ageClass,sex,mode,value
-ageClass1,M,walk,10
-ageClass2,M,walk,9
+disease,sex,ageClass,burdenType,value
+BreastCancer,M,ageClass1,dproj,0
+BreastCancer,M,ageClass2,dproj,0
+BreastCancer,M,ageClass3,dproj,0
 ...
-ageClass7,F,cycle,15
-ageClass8,F,cycle,13
+RTIs,F,ageClass6,daly,49108.81151
+RTIs,F,ageClass7,daly,29852.3031
+RTIs,F,ageClass8,daly,15942.95826
 ```
-The rows must be ordered by ageClass.
-#### GBD File Format
-The GBD file format has not been normalized like the others yet.
-```
-disease,sex,ageClass,dproj,yll,yld,daly
-BreastCancer,M,ageClass1,0,0,0,0
-BreastCancer,M,ageClass2,0,0,0,0
-...
-RTIs,F,ageClass7,1486.647682,21930.32348,7921.96966,29852.3031
-RTIs,F,ageClass8,1957.865757,10720.47109,5222.446302,15942.95826
-```
-RTIs stands for road traffic injuries.
+
+Notes about file format:
+* RTIs stands for road traffic injuries
+* Only _ITHIM_ diseases can be included, namely "BreastCancer",
+  "ColonCancer", "CVD", "Dementia", "Diabetes", "Depression" and "RTIs"
 
@@
 
-### Road Injury Data
+#### Road Injuries File Format
 
@@
 ```{r misc220, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile, GBDFile = GBDFile, roadInjuriesFile = roadInjuriesFile)
+ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
+                              GBDFile = GBDFile,
+                              roadInjuriesFile = roadInjuriesFile)
 ```
 
-#### Road Injury File Format
 * [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
 
-#### Road Injuries File Format
 ```
@@
 Fatal,Local,cycle,walk,13
+Fatal,Local,bus,walk,213
 ...
+Serious,Highway,HGV,NOV,0
 Serious,Highway,LGV,NOV,0
@@
 ```
-ebikes are not included
+
+Notes about file format:
+* ebikes are not included
 
@@
 
+## Built-In Example
+
+An example is contained within the _ITHIM_ package.
+
+```{r misc225, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+example("ITHIM")
+```
+
 ## Appendix

### 2017-04-25 R/ITHIM.Rmd
Changed state ITHIM list to use deciles instead of percentiles
Fixed a few bugs
Most of the problem is with external data.  Need to improve that.

@@
                                SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
-                               AGE = as.factor(convertToAgeClass(R_AGE)),
+                               AGE = as.factor(ITHIM:::convertToAgeClass(R_AGE)),
                                MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
@@
 ```{r loadNHTS, eval = !compute}
-NHTS <- readRDS(file = "./data/NHTS.rds")
+NHTS <- readRDS(file = "~/GHI/R/data/NHTS.rds")
 ```
@@
 ```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
-newMETsFile <- "ATUS_to_METS_key.csv"
+newMETsFile <- "~/GHI/data/ATUS_to_METS_key.csv"
 
-newMETsFile <- system.file(newMETsFile, package="ITHIM")
+#newMETsFile <- system.file(newMETsFile, package="ITHIM")
 
@@
 
-gbdRDS <-  readRDS(file = "./data/stateMetroBurdens2014.rds")
+gbdRDS <-  readRDS(file = "~/GHI/R/data/stateMetroBurdens2014.rds")
 gbdRDS <- within(gbdRDS, state <- stateNameToAbbrev(state))
@@
 
-ITHIM.default <- update(ITHIM.default, list( cvNonTravel = 3, cv = 1.65, muNonTravel = 4, muNonTravelMatrix = METmatrix, quantiles = (1:99)/100 ) )
+ITHIM.default <- update(ITHIM.default, list( cvNonTravel = 3, cv = 1.65, muNonTravel = 4, muNonTravelMatrix = METmatrix, quantiles = (1:9)/10 ) )
 

### 2017-04-25 tomek/roadTest.R

@@ -8,15 +8,15 @@ ITHIM.baseline <- createITHIM()
 
 # Scenario DistByRoadType
 
-ITHIM.scenario <- update(ITHIM.baseline, list(distRoadType = readDistByRoadType(system.file("distByRoadTypeScenario.csv", package = "ITHIM"))))
+ITHIM.scenario <- update(ITHIM.baseline, list(distRoadType = ITHIM:::readDistByRoadType(system.file("distByRoadTypeScenario.csv", package = "ITHIM"))))
 
 # RoadInjuries for baseline
 
-ITHIM.baseline <- update(ITHIM.baseline, list(roadInjuries = readRoadInjuries(system.file("roadInjuriesBaseline.csv", package = "ITHIM"))))
+ITHIM.baseline <- update(ITHIM.baseline, list(roadInjuries = ITHIM:::readRoadInjuries(system.file("roadInjuries.csv", package = "ITHIM"))))
 
 # RoadInjuries for scenario using scenario multiplier and baseline RoadInjuries
 
-ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
+ITHIM.scenario <- ITHIM:::updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
 
 ## rm(list=ls())
 ## library("ITHIM") # Make sure you're on the devel branch

### 2017-04-26 R/ITHIM.Rmd
A few tweaks to ITHIM.Rmd

@@
 date()
-compute <- TRUE
+compute <- FALSE
 ```
@@
 p <- ggplot(D, aes(state, value/1e3, fill = variable))
-p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip() # + coord_cartesian(ylim = c(175, 600))
+p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
 ```
@@
 
-p <- ggplot(subset(D, ageClass != "ageClass1"), aes(x = ageClass, y = mu, fill = state)) + geom_bar(stat= "identity", position = "dodge")
+p <- ggplot(subset(D, ageClass != "ageClass1"), aes(x = ageClass, y = mean, fill = state)) + geom_bar(stat= "identity", position = "dodge")
 p + theme_bw() + facet_grid( .
 ~ sex) + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Time Walking per Week (minutes)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + scale_fill_manual(values=cbPalette[-3])
+
 ```
@@
 ITHIM.CA <- update(ITHIM.CA, list(muNonTravel = 10))
-foo1 <- compareModels(ITHIM.national,ITHIM.CA)
+foo1 <- ITHIM:::compareModels(ITHIM.national,ITHIM.CA)
 foobar1 <- data.frame(cbind(melt(foo1$daly.delta),paste0("ageClass",1:8)),State = "CA")
@@
 ITHIM.NY <- update(ITHIM.NY, list(muNonTravel = 10))
-foo2 <- compareModels(ITHIM.national,ITHIM.NY)
+foo2 <- ITHIM:::compareModels(ITHIM.national,ITHIM.NY)
 foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
@@
     for( state2 in state2Vec ){
-        foo <- data.frame(lapply(compareModels(ITHIMList[[state1]],ITHIMList[[state2]])$daly.delta, function(x) sum(x$M + x$F)*prop), baseline = state1, scenario = state2)
+        foo <- data.frame(lapply(ITHIM:::compareModels(ITHIMList[[state1]],ITHIMList[[state2]])$daly.delta, function(x) sum(x$M + x$F)*prop), baseline = state1, scenario = state2)
         foobar <- rbind(foobar,foo)

### 2017-04-28 tomek/roadTest.R
Update to roadTest.R
Some of the functions Tomek were using are no longer exported.  Some other changes were made.

@@ -1,29 +1,19 @@
 rm(list=ls())
+
 library("ITHIM") # Make sure you're on the devel branch
 
-ITHIM <- createITHIM()
+# SafetyInNumbers and DistByRoadType Baseline data already included via default input files
 
-# DistByRoadType
+ITHIM.baseline <- createITHIM()
 
-ITHIM.baseline <- readDistByRoadType(ITHIM, system.file("distByRoadTypeBaseline.csv", package = "ITHIM"))
-ITHIM.scenario <- readDistByRoadType(ITHIM, system.file("distByRoadTypeScenario.csv", package = "ITHIM"))
+# Scenario DistByRoadType
 
-# SafetyInNumbers
-
-ITHIM.baseline <- readSafetyInNumbers(ITHIM.baseline, system.file("SiN.csv", package = "ITHIM"))
-ITHIM.scenario <- readSafetyInNumbers(ITHIM.scenario, system.file("SiN.csv", package = "ITHIM"))
+ITHIM.scenario <- update(ITHIM.baseline, list(distRoadType = ITHIM:::readDistByRoadType(system.file("distByRoadTypeScenario.csv", package = "ITHIM"))))
 
 # RoadInjuries for baseline
 
-ITHIM.baseline <- readRoadInjuries(ITHIM.baseline, system.file("roadInjuriesBaseline.csv", package = "ITHIM"))
+ITHIM.baseline <- update(ITHIM.baseline, list(roadInjuries = ITHIM:::readRoadInjuries(system.file("roadInjuries.csv", package = "ITHIM"))))
 
 # RoadInjuries for scenario using scenario multiplier and baseline RoadInjuries
 
-ITHIM.scenario <- updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
-
-##################################################################
-# refactored functions - individual call
-
-testComputeMultiplier <- computeMultiplier(baseline = ITHIM.baseline, scenario = ITHIM.scenario, safetyInNumbers = getSiN(ITHIM.baseline))
-
-testMultiplyInjuries <- multiplyInjuries(ITHIM.baseline = ITHIM.baseline, ITHIM.scenario = ITHIM.scenario)
+ITHIM.scenario <- ITHIM:::updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)

### 2017-05-01 R/ITHIM.Rmd
Restructuring GBD by State data
Debugging ITHIM.Rmd

@@
   gbdListTemp <- split(gbd,gbd$disease)
-  gbdListTemp[["CVD"]] <- data.frame(disease = "CVD", gbdListTemp$IHD[,c("sex",  "ageClass")], gbdListTemp$IHD[,c("dproj","yll","yld","daly")] + gbdListTemp$InflammatoryHD[,c("dproj","yll","yld","daly")] + gbdListTemp$HHD[,c("dproj","yll","yld","daly")])
   gbdListTemp2 <- lapply(gbdListTemp,function(x) split(x,as.factor(x$sex)))
@@
 
-gbdByState <- split(gbdRDS,gbdRDS$state)
+GBD <- bind_rows(
+  gbdRDS %>%
+  melt() %>%
+  filter(disease %in% c("IHD", "HHD", "InflammatoryHD")) %>%
+  group_by(state, sex, ageClass, variable) %>% summarise(value = sum(value)) %>%
+  mutate(disease = "CVD"),
+  gbdRDS %>%
+  melt() %>%
+  filter(disease %in% c("BreastCancer","ColonCancer","Dementia", "Diabetes", "Depression", "RTIs"))
+)
+
+GBD <- within(GBD, {
+  variable <- ifelse(variable == "dproj","deaths", as.character(variable))
+})
+
+GBD <- with(GBD, data.frame(state, disease, sex, ageClass, variable, value, stringsAsFactors = FALSE)) %>% arrange(state,disease,sex,ageClass,variable)
+
+gbdByState <- split(GBD,GBD$state)
 gbdByState <- lapply(gbdByState, function(x) x[,!names(x)=="state"])
@@
 
-ITHIM.default <- update(ITHIM.default, list( cvNonTravel = 3, cv = 1.65, muNonTravel = 4, muNonTravelMatrix = METmatrix, quantiles = (1:9)/10 ) )
+ITHIM.default <- update(ITHIM.default, list( cvNonTravel = 3, cv = 1.65, muNonTravel = 4, muNonTravelMatrix = METmatrix, quantiles = (1:99)/100 ) )
 
@@
 
+ITHIMList <- ITHIMList[names(ITHIMList) != "Natl."] # The national ITHIM obnject needs to have GBD updated.
+
 ITHIMListScenario <- lapply(ITHIMList, function(ithim) update(ithim, list(muwt = ithim@parameters@muwt + 10*7)))
@@
 
-getDALY <- function( ITHIM ){
-foo <- data.frame(lapply(ITHIM@parameters@GBD, function(x) x$M$daly + x$F$daly))
-sum(foo[,names(foo) %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")])
-}
+## getDALY <- function( ITHIM ){
+## foo <- data.frame(lapply(ITHIM@parameters@GBD, function(x) x$M$daly + x$F$daly))
+## sum(foo[,names(foo) %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")])
+## }
 
-burdenList <- lapply(ITHIMList, getDALY)
-foo <- data.frame(daly = unlist(burdenList), delta = deltaDALYList)
-foo <- within(foo, {
-  daly <- daly + delta
-  delta <- -delta
-  state <- names(burdenList)
-})
+## burdenList <- lapply(ITHIMList, getDALY)
+## foo <- data.frame(daly = unlist(burdenList), delta = deltaDALYList)
+## foo <- within(foo, {
+##   daly <- daly + delta
+##   delta <- -delta
+##   state <- names(burdenList)
+## })
 
-foo <- subset(foo, state %in% c("MI","OH"))
+## foo <- subset(foo, state %in% c("MI","OH"))
 
-D <- melt(foo)
+## D <- melt(foo)
 
@@
 
-p <- ggplot(D, aes(state, value/1e3, fill = variable))
-p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
+## p <- ggplot(D, aes(state, value/1e3, fill = variable))
+## p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
 ```

### 2017-05-03 R/ITHIM.Rmd
Fixed a bug in GBDList creation

@@
 
-GBD <- with(GBD, data.frame(state, disease, sex, ageClass, variable, value, stringsAsFactors = FALSE)) %>% arrange(state,disease,sex,ageClass,variable)
+GBD <- with(GBD, data.frame(state, disease, sex, ageClass, burdenType = variable, value, stringsAsFactors = FALSE)) %>% arrange(state,disease,sex,ageClass,burdenType)
 
@@
 
-```{r statewiseGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r statewiseGBD, eval = FALSE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
 D <- melt(lapply(ITHIMList, function(ITHIM) return(ITHIM@parameters@GBD)))
+
+D <- do.call(rbind,lapply(ITHIMList, getGBD))
+
 names(D) <- c("disease","sex","ageClass","type", "burden", "sex2", "disease2", "state")
@@
 
-```{r displayMeans, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
+```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
 

### 2017-05-03 R/ITHIM.Rmd
Fixed bugs
Rwt and Rct now muse be all NA.  We use 1e-6 for ageClass1.  Using NA was better.

@@
 ITHIM.national <- ITHIM.default
+
+RwtNational[1,] <- c(1e-6,1e-6)
+RctNational[1,] <- c(1e-6,1e-6)
+
 ITHIM.national <- update(ITHIM.national, list(muwt = muwtNational, muct = muctNational,Rwt = RwtNational,Rct = RctNational))
@@
 
-p <- ggplot(D, aes(x = ageClass, y = mu, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+p <- ggplot(D, aes(x = ageClass, y = mean, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
 p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8)) + facet_grid( type ~ ., scales = "free")
@@
 ```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 1.2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
 D <- melt(ITHIM.national@parameters@GBD)
-names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")
+
+D <- getGBD(ITHIM.national)
+
+names(D) <- c("disease","sex","ageClass","type","burden")
 D <- within(D, sex <- factor(sex, levels = c("F","M")))
@@
 p <- ggplot(subset(D, disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD") & type == "daly"), aes(x = ageClass, y = burden/1e3, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
+
 p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="DALYs Nationwide (thousands)") + theme(axis.text.x  = element_text(angle = 270, vjust = 0.5, size = 8)) + facet_grid( .
@@
 ```{r doseresponse, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
-RR1vec <- unique(unlist(createActiveTransportRRs()))
-lapply(createActiveTransportRRs(),function(x) lapply(x, function(x) unique(c(x))))
+RR1vec <- unique(unlist(ITHIM:::createActiveTransportRRs()))
+lapply(ITHIM:::createActiveTransportRRs(),function(x) lapply(x, function(x) unique(c(x))))
 #~~~~~~~~~~~~~~~~~~~~~~~

### 2017-05-04 R/ITHIM.Rmd
ITHIM.Rmd is now running without error again

@@
 
-```{r statewiseGBD, eval = FALSE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
-D <- melt(lapply(ITHIMList, function(ITHIM) return(ITHIM@parameters@GBD)))
-
-D <- do.call(rbind,lapply(ITHIMList, getGBD))
-
-names(D) <- c("disease","sex","ageClass","type", "burden", "sex2", "disease2", "state")
+```{r statewiseGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+D <- do.call(rbind, mapply(function(df,state) data.frame(df, state = state), lapply(ITHIMList, getGBD), names(ITHIMList), SIMPLIFY = FALSE))
+names(D) <- c("disease","sex","ageClass","type", "burden", "state")
 
@@
 deltaDALYList <- mapply(deltaBurden, ITHIMList, ITHIMListScenario)
-
-## getDALY <- function( ITHIM ){
-## foo <- data.frame(lapply(ITHIM@parameters@GBD, function(x) x$M$daly + x$F$daly))
-## sum(foo[,names(foo) %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")])
-## }
-
-## burdenList <- lapply(ITHIMList, getDALY)
-## foo <- data.frame(daly = unlist(burdenList), delta = deltaDALYList)
-## foo <- within(foo, {
-##   daly <- daly + delta
-##   delta <- -delta
-##   state <- names(burdenList)
-## })
-
-## foo <- subset(foo, state %in% c("MI","OH"))
-
-## D <- melt(foo)
-
-cbPalette <- c("#0072B2", "#56B4E9", "#999999", "#E69F00", "#009E73", "#F0E442", "#D55E00", "#CC79A7")
-
-## p <- ggplot(D, aes(state, value/1e3, fill = variable))
-## p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
 ```
@@
 ```{r rustBelt2, eval = TRUE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+
+cbPalette <- c("#0072B2", "#56B4E9", "#999999", "#E69F00", "#009E73", "#F0E442", "#D55E00", "#CC79A7")
 
@@
 ```
+
+```{r rustBeltOld, eval = FALSE, dev=c('pdf', 'png'), fig.width = 2*fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
+## getDALY <- function( ITHIM ){
+## foo <- data.frame(lapply(ITHIM@parameters@GBD, function(x) x$M$daly + x$F$daly))
+## sum(foo[,names(foo) %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes", "CVD")])
+## }
+
+## burdenList <- lapply(ITHIMList, getDALY)
+## foo <- data.frame(daly = unlist(burdenList), delta = deltaDALYList)
+## foo <- within(foo, {
+##   daly <- daly + delta
+##   delta <- -delta
+##   state <- names(burdenList)
+## })
+
+## foo <- subset(foo, state %in% c("MI","OH"))
+
+## D <- melt(foo)
+
+## cbPalette <- c("#0072B2", "#56B4E9", "#999999", "#E69F00", "#009E73", "#F0E442", "#D55E00", "#CC79A7")
+
+## p <- ggplot(D, aes(state, value/1e3, fill = variable))
+## p + geom_bar(stat= "identity") + scale_x_discrete(name="") + scale_y_continuous(name="Thousands of DALYs") + guides(fill=FALSE) + scale_fill_manual(values=cbPalette) + theme(axis.text.y  = element_text(size = 16)) + theme(axis.text.x  = element_text(size = 16), axis.title.x  = element_text(size = 16)) + coord_flip()
+```

### 2017-05-09 R/ITHIM.Rmd
Created NHTS.Rmd
Also split on MSA instead of state

--- 2017-05-03 R/ITHIM.Rmd
@@
 
-## HHS Metro Region Parameter Sets
-
-First we use NHTS to estimate mean walking and cycling times.
-
-### NHTS
-
-```{r compute, echo = FALSE, eval = TRUE}
-date()
-compute <- FALSE
-```
-
-#### Create a handful of functions to help parse data
-
-```{r parsely, eval = TRUE, echo = FALSE}
-convertToHHS <- function(states){
-  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
-  StateToHHS <- HHSRegions[,3]
-  names(StateToHHS) <- HHSRegions[,1]
-  return(StateToHHS[states])
-}
-
-countPeople <- function(data){
- nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
-}
-
-tripTime <- function(data, mode = "walk"){
- return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
-}
-
-reformatData <- function(data){
-foo <- matrix(unlist(data), nrow = 7, ncol = 2)
-foo <- rbind( c(0, 0), foo)
-dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
-return(foo[,c("M","F")])
-}
-
-```
-
-#### Read Two of the Four NHTS Data Tables
-
-```{r NHTS, eval = compute, echo = FALSE}
-DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
-names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")
-
-NHTS.trips <- with(DAYV2PUB,{
-    data.frame(HOUSEID = HOUSEID,
-                   PERSONID = PERSONID,
-                   TDCASEID = TDCASEID,
-                   AT = ifelse(TRPTRANS == "22", "cycle",
-                        ifelse(TRPTRANS == "23", "walk",
-                        ifelse(!is.na(TRPTRANS), "other", NA))),
-                   TRVL_MIN = as.numeric(TRVL_MIN),
-                   #TRVLCMIN = as.numeric(TRVLCMIN),
-                   URBRUR  = as.factor(URBRUR),
-                   MSACAT = as.factor(MSACAT),
-                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
-                   TRPTRANS = as.factor(TRPTRANS),
-                   TRPMILES = as.numeric(TRPMILES)
-)})
-
-PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
-names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
-
-NHTS.person <- with(PERV2PUB,{
-                    data.frame(HOUSEID = HOUSEID,
-                               PERSONID = PERSONID,
-                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
-                               AGE = as.factor(ITHIM:::convertToAgeClass(R_AGE)),
-                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
-                               HHSTATE = as.character(HHSTATE),
-                               HHS = as.character(convertToHHS(as.character(HHSTATE))), stringsAsFactors = FALSE
-
-)
-})
-
-```
-
-#### Subset and Merge
-
-We now correct for zero trip times in NHTS.
-
-```{r subset, eval = compute, fig.width = 10, echo = FALSE}
-NHTS.person <- subset( NHTS.person, MSA == TRUE )
-NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))
-
-NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
-NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
-saveRDS(NHTS, file = "./data/NHTS.rds")
-
-# write.csv(with(NHTS,data.frame(HOUSEID, PERSONID, SEX, AGE, MODE = AT, TRVL_MIN, TRPMILES)), file = "./misc/NHTS.csv", quote = FALSE, row.names = FALSE)
-
-```
-
-```{r loadNHTS, eval = !compute}
+```{r loadNHTS, eval = TRUE}
 NHTS <- readRDS(file = "~/GHI/R/data/NHTS.rds")
@@
 ```{r count, eval = TRUE, echo = FALSE}
-listOfLists <- lapply(split(NHTS, NHTS$HHSTATE),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
+listOfLists <- lapply(split(NHTS, NHTS$MSA_ID),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
+#listOfLists <- lapply(split(NHTS, NHTS$HHSTATE),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
 counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
@@
 p <- ggplot(foo, aes( x = state, y = sampleSize))
-p + geom_bar(stat = "identity") + scale_y_log10() + labs(y = "Sample Size (Referent Group)", x = "State") + theme_bw()
+p + geom_bar(stat = "identity") + labs(y = "Sample Size (Referent Group)", x = "State") + theme_bw()
 ```

### 2017-05-09 R/NHTS.Rmd
Created NHTS.Rmd
Also split on MSA instead of state

# Integrated Transport and Health Impacts Model

## R Package _ITHIM_

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
#install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
document("~/ITHIM/")
install("~/ITHIM/")
#install_github("syounkin/ITHIM", ref="devel")
install_github("smbache/magrittr", ref="master")

library("magrittr")
library("ITHIM")

fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

## HHS Metro Region Parameter Sets

First we use NHTS to estimate mean walking and cycling times.

### NHTS

```{r compute, echo = FALSE, eval = TRUE}
date()
compute <- TRUE
```

#### Create a handful of functions to help parse data

```{r parsely, eval = TRUE, echo = FALSE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

countPeople <- function(data){
 nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
}

tripTime <- function(data, mode = "walk"){
 return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
}

reformatData <- function(data){
foo <- matrix(unlist(data), nrow = 7, ncol = 2)
foo <- rbind( c(0, 0), foo)
dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
return(foo[,c("M","F")])
}

```

#### Read Two of the Four NHTS Data Tables

```{r NHTS, eval = compute, echo = FALSE}
DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")

NHTS.trips <- with(DAYV2PUB,{
    data.frame(HOUSEID = HOUSEID,
                   PERSONID = PERSONID,
                   TDCASEID = TDCASEID,
                   AT = ifelse(TRPTRANS == "22", "cycle",
                        ifelse(TRPTRANS == "23", "walk",
                        ifelse(!is.na(TRPTRANS), "other", NA))),
                   TRVL_MIN = as.numeric(TRVL_MIN),
                   #TRVLCMIN = as.numeric(TRVLCMIN),
                   URBRUR  = as.factor(URBRUR),
                   MSACAT = as.factor(MSACAT),
                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                   TRPTRANS = as.factor(TRPTRANS),
                   TRPMILES = as.numeric(TRPMILES)
)})

PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")

NHTS.person <- with(PERV2PUB,{
                    data.frame(HOUSEID = HOUSEID,
                               PERSONID = PERSONID,
                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
                               AGE = as.factor(ITHIM:::convertToAgeClass(R_AGE)),
                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                               HHSTATE = as.character(HHSTATE),
                               HHS = as.character(convertToHHS(as.character(HHSTATE))),
                               MSA_ID = HHC_MSA, stringsAsFactors = FALSE)
})

```

#### Subset and Merge

We now correct for zero trip times in NHTS.

```{r subset, eval = compute, fig.width = 10, echo = FALSE}

NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))

NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
saveRDS(NHTS, file = "./data/NHTS.rds")

```{r msaMatrix, eval = compute, fig.width = 10, echo = FALSE}
msaMatrix <- matrix(c(5602,4472,1922,7320,3120,6200,3362,7362,5720,4992,8872,
8280,7240,0640,6922,5960,3600,0520,8960,3480,1602,6840,
1280,1520,5360,6640,1122,5082,6162,2162,4920,7602,6480,
5120,1692,7040,7160,6442,4120,4520,6280,2082,1642,3760,
3280,1840,5880,3000,5560,
"NY/NJ Metro","Los Angeles-Riverside-Orange County","Dallas-Fort Worth","San Diego","Greensboro-Winston Salem", "Phoenix/Mesa","Houston-Galveston-Brazoria", "San Francisco-Oakland-San Jose", "Norfolk-Virginia Beach", "Miami-Ft. Lauderdale", "Washington-Baltimore", rep("no label yet", 38)), ncol = 2)
```

### 2017-05-09 R/NHTS.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("NHTS.Rmd","NHTS.html")

### 2017-05-17 R/NHTS.Rmd
Updates to NHTS data formatting
Now uses TravelSurvey package.

@@
 library("dplyr")
+library("TravelSurvey")
 ```
@@
 
+#saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
+#saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")
+
 NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
@@
 ```
+
+```{r TravelSurvey, eval = compute, fig.width = 10, echo = FALSE}
+
+person.df <- with(NHTS.person,{
+    data.frame(houseID=HOUSEID, subjectID=PERSONID, sex = SEX, age = AGE, location = HHS)
+    })
+
+trip.df <- with(NHTS.trips,{
+    data.frame(houseID=HOUSEID, subjectID=PERSONID, duration = TRVL_MIN, mode = AT)
+})
+
+person.df <- subset(person.df,apply(person.df,MARGIN=1,FUN=function(x)all(!is.na(x))))
+trip.df <- subset(trip.df,apply(trip.df,MARGIN=1,FUN=function(x)all(!is.na(x))))
+
+NHTS.TravelSurvey <- new("TravelSurvey", person = person.df, trip = trip.df)
+saveRDS(NHTS.TravelSurvey, file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
+
+NHTS.df <- dplyr::left_join(NHTS.TravelSurvey@person, NHTS.TravelSurvey@trip, by = c("houseID","subjectID"))
+
+NHTS.df <- within(NHTS.df, duration <- ifelse(is.na(duration), 0, duration))
+
+```

### 2017-05-22 R/NHTS.Rmd
Created TravelSurvey Rmd
I am breaking ITHIM.Rmd into smaller pieces that can be managed with the TravelSurvey class

@@
 ```
-
-#### Create a handful of functions to help parse data
 
@@
                    TRVL_MIN = as.numeric(TRVL_MIN),
-                   #TRVLCMIN = as.numeric(TRVLCMIN),
                    URBRUR  = as.factor(URBRUR),
@@
 
-#### Subset and Merge
-
-We now correct for zero trip times in NHTS.
-
-```{r subset, eval = compute, fig.width = 10, echo = FALSE}
-
-NHTS.person <- subset( NHTS.person, MSA == TRUE )
-NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))
-
-#saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
-#saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")
-
-NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
-NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
-saveRDS(NHTS, file = "./data/NHTS.rds")
-
-```{r msaMatrix, eval = compute, fig.width = 10, echo = FALSE}
-msaMatrix <- matrix(c(5602,4472,1922,7320,3120,6200,3362,7362,5720,4992,8872,
-8280,7240,0640,6922,5960,3600,0520,8960,3480,1602,6840,
-1280,1520,5360,6640,1122,5082,6162,2162,4920,7602,6480,
-5120,1692,7040,7160,6442,4120,4520,6280,2082,1642,3760,
-3280,1840,5880,3000,5560,
-"NY/NJ Metro","Los Angeles-Riverside-Orange County","Dallas-Fort Worth","San Diego","Greensboro-Winston Salem", "Phoenix/Mesa","Houston-Galveston-Brazoria", "San Francisco-Oakland-San Jose", "Norfolk-Virginia Beach", "Miami-Ft.
-Lauderdale", "Washington-Baltimore", rep("no label yet", 38)), ncol = 2)
-```
+#### Create TravelSurvey Object
 
@@
 person.df <- with(NHTS.person,{
-    data.frame(houseID=HOUSEID, subjectID=PERSONID, sex = SEX, age = AGE, location = HHS)
+    data.frame(houseID=HOUSEID, subjectID=PERSONID, sex = SEX, age = AGE, location = MSA_ID)
     })
@@
 NHTS.TravelSurvey <- new("TravelSurvey", person = person.df, trip = trip.df)
+
 saveRDS(NHTS.TravelSurvey, file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
+```
+
+```{r TravelSurvey2, eval = compute, fig.width = 10, echo = FALSE}
+NHTS.TravelSurvey <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
 
@@
 ```
+
+```{r msaMatrix, eval = FALSE, fig.width = 10, echo = FALSE}
+msaMatrix <- matrix(c(5602,4472,1922,7320,3120,6200,3362,7362,5720,4992,8872,
+8280,7240,0640,6922,5960,3600,0520,8960,3480,1602,6840,
+1280,1520,5360,6640,1122,5082,6162,2162,4920,7602,6480,
+5120,1692,7040,7160,6442,4120,4520,6280,2082,1642,3760,
+3280,1840,5880,3000,5560,
+"NY/NJ Metro","Los Angeles-Riverside-Orange County","Dallas-Fort Worth","San Diego","Greensboro-Winston Salem", "Phoenix/Mesa","Houston-Galveston-Brazoria", "San Francisco-Oakland-San Jose", "Norfolk-Virginia Beach", "Miami-Ft.
+Lauderdale", "Washington-Baltimore", rep("no label yet", 38)), ncol = 2)
+```
+
+```{r foob, eval = FALSE, echo = FALSE}
+#NHTS.person <- subset( NHTS.person, MSA == TRUE )
+#NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))
+
+#saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
+#saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")
+
+#NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
+#NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
+#saveRDS(NHTS, file = "./data/NHTS.rds")
+```

### 2017-05-22 R/TravelSurvey.Rmd
Created TravelSurvey Rmd
I am breaking ITHIM.Rmd into smaller pieces that can be managed with the TravelSurvey class

# Travel Survey Exploratory Analysis

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
library("tidyverse")
library("ITHIM")
library("TravelSurvey")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
document("~/ITHIM/")
install("~/ITHIM/")
#install_github("syounkin/ITHIM", ref="devel")
install_github("smbache/magrittr", ref="master")

library("magrittr")
library("ITHIM")

fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

```{r TravelSurvey, eval = compute, fig.width = 10, echo = FALSE}
NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
NHTS.df <- dplyr::left_join(NHTS.TravelSurvey@person, NHTS.TravelSurvey@trip, by = c("houseID","subjectID"))
NHTS.df <- within(NHTS.df, {
    duration <- ifelse(is.na(duration), 0, duration)
    mode <- factor(mode, levels = c("walk","cycle","other"))
}
)

```

## Mode share of walk/bike by MSA

```{r plot1, eval = TRUE, fig.width = 20, echo = FALSE}
total <- NHTS.df %>% group_by(location) %>% summarise(total = sum(duration) )
byMode <- NHTS.df %>% group_by(location, mode) %>% summarise(byMode = sum(duration) )
NHTS.table1 <- dplyr::full_join(byMode, total, by = "location") %>% mutate(propMode = byMode/total) %>% dplyr::filter(mode %in% c("walk","cycle"))

p <- ggplot(data = NHTS.table1, aes(x = location, y = propMode, fill = mode))
p + geom_bar(stat="identity") + theme_bw()
```

## Distribution of Trip Duration by MSA

```{r plot2, eval = TRUE, fig.width = 20, echo = FALSE}
p <- ggplot(subset(NHTS.df,!is.na(mode)), aes(x = location, y = duration, fill = location))
p + geom_boxplot(outlier.shape = NA) + ylim(0,60)
p + geom_violin(scale = "width" ) + ylim(0,60)
```

### Distribution of Trip Duration for NY MSA (5602)

```{r plot3, eval = TRUE, fig.width = 10, echo = FALSE}
p <- ggplot(subset(NHTS.df,!is.na(mode) & location == "5602"), aes(x = mode, y = duration, fill = mode))
p + geom_boxplot(outlier.shape = NA)
p + geom_violin(scale = "width", adjust = 4 )
```

### 2017-05-22 R/TravelSurvey.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("TravelSurvey.Rmd","TravelSurvey.html")

### 2017-05-22 R/TravelSurvey.Rmd
Updates to TravelSurvey.Rmd

@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-document("~/ITHIM/")
-install("~/ITHIM/")
-#install_github("syounkin/ITHIM", ref="devel")
-install_github("smbache/magrittr", ref="master")
-
-library("magrittr")
-library("ITHIM")
-
 fig.width <- 6 # inches
@@
 NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
-NHTS.df <- dplyr::left_join(NHTS.TravelSurvey@person, NHTS.TravelSurvey@trip, by = c("houseID","subjectID"))
+NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
 NHTS.df <- within(NHTS.df, {
@@
 
-```{r plot3, eval = TRUE, fig.width = 10, echo = FALSE}
+```{r plot3, eval = TRUE, echo = FALSE, fig.width = 20, fig.asp = 1/phi}
 p <- ggplot(subset(NHTS.df,!is.na(mode) & location == "5602"), aes(x = mode, y = duration, fill = mode))

### 2017-05-23 R/NHTS.Rmd
Updates to TravelSurvey.Rmd and NHTS.Rmd

@@
-# Integrated Transport and Health Impacts Model
-
-## R Package _ITHIM_
+# U.S. National Household Travel Surevy
 
@@
 #install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
-library("ggplot2")
+library("tidyverse")
 library("devtools")
-library("reshape2")
-library("plyr")
-library("dplyr")
+#library("reshape2")
+#library("plyr")
+#library("dplyr")
 library("TravelSurvey")
-```
-
-```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-document("~/ITHIM/")
-install("~/ITHIM/")
-#install_github("syounkin/ITHIM", ref="devel")
-install_github("smbache/magrittr", ref="master")
-
-library("magrittr")
 library("ITHIM")
-
 fig.width <- 6 # inches
@@
 ```
-
-## HHS Metro Region Parameter Sets
-
-First we use NHTS to estimate mean walking and cycling times.
-
-### NHTS
 
@@
 
-```{r parsely, eval = TRUE, echo = FALSE}
-convertToHHS <- function(states){
-  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
-  StateToHHS <- HHSRegions[,3]
-  names(StateToHHS) <- HHSRegions[,1]
-  return(StateToHHS[states])
-}
+### Read Two of the Four NHTS Data Tables
 
-countPeople <- function(data){
- nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
-}
-
-tripTime <- function(data, mode = "walk"){
- return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
-}
-
-reformatData <- function(data){
-foo <- matrix(unlist(data), nrow = 7, ncol = 2)
-foo <- rbind( c(0, 0), foo)
-dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
-return(foo[,c("M","F")])
-}
-
-```
-
-#### Read Two of the Four NHTS Data Tables
-
-```{r NHTS, eval = compute, echo = FALSE}
+```{r NHTS, eval = compute, echo = TRUE}
 DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
@@
                                HHSTATE = as.character(HHSTATE),
-                               HHS = as.character(convertToHHS(as.character(HHSTATE))),
+                               # HHS = as.character(convertToHHS(as.character(HHSTATE))),
                                MSA_ID = HHC_MSA, stringsAsFactors = FALSE)
@@
 
-#### Create TravelSurvey Object
+### Create TravelSurvey Object
 
-```{r TravelSurvey, eval = compute, fig.width = 10, echo = FALSE}
+```{r TravelSurvey, eval = compute, fig.width = 10, echo = TRUE}
 
@@
 ```
-
-```{r TravelSurvey2, eval = compute, fig.width = 10, echo = FALSE}
-NHTS.TravelSurvey <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
-
-NHTS.df <- dplyr::left_join(NHTS.TravelSurvey@person, NHTS.TravelSurvey@trip, by = c("houseID","subjectID"))
-
-NHTS.df <- within(NHTS.df, duration <- ifelse(is.na(duration), 0, duration))
-
-```
-
-```{r msaMatrix, eval = FALSE, fig.width = 10, echo = FALSE}
-msaMatrix <- matrix(c(5602,4472,1922,7320,3120,6200,3362,7362,5720,4992,8872,
-8280,7240,0640,6922,5960,3600,0520,8960,3480,1602,6840,
-1280,1520,5360,6640,1122,5082,6162,2162,4920,7602,6480,
-5120,1692,7040,7160,6442,4120,4520,6280,2082,1642,3760,
-3280,1840,5880,3000,5560,
-"NY/NJ Metro","Los Angeles-Riverside-Orange County","Dallas-Fort Worth","San Diego","Greensboro-Winston Salem", "Phoenix/Mesa","Houston-Galveston-Brazoria", "San Francisco-Oakland-San Jose", "Norfolk-Virginia Beach", "Miami-Ft.
-Lauderdale", "Washington-Baltimore", rep("no label yet", 38)), ncol = 2)
-```
-
-```{r foob, eval = FALSE, echo = FALSE}
-#NHTS.person <- subset( NHTS.person, MSA == TRUE )
-#NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))
-
-#saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
-#saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")
-
-#NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
-#NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
-#saveRDS(NHTS, file = "./data/NHTS.rds")
-```

### 2017-05-23 R/TravelSurvey.Rmd
Updates to TravelSurvey.Rmd and NHTS.Rmd

@@
 ```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
-options(repos='http://cran.rstudio.com/')
 library("tidyverse")
@@
     mode <- factor(mode, levels = c("walk","cycle","other"))
-}
-)
-
+})
 ```
@@
 ```{r plot1, eval = TRUE, fig.width = 20, echo = FALSE}
-total <- NHTS.df %>% group_by(location) %>% summarise(total = sum(duration) )
-byMode <- NHTS.df %>% group_by(location, mode) %>% summarise(byMode = sum(duration) )
-NHTS.table1 <- dplyr::full_join(byMode, total, by = "location") %>% mutate(propMode = byMode/total) %>% dplyr::filter(mode %in% c("walk","cycle"))
+total <- NHTS.df %>% group_by(location) %>% summarise(mean = mean(duration), total = sum(duration), sdTotal = sd(duration), nTotal = n())
+
+byMode <- NHTS.df %>% group_by(location, mode) %>% summarise(byMode = mean(duration), sdMode = sd(duration), nMode=n() )
+NHTS.table1 <- dplyr::full_join(byMode, total, by = "location") %>% mutate(propMode = byMode/total) %>% dplyr::filter(mode %in% c("walk","cycle")) %>% mutate(ciModeLeft = byMode - 1.96*sdMode/sqrt(nMode), ciModeRight = byMode + 1.96*sdMode/sqrt(nMode))
 
@@
 p + geom_bar(stat="identity") + theme_bw()
+
+NHTS.table2 <- NHTS.table1 %>% select(location, mode, ciModeLeft, ciModeRight) %>% gather(ciMode, value, 3:4)
+NHTS.table2 <- within(NHTS.table2, {
+    location <- factor(location, levels=unique(location[order(value)]))
+    })
+
+p <- ggplot(NHTS.table2, aes(x = location, y = value, colour = mode))
+p + geom_line(aes(group = ciMode))# + facet_grid(mode ~ . )
+
 ```

### 2017-05-23 R/appendix.Rmd
Updates to TravelSurvey.Rmd and NHTS.Rmd

@@
 ### Appendix
+### Appendix
+
+```{r msaMatrix, eval = FALSE, fig.width = 10, echo = FALSE}
+msaMatrix <- matrix(c(5602,4472,1922,7320,3120,6200,3362,7362,5720,4992,8872,
+8280,7240,0640,6922,5960,3600,0520,8960,3480,1602,6840,
+1280,1520,5360,6640,1122,5082,6162,2162,4920,7602,6480,
+5120,1692,7040,7160,6442,4120,4520,6280,2082,1642,3760,
+3280,1840,5880,3000,5560,
+"NY/NJ Metro","Los Angeles-Riverside-Orange County","Dallas-Fort Worth","San Diego","Greensboro-Winston Salem", "Phoenix/Mesa","Houston-Galveston-Brazoria", "San Francisco-Oakland-San Jose", "Norfolk-Virginia Beach", "Miami-Ft.
+Lauderdale", "Washington-Baltimore", rep("no label yet", 38)), ncol = 2)
+```
+
+```{r foob, eval = FALSE, echo = FALSE}
+#NHTS.person <- subset( NHTS.person, MSA == TRUE )
+#NHTS.trips <- subset(NHTS.trips, !is.na(TRVL_MIN))
+
+#saveRDS(NHTS.person, file = "./data/NHTS.person.rds")
+#saveRDS(NHTS.trips, file = "./data/NHTS.trips.rds")
+
+#NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"), all.x = TRUE)
+#NHTS <- within(NHTS, TRVL_MIN <- ifelse(is.na(TRVL_MIN), 0, TRVL_MIN))
+#saveRDS(NHTS, file = "./data/NHTS.rds")
+```
+
+```{r parsely, eval = FALSE, echo = FALSE}
+convertToHHS <- function(states){
+  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
+  StateToHHS <- HHSRegions[,3]
+  names(StateToHHS) <- HHSRegions[,1]
+  return(StateToHHS[states])
+}
+
+countPeople <- function(data){
+ nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
+}
+
+tripTime <- function(data, mode = "walk"){
+ return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
+}
+
+reformatData <- function(data){
+foo <- matrix(unlist(data), nrow = 7, ncol = 2)
+foo <- rbind( c(0, 0), foo)
+dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
+return(foo[,c("M","F")])
+}
+
+```
 

### 2017-05-24 R/TravelSurvey.Rmd
Updates to TravelSurvey.Rmd

@@
 
-```{r TravelSurvey, eval = compute, fig.width = 10, echo = FALSE}
+```{r TravelSurvey, eval = compute, fig.width = fig.width, echo = FALSE}
 NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
@@
 
-```{r plot1, eval = TRUE, fig.width = 20, echo = FALSE}
+```{r plot1, eval = TRUE, fig.width = fig.width, echo = FALSE, fig.asp = 1/phi}
 total <- NHTS.df %>% group_by(location) %>% summarise(mean = mean(duration), total = sum(duration), sdTotal = sd(duration), nTotal = n())
@@
 p <- ggplot(NHTS.table2, aes(x = location, y = value, colour = mode))
-p + geom_line(aes(group = ciMode))# + facet_grid(mode ~ . )
+p + geom_line(aes(group = ciMode)) + facet_grid(mode ~ . )
 
@@
 
-```{r plot2, eval = TRUE, fig.width = 20, echo = FALSE}
+```{r plot2, eval = TRUE, fig.width = fig.width, echo = FALSE, fig.asp = 1/phi}
 p <- ggplot(subset(NHTS.df,!is.na(mode)), aes(x = location, y = duration, fill = location))
@@
 
-```{r plot3, eval = TRUE, echo = FALSE, fig.width = 20, fig.asp = 1/phi}
+```{r plot3, eval = TRUE, echo = FALSE, fig.width = fig.width, fig.asp = 1/phi}
 p <- ggplot(subset(NHTS.df,!is.na(mode) & location == "5602"), aes(x = mode, y = duration, fill = mode))
-p + geom_boxplot(outlier.shape = NA)
-p + geom_violin(scale = "width", adjust = 4 )
+p + geom_boxplot(outlier.shape = NA) + ylim(0,60)
+p + geom_violin(scale = "width", adjust = 4 ) + ylim(0,60)
 ```

### 2017-05-30 R/TravelSurvey.Rmd
Updates to TravelSurvey.Rmd
Trying to incorporate MSA names
Having difficulty.  It's not working yet.

@@
 ```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
-fig.width <- 6 # inches
+fig.width <- 24 # inches
 phi <- (1+sqrt(5))/2 # the golden ratio ;)
@@
 
-```{r TravelSurvey, eval = compute, fig.width = fig.width, echo = FALSE}
+```{r msa, eval = TRUE, fig.width = fig.width, echo = FALSE}
+foo <- read.csv(file ="./../data/cbsatocountycrosswalk.csv", colClasses = "character")
+foo <- unique(foo[,c("msa","msaname")])
+msa <- foo$msa
+msaname <- foo$msaname
+names(msaname) <- msa
+
+```
+
+```{r TravelSurvey, eval = TRUE, fig.width = fig.width, echo = FALSE}
 NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
@@
     mode <- factor(mode, levels = c("walk","cycle","other"))
+    location <- msaname[location]
 })
@@
 p <- ggplot(data = NHTS.table1, aes(x = location, y = propMode, fill = mode))
-p + geom_bar(stat="identity") + theme_bw()
+p + geom_bar(stat="identity") + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 p <- ggplot(NHTS.table2, aes(x = location, y = value, colour = mode))
-p + geom_line(aes(group = ciMode)) + facet_grid(mode ~ . )
+p + geom_line(aes(group = ciMode)) + facet_grid(mode ~ . ) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 p <- ggplot(subset(NHTS.df,!is.na(mode)), aes(x = location, y = duration, fill = location))
-p + geom_boxplot(outlier.shape = NA) + ylim(0,60)
-p + geom_violin(scale = "width" ) + ylim(0,60)
+p + geom_boxplot(outlier.shape = NA) + ylim(0,60) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
+p + geom_violin(scale = "width" ) + ylim(0,60) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
@@
 p <- ggplot(subset(NHTS.df,!is.na(mode) & location == "5602"), aes(x = mode, y = duration, fill = mode))
-p + geom_boxplot(outlier.shape = NA) + ylim(0,60)
-p + geom_violin(scale = "width", adjust = 4 ) + ylim(0,60)
+p + geom_boxplot(outlier.shape = NA) + ylim(0,60) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
+p + geom_violin(scale = "width", adjust = 4 ) + ylim(0,60) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
+
+```{r means, eval = TRUE, echo = FALSE, fig.width = fig.width, fig.asp = 1/phi}
+meanDuration <- NHTS.df %>% group_by(location,sex,age,mode) %>% summarise(meanDuration = mean(duration))
+
+%>% complete(nesting(location,sex,age,mode))
+
+```

### 2017-06-01 R/TravelSurvey.Rmd
Update to Oregon Tutorial
and also TravelSurvey.Rmd

@@
 ```{r msa, eval = TRUE, fig.width = fig.width, echo = FALSE}
-foo <- read.csv(file ="./../data/cbsatocountycrosswalk.csv", colClasses = "character")
-foo <- unique(foo[,c("msa","msaname")])
+#foo <- read.csv(file ="./../data/cbsatocountycrosswalk.csv", colClasses = "character")
+foo <- read.csv(file = "~/GHI/data/MSA.csv", colClasses = c("character","character"))
+names(foo) <- c("msa","msaname")
+#foo <- unique(foo[,c("msa","msaname")])
 msa <- foo$msa
@@
 NHTS.df <- within(NHTS.df, {
+    id <- paste0(houseID,"-",subjectID)
     duration <- ifelse(is.na(duration), 0, duration)

### 2017-06-01 R/oregon.Rmd
Update to Oregon Tutorial
and also TravelSurvey.Rmd

--- 2017-03-16 R/tutorial.Rmd
@@
-# ITHIM Package Tutorial
+# ITHIMr Demonstration for Portland, Oregon
 
@@
 ```
-To begin using ITHIM one first must install it.  The package
-_devtools_ makes this is easy.  Install _devtools_ using
-`install.packages("devtools")` or some other way in _RStudio_.
 
@@
 
-We use the function `install_github` to install the _master_ branch of
-the _ITHIM_ package repository.
+The package _devtools_ makes installation of the _ITHIM_ package easy.
+Install _devtools_ using `install.packages("devtools")` or some other
+way in _RStudio_.  We use the function `install_github` from
+_devtools_ to install the _devel_ branch of the _ITHIM_ package
+repository.
 
@@
 library("devtools")
-install_github("syounkin/ITHIM", ref="master")
-library("ITHIM")
+install_github("ITHIM/ITHIM", ref="devel")
 ```
 
-## Creating a Baseline ITHIM object
+```{r libs2, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
+library("ITHIM")
+library("tidyverse")
+```
+
+```{r figwidth, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
+fig.width <- 24 # inches
+phi <- (1+sqrt(5))/2 # the golden ratio ;)
+```
+
+## Create a Baseline ITHIM object
 
@@
 
-For a description of the file format see the help page for
-`createITHIM` using `help("createITHIM")`.
+The file_activeTransprtTime.csv_ contains weekly mean travel times in
+minutes, stratified by age, sex and mode (walk,cycle). Currently eight
+age classes are used for the _ITHIM_ model.  For a more detailed
+description of the file format see the help page for `createITHIM`
+using `help("createITHIM")`.
 
@@
 
+Sample Files
 * [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
 * [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)
-
-### File Formats
-#### Active Transport File Format
-
-The active transport file contains mean walking and cycling times
-stratified by sex and _ITHIM_ age class.  The default file is located
-in _inst/_.
-
-```
-ageClass,sex,mode,value
-ageClass1,M,walk,21.47
-ageClass2,M,walk,53.11
-ageClass3,M,walk,50.13
-...
-ageClass6,F,cycle,1.53
-ageClass7,F,cycle,0.53
-ageClass8,F,cycle,0.23
-```
-
-Notes about file format:
-* The rows must be ordered by ageClass.
-* The mean values are in minutes per week
-
-#### Disease Burden File Format
-
-The disease burden file contains estimates for disease burden
-stratified by _ITHIM_ age class and sex.
-
-```
-disease,sex,ageClass,burdenType,value
-BreastCancer,M,ageClass1,dproj,0
-BreastCancer,M,ageClass2,dproj,0
-BreastCancer,M,ageClass3,dproj,0
-...
-RTIs,F,ageClass6,daly,49108.81151
-RTIs,F,ageClass7,daly,29852.3031
-RTIs,F,ageClass8,daly,15942.95826
-```
-
-Notes about file format:
-* RTIs stands for road traffic injuries
-* Only _ITHIM_ diseases can be included, namely "BreastCancer",
-  "ColonCancer", "CVD", "Dementia", "Diabetes", "Depression" and "RTIs"
-
-```{r misc22, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
-purl("./tutorial.Rmd","./tutorialScript.R", documentation = 0)
-```
-
-#### Road Injuries File Format
-
-If road injury data is present it can be included in the same manner.
-
-```{r misc123, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-roadInjuriesFile <- "~/ITHIM/inst/roadInjuries.csv"
-```
-
-```{r misc220, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
-                              GBDFile = GBDFile,
-                              roadInjuriesFile = roadInjuriesFile)
-```
-
-* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
-
-```
-severity,roadType,strikingMode,victimMode,value
-Fatal,Local,walk,walk,0
-Fatal,Local,cycle,walk,13
-Fatal,Local,bus,walk,213
-...
-Serious,Highway,HGV,NOV,0
-Serious,Highway,LGV,NOV,0
-Serious,Highway,mbike,NOV,0
-```
-
-Notes about file format:
-* ebikes are not included
 
@@
 
-## Built-In Example
+## Active Travel Estimates for Portland (MSA 6442)
+
+We use the NAtional Household Transportation Survey data stored as a _TravelSurvey_ R object to estimate mean walk and cycle times.
+
+```{r oregon23, eval = TRUE, echo = FALSE, message = FALSE, warnings = FALSE, results = "hide", fig.width = fig.width, fig.asp = 1/phi}
+install_github("syounkin/TravelSurvey", ref="master")
+library("TravelSurvey")
+NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
+NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
+```
+
+```{r oregon333, eval = TRUE, echo = FALSE, fig.width = fig.width, fig.asp = 1/phi}
+NHTS.df <- within(NHTS.df, {
+    id <- paste0(houseID,"-",subjectID)
+    duration <- ifelse(is.na(duration), 0, duration)
+    mode <- factor(mode, levels = c("walk","cycle","other"))
+    age <- factor(age, levels = c("00-04","05-14","15-29","30-44","45-59","60-69","70-79","80+"))
+    sex <- factor(sex, levels = c("M","F"))
+})
+```
+
+```{r oregon2, eval = TRUE, echo = TRUE, fig.width = fig.width, fig.asp = 1/phi}
+sampleSize <- NHTS.df %>% filter(location == "6442") %>% group_by(sex,age) %>% summarise(n=n_distinct(id))
+
+tripDuration <- NHTS.df %>% filter(location == "6442") %>% group_by(mode,sex,age) %>% summarise(total=sum(duration))
+
+activeTravelOregon <- right_join(sampleSize,tripDuration, by = c("sex","age")) %>% ungroup() %>% mutate(mean = total/n*7) %>% select(mode,sex,age,mean) %>% complete(mode,sex,age) %>% filter(mode %in% c("walk","cycle")) %>% as.data.frame()
+
+activeTravelOregon
+```
+
+## Appendix
+
+### Built-In Example
 
@@
 
-## Appendix
+### File Formats
+#### Active Transport File Format
+
+The active transport file contains mean walking and cycling times
+stratified by sex and _ITHIM_ age class.  The default file is located
+in _inst/_.
+
+```
+ageClass,sex,mode,value
+ageClass1,M,walk,21.47
+ageClass2,M,walk,53.11
+ageClass3,M,walk,50.13
+...
+ageClass6,F,cycle,1.53
+ageClass7,F,cycle,0.53
+ageClass8,F,cycle,0.23
+```
+
+Notes about file format:
+* The rows must be ordered by ageClass.
+* The mean values are in minutes per week
+
+#### Disease Burden File Format
+
+The disease burden file contains estimates for disease burden
+stratified by _ITHIM_ age class and sex.
+
+```
+disease,sex,ageClass,burdenType,value
+BreastCancer,M,ageClass1,dproj,0
+BreastCancer,M,ageClass2,dproj,0
+BreastCancer,M,ageClass3,dproj,0
+...
+RTIs,F,ageClass6,daly,49108.81151
+RTIs,F,ageClass7,daly,29852.3031
+RTIs,F,ageClass8,daly,15942.95826
+```
+
+Notes about file format:
+* RTIs stands for road traffic injuries
+* Only _ITHIM_ diseases can be included, namely "BreastCancer",
+  "ColonCancer", "CVD", "Dementia", "Diabetes", "Depression" and "RTIs"
+
+### Road Injuries
+#### Road Injuries File Format
+
+If road injury data is present it can be included in the same manner.
+
+```{r misc123, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+roadInjuriesFile <- "~/ITHIM/inst/roadInjuries.csv"
+```
+
+```{r misc220, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
+                              GBDFile = GBDFile,
+                              roadInjuriesFile = roadInjuriesFile)
+```
+
+* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
+
+```
+severity,roadType,strikingMode,victimMode,value
+Fatal,Local,walk,walk,0
+Fatal,Local,cycle,walk,13
+Fatal,Local,bus,walk,213
+...
+Serious,Highway,HGV,NOV,0
+Serious,Highway,LGV,NOV,0
+Serious,Highway,mbike,NOV,0
+```
+
+Notes about file format:
+* ebikes are not included
 
 ### Session Info & Package Test
-```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
+```{r sessionInfo, eval = FALSE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
 library("testthat")

### 2017-06-01 R/oregon.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("oregon.Rmd","oregon.html")

### 2017-06-01 R/oregon.Rmd
edits to Oregon Markdown

@@
 
-```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-activeTransportFile <- "~/ITHIM/inst/activeTransportTime.csv"
-GBDFile <- "~/ITHIM/inst/gbd.csv"
-```
+Sample Files
+* [activeTransportTime.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/activeTransportTime.csv)
+* [gbd.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/gbd.csv)
 
@@
 
+Please update the path and names in the code below to match the csv files you wish to use.
+
 ```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
-                              GBDFile = GBDFile)
+ITHIM.baseline <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime.csv",
+                              GBDFile = "~/Downloads/gbd.csv")
 ```
-
-Sample Files
-* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
-* [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)
 
@@
 parameters.
+(note: this is the overall mean for the whole population. From this, _ITHIM_ will create means for each age & gender strata)
 
@@
 Currently the list of parameters resides in the help page for the
-ITHIM class (`help("ITHIM")`).  Alternatively one can use
+ITHIM class (`help("ITHIM")`).  Alternatively one can use the method
 `getParameterNames`.
@@
 Now we need to retrieve the walking and cycling means from the
-baseline model.  To do so we use the method `getMeans`.  To find a
+baseline model.  To do so we use the method `getMeans`, which returns a vector of means.  To find a
 list of available methods use `getMethodsITHIM` with no arguments.
@@
 getMethodsITHIM()
+```
+
+We'll store the vector of means in a variable called _meanVec_ and then view that vector.
+```{r misc322, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
 meanVec <- getMeans(ITHIM.baseline)
 meanVec
-meanWalkingTime <- meanVec$walk; meanCyclingTime <- meanVec$cycle
+```
+
+We can also pull out the mode-specific means from that vector and assign them to their own variables.
+
+```{r misc3322, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+meanWalkingTime <- meanVec$walk
+meanCyclingTime <- meanVec$cycle
+```
+
+Next we'll create a list of the parameters we wish to change in our scenario.
+Recall, we're interested in a scenario that has double the mean walking time and double the mean cycling time of the baseline scenario.
+
+```{r misc3242, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
 scenarioParameters <- list(muwt = 2*meanWalkingTime, muct = 2*meanCyclingTime)
@@
 
-Now we use the list of parameters and the baseline ITHIM object to
-create the scenario ITHIM object.
+_ITHIM_'s comparative risk assessment framework uses two models to describe the potential health impacts for small differences between them.
+In our case the two models differ in the active travel times of the population.
+Above we created the baseline model, and now we use the list of changed parameters to `update` the baseline ITHIM object.
+This method creates a copy of the baseline ITHIM object with the changes we'd like to see in the new ITHIM object.
+We call this new object the scenario.
 
@@
 
-Let's inspect it to be sure.
+Let's inspect the scenario ITHIM object to be sure it includes the changes we expect.
 
@@
 Once the baseline and scenario ITHIM objects are defined, we may
-compute the change in burden due to physical activity as follows.
+compute the change in burden due to physical activity as follows. We provide a few examples to give you an idea of options for diseases and burden.
 
@@
 
-We use the NAtional Household Transportation Survey data stored as a _TravelSurvey_ R object to estimate mean walk and cycle times.
+We use the National Household Transportation Survey (NHTS) data stored as a _TravelSurvey_ R object to estimate mean walk and cycle times.
 
@@
 
-* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)
+* [roadInjuries.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/roadInjuries.csv)
 

### 2017-06-01 R/oregon.Rmd
Cleaned up edits from vargo a little bit
Removed RI bit from Oregon demo

@@
 files by simply updating the baseline object with a few new
-parameters.
-(note: this is the overall mean for the whole population. From this, _ITHIM_ will create means for each age & gender strata)
+parameters.  (note: this is the overall mean for the whole
+population. From this, _ITHIM_ will create means for each age & gender
+strata)
 
@@
 We'll store the vector of means in a variable called _meanVec_ and then view that vector.
+
 ```{r misc322, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
@@
 
-Next we'll create a list of the parameters we wish to change in our scenario.
-Recall, we're interested in a scenario that has double the mean walking time and double the mean cycling time of the baseline scenario.
+Next we'll create a list of the parameters we wish to change in our
+scenario. Recall,
+we're interested in a scenario that has double the mean walking time and double the mean cycling time of the baseline scenario.
 
@@
 Once the baseline and scenario ITHIM objects are defined, we may
-compute the change in burden due to physical activity as follows. We provide a few examples to give you an idea of options for diseases and burden.
+compute the change in burden due to physical activity as follows. We
+provide a few examples to give you an idea of options for diseases and
+burden.
 
@@
 
-We use the National Household Transportation Survey (NHTS) data stored as a _TravelSurvey_ R object to estimate mean walk and cycle times.
+We use the National Household Transportation Survey (NHTS) data stored
+as a _TravelSurvey_ R object to estimate mean walk and cycle times.
 
@@
 
-### Road Injuries
 #### Road Injuries File Format
-
-If road injury data is present it can be included in the same manner.
-
-```{r misc123, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-roadInjuriesFile <- "~/ITHIM/inst/roadInjuries.csv"
-```
-
-```{r misc220, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-ITHIM.baseline <- createITHIM(activeTransportFile = activeTransportFile,
-                              GBDFile = GBDFile,
-                              roadInjuriesFile = roadInjuriesFile)
-```
 
@@
 ### Session Info & Package Test
-```{r sessionInfo, eval = FALSE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
+```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
 library("testthat")

### 2017-06-02 R/oregon.Rmd
Updates to oregon.Rmd

@@
 
-```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+```{r misc320, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
 ITHIM.baseline <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime.csv",
@@
 
-## Creating Scenarios Using the `update` Method
+The contents of the _ITHIM_ object are displayed in summary when the
+object is printed.  See below.
 
-Suppose we wanted to create a scenario in which the mean cycling time
-and mean walking time are doubled.  We need only to change two means
-in the model and nothing else.  To do so we could avoid creating input
-files by simply updating the baseline object with a few new
-parameters.  (note: this is the overall mean for the whole
-population. From this, _ITHIM_ will create means for each age & gender
-strata)
+```{r show, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM.baseline
+```
 
-First we need to know the name of the parameters we want to update.
-Currently the list of parameters resides in the help page for the
+Currently the list of internal model parameters resides in the help page for the
 ITHIM class (`help("ITHIM")`).  Alternatively one can use the method
-`getParameterNames`.
+`getParameterNames` to display the names.
 
@@
 
-Now we need to retrieve the walking and cycling means from the
-baseline model.  To do so we use the method `getMeans`, which returns a vector of means.  To find a
-list of available methods use `getMethodsITHIM` with no arguments.
+If we wanted to examine the overal means in the model we could use the
+method `getMeans`, which returns a vector of means.  To find a list of
+available methods use `getMethodsITHIM` with no arguments.
 
@@
 
-We can also pull out the mode-specific means from that vector and assign them to their own variables.
+## Creating a Scenario ITHIM object
+
+Suppose we wanted to create a scenario in which the mean cycling time
+and mean walking time are doubled.  We simply need to replace the
+active transport file with one that has the desired values.
+
+Sample Files
+* [activeTransportTime_scenario.csv](https://ithim.wiscweb.wisc.edu/wp-content/uploads/sites/110/2017/06/activeTransportTime_scenario.csv)
+
+Please update the path and names in the code below to match the csv
+files you wish to use.
+
+```{r misc20, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+ITHIM.scenario <- createITHIM(activeTransportFile = "~/Downloads/activeTransportTime_scenario.csv",
+                              GBDFile = "~/Downloads/gbd.csv")
+```
+
+```{r misc7620, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
+modelMeans <- rbind(cbind(rbind(getWalkTime(ITHIM.baseline), getCycleTime(ITHIM.baseline), getNonTravelMETs(ITHIM.baseline)),vision = "baseline"), cbind(rbind(getWalkTime(ITHIM.scenario), getCycleTime(ITHIM.scenario), getNonTravelMETs(ITHIM.scenario)),vision = "scenario"))
+
+p <- ggplot(modelMeans, aes(x = ageClass, y = mean, fill = vision))
+p + geom_bar(stat = "identity", position = "dodge") + facet_grid(type ~ sex, scales = "free_y")  + theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+## Physical Activity Component
+
+Once the baseline and scenario ITHIM objects are defined, we may
+compute the change in burden due to physical activity as follows. We
+provide a few examples to give you an idea of options for diseases and
+burden.
+
+```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "all", bur = "deaths")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "BreastCancer", bur = "deaths")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "deaths")
+deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "daly")
+```
+
+## Creating Scenarios Using the `update` Method
+
+In the above example we needed only to change two means in the model
+and nothing else.  To do so we could avoid creating input files by
+simply updating the baseline object with a few new parameters.  (note:
+this is the overall mean for the whole population. From this, _ITHIM_
+will create means for each age & gender strata)
+
+We can use the mode-specific means from the _meanVec_ vector and
+assign them to their own variables.
 
@@
 Next we'll create a list of the parameters we wish to change in our
-scenario. Recall,
-we're interested in a scenario that has double the mean walking time and double the mean cycling time of the baseline scenario.
+scenario. Recall, we're interested in a scenario that has double the
+mean walking time and double the mean cycling time of the baseline
+scenario.
 
@@
 
-## Physical Activity Component
+## Creating an ITHIM object for Portland (MSA 6442)
 
-Once the baseline and scenario ITHIM objects are defined, we may
-compute the change in burden due to physical activity as follows. We
-provide a few examples to give you an idea of options for diseases and
-burden.
-
-```{r misc25, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
-deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "all", bur = "deaths")
-deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "BreastCancer", bur = "deaths")
-deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "deaths")
-deltaBurden(ITHIM.baseline, ITHIM.scenario, dis = "Depression", bur = "daly")
-```
-
-## Active Travel Estimates for Portland (MSA 6442)
-
+### Active Transport Means
 We use the National Household Transportation Survey (NHTS) data stored
@@
 activeTravelOregon
+```
+
+### Disease Burden
+
+```{r gbd, eval = TRUE, echo = TRUE, fig.width = fig.width, fig.asp = 1/phi}
 ```

### 2017-06-05 tomek/roadTest.R
Updated Tomek's road test code

@@ -1,49 +1,34 @@
 rm(list=ls())
 
-library("ITHIM") # Make sure you're on the devel branch
-
-# SafetyInNumbers and DistByRoadType Baseline data already included via default input files
+library("ITHIM")
 
 ITHIM.baseline <- createITHIM()
 
-# Scenario DistByRoadType
+source("~/ITHIM/tests/testthat/helper-RI-usage.R")
 
-ITHIM.scenario <- update(ITHIM.baseline, list(distRoadType = ITHIM:::readDistByRoadType(system.file("distByRoadTypeScenario.csv", package = "ITHIM"))))
+ITHIM.baseline <- update(ITHIM.baseline, list(distRoadType = ITHIM:::helperCreateArray(excelDistByRoadTypeBaselineWithoutNOV)$createdArray))
 
-# RoadInjuries for baseline
+ITHIM.baseline <- update(ITHIM.baseline, list(safetyInNumbers = ITHIM:::helperCreateArray(excelSafetyInNumbersWithoutNOV)$createdArray))
 
-ITHIM.baseline <- update(ITHIM.baseline, list(roadInjuries = ITHIM:::readRoadInjuries(system.file("roadInjuries.csv", package = "ITHIM"))))
 
-# RoadInjuries for scenario using scenario multiplier and baseline RoadInjuries
+ITHIM.scenario <- update(ITHIM.baseline, list(distRoadType = ITHIM:::helperCreateArray(excelDistByRoadTypeScenarioWithoutNOV)$createdArray))
+
+
+ITHIM.baseline <- update(ITHIM.baseline, list(roadInjuries = ITHIM:::helperCreateArray(excelRoadInjuriesBaselineWithoutNOV)$createdArray))
+
 
 ITHIM.scenario <- ITHIM:::updateRoadInjuries(ITHIM.baseline, ITHIM.scenario)
 
-## rm(list=ls())
-## library("ITHIM") # Make sure you're on the devel branch
 
-## ITHIM <- createITHIM()
 
-## # DistByRoadType
+# Make sure you're on the devel branch
 
-## ITHIM.baseline <- readDistByRoadType(ITHIM, system.file("distByRoadTypeBaseline.csv", package = "ITHIM"))
-## ITHIM.scenario <- readDistByRoadType(ITHIM, system.file("distByRoadTypeScenario.csv", package = "ITHIM"))
+  # init ITHIM baseline and scenario using data for helper
 
-## # SafetyInNumbers
[…]

### 2017-06-06 vargo/getF.R
Added Vargos census script
Renamed it getF.R

@@ -1,0 +1,122 @@
+library(jsonlite)
+library(tidyverse)
+
+getTractAgeSex <- function(state, county) {
+  varString <-
+    "B01001_003E,B01001_004E,B01001_005E,B01001_006E,B01001_007E,B01001_008E,B01001_009E,B01001_010E,B01001_011E,B01001_012E,B01001_013E,B01001_014E,B01001_015E,B01001_016E,B01001_017E,B01001_018E,B01001_019E,B01001_020E,B01001_021E,B01001_022E,B01001_023E,B01001_024E,B01001_025E,B01001_027E,B01001_028E,B01001_029E,B01001_030E,B01001_031E,B01001_032E,B01001_033E,B01001_034E,B01001_035E,B01001_036E,B01001_037E,B01001_038E,B01001_039E,B01001_040E,B01001_041E,B01001_042E,B01001_043E,B01001_044E,B01001_045E,B01001_046E,B01001_047E,B01001_048E,B01001_049E"
+
+  ACSpop <-
+    as.data.frame(fromJSON(
+      paste(
+        "https://api.census.gov/data/2015/acs5?get=NAME,",
+        varString,
+        "&for=tract:*&in=state:",
+        state,
+        "+county:",
+        county,
+        "&key=f78d6b6c18608edc379b5a06c55407ceb45e7038",
+        sep = ""
+      )
+    ), stringsAsFactors = FALSE)
+  ACSpop <- ACSpop[-1,]
+
+  colnames(ACSpop) <-
+    c(
+      "name",
+      "M_Under 5 years",
+      "M_5 to 9 years",
+      "M_10 to 14 years",
+      "M_15 to 17 years",
+      "M_18 and 19 years",
+      "M_20 years",
+      "M_21 years",
+      "M_22 to 24 years",
+      "M_25 to 29 years",
+      "M_30 to 34 years",
+      "M_35 to 39 years",
+      "M_40 to 44 years",
+      "M_45 to 49 years",
+      "M_50 to 54 years",
+      "M_55 to 59 years",
+      "M_60 and 61 years",
+      "M_62 to 64 years",
+      "M_65 and 66 years",
+      "M_67 to 69 years",
+      "M_70 to 74 years",
+      "M_75 to 79 years",
+      "M_80 to 84 years",
+      "M_85 years and over",
+      "F_Under 5 years",
+      "F_5 to 9 years",
+      "F_10 to 14 years",
+      "F_15 to 17 years",
[…]

### 2017-06-07 R/TravelSurvey.Rmd
Created msa.Rmd and updated TravelSurvey.Rmd

@@
     mode <- factor(mode, levels = c("walk","cycle","other"))
-    location <- msaname[location]
+    location <- location #msaname[location]
 })
@@
 
-### Distribution of Trip Duration for NY MSA (5602)
+### Distribution of Trip Duration for Portland, OR MSA (6442)
 
 ```{r plot3, eval = TRUE, echo = FALSE, fig.width = fig.width, fig.asp = 1/phi}
-p <- ggplot(subset(NHTS.df,!is.na(mode) & location == "5602"), aes(x = mode, y = duration, fill = mode))
+p <- ggplot(subset(NHTS.df,!is.na(mode) & location == "6442"), aes(x = mode, y = duration, fill = mode))
 p + geom_boxplot(outlier.shape = NA) + ylim(0,60) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

### 2017-06-07 R/msa.Rmd
Created msa.Rmd and updated TravelSurvey.Rmd

foo <- read.csv(file="~/GHI/data/join2.csv", stringsAsFactors = FALSE)

msaTest <- 6442

msaMatrix <- foo %>% filter(msaNHTS == msaTest) %>%
select(stateFIPS,countyFIPS)

getPopulation(state=msaMatrix$stateFIPS, county =
msaMatrix$countyFIPS)

### 2017-06-08 R/msa.Rmd
Created msa.Rmd and added msa to Makefile

@@
-foo <- read.csv(file="~/GHI/data/join2.csv", stringsAsFactors = FALSE)
+# MSA Population Size (F)
 
-msaTest <- 6442
+```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
+library("devtools")
+install_github("ITHIM/ITHIM", ref="devel")
+library("ITHIM")
+```
+## List of Population Size Matrices
 
-msaMatrix <- foo %>% filter(msaNHTS == msaTest) %>%
-select(stateFIPS,countyFIPS)
+```{r msa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
+msaIDs <- msaIDFile$msaNHTS %>% unique
+msaIDs <- msaIDs[!is.na(msaIDs)] %>% as.list
+names(msaIDs) <- paste0("MSA-NHTS-", msaIDs)
+populationList <- getPopulation(msaIDs)
 
-getPopulation(state=msaMatrix$stateFIPS, county =
-msaMatrix$countyFIPS)
+length(populationList)
+names(populationList)
+head(populationList, 3)
+```
+
+## Top Three MSAa
+
+Let's inspect the top three MSAs by population size.
+
+```{r msa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+totalPop <- unlist(lapply(populationList,sum))
+head(totalPop[order(totalPop, decreasing = TRUE)],3)
+msaIDFile %>% filter(msaNHTS %in% c(4472,5602,1602)) %>% select(msaName,msaNHTS) %>% unique
+```
+
+Notice the non-uniqueness of `msaNHTS` in the data.

### 2017-06-08 R/msa.R

@@ -1,0 +1,2 @@
+library("knitr")
+knit2html("msa.Rmd","msa.html")

### 2017-06-09 R/TravelSurvey.Rmd
Update to TravelSurvey.Rmd and msa.Rmd

@@
     mode <- factor(mode, levels = c("walk","cycle","other"))
-    location <- location #msaname[location]
+    location <- as.integer(as.character(location)) #msaname[location]
 })
+saveRDS(NHTS.df, file = "~/GHI/R/data/NHTS.df.rds")
 ```

### 2017-06-09 R/msa.Rmd
Update to TravelSurvey.Rmd and msa.Rmd

--- 2017-06-07 R/msa.Rmd
@@
-foo <- read.csv(file="~/GHI/data/join2.csv", stringsAsFactors = FALSE)
+# MSA Population Size (F)
 
-msaTest <- 6442
+```{r msalibs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
+library("devtools")
+install_github("ITHIM/ITHIM", ref="devel")
+library("ITHIM")
+```
+## List of Population Size Matrices
 
-msaMatrix <- foo %>% filter(msaNHTS == msaTest) %>%
-select(stateFIPS,countyFIPS)
+```{r msamsa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
+msaIDs <- msaIDFile$msaNHTS %>% unique
+msaIDs <- msaIDs[!is.na(msaIDs)] %>% as.list
+names(msaIDs) <- paste0("MSA-NHTS-", msaIDs)
+populationList <- getPopulation(msaIDs)
 
-getPopulation(state=msaMatrix$stateFIPS, county =
-msaMatrix$countyFIPS)
+length(populationList)
+names(populationList)
+head(populationList, 3)
+
+saveRDS(populationList, file = "~/GHI/R/data/populationList.rds")
+```
+
+## Top Three MSAs
+
+Let's inspect the top three MSAs by population size.
+
+```{r msa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+totalPop <- lapply(populationList,sum) %>% unlist
+totalPop[totalPop %>% order(., decreasing = TRUE)] %>% head(., 3)
+msaIDFile %>% filter(msaNHTS %in% c(4472,5602,1602)) %>% select(msaName,msaNHTS) %>% unique
+```
+
+It's L.A. New York and Chicago.  Seems right.
+
+Notice the non-uniqueness of `msaNHTS` in the data.
+
+## MSA ID check
+
+```{r msa3, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+msaIDs <- msaIDFile$msaNHTS %>% unique
+msaFromFile <- data.frame(msa = unlist(msaIDs)) %>% filter(!is.na(msa) & !(msa == -1)) %>% unlist
+
+NHTS.df <- readRDS(file = "~/GHI/R/data/NHTS.df.rds")
+msaIDs <- NHTS.df$location %>% unique
+msaFromNHTS <- data.frame(msa = unlist(msaIDs)) %>% filter(!is.na(msa) & !(msa == -1)) %>% unlist
+
+data.frame(msaFromFile[order(msaFromFile)],msaFromNHTS[order(msaFromNHTS)])
+```
+
+Sweet.

### 2017-06-13 R/NHTS.Rmd
Trying to incorporate MSA names
Having difficulty.  It's not working yet.
Created msa2.R
This is a temporary name.  This script creates the active transport mean list by MSA.  msa.Rmd creates the list for F, the population

@@
 options(repos='http://cran.rstudio.com/')
+rm(list=ls())
 #install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
@@
 
-PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
+PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9", stringsAsFactors = FALSE)
 names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
@@
                                # HHS = as.character(convertToHHS(as.character(HHSTATE))),
-                               MSA_ID = HHC_MSA, stringsAsFactors = FALSE)
+                               MSA_ID = as.character(as.integer(HHC_MSA)), stringsAsFactors = FALSE)
 })

### 2017-06-13 R/TravelSurvey.Rmd
Created msa2.R
This is a temporary name.  This script creates the active transport mean list by MSA.  msa.Rmd creates the list for F, the population

--- 2017-06-07 R/TravelSurvey.Rmd
@@
 ```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
+rm(list=ls())
 library("tidyverse")
@@
     mode <- factor(mode, levels = c("walk","cycle","other"))
-    location <- location #msaname[location]
+    location <- location
 })
+saveRDS(NHTS.df, file = "~/GHI/R/data/NHTS.df.rds")
 ```

### 2017-06-13 R/msa.Rmd
Created msa2.R
This is a temporary name.  This script creates the active transport mean list by MSA.  msa.Rmd creates the list for F, the population

--- 2017-06-07 R/msa.Rmd
@@
-foo <- read.csv(file="~/GHI/data/join2.csv", stringsAsFactors = FALSE)
+# MSA Population Size (F)
 
-msaTest <- 6442
+```{r msalibs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
+rm(list=ls())
+library("devtools")
+#install_github("ITHIM/ITHIM", ref="devel")
+install("~/ITHIM/")
+library("ITHIM")
+```
+## List of Population Size Matrices
 
-msaMatrix <- foo %>% filter(msaNHTS == msaTest) %>%
-select(stateFIPS,countyFIPS)
+```{r msamsa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
+msaIDs <- msaIDFile$msaNHTS %>% unique
+msaIDs <- msaIDs[!is.na(msaIDs)] %>% as.list
+names(msaIDs) <- msaIDs
+populationList <- getPopulation(msaIDs)
 
-getPopulation(state=msaMatrix$stateFIPS, county =
-msaMatrix$countyFIPS)
+length(populationList)
+names(populationList)
+head(populationList, 3)
+
+saveRDS(populationList, file = "~/GHI/R/data/populationList.rds")
+```
+
+## Top Three MSAs
+
+Let's inspect the top three MSAs by population size.
+
+```{r msa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+totalPop <- lapply(populationList,sum) %>% unlist
+totalPop[totalPop %>% order(., decreasing = TRUE)] %>% head(., 3)
+msaIDFile %>% filter(msaNHTS %in% c(4472,5602,1602)) %>% select(msaName,msaNHTS) %>% unique
+```
+
+It's L.A. New York and Chicago.  Seems right.
+
+Notice the non-uniqueness of `msaNHTS` in the data.

### 2017-06-13 R/msa2.R

@@ -1,0 +1,28 @@
+NHTS.df <- readRDS(file = "~/GHI/R/data/NHTS.df.rds")
+FList <- readRDS(file = "~/GHI/R/data/populationList.rds")
+
+formatMeanMatrix <- function(meanTable){
+    meanMatrix <- meanTable %>% select(age,sex,mean) %>% ungroup() %>% spread(sex,mean) %>% select(M,F) %>% as.matrix()
+    row.names(meanMatrix) <- paste0("ageClass",1:8)
+    return(meanMatrix)
+}
+
+msaIDs <- unique(NHTS.df$location)
+msaIDs <- as.character(msaIDs[msaIDs!="-1"])
+
+sampleSize <- NHTS.df %>% group_by(location,sex,age) %>% summarise(n=n_distinct(id))
+tripDuration <- NHTS.df %>% group_by(location,mode,sex,age) %>% summarise(total=sum(duration))
+foo <- right_join(sampleSize,tripDuration, by = c("location","sex","age")) %>% ungroup() %>% mutate(mean = total/n*7) %>% select(location,mode,sex,age,mean) %>% complete(location,mode,sex,age)
+
+meanByLocationByMode <- lapply(split(foo,foo$mode), function(x) split(x, x$location))
+
+MwtList <- lapply(meanByLocationByMode$walk, formatMeanMatrix)
+MctList <- lapply(meanByLocationByMode$cycle, formatMeanMatrix)
+RwtList <- lapply(MwtList, function(x) x/x[3,2])
+RctList <- lapply(MctList, function(x) x/x[3,2])
+
+meanWalkList <- mapply(function(x,y) sum(x*y/sum(y), na.rm = TRUE), MwtList[msaIDs], FList[msaIDs], SIMPLIFY = FALSE)
+meanCycleList <- mapply(function(x,y) sum(x*y/sum(y), na.rm = TRUE), MctList[msaIDs], FList[msaIDs], SIMPLIFY = FALSE)
+
+cbind(data.frame(walk = unlist(meanWalkList)), data.frame(cycle = unlist(meanCycleList)))
+

### 2017-06-13 R/NHTS.Rmd
Created moves-json.R
This script will parse JSON data from Moves for a study by Maggie

@@
 #install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
-library("tidyverse")
 library("devtools")
+library("TravelSurvey")
 #library("reshape2")
@@
 #library("dplyr")
-library("TravelSurvey")
+install("~/ITHIM/")
 library("ITHIM")
@@
 ```
+
+```{r TravelSurvey, eval = TRUE, fig.width = fig.width, echo = FALSE}
+NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
+NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
+NHTS.df <- within(NHTS.df, {
+    id <- paste0(houseID,"-",subjectID)
+    duration <- ifelse(is.na(duration), 0, duration)
+    mode <- factor(mode, levels = c("walk","cycle","other"))
+    location <- location
+})
+saveRDS(NHTS.df, file = "~/GHI/R/data/NHTS.df.rds")
+```

### 2017-06-13 R/TravelSurvey.Rmd
Created moves-json.R
This script will parse JSON data from Moves for a study by Maggie

@@
 phi <- (1+sqrt(5))/2 # the golden ratio ;)
-```
-
-```{r msa, eval = TRUE, fig.width = fig.width, echo = FALSE}
-#foo <- read.csv(file ="./../data/cbsatocountycrosswalk.csv", colClasses = "character")
-foo <- read.csv(file = "~/GHI/data/MSA.csv", colClasses = c("character","character"))
-names(foo) <- c("msa","msaname")
-#foo <- unique(foo[,c("msa","msaname")])
-msa <- foo$msa
-msaname <- foo$msaname
-names(msaname) <- msa
-
-```
-
-```{r TravelSurvey, eval = TRUE, fig.width = fig.width, echo = FALSE}
-NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
-NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
-NHTS.df <- within(NHTS.df, {
-    id <- paste0(houseID,"-",subjectID)
-    duration <- ifelse(is.na(duration), 0, duration)
-    mode <- factor(mode, levels = c("walk","cycle","other"))
-    location <- location
-})
-saveRDS(NHTS.df, file = "~/GHI/R/data/NHTS.df.rds")
 ```
@@
 ```
+
+```{r msa, eval = TRUE, fig.width = fig.width, echo = FALSE}
+#foo <- read.csv(file ="./../data/cbsatocountycrosswalk.csv", colClasses = "character")
+foo <- read.csv(file = "~/GHI/data/MSA.csv", colClasses = c("character","character"))
+names(foo) <- c("msa","msaname")
+#foo <- unique(foo[,c("msa","msaname")])
+msa <- foo$msa
+msaname <- foo$msaname
+names(msaname) <- msa
+
+```

### 2017-06-13 R/msa.Rmd
Created moves-json.R
This script will parse JSON data from Moves for a study by Maggie

@@
 
-```{r msamsa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+```{r msamsa, eval = FALSE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
 msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
@@
 names(msaIDs) <- msaIDs
-populationList <- getPopulation(msaIDs)
+FList <- getPopulation(msaIDs)
 
-length(populationList)
-names(populationList)
-head(populationList, 3)
+length(FList)
+names(FList)
+head(FList, 3)
 
-saveRDS(populationList, file = "~/GHI/R/data/populationList.rds")
+saveRDS(FList, file = "~/GHI/R/data/FList.rds")
+```
+
+```{r msamsa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+NHTS.df <- readRDS(file = "~/GHI/R/data/NHTS.df.rds")
+FList <- readRDS(file = "~/GHI/R/data/FList.rds")
+
+formatMeanMatrix <- function(meanTable){
+    meanMatrix <- meanTable %>% select(age,sex,mean) %>% ungroup() %>% spread(sex,mean) %>% select(M,F) %>% as.matrix()
+    row.names(meanMatrix) <- paste0("ageClass",1:8)
+    return(meanMatrix)
+}
+
+#msaIDs <- unique(NHTS.df$location)
+#msaIDs <- as.character(msaIDs[msaIDs!="-1"])
+
+sampleSize <- NHTS.df %>% group_by(location,sex,age) %>% summarise(n=n_distinct(id))
+tripDuration <- NHTS.df %>% group_by(location,mode,sex,age) %>% summarise(total=sum(duration))
+foo <- right_join(sampleSize,tripDuration, by = c("location","sex","age")) %>% ungroup() %>% mutate(mean = total/n*7) %>% select(location,mode,sex,age,mean) %>% complete(location,mode,sex,age)
+
+meanByLocationByMode <- lapply(split(foo,foo$mode), function(x) split(x, x$location))
+
+MwtList <- lapply(meanByLocationByMode$walk, formatMeanMatrix)
+MctList <- lapply(meanByLocationByMode$cycle, formatMeanMatrix)
+RwtList <- lapply(MwtList, function(x) x/x[3,2])
+RctList <- lapply(MctList, function(x) x/x[3,2])
+
+meanWalkList <- mapply(function(x,y) sum(x*y/sum(y), na.rm = TRUE), MwtList[msaIDs], FList[msaIDs], SIMPLIFY = FALSE)
+meanCycleList <- mapply(function(x,y) sum(x*y/sum(y), na.rm = TRUE), MctList[msaIDs], FList[msaIDs], SIMPLIFY = FALSE)
+
+cbind(data.frame(walk = unlist(meanWalkList)), data.frame(cycle = unlist(meanCycleList)))
+
 ```
@@
 ```{r msa2, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
-totalPop <- lapply(populationList,sum) %>% unlist
+totalPop <- lapply(FList,sum) %>% unlist
 totalPop[totalPop %>% order(., decreasing = TRUE)] %>% head(., 3)

### 2017-06-13 R/moves-json.R

@@ -1,0 +1,18 @@
+foo <- fromJSON("~/Downloads/moves-export (1).json")
+masterList <- foo[[1]]
+dfTotal <- data.frame()
+nDays <- length(masterList) - 1
+
+for(i in 1:nDays){
+    nSegments <- length(masterList[[i]]$segments)
+    for( j in 1:nSegments ){
+        nActivities <- length(masterList[[i]]$segments[[j]]$activities)
+        if( masterList[[i]]$segments[[j]]$type == "move" ){
+            for(k in 1:nActivities){
+                subList <- masterList[[i]]$segments[[j]]$activities[[k]]
+                df <- data.frame(activity = subList$activity, duration = subList$duration, distance = subList$distance)
+                dfTotal <- rbind(dfTotal, df)
+            }
+        }
+    }
+}

### 2017-06-14 R/moves-json.R
Updated moves-json.R

@@ -1,4 +1,10 @@
-foo <- fromJSON("~/Downloads/moves-export (1).json")
+# Connect and export JSON of your data from this site
+# http://moves-export.herokuapp.com/
+library(RJSONIO)
+library(tidyverse)
+
+rm(list=ls())
+foo <- fromJSON("/home/sgy/GHI/data/sgy-moves.json")
 masterList <- foo[[1]]
 dfTotal <- data.frame()
 nDays <- length(masterList) - 1
@@ -7,12 +13,41 @@ for(i in 1:nDays){
     nSegments <- length(masterList[[i]]$segments)
     for( j in 1:nSegments ){
         nActivities <- length(masterList[[i]]$segments[[j]]$activities)
-        if( masterList[[i]]$segments[[j]]$type == "move" ){
+        if( !is.null(masterList[[i]]$segments) && masterList[[i]]$segments[[j]]$type == "move" ){
             for(k in 1:nActivities){
                 subList <- masterList[[i]]$segments[[j]]$activities[[k]]
-                df <- data.frame(activity = subList$activity, duration = subList$duration, distance = subList$distance)
+                df <- data.frame(activity = subList$activity, duration = subList$duration, distance = subList$distance, starttime = subList$startTime)
                 dfTotal <- rbind(dfTotal, df)
             }
         }
     }
 }
+
+dfTotal <- dfTotal %>% separate(col = starttime, into = c("date", "time"), sep = "T") %>%  mutate(date = as.Date(date,"%Y%m%d")) %>%
+    mutate(dow = factor(weekdays(date),levels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")),
+           day = format(date, "%d"),
+           month = format(date, "%B"))
+
+
+D <- dfTotal %>% filter(activity != "trp") %>% group_by(date, activity) %>% summarise(dailyMiles=sum(duration)) %>% ungroup()
+
+p <- ggplot(D, aes(x=activity, y = dailyMiles))
+p + geom_boxplot()#  +   scale_size_area() + xlab("") + ylab("Total Distance (miles)") + ggtitle("")
+
+
+
+## ggplot(aes(x=duration/60, fill=factor(activity)))+ geom_histogram(position="dodge") + facet_grid(.~dow)
+
+
+## dfTotal %>% filter(activity != "trp") %>%
[…]

### 2017-06-28 R/msa.Rmd

@@
 
-```{r msamsa, eval = FALSE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
+```{r msamsa, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
 msaIDFile <- read.csv(system.file("join2.csv", package = "ITHIM"), stringsAsFactors = FALSE)
@@
 
-#msaIDs <- unique(NHTS.df$location)
-#msaIDs <- as.character(msaIDs[msaIDs!="-1"])
+msaIDs <- unique(NHTS.df$location)
+msaIDs <- as.character(msaIDs[msaIDs!="-1"])
 
@@
 
+msaID.portland <- "6442"
+FList[[msaID.portland]] %>% gather(sex,value,1:2) %>% mutate(ageClass=rep(rownames(FList[[msaID.portland]]),2)) %>% select(ageClass,sex,value) %>% write.csv(., file = "~/ITHIM/inst/F.portland.csv", quote = FALSE, row.names = FALSE)
+
+rbind(MwtList[[msaID.portland]] %>% as.data.frame() %>% gather(sex,value,1:2) %>% mutate(ageClass=rep(rownames(FList[[msaID.portland]]),2)) %>% select(ageClass,sex,value) %>% mutate(mode = "walk"), MctList[[msaID.portland]] %>% as.data.frame() %>% gather(sex,value,1:2) %>% mutate(ageClass=rep(rownames(FList[[msaID.portland]]),2)) %>% select(ageClass,sex,value) %>% mutate(mode = "cycle")) %>% select(ageClass,sex,mode,value) %>% write.csv(., file = "~/ITHIM/inst/activeTransport.portland.csv", quote = FALSE, row.names = FALSE)
+
 ```

### 2017-08-09 R/NHTS.Rmd
Removed TravelSurvey step
No longer dependent on the package TravelSurvey

@@
-# U.S. National Household Travel Surevy
+ U.S. National Household Travel Surevy
 
@@
                                # HHS = as.character(convertToHHS(as.character(HHSTATE))),
-                               MSA_ID = as.character(as.integer(HHC_MSA)), stringsAsFactors = FALSE)
+                               MSA_ID = as.character(as.integer(HHC_MSA)), stringsAsFactors = FALSE,
+                               HHFAMINC = as.factor(HHFAMINC))
 })
@@
 person.df <- with(NHTS.person,{
-    data.frame(houseID=HOUSEID, subjectID=PERSONID, sex = SEX, age = AGE, location = MSA_ID)
+    data.frame(houseID=HOUSEID, subjectID=PERSONID, sex = SEX, age = AGE, location = MSA_ID, income = HHFAMINC)
     })
@@
 
-NHTS.TravelSurvey <- new("TravelSurvey", person = person.df, trip = trip.df)
+#NHTS.TravelSurvey <- new("TravelSurvey", person = person.df, trip = trip.df)
 
-saveRDS(NHTS.TravelSurvey, file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
+#saveRDS(NHTS.TravelSurvey, file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
 ```
 
-```{r TravelSurvey, eval = TRUE, fig.width = fig.width, echo = FALSE}
-NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
-NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
+```{r TravelSurvey2, eval = TRUE, fig.width = fig.width, echo = FALSE}
+#NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
+
+#NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
+
+NHTS.df <- dplyr::left_join(person.df, trip.df, by = c("houseID","subjectID"))
 NHTS.df <- within(NHTS.df, {

### 2017-08-11 R/NHTS.Rmd
Added nwalktrp and nbiketrp to NHTS.df
These may be usefule for estimating the proportion of inactive travelers and f_AT

@@
- U.S. National Household Travel Surevy
+# U.S. National Household Travel Surevy
 
@@
                                MSA_ID = as.character(as.integer(HHC_MSA)), stringsAsFactors = FALSE,
-                               HHFAMINC = as.factor(HHFAMINC))
+                               HHFAMINC = as.factor(HHFAMINC),
+                               NWALKTRP = as.numeric(NWALKTRP),
+                               NBIKETRP = as.numeric(NBIKETRP))
 })
@@
 person.df <- with(NHTS.person,{
-    data.frame(houseID=HOUSEID, subjectID=PERSONID, sex = SEX, age = AGE, location = MSA_ID, income = HHFAMINC)
+    data.frame(houseID=HOUSEID, subjectID=PERSONID, sex = SEX, age = AGE, location = MSA_ID, income = HHFAMINC, nwalktrp = NWALKTRP, nbiketrp = NBIKETRP)
     })

### 2017-08-16 R/NHTS.Rmd
Updated NWALKTRP and NBIKETRP
Also removed NA constraint.

@@
 rm(list=ls())
-#install.packages(c("ggplot2","dplyr","openssl","httr", "git2r", "devtools", "roxygen2"))
-library("devtools")
-library("TravelSurvey")
-#library("reshape2")
-#library("plyr")
-#library("dplyr")
-install("~/ITHIM/")
-library("ITHIM")
+#library("devtools")
+#install("~/ITHIM/")
+#library("ITHIM")
 fig.width <- 6 # inches
@@
 
-PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9", stringsAsFactors = FALSE)
+PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = c("-7","-8","-9"), stringsAsFactors = FALSE)
 names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")
@@
                                HHFAMINC = as.factor(HHFAMINC),
-                               NWALKTRP = as.numeric(NWALKTRP),
-                               NBIKETRP = as.numeric(NBIKETRP))
+                               #NWALKTRP = as.numeric(if(NWALKTRP < 0, NA, NWALKTRP)),
+                               #NBIKETRP = as.numeric(if(NBIKETRP < 0, NA, NBIKETRP)))
+                               NWALKTRP = NWALKTRP,
+                               NBIKETRP = NBIKETRP)
 })
@@
 ```
-
-### Create TravelSurvey Object
 
@@
 
-person.df <- subset(person.df,apply(person.df,MARGIN=1,FUN=function(x)all(!is.na(x))))
-trip.df <- subset(trip.df,apply(trip.df,MARGIN=1,FUN=function(x)all(!is.na(x))))
+#person.df <- subset(person.df,apply(person.df,MARGIN=1,FUN=function(x)all(!is.na(x))))
+#trip.df <- subset(trip.df,apply(trip.df,MARGIN=1,FUN=function(x)all(!is.na(x))))
 
-#NHTS.TravelSurvey <- new("TravelSurvey", person = person.df, trip = trip.df)
-
-#saveRDS(NHTS.TravelSurvey, file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
 ```
@@
 ```{r TravelSurvey2, eval = TRUE, fig.width = fig.width, echo = FALSE}
-#NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
-
-#NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
-
 NHTS.df <- dplyr::left_join(person.df, trip.df, by = c("houseID","subjectID"))

### 2017-08-16 methodsPaper/methodsPaper.tex
Made a few changes to get running on new iMac
Created two new figures
Also created appendix.Rnw to store old R code.
Deleted sgy.R and created methodsPaper.tex

@@
-\documentclass[10pt,letterpaper]{article}
+\documentclass[10pt,letterpaper]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
@@
 \usepackage{url}
+\usepackage{mathtools}
+\usepackage{rotating}
 \input{../tex/sgy}
@@
 %% END MACROS SECTION
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 Neil Maizlish\textsuperscript{3},\ %\Yinyang},
+\ldots,\
 Jonathan Patz\textsuperscript{1, 2}%,3\textcurrency},
@@
 \section*{Abstract}
-\input{tex/abstract}
+
+Comparative risk assessments to quantify the health potential
+of active transportation, e.g., walking and cycling, have been
+performed in the United States on a limited basis. Current
+methodologies, which rely on extensive data collection and user
+expertise, have estimated health benefits for select metropolitan
+areas under a few alternative scenarios. These analyses have largely
+been conducted using the Integrated Transport Health Impacts Model
+(ITHIM) which was developed in the UK and has been adapted for use in
+the US, primarily California. Here we present an implementation of the
+active transportation component of ITHIM to model the health impacts
+of modified active transportation levels. Using the newly developed
+software package \package{} written in R, along with publicly
+available national datasets, we quantify the change in disability
+adjusted life years over a wide range of scenarios.  The R package
+implemention of ITHIM's physical activity component provides a robust
+environment for exploring the relationship between active
+transportation and positive health outcomes.  A web-based
+user-interface is available for users who want a first estimate for
+the impact active transportation has on their
+state/region.
+
+%\input{lib/abstract}
 
@@
 % \section*{Author Summary}
-% \input{tex/authorSummary}
+% \input{lib/authorSummary}
 
@@
 \section*{Introduction}
-\input{tex/introduction.tex}
+
+Regular exercise has been shown to provide health benefits in terms of
+chronic disease prevention \cite{warburton2006}, cognitive function
+\cite{hillman2008}, and overall well being and satisfaction with life
+\cite{maher2013}. Still only half of Americans obtain the recommended
+levels of aerobic activity of 2.5 hours per week. There is
+considerable variability in the attainment of this goal by location
+within the country. Supportive local built and policy environments are
+important determinants of population-level activity
+patterns. Particularly, as they facilitate and foster routine and
+moderate physical activity, such as walking, local decisions on design
+and policy can play a huge role in improving the nation's health.
+
+Given the constraints and complexity of local governance and
+infrastructure investments, land use, and human behavior there is a
+need for high quality and reliable data and tools for decision-making
+related to urban transportaiton systems. The lack of accurate and
+availbale estimates of the health impacts of such decisions has
+limited the effectiveness of efforts to encourage non-motorized
+transport in the US and in other countries. In the last decade new
+tools, applying comparative risk assessment methodologies, have made
+progress toward quantifying the multple health impacts of changes to
+the transportation system and behaviors.
+
+ITHIM is one such statistical model that integrates data on active
+transport, physical activity, fine particulate matter and greenhouse
+gas emissions to provide estimates for the proportional change in
+mortality and morbidity for given baseline and alternate travel
+scenarios. The model has been used to calculate the health impacts of
+walking and bicycling short distances usually traveled by car or
+driving low-emission automobiles \cite{woodcock2013,maizlish2013}.
+
+ITHIM uses a comparative risk assessment framework for the active
+transport component of the model.  We improve on the existing
+implementation of the active transport component by including the
+distribution of non-travel-related activity, improving numerical
+precision when computing the population attributable fraction and
+creating a simple user-interface for the custom-built \R{} package
+\package{}.  We present this implementation here by using it to
+estimate the relationship between overall, nationwide disease burden
+and active transportation time in the United States.
 
 \section*{Materials and Methods}
-\input{tex/materialsAndMethods}
+
+As with the original ITHIM implementation we, employed a comparative
+risk assessment (CRA) across scenarios defined by mean active
+transportation time, i.e., combined walking and cycling.  To do so we
+investigate the proportional change in national disease burden between
+the baseline model and a given alternative scenario, i.e., the
+population attributable fraction, $\af$.  The population attributable
+fraction is defined below and computed for each disease (ICD-10)
+included in the model, namely Breast Cancer (C50), Colon Cancer
+(C18-C21), Dementia (G30), Diabetes (E10-E14) and Cardiovascular
+disease (I10, I12, I15, I20-I25, I30-I31, I33, I40).  We also include
+the disability impacts of depression, though it is not listed as a
+cause of death in CDC's data. For readability we omit an index
+indicating disease.
+
+\begin{equation}
+\af_j = \frac{\int \! R_j(x)P(x) \, \mathrm{d}x  - \int \! R_j(x)Q(x) \,
+  \mathrm{d}x}{\int \!
+R_j(x)P(x) \, \mathrm{d}x} \approx 1 - \frac{{\displaystyle \sum_{i=1}^n} R_j(x_i^\prime)}{{\displaystyle \sum_{i=1}^n} R_j(x_i)}  = 1 - \delta_j % = 1 - \frac{{\displaystyle \sum_{i=1}^n} e^{-\alpha \sqrt{x_i^\prime}}}{{\displaystyle \sum_{i=1}^n} e^{-\alpha \sqrt{x_i}}}
+  \label{paf}
+\end{equation}
+
+Here $R(x)$ represents the risk when the individual has exposure $x$,
+physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
+density of individuals with exposure $x$, in the baseline and
+alternate scenarios, respectively.  The exposure variable, $x$, is
+modeled as the sum of two independent random variables, travel-related
+and non-travel-related physical activity.  $\mathbf{x}$ and
+$\mathbf{x}^\prime$ represent the quantiles for the exposure
+distribution in baseline and alternative scenarios, respectively.
+$\delta_j$ is the proportional burden for disease $j$ in the alternate
+scenario.
+
+\paragraph{Non-Travel-Related Physical Activity}
+
+\begin{equation}
+X_{ij}^{\mathrm{non-travel}} \sim \logNormalMath\left(\mu_{ij}, \gamma \mu_{ij}\right)
+\end{equation}
+
+Within each age-sex class the non-travel-related exposure is modeled
+with a \logNormal{} distribution.  The age-sex specific mean ratios,
+$r_{ij} = \frac{\mu_{ij}}{\mu_0}$ and $\gamma$ are estimated from the
+American Time Use Survey data \cite{ATUS}.
+
+\paragraph{Travel-Related Physical Activity}
+
+The distribution for travel-related exposure is estimated from the
+time spent walking or cycling, i.e., active transport time, which as
+in the original ITHIM model is assigned a \logNormal{} distribution
+with constant coefficient of variation across age-sex classes.
+Travel-related exposure is estimated from active transport time using
+assumptions about how much physical activity is required for cycling
+and walking.
+
+%To estimate the disease burden we first approximate the population
+%attributable fraction through a discretization of the integral in
+%equation \ref{paf}.
+
+\paragraph{Non-Travel-Related Physical Activity}
+
+With a simulated distribution for total physical activity in hand we
+compute the empirical percentiles in the baseline and alternate
+scenarios, $\mathbf{x}$ and $\mathbf{x}^\prime$. The proportional
+change in disease burden, $\delta$, may be approximated as follows,
+where $R$ is the risk function for total activity.
+
+%% \begin{equation}
+%% \delta = 1 - \af \approx \frac{{\displaystyle \sum_{i=1}^{n}} R(x_i)}{{\displaystyle \sum_{i=1}^{n}} R(x_i^\prime)}
+%% \end{equation}
+
+\paragraph{Estimation of Disease Burden}
+
+The disease burden, in this case DALYs, is found for each of the
+quantiles using equation \ref{burden} using $B_0$, the overall
+nation-wide burden for the baseline.  The overall burden is then
+computed as the sum across quintiles, $B=\sum_{j=1}^n B_j$.
+
+\begin{equation}
+B_j = \delta B_0\frac{\RR_j}{\sum_{k=1}^n \RR_j}\label{burden}
+\end{equation}
+
+\paragraph{Risk Functions}
+
+We assign risk based on dose-response curves for each of the diseases
+included in the model.
+
+\begin{equation}
+\RR_j(x) = \mathrm{e}^{-\alpha_j \tilde{x}}
+\end{equation}
+
+where $\tilde{x} = x^k$, $k=\frac{1}{2}$ and $r$ varies by age-sex
+class and disease.  For a table of $r$ values see supplementary
+material.
+
+\paragraph{Nationwide Health Benefits to Active Transport}
+
+For the purposes of this implementation we limited our analyses to
+metropolitan regions of the United States. To idenfity urban
+populations we used the National Center for Health Statistics 2013
+Urban-Rural Classification Scheme for Counties \cite{ingram2014}. We
+eliminated counties in the US with the classifcation of
+`nonmetropolitan' (noncore or micropolitan). We used this and similar
+classification schemes as well as county and state identifiers in
+other datsets (NHTS, ATUS, CDC WONDER) to ensure consistency in our
+estimates for input parameters.
+
+To test the new implementation of ITHIM's active transportation
+component, we used the US national metro population estimates for
+active travel time (walking and cycling) as the baseline. We obtained
+active travel time input parameters for alternate scenarios using
+states with large sample sizes. From these examples we were able to
+approximate low and high active travel scenarios. Finally, we applied
+these active travel times to the US metro population to perform the
+comparative risk assessment against the baseline US numbers. We then
+examined the sensitivity of disease burden estimates to input
+parameters including the active travel time and non-travel activity.
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig1}}
+    \caption{}\label{fig1}
+\end{figure}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig2.pdf}}
+    \caption{}\label{fig2}
+\end{figure}
+
+\begin{sidewaysfigure}[t]
+  \centerline{\includegraphics[width=\textwidth]{./figures/fig3.pdf}}
+  \caption{Estimates for disease burden nationwide in terms of thousands of
+    DALYs for each of the six diseases included in the model.
+  }\label{fig3}
+\end{sidewaysfigure}
+
+%\input{lib/materialsAndMethods}
 
@@
 
-\begin{figure}[t]
-  \centerline{\includegraphics[width=\textwidth]{./figures/fig1}}
-    \caption{}\label{dalyFigureQuintiles}
-\end{figure}
+We report total change (averted or increased) in diability adjusted
+life years (DALYs) as an estimate of disease burden change with
+alternative scenarios.
 
-%% \begin{figure}[t]
+To estimate DALYs we use a methodology adopted from previous US
+implementations of ITHIM which scale Global Budern of Disease
+Estimates for the US to smaller populations using mortality rate
+ratios \cite{maizlish2013}. Age, sex, and casuse specific mortalities
+were obtained form the CDC WONDER database for US metro counties for
+the years 2010-2014. Cells with mortality counts less than 10 deaths
+were suppressed for privacy and imputed (0 to 9) using R's random
+integer generator.
 
-%%     \centerline{\includegraphics[width=\textwidth]{./figures/fig2}}
-%%     \caption{}\label{dalyFigure}
-%% \end{figure}
+Our subset of the US population to only metropolitan counties
+consisted of 1,090 of 3,141 (34.7\%) counties, or approximately 272M
+of 312M (87\%) of the total US population. The age and gender
+distribution of our subset population was slightly younger (???making
+this up???) than the national population. Baseline active transport
+times (walking and cycling) for the US across age and sex groupings
+was found to be approximately 43 min/week (???). Baseline non-travel
+activity time was found to be ???X??? (min/week) from the ATUS for US
+metropolitan populations.
 
-%% \begin{figure}[t]
-%%   \centerline{\includegraphics[width=\textwidth]{./figures/fig4.pdf}}
-%%     \caption{}\label{California}
-%% \end{figure}
+Adapting ITHIM's methods to the R platform enables the rapid and
+thorough assessment of parameter influence within the model. Another
+important improvement enabled by the shift to a more computationaly
+robust platform was to examine methodological assumptions of former
+implementations. One critical assumption invovles the approximation of
+??? used to calculate the attirbutable fraction (see Figure
+\ref{fig:overview}). Previous implementations used quitiles to
+approximate this value, essentailly the area under the ???? curve. We
+improved upon those methods to use percentiles or 100 points to
+estimate the same values. We find our estimates to be much more stable
+and reliable at all values of active travel and non-travel activity
+parameters.
 
-\input{tex/results}
+As exhibited in Figure \ref{fig1}, the variability in disease burden
+benefits varies to a much greater degree when using fewer points in
+the attirbutable fraction approximation. Importantly, we see greater
+stability when using the package and percentiles to approximate
+attirbutable fraction.
+
+Comparing baseline to scenarios with increased active travel produced
+estimates of up to 360,000??? fewer DALYs (with 300\% increase in mean
+active travel time) for our US population subset. We then varied the
+mean non-travel activity time and active travel time to observe how
+this chaged results (\ref{fig1}). We observe the most dramatic
+responses in disease burden estimates to changes in active travel time
+when non-travel activity times are lowest. This result is consistent
+with the non-linear dose-response curve for physical activity and
+several health outcomes.
+
+%\input{lib/results}
 
 \section*{Discussion}
-\input{tex/discussion}
+%\input{lib/discussion}
 
 \section*{Conclusion}
-\input{tex/conclusion}
+%\input{lib/conclusion}
 
@@
 
-\input{tex/supportingInformation}
+%\input{lib/supportingInformation}
 

### 2017-08-16 misc/sgy.R

@@ -1,4 +1,0 @@
-DaneTravelMeansByTime <- getCountyMeansTravelByTime(state = 55,county = 025)
-DaneTravelMeansByTime <- within(DaneTravelMeansByTime, travelTime <- factor(travelTime, levels = c("minLT10", "min10to14", "min15to19","min20to24","min25to29","min30to34","min35to44","min45to59","minOver60")))
-DaneTravelMeansByTime <- within(DaneTravelMeansByTime, mode <- factor(mode, levels = c("drive","walk","bicycle","transit")))
-ggplot(DaneTravelMeansByTime, aes(x = travelTime, y = estimate, fill = mode)) + geom_bar(stat = "identity", position = "dodge") + theme_bw() + ylab("Number of Commuters") + xlab("Duration of Commute")

### 2017-10-17 methodsPaper/lib/abstract.Rnw
Update abstract.Rnw

@@
 Comparative risk assessments to quantify the health potential
-of active transportation, e.g., walking and cycling, have been
+of active transportation (e.g., walking and cycling) have been
 performed in the United States on a limited basis. Current
@@
 expertise, have estimated health benefits for select metropolitan
-areas under a few alternative scenarios. These analyses have largely
+areas under a few alternative scenarios. Such analyses have
 been conducted using the Integrated Transport Health Impacts Model
-(ITHIM) which was developed in the UK and has been adapted for use in
+(ITHIM), which was developed in the UK and has been adapted for use in
 the US, primarily California. Here we present an implementation of the
@@
 transportation and positive health outcomes.  A web-based
-user-interface is available for users who want a first estimate for
+user-interface based on the package is being developed for users who want a first estimate for
 the impact active transportation has on their

### 2017-10-18 methodsPaper/lib/introduction.Rnw
Update introduction.Rnw

@@
 important determinants of population-level activity
-patterns. Particularly, as they facilitate and foster routine and
-moderate physical activity, such as walking, local decisions on design
-and policy can play a huge role in improving the nation's health.
+patterns. Local decisions on design
+and policy can play a huge role in improving the nation's health,
+particularly, as they facilitate and foster routine and
+moderate physical activity, such as walking.
 
@@
 need for high quality and reliable data and tools for decision-making
-related to urban transportaiton systems. The lack of accurate and
+related to urban transportaiton systems. The absence of accurate and
 availbale estimates of the health impacts of such decisions has
 limited the effectiveness of efforts to encourage non-motorized
-transport in the US and in other countries. In the last decade new
+transport in the US and in other countries. Despite robust evidence showing 
+the mental and physical health benefits of physical activity, the quantification of health 
+benefits tied to proposed transportation funding and planning is not common. In the last decade new
 tools, applying comparative risk assessment methodologies, have made
